---
author: "Joshua Born"
#bibliography: "../bibliographies/myth-of-left-and-right.bib"
csl: "../bibliographies/chicago-author-date.csl"
date: 2023-11-09
draft: true
implicit-figures: true
include-before: |
  ## Table of Contents
output:
  blogdown::html_page:
    number_sections: false
    toc: true
    toc_depth: 5
summary: '[Put summary sentence here.]'
tags: ["analysis", "induced abortion", "politics", "first-person"]
title: 'Untitled Article'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)
```


## Introduction

I do not typically write much about abortion[^definition] politics. However, this past week there were elections in the Virginia, where I live, and in Ohio. Several Democrats won elections in Virginia after focusing their campaigns on keeping abortion as legal as possible, and a referendum labeled _"Issue 1"_ was passed in Ohio that essentially replicated the policy instituted by the _Roe vs. Wade_ and _Doe vs. Bolton_ decisions of the Supreme Court of the United States.

[^definition]: Unless otherwise noted, in this article the word "abortion" is used to mean induced abortion of pregnancy.

    Technically, the word "abortion" is a generic term. Any process that is aborted before it comes to completion can, in theory, be labeled "abortion." However, because of its association with abortion of pregnancy and the emotional weight of such occurrence, the word "abortion" is usually used to mean abortion of pregnancy.

    Furthermore, even if we just consider "abortion" to mean abortion of pregnancy, there is ambiguity because in the medical literature the word "abortion" is used to mean two different things: spontaneous abortion, which is commonly called "miscarriage" in the vernacular, occurs when a pregnancy terminates without anyone's intervention; induced abortion occurs when a pregnancy is terminated on purpose. When "abortion" is used in the vernacular it is commonly used to mean induced abortion.

    This ambiguity can lead to misinterpretation. For instance, if a study were to report on abortions in a given population, it could be including both spontaneous and induced abortions if it were using the medical literature definition, but it could be excluding what are commonly called "miscarriages" if it were using the common definition.


In the wake of this past week's elections, there have been opinion pieces interpreting the results as a resounding win for the abortion lobby and indicative of abortion policy being a weak spot for Republican candidates to political office. (In 2023, there is a strong correlation between opposition to abortion and Republican politicians and between support for abortion and Democrats, though this was not always the case.)

However, as I show in this article, an analysis of Gallup data on the views of Americans on abortion does not support the interpretation of abortion policy being a strong point for Democrats. Instead, it shows that the abortion policy pursued by Democrats represents a minority of Americans.


## Viewpoints on Abortion in the United States

The main question asked of Americans by Gallup polling since 1975 about viewpoints on the legality of abortion gives the respondent three major options and is phrased as "Do you think abortions should be legal under any circumstances, legal only under certain circumstances or illegal in all circumstances?"

Throughout the past several decades, there has been a majority of Americans who think abortion should be legal only under certain circumstances, flanked by two minorities: one that thinks abortion should legal under any circumstances and one that abortion should be illegal in all circumstances.[^all]

[^all]: The "illegal in all circumstances" verbiage probably makes this question biased against those respondents who wish to express disapproval of abortion, since even those most emphatically against abortion usually allow for at least _some_ exceptional cases. Starting in the 1990s, Gallup has asked a follow-up question for those who reply "legal only under certain circumstances," and the follow-up question has revealed that a supermajority of these respondents intend for abortion to be legal in few or limited circumstances.


```{r}

```

The minority of Americans that think abortion should be legal under any circumstances peaked in 2022 at 35%. It has never been larger than this. In contrast, if we take those who answer "legal only under certain circumstances" and "illegal in all circumstances" as constituting those who want there to be _at least some_ restrictions on abortion, then there is now and historically has been a supermajority of Americans who want _at least some_ restrictions on abortion.

```{r}

```

Additionally, a large supermajority of Americans, according to Gallup polling, think that abortion in the third trimester should be generally illegal.

```{r}

```

Furthermore, a simple majority of Americans think that abortion in the second trimester should be generally illegal. 

```{r}

```


### Policy of _Roe vs. Wade,_ _Doe vs. Bolton,_ and Ohio's _Issue 1_

The abortion laws of all fifty states of the United States from 1973 to 2022 had been overruled by the _Roe vs. Wade_ and _Doe vs. Bolton_ decisions. _Roe vs. Wade_ overturned any laws that prohibited abortion _for any reason_ up until "viability," which was stipulated to be the time at which a fetus can live outside his or her mother's womb and was interpreted in _Roe vs. Wade_ as "at about seven months (28 weeks) but may occur earlier, even at 24 weeks." [] Thus, "viability" according to _Roe vs. Wade_ includes the first two trimesters of pregnancy and some unspecified amount of the first month of the third trimester of pregnancy.

Since the vast majority of abortions are induced before the twenty-eighth week of gestational age, [] _Roe vs. Wade_ already made abortion policy in all fifty states very close to legal abortion under any circumstances. Additionally, the companion decision _Doe vs. Bolton_ overturned any law that prohibited abortion, even only abortions after viability, that did not have an exception for "health" reasons, and "health" was interpreted so broadly as to include things such as emotional distress. _Doe vs. Bolton_ would later be used to strike down laws prohibiting abortion even after the viability threshold. Thus, _Roe_ and _Doe_ together legalized abortion _under any circumstances_ and _at any time throughout pregnancy._ []

The text of Ohio's _Issue 1_ mirrors _Roe vs. Wade_ and _Doe vs. Bolton_ very closely, again legalizing abortion for any reason up until fetal viability, backed up a very broad "health" exception that could otherwise be used to strike down laws pertaining to any stage of pregnancy.[] 

The policy actually enforced by _Roe vs. Wade_ and _Doe vs. Bolton_ and now enforced once again by Ohio's _Issue 1_ is not the policy that the majority of Americans want. _Roe vs. Wade_ and Ohio's _Issue 1_ legalize abortion _for any reason,_ but as we have seen, a majority of Americans think abortion should be legal only in certain circumstances. _Roe _ and _Issue 1_ legalize abortion for any reason up until viability, which at least partially includes the _third trimester_ of pregnancy, but as was seen in the previous section, a supermajority of Americans think that abortions in the third trimester should be generally illegal. Finally, _Roe_ and _Issue 1_ legalize abortion for any reason throughout all of the _second trimester_ of pregnancy, but as was seen, a simple majority of Americans think that abortion should be generally illegal in the second trimester. 


### How the Abortion Lobby Succeeds

The abortion lobby that campaigns for the legality of abortion under any circumstances represents only a minority of Americans. The majority of Americans think that abortion should be legal only in certain circumstances and think that second and third trimester abortions should be generally illegal. How, then, has the abortion lobby succeeded?

With _Roe_ and _Doe_, the answer is obvious. The _Roe_ and _Doe_ policy was instituted by just a half dozen unelected judges[^burger] who were not subject to popular vote and hence public opinion.

[^burger]: The Supreme Court vote for _Roe_ and _Doe_ was actually seven to two, not six to three, but Chief Justice Burger may have voted with the majority once he saw that a majority would vote for _Roe_ and _Doe_ just so that he could choose which Justice to assign the writing of the opinion.

With Ohio's _Issue 1,_ which was put to a popular vote, the answer is more subtle. The 2023 Ohio election occurred during a year when there was not a vote for President of the United States. Historically, when an election occurs in one of the several states of the United States when there is no vote for President, voter turnout is lower.

```{r}

```

While the 2023 Ohio election has yet to be certified at the time of the writing of this article, the unofficial rate of voter turnout for the 2023 Ohio election is currently at 48.9%, which is consistent with previous Ohio elections in which there was no vote for President. [] The unofficial rate of those who voted "yes" on _Issue 1_ is currently at 56.6%.[] Thus, to put things in perspective, only about 27.7% of registered voters in the state of Ohio voted for _Issue 1._


#### Median Voter vs. Polar Voter Strategies

The fact that only 27.7% of registered voters in a state can win a referendum in an off-year election is striking, and it highlights two alternative strategies for winning elections.

The more obvious strategy is to appeal to the _median voter._ This is the strategy of figuring out what the majority of voters in an electorate want and to match a campaign's policy platform to this majority position. All other things being equal, if a candidate's platform matches the views of the majority of voters, then that candidate should win.

However, the success of the median voter strategy is contingent on everyone voting. When only 48.86% of registered voters -- and thus an even lower rate of _eligible_ voters -- vote in an election, then for close votes like Ohio's _Issue 1,_ _which voters_ actually do vote in an election can determine the outcome of the election.

An alternative strategy to the median voter strategy is to target the _polar voter._ This strategy explicitly targets viewpoints that are in the minority in an electorate, but that constitute a very _passionate_ minority. The goal here is to target a cohort that is more likely to vote in an election and to excite others to vote in the election than the median voter.

This is likely what occurred in the Ohio 2023 election. With _Issue 1_ on the ballot, a passionate minority of voters who would otherwise not have bothered voting in the 2023 election went to vote specifically for abortion, which was enough to tip the voting for _Issue 1_ to a 56.6% to 43.4% majority during an off year when turnout was less than 50%.


#### Lack of Understanding of Actual Policy

It is clear that a sizable number of people in the United States do not know what it means for abortion to be legal for any reason up until the point of fetal viability, which is the policy of _Roe vs. Wade_ and of Ohio's _Issue 1._ As we have seen, during the time _Roe vs. Wade_ was in force, a supermajority of Americans thought (and still think) that abortion in the third trimester should be generally illegal and simple majority thought (and still think) that abortion in the second trimester should be generally illegal. At the same time, a majority of Americans did not support overturning _Roe vs. Wade_, and a majority of Americans think that overturning _Roe vs. Wade_ was a bad thing.

```{r}

```

How can a majority of Americans support _Roe vs. Wade,_ which legalizes abortion for any reason into the early third trimester, and at the same time a majority of Americans think that abortion in the second and third trimester should be generally illegal?

The only way this can be the case is for some nontrivial number of Americans to be in a state of self-contradiction, both supporting policies that legalize second and third trimester abortions and supporting the general prohibition of second and third trimester abortions. This indicates that there is a nontrivial number of Americans who either do not understand that _Roe vs. Wade_ implies abortion is legal for any reason up until fetal viability or do not understand what legal abortion for any reason until fetal viability means.


#### Capture of Democrats

While those who think abortion should be legal in all circumstances constitute a minority of Americans, something important for politics in the United States has occurred in recent years. From 1975 to 2013, among Democrats, legal abortion in all circumstances was also a minority viewpoint, just it was for all Americans regardless of party affiliation. However, starting in 2014, about an equal number of Democrats supported legal abortion in all circumstances as did those who supported legal abortion only in certain circumstances. Now, as of 2023, a clear _majority_ of Democrats, about 60%, support legal abortion _in all circumstances._

```{r}

```

Thus, while the viewpoint that abortion should be legal in all circumstances is a minority viewpoint among Americans, this viewpoint now dominates the Democrats. Since there are only two political parties in the United States that regularly win elections, the legal-abortion-in-all-cases platform of the abortion lobby has an outsized impact on American politics because it controls one of the two main political parties.


## Recommendations for Median Voter Strategy

Democrats as a party have been captured by the minority viewpoint that abortion should be legal in all circumstances. As we have seen, the majority of Americans want abortion to be legal only in certain circumstances, and the majority of Americans think that abortion should be generally illegal in the second and third trimesters. The abortion lobby, on the other hand, has legalized abortion for any reason into the third trimester with _Roe vs. Wade_ and with Ohio's _Issue 1._

If Republicans can successfully campaign for the median voter, then abortion policy will be an issue of strength for them. However, Democrats have the advantage of being able to excite a very passionate minority, which can cause an increase in voter turnout and be successful in elections by way of the polar voter strategy.

Next year during the 2024 elections, Americans will vote for a President of the United States. Such elections usually have higher voter turnout, which may decrease the effectiveness of the polar voter strategy and increase the effectiveness of the median voter strategy.

Regardless of whether any given election is a Presidential election or not, candidates using the median voter strategy must adhere to several strategies in order for abortion policy to be an asset, not a detriment, to their campaigns.


### Reject an All-or-Nothing Framing

A tactic that the abortion lobby has been successful in using is to frame the abortion issue as an all-or-nothing question between abortion being illegal in all circumstances or abortion being legal in all circumstances. As Gallup polling indicates, a minority of Americans think that abortion should be illegal _in all circumstances._ Thus, if the way the issue is framed to voters is this dichotomy, then abortion advocates can cobble together a majority to vote for abortion being legal in all circumstances. However, a minority of Americans _actually want_ abortion to be legal in all circumstances.

Politicians campaigning for the median voter must be very careful, therefore, not to allow the issue to be framed as whether abortion should be legal or be banned. Instead, they must frame abortion policy as being a question of _under what circumstances_ abortion is to be legal and _under what circumstances_ it is to be illegal.


### Find the Median Voter

This leads to the next important point. It is imperative for politicians campaigning for the median voter on abortion policy to _discover_ under what circumstances a majority in their electorate want abortion to be illegal. Since abortion policy is made by the legislatures of the several states in the United States, this must be done at the state level.

This is admittedly more challenging than the all-or-nothing framing that is working for the abortion lobby, but it can be accomplished simply by taking a survey of a representative random sample of registered voters in a given state. Furthermore, polling at the national level can give a political campaign hints on where to begin. Since a majority of Americans think that second and third trimester abortions should be generally illegal, it is likely that the electorate in many states want some kind of gestational age limit on abortion. It is therefore a matter of polling to find out the specific gestational age limit is desired by majority of voters in a given state.

This was actually a process that began back in the 1960s in the state legislatures, but was overridden by the _Roe vs. Wade_ and _Doe vs. Bolton_ decisions. For instance, several states had allowed for exceptions to their criminal abortion statutes for cases such as rape or severe fetal deformity, while rejecting attempts to legalize abortion as a method of fertility control. [] It remains to be seen what are the circumstances under which majorities in various states want abortion to be prohibited, but this can be discovered.

In addition to outright prohibition, there are likely many _regulations_ of abortion that a majority of voters in a given state approve.[^regulations] For instance, Gallup polling suggests that a policy requiring abortion providers to provide information about alternatives to abortion has historically been quite popular at the national level in the United States. There are probably numerous such regulations that would be found to be popular in any given state.

[^regulations]: In addition to legalizing abortion until fetal viability, _Roe vs. Wade_ overturned any _regulations_ of abortions done in the first trimester. 

```{r}

```


### Create Simple Messaging

The abortion lobby has a stupidly simple message that they are promulgating that abortion is a "right" and that nefarious evildoers want to "take away your rights." This is an easy message on which to campaign. On the other hand, the Combined Argument and Explanation Against _Issue 1_ included in the Ohio ballots was scattered and unfocused. Campaigning for the median voter on abortion policy points the way to a much more streamlined message.

This message should be something to the effect of, "Elites used unelected judges who abused their Constitutional authority in order to override the will of the American people and overturn laws in all fifty states pertaining to abortion policy. More recently, the Supreme Court returned matters of abortion policy back to American people and their duly elected representatives. I, as a candidate for ______ office, will fulfill my duty to represent the people of the state of ______. The majority of voters in the state of ______ want..."

Knowledge of the circumstances under which the the majority of voters in a given state want abortion to be illegal, as discussed in the previous section, could then be used to substantiate this message.


## Consideration of Polar Voter Strategy

Some politicians opposed to abortion argue that election losses such as those experienced in Virginia are due to use of the median voter strategy, which has led to losses in the turnout aspect of elections. There is logic to this. It is highly likely that the abortion lobby, where it has won, has won based on voter turnout.

If the abortion lobby has been so successful using the polar voter strategy, why not adopt similar tactics? The abortion lobby has not represented the majority of Americans, but has had political successes. Indeed, rather than following public opinion, the abortion lobby appears to have changed public opinion in its favor by first imposing abortion policy against the viewpoints of the majority of Americans and then worrying about public opinion afterwards.

As previously mentioned, elections in 2024 will be during a Presidential election, which typically has much higher turnout. Therefore, in the short term, strategies which aim to increase turnout at the expense of attracting median voters may be not effective next year.

In the long term, the main issue with using the polar voter strategy _in opposition_ to abortion is its strong potential to backfire. Indeed, the fact that the abortion lobby has had success using the polar voter strategy is indicative _both_ of how it can be used successfully _and_ of how the strategy must be used by candidates opposed to abortion _with caution._ Framing abortion policy as an all-or-nothing question is a tactic with which the abortion lobby has had demonstrable success. Appealing to enthusiastic minorities opposed to abortion risks creating an artificial majority for the abortion lobby's platform of legal abortion in all circumstances.

The polar voter strategy trades in broader appeal among the electorate for increases in turnout among passionate voters. Whether or not this trade is a net positive for a candidate is heavily contingent on the views of voters in the candidate's electorate and the propensity of voters to vote. Thus, it is important for the candidates opposed to abortion to understand the views on abortion policy of the voters in the candidates' electorate and adopt a strategy accordingly.


## Do Not Be Scared of Abortion Policy

Regardless of whether the median voter or polar voter strategy is being used, candidates for political office who hope to succeed should run their campaigns with confidence. I have noticed several Republican politicians approach the abortion issue with trepidation and almost apologetically. This is a bad appearance for any candidate for public office. People can sense when others lack confidence, and this lack of confidence does not inspire confidence in others.

One of the advantages of the median voter strategy is that a candidate is armed with the knowledge that the candidate's platform is aligned with the majority of voters in the candidate's state. Candidates using the polar voter strategy must express personal conviction genuinely and appeals. Either way, a candidate hemming and hawing about one's platform on abortion does not flatter a candidate.


## Conclusion

A majority of Democrats want abortion to be legal in all circumstances, which is a view shared by only a minority of Americans. The majority of Americans want abortion to be legal only in certain circumstances and want abortion to be generally illegal in the second and third trimester. Any candidate for public office who can successfully campaign for the median voter by rejecting an all-or-nothing framing, by identifying the abortion policy desired by a majority of voters in the candidate's electorate, and by messaging this policy in a straightforward and confident manner can exploit this as a weakness in Democrats' platform.

It may be the case that, in certain constituencies, candidates opposed to abortion can succeed by using a strategy of campaigning for the polar voter, in which a passionate minority wins an election by way of increasing voter turnout. However, this has a strong potential to backfire, given the recent successes of the abortion lobby itself using such a strategy.

Regardless of the particular strategy used to win elections, it is important to remember that politics is only aspect of human civilization and only one aspect of opposition to abortion. This is ultimately not a struggle over the laws of any one country in any one time and place, but a struggle for the hearts and minds of humanity. Infanticide was not pushed to the extreme margins of society by laws alone. There is a danger in getting tunnel vision around the politics of abortion. Such narrow focus can cause us to ignore better opportunities to use limited resources that would have made more progress toward putting abortion into the history books.

Nonetheless, politics does have _some_ role to play in this process, and unlike other fields, politics only works for your social goals if you win. Therefore, those opposed to abortion who are actively engaged in the political sphere -- politicians, lobbyists, lawyers -- must be very careful how they approach political strategy.

Finally, one of the more overlooked functions of the role of those engaged in the political realm who do win is funding those aspects of society that prevent abortions and that are in need of much more funding, the most prominent being alloparenting support and prevention of unintended pregnancy. This is particularly relevant for those elected to legislative bodies in the United States, because in the United States legislatures control the budget and appropriate funds. 

"Alloparenting" refers to parental care given by anyone who is not a biological parent of a child, and can constitute anything from prenatal care to day care to adoption. Better support for parents, especially those with unexpected pregnancies, is a vital need both in the short term for any individual child and in long term for changing attitudes toward abortion.

The birth control movement in the United States was founded in direct opposition to abortion as a method of fertility control. However, birth control organizations such as Planned Parenthood have been stolen by the very abortion industry [their founders were attempting to put out of business](/margaret-sanger-on-abortion-in-her-own-words.html). There have been some movements to stop giving government funding to Planned Parenthood. An even more pressing need is to have an organization to fund instead, one dedicated to preventing unintended pregnancy exclusively through nonviolent means, such as through contraception and education. Every two unintended pregnancies prevents, on average, one abortion, and _every_ unintended pregnancy prevented attenuates the motivation used by the abortion lobby in order to generate support for abortion.


## Footnotes
