---
author: "Joshua Born"
bibliography: "../bibliographies/untitled.json"
csl: "../bibliographies/chicago-author-date.csl"
date: 2023-12-16
draft: true
implicit-figures: true
include-before: |
  ## Table of Contents
output:
  blogdown::html_page:
    number_sections: false
    toc: true
    toc_depth: 5
summary: '[Put summary sentence here.]'
tags: ["analysis", "induced abortion", "politics", "first-person"]
title: 'Untitled Article'
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(showtext)
library(tibble)
library(tidyr)

font_add_google('Open Sans', 'open')
showtext_auto()
showtext_opts(dpi = 200)
theme_update(text = element_text(family = 'open'))

knitr::opts_chunk$set(
  cache = FALSE,
  dpi = 300,
  echo = FALSE,
  fig.align = 'center',
  fig.height = 3.75,
  fig.retina = 1,
  fig.width = 5
)
```


## Introduction

I do not typically write much about abortion[^definition] politics. However, recently there were elections in the Virginia, where I live, and in Ohio. Several Democrats won elections in Virginia after focusing their campaigns on keeping abortion as legal as possible, and a referendum labeled _"State Issue 1 A Self-Executing Amendment Relating to Abortion and Other Reproductive Decisions"_ was passed in Ohio that replicated the policy instituted by the _Roe v. Wade_ and _Doe v. Bolton_ decisions of the Supreme Court of the United States.

[^definition]: Unless otherwise noted, in this article the word "abortion" is used to mean induced abortion of pregnancy.

    Technically, the word "abortion" is a generic term. Any process that is aborted before it comes to completion can, in theory, be labeled "abortion." However, because of its association with abortion of pregnancy and the emotional weight of such occurrence, the word "abortion" is usually used to mean abortion of pregnancy.

    Furthermore, even if we just consider "abortion" to mean abortion of pregnancy, there is ambiguity because in the medical literature the word "abortion" is used to mean two different things: spontaneous abortion, which is commonly called "miscarriage" in the vernacular, occurs when a pregnancy terminates without anyone's intervention; induced abortion occurs when a pregnancy is terminated on purpose. When "abortion" is used in the vernacular it is commonly used to mean induced abortion.

    This ambiguity can lead to misinterpretation. For instance, if a study were to report on abortions in a given population, it could be including both spontaneous and induced abortions if it were using the medical literature definition, but it could be excluding what are commonly called "miscarriages" if it were using the common definition.

Because of these and other election results in the past year since the _Dobbs v. Jackson Women's Health Organization_ decision of the Supreme Court that overturned _Roe_ and _Doe,_ there have been numerous opinion articles interpreting the results as indicating that abortion policy is strong point for Democrat campaigns and a weak spot for Republican campaigns. (Before the 1990s, there was not a strong correlation between political party and views on abortion, but in 2023, there is a strong correlation between Democrats and support for legalization of abortion in all circumstances and between Republicans and support for at least some restrictions on abortion.)

However, an analysis of Gallup polling on the views of Americans on abortion does not support this interpretation. A minority of Americans actually support the abortion policy advocated by Democrats. To understand how campaigning on a platform that only a minority of voters agree with can lead to victories in elections, an understanding of voting dynamics in the United States is necessary.


### Majority Voters versus Polar Voters in the United States

The most obvious way to win an election when there is a popular vote is to find out what policy position a majority of voters in the electorate support and to campaign on that. Call this the **"majority voter strategy."** If there were only one issue, and if there were a majority position on this one issue, and if a candidate for public office were to campaign on the majority position, then it seems like a safe assumption that the that candidate would win.

For instance, suppose there were only one issue and on this issue, 34% of registered voters support Policy A, 51% support Policy B, and 13% support policy B. If you were a political candidate who wanted to win an election, it seems like Policy B would be the best platform. Even if all of the supporters of Policy A and Policy C voted for an alternative candidate, you would still get 51% of the vote and win the election.

However, there is a tacit assumption here that _all registered voters will vote._ This is simply not the case in the United States. Voter turnout in the United States is actually quite low. It is better in elections that include voting for the President of the United States, but in elections that do not include a vote for President, voter turnout is particularly low.

To use a concrete example, the official voter turnout for Ohio's November 7, 2023 general election, in which there was not a vote for President, was 49.63%. [@ohio_election_2023] A _minority_ of registered voters actually voted in the 2023 Ohio election. (If you were to quantify the proportion of _eligible_ voters who actually voted, it would be even lower.)

This phenomenon is what allows political polarization to thrive in the United States. You can win an election in the United States by campaigning on a platform that is a minority position, provided that the minority is excited enoguh about the position that turnout will be higher among the minority than among the rest of the electorate. Call this the **"polar voter strategy."**

For instance, suppose that those who support Policy A in the thought experiment are very excited by a candidate campaigning on Policy A, but supporters of Policy B and of Policy C are less excited about their respective positions and so are more inclined to spend the election day on working or pursuing other interests.

If the difference in turnout is drastic enough -- say, 80% turnout for supporters of Policy A compared with 40% turnout for supporters of Policy B or Policy C -- then a candidate running on Policy A will win the election, even though Policy A is actually supported by a minority of the electorate. In such a scenario, voter turnout would be only $(80\% \times 34\%) + 40\% \times (51\% + 13\%) = 52.8\%$, and $\frac{(80\% \times 34\%)}{58.8\%} = 51.5\%$ of voters _who showed up to the election_ would vote for Policy A, even though only 34% of the electorate actually support Policy A.

This is how running on minority positions succeeds in elections in the United States and is a big reason why political polarization is so abundant in the United States.

This is, also, how Democrats win elections by campaigning on abortion. To put things in perspective, the official voter turnout in the 2023 Ohio election was 49.63%, and 56.78% of voters in the election voted for _Issue 1._ [@ohio_election_2023] Therefore, only $49.63\% \times 56.78\% = 28.18\%$ of registered voters in Ohio voted for _Issue 1._ To understand why this is the case with the specific issue of abortion policy, an understanding of public opinion on abortion in the United States is needed.


## Viewpoints on Abortion in the United States

Gallup has been polling Americans on abortion more or less annually since 1975. The main question asked about the legality of abortion gives the respondent three options and is phrased as "Do you think abortions should be legal under any circumstances, legal only under certain circumstances or illegal in all circumstances?"

Throughout the past several decades, there has been a majority[^sampling-error] of Americans who respond that abortion should be legal only under certain circumstances, flanked by two minorities: one that respond that abortion should be legal under any circumstances and one that responds that abortion should be illegal in all circumstances.[^illegal-all-circumstances]

[^sampling-error]: Sometimes the percent of respondents who reply "legal only under certain circumstances" dips below 51%, but in these cases, the the sample proportion responding "legal only under certain circumstances" is still within margin of error of 51% and may not represent actual changes in the population, but could be fluctuations due to random sampling.

[^illegal-all-circumstances]: The "illegal in all circumstances" response was a poor construction on the part of Gallup. Even those who are very strongly against abortion usually admit of exception cases in which it should be legal, and so would be less likely to respond "illegal in all circumstances." Starting in the 1990s, Gallup has asked a follow-up question for those who reply "legal only under certain circumstances," and the follow-up question has revealed that most of these respondents intend for abortion to be legal in few or limited circumstances. Thus, those strongly against abortion are split between "legal only under certain circumstances" and "illegal in all circumstances" responses in Gallup polling.

```{r ggplot-theme}
theme_set(
  theme_classic() +
  theme(
    axis.line.x.bottom = element_line(
      color = 'black',
      linewidth = 0.2
    ),
    axis.line.y.left = element_line(
      color = 'black',
      linewidth = 0.2
    ),
    axis.text.x = element_text(
      angle = 45,
      color = 'black',
      hjust = 0.5,
      vjust = 0.65
    ),
    axis.text.y = element_text(
      color = 'black',
      margin = margin(r = 0)
    ),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_line(
      color = 'black',
      linewidth = 0.2
    ),
    axis.title.x = element_text(
      margin = margin(t = 2)
    ),
    axis.title.y = element_text(
      margin = margin(r = 10)
    ),
    legend.margin = margin(t = 0, b = 0),
    legend.position = 'top',
    legend.title = element_blank(),
    panel.grid.major = element_line(
      color = 'gray',
      linetype = 3,
      linewidth = 0.25
    ),
    panel.grid.minor = element_line(
      color = 'gray',
      linetype = 3,
      linewidth = 0.25
    ),
    plot.margin = margin(t = 0, r = 10)
  )
)
```

```{r data-Gallup-abortions-should-be-legal}
Gallup_abortions_should_be_legal <- tribble(
  ~Date, ~legal_any, ~legal_only_certain, ~illegal_all, ~no_opinion, ~based_on_registered,
  "2023 May 1", 34,	51,	13,	2, FALSE,
  "2022 May 2", 35,	50,	13,	2, FALSE,
  "2021 May 3", 32,	48,	19,	2, FALSE,
  "2020 May 1", 29,	50,	20,	2, FALSE,
  "2019 May 1", 25,	53,	21,	2, FALSE,
  "2018 May 1", 29,	50,	18,	2, FALSE,
  "2017 May 3", 29,	50,	18,	3, FALSE,
  "2016 May 4", 29,	50,	19,	2, FALSE,
  "2015 May 6", 29,	51,	19,	1, FALSE,
  "2014 May 8", 28,	50,	21,	2, FALSE,
  "2013 May 2", 26,	52,	20,	2, FALSE,
  "2012 Dec 27", 28,	52,	18,	3, FALSE,
  "2012 May 3", 25,	52,	20,	3, FALSE,
  "2011 Jul 15", 26,	51,	20,	3, FALSE,
  "2011 Jun 9", 26,	52,	21,	1, FALSE,
  "2011 May 5", 27,	50,	22,	2, FALSE,
  "2010 May 3", 24,	54,	19,	3, FALSE,
  "2009 Jul 17", 21,	57,	18,	4, FALSE,
  "2009 May 7", 22,	53,	23,	2, FALSE,
  "2008 May 8", 28,	54,	17,	2, FALSE,
  "2007 May 10", 26,	55,	18,	1, FALSE,
  "2006 May 8", 30,	53,	15,	2, FALSE,
  "2005 Nov 11", 26,	56,	16,	2, FALSE,
  "2005 Jun 24", 24,	55,	20,	1, FALSE,
  "2005 May 2", 23,	53,	22,	2, FALSE,
  "2005 Mar 21", 23,	55,	20,	2, FALSE,
  "2004 May 2", 24,	56,	19,	1, FALSE,
  "2003 Oct 24", 26,	55,	17,	2, FALSE,
  "2003 May 5", 23,	57,	19,	1, FALSE,
  "2003 Jan 10", 24,	57,	18,	1, FALSE,
  "2002 May 6", 25,	51,	22,	2, FALSE,
  "2002 Mar 22", 27,	53,	19,	1, FALSE,
  "2002 Feb 8", 26,	54,	18,	2, FALSE,
  "2001 Aug 10", 26,	56,	17,	1, FALSE,
  "2001 May 10", 26,	58,	15,	1, FALSE,
  "2001 Mar 26", 26,	51,	18,	5, FALSE,
  "2000 Mar 30", 28,	51,	19,	2, FALSE,
  "2000 Jan 13", 26,	56,	15,	3, FALSE,
  "1999 Apr 30", 27,	55,	16,	2, FALSE,
  "1998 Jan 16", 23,	59,	17,	1, FALSE,
  "1997 Nov 6", 26,	55,	17,	2, FALSE,
  "1997 Aug 12", 22,	61,	15,	2, FALSE,
  "1996 Sep 20", 24,	52,	17,	7, TRUE,
  "1996 Jul 25", 25,	58,	15,	2, FALSE,
  "1995 Sep 22", 31,	54,	12,	3, FALSE,
  "1995 Feb 24", 33,	50,	15,	2, FALSE,
  "1994 Sep 6", 33,	52,	13,	2, FALSE,
  "1994 Mar 28", 31,	51,	15,	3, FALSE,
  "1993 Mar 12", 32,	51,	13,	4, FALSE,
  "1992 Jul 23", 34,	48,	15,	3, TRUE,
  "1992 Jan 16", 31,	53,	14,	2, FALSE,
  "1991 Sep 5", 33,	49,	14,	4, FALSE,
  "1991 May 30", 32,	50,	17,	1, FALSE,
  "1990 Apr 5", 31,	53,	12,	4, FALSE,
  "1989 Jul 6", 29,	51,	17,	3, FALSE,
  "1989 Jul 6", 29,	51,	17,	3, FALSE,
  "1989 Apr 12", 27,	50,	18,	5, FALSE,
  "1988 Sep 25", 24,	57,	17,	2, FALSE,
  "1985 Jan 3", 21,	55,	21,	3, FALSE,
  "1983 Jun 24", 23,	58,	16,	3, FALSE,
  "1981 May 8", 23,	52,	21,	4, FALSE,
  "1980 Jul 11", 25,	53,	18,	4, FALSE,
  "1979 Feb 23", 22,	54,	19,	5, FALSE,
  "1977 Dec 9", 22,	55,	19,	4, FALSE,
  "1975 Apr 4", 21,	54,	22,	3, FALSE
) |>
  mutate(
    at_least_some = legal_only_certain + illegal_all
  ) |>
  mutate(
    across(Date, ymd)
  )
```

```{r plot-Gallup-abortions-should-be-legal, cache=TRUE, fig.cap="Distributions of responses over time to question 'Do you think abortions should be legal under any circumstances, legal only under certain circumstances or illegal in all circumstances?' in Gallup polling. [@gallup_abortion_in_depth]"}
Gallup_abortions_should_be_legal |>
  mutate(
    across(c(legal_any, legal_only_certain, illegal_all, no_opinion), \(x) x / 100)
  ) |>
  pivot_longer(
    cols = c(legal_any, legal_only_certain, illegal_all, no_opinion)
  ) |>
  ggplot(
    mapping = aes(
      x = Date,
      y = value,
      color = name,
      shape = name
    )
  ) +
    geom_line(
      alpha = 0.4,
      linewidth = 1
    ) +
    geom_point(
      fill = 'white',
      size = 0.8
    ) +
    scale_y_continuous(
      breaks = \(x) seq(0, x[2], 0.1),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_percent(),
      limits = c(0, NA)
    ) +
    scale_x_date(
      breaks = scales::breaks_width('2 years', offset = '1 year'),
      expand = expansion(add = 365),
      labels = scales::label_date(format = '%Y')
    ) +
    scale_color_hue(
      label = c(
        illegal_all = 'Illegal in all',
        legal_any = 'Legal under any',
        legal_only_certain = 'Legal only under certain',
        no_opinion = 'No opinion'
      )
    ) +
    scale_shape_manual(
      label = c(
        illegal_all = 'Illegal in all',
        legal_any = 'Legal under any',
        legal_only_certain = 'Legal only under certain',
        no_opinion = 'No opinion'
      ),
      values = c(21, 22, 23, 24)
    ) +
    labs(
      x = 'Year',
      y = 'Percent of respondents'
    )
```

The proportion of respondents that think abortion should be legal under any circumstances has fluctuated since 1975 with low of `r paste0(min(Gallup_abortions_should_be_legal$legal_any), '%')` and a high of `r paste0(max(Gallup_abortions_should_be_legal$legal_any), '%')`, which occurred last year in 2022 exceeding a similar maximum in 1992. 

In contrast, if we take those who answer "legal only under certain circumstances" and "illegal in all circumstances" as constituting those who want there to be _at least some_ restrictions on abortion, then there is now and historically has been a majority of Americans who want _at least some_ restrictions on abortion.

```{r plot-Gallup-at-least-some-restrictions, cache=TRUE, fig.cap="Distributions of responses over time to question 'Do you think abortions should be legal under any circumstances, legal only under certain circumstances or illegal in all circumstances?' with 'legal only under certain circumstances' and 'illegal in all circumstances' responses summed into an 'at least some restrictions' category. [@gallup_abortion_in_depth]"}
Gallup_abortions_should_be_legal |>
  mutate(
    across(c(legal_any, at_least_some, no_opinion), \(x) x / 100)
  ) |>
  pivot_longer(
    cols = c(legal_any, at_least_some, no_opinion)
  ) |>
  ggplot(
    mapping = aes(
      x = Date,
      y = value,
      color = name,
      shape = name
    )
  ) +
    geom_line(
      alpha = 0.4,
      linewidth = 1
    ) +
    geom_point(
      fill = 'white',
      size = 0.8
    ) +
    scale_y_continuous(
      breaks = \(x) seq(0, x[2], 0.1),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_percent(),
      limits = c(0, NA)
    ) +
    scale_x_date(
      breaks = scales::breaks_width('2 years', offset = '1 year'),
      expand = expansion(add = 365),
      labels = scales::label_date(format = '%Y')
    ) +
    scale_color_hue(
      label = c(
        legal_any = 'Legal under any',
        at_least_some = 'At least some restrictions',
        no_opinion = 'No opinion'
      )
    ) +
    scale_shape_manual(
      label = c(
        legal_any = 'Legal under any',
        at_least_some = 'At least some restrictions',
        no_opinion = 'No opinion'
      ),
      values = c(21, 22, 23)
    ) +
    labs(
      x = 'Year',
      y = 'Percent of respondents'
    )
```

The proportion of respondents that think there should be at least some restrictions on abortion has fluctuated since 1975 with a low of `r paste0(min(Gallup_abortions_should_be_legal$at_least_some), '%')` and a high of `r paste0(max(Gallup_abortions_should_be_legal$at_least_some), '%')`.

Additionally, a large majority of Americans, according to Gallup polling that has been done intermittently since 1996, think that abortion in the third trimester should be generally illegal.

```{r data-Gallup-third-trimester-legal}
Gallup_third_trimester_legal <- tribble(
  ~Date, ~legal, ~illegal, ~depends, ~no_opinion,
  '2023 May 1',   22,	70,	7, 1,
  '2022 May 2',   20,	71,	8, 1,
  '2018 May 1',   13,	81,	4, 2,
  '2012 Dec 27',	14,	80,	4, 2,
  '2011 Jun 9',	  10,	86,	2, 1,
  '2003 Jan 10',	10,	84, 4, 2,
  '2000 Mar 30',	8, 	86, 3, 3,
  '1996 Jul 25',	13,	82, 3, 2
) |>
  mutate(
    across(Date, ymd)
  )
```

```{r plot-Gallup-third-trimester-legal, cache=TRUE, fig.cap="Distributions of responses over time to question 'Thinking more generally, do you think abortion should generally be legal or generally illegal during each of the following stages of pregnancy. How about in the last three months of pregnancy?' [@gallup_abortion_in_depth]"}
Gallup_third_trimester_legal|>
  mutate(
    across(c(legal, illegal, depends, no_opinion), \(x) x / 100)
  ) |>
  pivot_longer(
    cols = c(legal, illegal, depends, no_opinion)
  ) |>
  ggplot(
    mapping = aes(
      x = Date,
      y = value,
      color = name,
      shape = name
    )
  ) +
    geom_line(
      alpha = 0.4,
      linewidth = 1
    ) +
    geom_point(
      fill = 'white',
      size = 1.5
    ) +
    scale_y_continuous(
      breaks = \(x) seq(0, x[2], 0.1),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_percent(),
      limits = c(0, NA)
    ) +
    scale_x_date(
      breaks = scales::breaks_width('2 years', offset = '1 year'),
      expand = expansion(add = 365),
      labels = scales::label_date(format = '%Y')
    ) +
    scale_color_hue(
      label = c(
        legal = 'Should be legal',
        illegal = 'Should be illegal',
        depends = 'Depends (volunteered)',
        no_opinion = 'No opinion'
      )
    ) +
    scale_shape_manual(
      label = c(
        legal = 'Should be legal',
        illegal = 'Should be illegal',
        depends = 'Depends (volunteered)',
        no_opinion = 'No opinion'
      ),
      values = c(21, 22, 23, 24)
    ) +
    labs(
      x = 'Year',
      y = 'Percent of respondents'
    )
```

The proportion of respondents that think third trimester abortions should be generally illegal has fluctuated since 1996 with a low of `r paste0(min(Gallup_third_trimester_legal$illegal), '%')` and a high of `r paste0(max(Gallup_third_trimester_legal$illegal), '%')`.[^trend]

Furthermore, a smaller majority of Americans think that abortion in the second trimester should be generally illegal. 

```{r data-Gallup-second-trimester-legal}
Gallup_second_trimester_legal <- tribble(
  ~Date, ~legal, ~illegal, ~depends, ~no_opinion,
  '2023 May 1', 37,	55,	6,	1,
  '2022 May 2', 36,	55,	7,	2,
  '2018 May 1', 28,	65,	4,	2,
  '2012 Dec 27',27,	64,	5,	4,
  '2011 Jun 9', 24,	71,	3,	2,
  '2003 Jan 10',25,	68,	4,	3,
  '2000 Mar 30',24,	69,	4,	3,
  '1996 Jul 25',26,	65,	7,	2
) |>
  mutate(
    across(Date, ymd)
  )
```

```{r plot-Gallup-second-trimester-legal, cache=TRUE, fig.cap="Distributions of responses over time to question 'Thinking more generally, do you think abortion should generally be legal or generally illegal during each of the following stages of pregnancy. How about in the second three months of pregnancy?' [@gallup_abortion_in_depth]"}
Gallup_second_trimester_legal|>
  mutate(
    across(c(legal, illegal, depends, no_opinion), \(x) x / 100)
  ) |>
  pivot_longer(
    cols = c(legal, illegal, depends, no_opinion)
  ) |>
  ggplot(
    mapping = aes(
      x = Date,
      y = value,
      color = name,
      shape = name
    )
  ) +
    geom_line(
      alpha = 0.4,
      linewidth = 1
    ) +
    geom_point(
      fill = 'white',
      size = 1.5
    ) +
    scale_y_continuous(
      breaks = \(x) seq(0, x[2], 0.1),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_percent(),
      limits = c(0, NA)
    ) +
    scale_x_date(
      breaks = scales::breaks_width('2 years', offset = '1 year'),
      expand = expansion(add = 365),
      labels = scales::label_date(format = '%Y')
    ) +
    scale_color_hue(
      label = c(
        legal = 'Should be legal',
        illegal = 'Should be illegal',
        depends = 'Depends (volunteered)',
        no_opinion = 'No opinion'
      )
    ) +
    scale_shape_manual(
      label = c(
        legal = 'Should be legal',
        illegal = 'Should be illegal',
        depends = 'Depends (volunteered)',
        no_opinion = 'No opinion'
      ),
      values = c(21, 22, 23, 24)
    ) +
    labs(
      x = 'Year',
      y = 'Percent of respondents'
    )
```

The proportion of respondents that think second trimester abortions should be generally illegal has fluctuated between a low of `r paste0(min(Gallup_second_trimester_legal$illegal), '%')` and a high of `r paste0(max(Gallup_second_trimester_legal$illegal), '%')`.[^trend]

[^trend]: While it may appear that there is a trend toward more public support for the legality of second and third trimester abortions, it should be noted that the polling with regard to trimesters of pregnancy began in 1996, which was just before the a peak in popular opinion supporting abortion restriction (Figure \@ref(fig:plot-Gallup-at-least-some-restrictions)). It may be that this apparent trend is just part of a cyclical fluctuation, it may be that this is a trend in its own right.

Thus, there has actually been a majority consensus among Americans for at least forty eight years. Throughout this time, a majority of Americans support at least some restrictions on abortion, thinking either that it should be legal only in certain circumstances or that it should be illegal in all circumstances, and a majority of Americans think that abortion in the second and third trimesters should be generally illegal.


## Policy of _Roe v. Wade,_ _Doe v. Bolton,_ and Ohio's _Issue 1_

The abortion laws of all fifty states of the United States from 1973 to 2022 had been overruled by the _Roe v. Wade_ and _Doe v. Bolton_ decisions. _Roe v. Wade_ overturned any laws that prohibited abortion _for any reason_ up until "viability," which was stipulated to be the time at which a fetus can live outside his or her mother's womb and was interpreted in _Roe v. Wade_ as "at about seven months (28 weeks) but may occur earlier, even at 24 weeks." [@roe_v_wade] Thus, the time before "viability" according to _Roe v. Wade_ includes the first two trimesters of pregnancy and some unspecified amount of the first month of the third trimester of pregnancy.

Since the vast majority of abortions are induced before the twenty-eighth week of gestational age, [@cdc_2020] _Roe v. Wade_ already made abortion policy in all fifty states very close to legal abortion under any circumstances. Additionally, the companion decision _Doe v. Bolton_ overturned any law that prohibited abortion, even only abortions after viability, that did not have an exception for "health" reasons, and "health" was interpreted so broadly as to include things such as emotional distress. _Doe v. Bolton_ would later be used to strike down laws prohibiting abortion even after the viability threshold. [@forsythe_2013, 150-152] Thus, _Roe_ and _Doe_ together effectively legalized abortion _under any circumstances_ and _at any time throughout pregnancy._

The text of Ohio's _Issue 1_ mirrors _Roe v. Wade_ and _Doe v. Bolton_ very closely, again legalizing abortion for any reason up until fetal viability, coupled with a very broad "health" exception that could otherwise be used to strike down laws pertaining to any stage of pregnancy. [@ohio_issues_1_and_2] 

The policy actually enforced by _Roe v. Wade_ and _Doe v. Bolton_ and now enforced once again by Ohio's _Issue 1_ is not the policy that the majority of Americans want.

_Roe v. Wade_ and Ohio's _Issue 1_ legalize abortion _for any reason,_ but as we saw in the previous section, a majority of Americans think abortion should be legal only in certain circumstances and support at least some restrictions on abortion.

_Roe _ and _Issue 1_ legalize abortion for any reason up until viability, which at least partially includes the _third trimester_ and includes all of the _second trimester_ of pregnancy, but as we saw in the previous section, a majority of Americans think that abortions in the third and second trimesters should be generally illegal.


## How the Abortion Lobby Succeeds

For the sake of brevity, let us use the term "abortion lobby" to refer to the elements in American society that labor to make abortion policy match the policy of _Roe_ and _Doe_ and that was reiterated in Ohio's _Issue 1._ The abortion lobby campaigns for a set of policies that is effectively equivalent to the response to Gallup polling that "abortions should be legal under any circumstances," and so support for the platform of the abortion lobby has fluctuated between 21% and 35% of Americans.

How, then, has the abortion lobby succeeded in instituting its policies when it represents a minority of Americans?


### Judicial Activism

With _Roe_ and _Doe_, the answer is obvious. The _Roe_ and _Doe_ policy was instituted by just a half dozen unelected judges[^burger] who were not subject to popular vote and hence public opinion.

[^burger]: The Supreme Court vote for _Roe_ and _Doe_ was actually seven to two, not six to three, but Chief Justice Burger may have voted with the majority once he saw that a majority would vote for _Roe_ and _Doe_ just so that he could choose which Justice to assign the writing of the opinion.

You could read numerous papers and books of arguments for and against doctrines such as _substantive due process_ and _selective fourteenth amendment incorporation_ on which the validity of _Roe_ and _Doe_ turn. You could also read numerous papers and books on the validity of the specific arguments and circumstances of the _Roe_ and _Doe_ decisions. Such legal aspect of the validity of the _Roe_ and _Doe_ decisions are beyond the scope of this article. 

What is not controversial, however, is that the judiciary in the United States is very much by design not subject to popular opinion. Therefore, the use of the judiciary to achieve policy changes, regardless if you think it is valid or invalid, explains how the abortion lobby succeeded in achieving its policy goals, despite representing a minority viewpoint, for the time period of 1973 to 2022.


### Polarization

Unlike _Roe_ and _Doe,_ Ohio's _Issue 1_ was put to a popular vote. Up until now, we have been discussing Gallup polling of a representative _nation-wide_ sample of the United States. The federal system of government in the United States entails that voting such as what was done for Ohio's _Issue 1_ is done on a state by state basis. Any one given state may have a different distribution of public opinion about abortion than what is observed nationally.

However, 62.41% of voters in Ohio's 2022 general election voted for current Republican governor Mike DeWine, who is publicly opposed to abortion, compared with just 37.38% for the Democrat running against him. [@ohio_election_2022] Furthermore, the general election occurred on November 3, 2020, a few months after the _Dobbs_ decision that overturned _Roe_ and _Doe._ If the distribution of viewpoints on abortion of the citizens of Ohio differed from the national distribution such that more Ohioans support legal abortion under any circumstances than do Americans generally, then we would have expected to have seen it affect the 2022 gubernatorial election in Ohio. Therefore, this is unlikely to be the case.

_Issue 1_ passed in Ohio because different people turned out to vote in the 2023 general election than did in the 2022 general election. Exit polling confirms that only 35% of voters in the 2023 Ohio general election were Republicans, the lowest proportion since 2012, and self-described "liberals" accounted for 34% of voters in the 2023 elections, compared with 20% in the 2022 elections. [@kgfw_issue_1]

In short, the abortion lobby succeeded in the 2023 Ohio general election by using the [polar voter strategy](#majority-voters-versus-polar-voters-in-the-united-states). As discussed in the preceding, only 49.63% of registered voters participated in the 2023 Ohio general election, and only 28.18% of registered voters in Ohio voted for _Issue 1._

The polar voter strategy only works in elections in which turnout is relatively low. Next year, the 2024 general elections will include voting for President of the United States, and such elections historically have had higher voter turnout. This is why Democratic strategist Ruy Teixeira commented on a recent podcast that Democrats risk becoming "low turnout election specialists" should they continue to emphasize issues like abortion.

> And we should always remember that, I mean, people are moderately pro-choice but are actually very leery about abortions after the first trimester, and then in the third trimester, they're just flat out against it. So, the real hardcore activist position on abortion rights, which obviously has a real presence to the Democratic Party, is abortion at any time for any reason, at will. That is not the position of the median voter who is quite a bit more conservative on this issue. They want abortion rights in some form maintained, but they don't want a sort of at will anytime kind of approach to abortion.
> 
> So, that may eventually come to be a bit more of a conflict and a problem for the Democrats. So far, they've been able to leverage the issue pretty easily in their favor by basically saying Republicans want to ban it completely. So, we'll see what happens with that, but even as it's played out so far, it hasn't been quite the magic bullet people think, and as I argue in a piece that's just about to come out from the _[Washington] Post,_ another important thing to remember here is Democrats are now in a sense low turnout election specialists. Right? Their coalition is engaged. It's more educated. It's more animated by an issue like abortion, and they're actually likely, because of the structure of their coalition, to do particularly well in off-year and special elections. And it's much more of a challenge in a big presidential election when the peripheral voters come out of the woodwork. [@teixeira_podcast]


### Lack of Understanding of Actual Policy

It is clear that a sizable number of people in the United States do not know what it means for abortion to be legal for any reason up until the point of fetal viability, which is the policy of _Roe v. Wade_ and of Ohio's _Issue 1._ As we have seen, during the time _Roe v. Wade_ was in force, a large majority of Americans thought (and still think) that abortion in the third trimester should be generally illegal and smaller majority thought (and still think) that abortion in the second trimester should be generally illegal. At the same time, a majority of Americans did not support overturning _Roe v. Wade._

```{r data-Gallup-overturn-roe-v-wade}
Gallup_overturn_roe_v_wade <- tribble(
  ~Date, ~yes_overturn, ~no_not_overturn, ~no_opinion, ~half_sample,
  "2023 May 1",	38,	61,	1, FALSE,
  "2022 May 2", 35,	58,	7, TRUE,
  "2021 May 3", 32,	58,	10, TRUE,
  "2019 Jun 3", 33,	60,	7, FALSE,
  "2018 Jul 2", 28,	64,	9, FALSE,
  "2012 Dec 27",29,	53,	18, FALSE,
  "2008 May 8", 33,	52,	15, FALSE,
  "2007 May 10",35,	53,	12, FALSE,
  "2006 May 8", 32,	55,	13, FALSE,
  "2006 Jan 20",25,	66,	9, FALSE,
  "2005 Jul 7", 28,	63,	9, TRUE,
  "2002 Mar 22",36,	60,	4, FALSE,
  "1992 Aug 13",34,	60,	6, FALSE,
  "1989 Oct 5", 33,	61,	6, FALSE,
  "1989 Jul 6", 31,	58,	11, FALSE
  
) |>
  mutate(
    across(Date, ymd)
  )
```

```{r plot-Gallup-overturn-roe-v-wade, cache=TRUE, fig.cap="Distributions of responses over time to question 'Would you like to see the Supreme Court overturn its 1973 Roe v. Wade decision concerning abortion, or not?' or in 2023 'Do you think overturning Roe versus Wade was a good thing or a bad thing?' [@gallup_abortion_in_depth]"}
Gallup_overturn_roe_v_wade |>
  mutate(
    across(c(yes_overturn, no_not_overturn, no_opinion), \(x) x / 100)
  ) |>
  pivot_longer(
    cols = c(yes_overturn, no_not_overturn, no_opinion)
  ) |>
  ggplot(
    mapping = aes(
      x = Date,
      y = value,
      color = name,
      shape = name
    )
  ) +
    geom_line(
      alpha = 0.4,
      linewidth = 1
    ) +
    geom_point(
      fill = 'white',
      size = 1.5
    ) +
    scale_y_continuous(
      breaks = \(x) seq(0, x[2], 0.1),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_percent(),
      limits = c(0, NA)
    ) +
    scale_x_date(
      breaks = scales::breaks_width('2 years', offset = '1 year'),
      expand = expansion(add = 365),
      labels = scales::label_date(format = '%Y')
    ) +
    scale_color_hue(
      label = c(
        yes_overturn = 'Yes, overturn / Good thing overturned',
        no_not_overturn = 'No, not overturn / Bad thing overturned',
        no_opinion = 'No opinion'
      )
    ) +
    scale_shape_manual(
      label = c(
        yes_overturn = 'Yes, overturn / Good thing overturned',
        no_not_overturn = 'No, not overturn / Bad thing overturned',
        no_opinion = 'No opinion'
      ),
      values = c(21, 22, 23, 24)
    ) +
    labs(
      x = 'Year',
      y = 'Percent of respondents'
    )
```

The proportion of respondents that did not want the Supreme Court to overturn its 1973 _Roe v. Wade_ decision fluctuated between 1989 and 2022 with a low of `r paste0(min(Gallup_overturn_roe_v_wade$no_not_overturn), '%')` and a high of `r paste0(max(Gallup_overturn_roe_v_wade$no_not_overturn), '%')`. As of May 2023, 61% of respondents thought that overturning _Roe v. Wade_ was a bad thing.

How can a majority of Americans support _Roe v. Wade,_ which legalizes abortion for any reason into the early third trimester, and at the same time, a majority of Americans think that abortion in the second and third trimester should be generally illegal and that abortion should only be legal in certain circumstances? 

The only way this can be the case is for some nontrivial number of Americans to be in a state of self-contradiction, both supporting policies that legalize second and third trimester abortions and supporting the general prohibition of second and third trimester abortions. This indicates that there is a nontrivial number of Americans who either do not understand that _Roe v. Wade_ implies abortion is legal for any reason up until fetal viability or do not understand what legal abortion for any reason until fetal viability means.


### Capture of Democrats

While those who think abortion should be legal in all circumstances constitute a minority of Americans, something important for politics in the United States has occurred in recent years. From 1975 to 2013, among Democrats, legal abortion in all circumstances was also a minority viewpoint, just it was for all Americans regardless of party affiliation. However, starting in 2014, about an equal number of Democrats supported legal abortion in all circumstances as did those who supported legal abortion only in certain circumstances. Now, as of 2023, a clear _majority_ of Democrats, about 60%, support legal abortion _in all circumstances._

```{r data-Gallup-abortion-legality-by-party}
Gallup_abortion_legality_by_party <- tribble(
  ~Date, ~value, ~party, ~viewpoint,
  "1975-04-04", 19, 'Democrats', 'legal_any',
  "1979-02-23", 20, 'Democrats', 'legal_any',
  "1980-07-11", 25, 'Democrats', 'legal_any',
  "1981-05-08", 21, 'Democrats', 'legal_any',
  "1983-06-24", 23, 'Democrats', 'legal_any',
  "1988-09-25", 23, 'Democrats', 'legal_any',
  "1989-07-06", 32, 'Democrats', 'legal_any',
  "1990-04-05", 34, 'Democrats', 'legal_any',
  "1991-05-30", 36, 'Democrats', 'legal_any',
  "1991-09-05", 36, 'Democrats', 'legal_any',
  "1992-01-16", 31, 'Democrats', 'legal_any',
  "1993-03-12", 35, 'Democrats', 'legal_any',
  "1994-03-28", 34, 'Democrats', 'legal_any',
  "1994-09-06", 39, 'Democrats', 'legal_any',
  "1995-02-24", 36, 'Democrats', 'legal_any',
  "1995-09-22", 40, 'Democrats', 'legal_any',
  "1996-07-25", 28, 'Democrats', 'legal_any',
  "1996-09-20", 28, 'Democrats', 'legal_any',
  "1997-08-12", 26, 'Democrats', 'legal_any',
  "1997-11-06", 28, 'Democrats', 'legal_any',
  "1998-01-16", 23, 'Democrats', 'legal_any',
  "1999-04-30", 31, 'Democrats', 'legal_any',
  "2000-01-13", 33, 'Democrats', 'legal_any',
  "2000-03-30", 36, 'Democrats', 'legal_any',
  "2001-05-10", 31, 'Democrats', 'legal_any',
  "2002-05-06", 31, 'Democrats', 'legal_any',
  "2003-05-05", 32, 'Democrats', 'legal_any',
  "2004-05-02", 35, 'Democrats', 'legal_any',
  "2005-05-02", 32, 'Democrats', 'legal_any',
  "2006-05-08", 37, 'Democrats', 'legal_any',
  "2007-05-10", 38, 'Democrats', 'legal_any',
  "2008-05-08", 37, 'Democrats', 'legal_any',
  "2009-05-07", 33, 'Democrats', 'legal_any',
  "2010-05-03", 33, 'Democrats', 'legal_any',
  "2011-05-05", 38, 'Democrats', 'legal_any',
  "2012-05-03", 40, 'Democrats', 'legal_any',
  "2013-05-02", 39, 'Democrats', 'legal_any',
  "2014-05-08", 43, 'Democrats', 'legal_any',
  "2015-05-06", 42, 'Democrats', 'legal_any',
  "2016-05-04", 45, 'Democrats', 'legal_any',
  "2017-05-03", 46, 'Democrats', 'legal_any',
  "2018-05-01", 46, 'Democrats', 'legal_any',
  "2019-05-01", 39, 'Democrats', 'legal_any',
  "2020-05-01", 49, 'Democrats', 'legal_any',
  "2021-05-03", 50, 'Democrats', 'legal_any',
  "2022-05-02", 57, 'Democrats', 'legal_any',
  "2023-05-01", 60, 'Democrats', 'legal_any',
  "1975-04-04", 18, 'Republicans', 'legal_any',
  "1979-02-23", 19, 'Republicans', 'legal_any',
  "1980-07-11", 22, 'Republicans', 'legal_any',
  "1981-05-08", 24, 'Republicans', 'legal_any',
  "1983-06-24", 21, 'Republicans', 'legal_any',
  "1988-09-25", 23, 'Republicans', 'legal_any',
  "1989-07-06", 27, 'Republicans', 'legal_any',
  "1990-04-05", 29, 'Republicans', 'legal_any',
  "1991-05-30", 24, 'Republicans', 'legal_any',
  "1991-09-05", 28, 'Republicans', 'legal_any',
  "1992-01-16", 27, 'Republicans', 'legal_any',
  "1993-03-12", 26, 'Republicans', 'legal_any',
  "1994-03-28", 21, 'Republicans', 'legal_any',
  "1994-09-06", 28, 'Republicans', 'legal_any',
  "1995-02-24", 25, 'Republicans', 'legal_any',
  "1995-09-22", 25, 'Republicans', 'legal_any',
  "1996-07-25", 19, 'Republicans', 'legal_any',
  "1996-09-20", 17, 'Republicans', 'legal_any',
  "1997-08-12", 18, 'Republicans', 'legal_any',
  "1997-11-06", 23, 'Republicans', 'legal_any',
  "1998-01-16", 15, 'Republicans', 'legal_any',
  "1999-04-30", 16, 'Republicans', 'legal_any',
  "2000-01-13", 17, 'Republicans', 'legal_any',
  "2000-03-30", 21, 'Republicans', 'legal_any',
  "2001-05-10", 14, 'Republicans', 'legal_any',
  "2002-05-06", 17, 'Republicans', 'legal_any',
  "2003-05-05", 15, 'Republicans', 'legal_any',
  "2004-05-02", 12, 'Republicans', 'legal_any',
  "2005-05-02", 12, 'Republicans', 'legal_any',
  "2006-05-08", 16, 'Republicans', 'legal_any',
  "2007-05-10", 13, 'Republicans', 'legal_any',
  "2008-05-08", 12, 'Republicans', 'legal_any',
  "2009-05-07", 11, 'Republicans', 'legal_any',
  "2010-05-03", 14, 'Republicans', 'legal_any',
  "2011-05-05", 13, 'Republicans', 'legal_any',
  "2012-05-03", 9,  'Republicans', 'legal_any',
  "2013-05-02", 13, 'Republicans', 'legal_any',
  "2014-05-08", 14, 'Republicans', 'legal_any',
  "2015-05-06", 14, 'Republicans', 'legal_any',
  "2016-05-04", 12, 'Republicans', 'legal_any',
  "2017-05-03", 14, 'Republicans', 'legal_any',
  "2018-05-01", 11, 'Republicans', 'legal_any',
  "2019-05-01", 12, 'Republicans', 'legal_any',
  "2020-05-01", 13, 'Republicans', 'legal_any',
  "2021-05-03", 15, 'Republicans', 'legal_any',
  "2022-05-02", 10, 'Republicans', 'legal_any',
  "2023-05-01", 8,  'Republicans', 'legal_any',
  "1975-04-04", 51, 'Democrats', 'legal_only_certain',
  "1979-02-23", 53, 'Democrats', 'legal_only_certain',
  "1980-07-11", 53, 'Democrats', 'legal_only_certain',
  "1981-05-08", 49, 'Democrats', 'legal_only_certain',
  "1983-06-24", 56, 'Democrats', 'legal_only_certain',
  "1988-09-25", 55, 'Democrats', 'legal_only_certain',
  "1989-07-06", 50, 'Democrats', 'legal_only_certain',
  "1990-04-05", 52, 'Democrats', 'legal_only_certain',
  "1991-05-30", 48, 'Democrats', 'legal_only_certain',
  "1991-09-05", 45, 'Democrats', 'legal_only_certain',
  "1992-01-16", 52, 'Democrats', 'legal_only_certain',
  "1993-03-12", 52, 'Democrats', 'legal_only_certain',
  "1994-03-28", 51, 'Democrats', 'legal_only_certain',
  "1994-09-06", 49, 'Democrats', 'legal_only_certain',
  "1995-02-24", 49, 'Democrats', 'legal_only_certain',
  "1995-09-22", 44, 'Democrats', 'legal_only_certain',
  "1996-07-25", 60, 'Democrats', 'legal_only_certain',
  "1996-09-20", 48, 'Democrats', 'legal_only_certain',
  "1997-08-12", 59, 'Democrats', 'legal_only_certain',
  "1997-11-06", 55, 'Democrats', 'legal_only_certain',
  "1998-01-16", 59, 'Democrats', 'legal_only_certain',
  "1999-04-30", 49, 'Democrats', 'legal_only_certain',
  "2000-01-13", 53, 'Democrats', 'legal_only_certain',
  "2000-03-30", 47, 'Democrats', 'legal_only_certain',
  "2001-05-10", 55, 'Democrats', 'legal_only_certain',
  "2002-05-06", 49, 'Democrats', 'legal_only_certain',
  "2003-05-05", 51, 'Democrats', 'legal_only_certain',
  "2004-05-02", 47, 'Democrats', 'legal_only_certain',
  "2005-05-02", 52, 'Democrats', 'legal_only_certain',
  "2006-05-08", 51, 'Democrats', 'legal_only_certain',
  "2007-05-10", 49, 'Democrats', 'legal_only_certain',
  "2008-05-08", 50, 'Democrats', 'legal_only_certain',
  "2009-05-07", 50, 'Democrats', 'legal_only_certain',
  "2010-05-03", 50, 'Democrats', 'legal_only_certain',
  "2011-05-05", 47, 'Democrats', 'legal_only_certain',
  "2012-05-03", 42, 'Democrats', 'legal_only_certain',
  "2013-05-02", 47, 'Democrats', 'legal_only_certain',
  "2014-05-08", 42, 'Democrats', 'legal_only_certain',
  "2015-05-06", 43, 'Democrats', 'legal_only_certain',
  "2016-05-04", 43, 'Democrats', 'legal_only_certain',
  "2017-05-03", 44, 'Democrats', 'legal_only_certain',
  "2018-05-01", 40, 'Democrats', 'legal_only_certain',
  "2019-05-01", 45, 'Democrats', 'legal_only_certain',
  "2020-05-01", 42, 'Democrats', 'legal_only_certain',
  "2021-05-03", 41, 'Democrats', 'legal_only_certain',
  "2022-05-02", 38, 'Democrats', 'legal_only_certain',
  "2023-05-01", 34, 'Democrats', 'legal_only_certain'
) |>
  mutate(
    across(Date, ymd)
  )
```

```{r plot-Gallup-abortion-legal-under-any-by-party, cache=TRUE, fig.cap="Proportions of Democrats, of Republicans, and of all Americans who responded 'abortions should be legal under any circumstances' to Gallup polling. [@gallup_abortion_by_party] [@gallup_abortion_in_depth]"}
Gallup_abortions_should_be_legal |>
  select(Date, legal_any) |>
  rename(value = legal_any) |>
  mutate(party = 'all', viewpoint = 'legal_any') |>
  bind_rows(Gallup_abortion_legality_by_party) |>
  filter(viewpoint == 'legal_any') |>
  mutate(
    value = value / 100
  ) |>
  ggplot(
    mapping = aes(
      x = Date,
      y = value,
      color = party,
      shape = party
    )
  ) +
    geom_line(
      alpha = 0.6,
      linewidth = 1
    ) +
    geom_point(
      alpha = 1,
      fill = 'white',
      size = 0.8
    ) +
    scale_y_continuous(
      breaks = \(x) seq(0, x[2], 0.1),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_percent(),
      limits = c(0, NA)
    ) +
    scale_x_date(
      breaks = scales::breaks_width('2 years', offset = '1 year'),
      expand = expansion(add = 365),
      labels = scales::label_date(format = '%Y')
    ) +
    scale_color_brewer(
      label = c(
        all = 'All Americans',
        Democrats = 'Democrats',
        Republicans = 'Republicans'
      ),
      type = 'qual'
    ) +
    scale_shape_manual(
      label = c(
        all = 'All Americans',
        Democrats = 'Democrats',
        Republicans = 'Republicans'
      ),
      values = 21:23
    ) +
    labs(
      x = 'Year',
      y = 'Percent of respondents'
    )
```

```{r plot-Gallup-abortion-legality-Democrats, cache=TRUE, fig.cap="Proportions of Democrats who responded either 'abortions should be legal under any circumstances' or 'legal only under certain circumstances' to Gallup polling. [@gallup_abortion_by_party] "}
Gallup_abortion_legality_by_party |>
  filter(party == 'Democrats') |>
  mutate(
    value = value / 100
  ) |>
  ggplot(
    mapping = aes(
      x = Date,
      y = value,
      color = viewpoint,
      shape = viewpoint
    )
  ) +
    geom_line(
      alpha = 0.6,
      linewidth = 1
    ) +
    geom_point(
      fill = 'white',
      size = 0.8
    ) +
    scale_y_continuous(
      breaks = \(x) seq(0, x[2], 0.1),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_percent(),
      limits = c(0, NA)
    ) +
    scale_x_date(
      breaks = scales::breaks_width('2 years', offset = '1 year'),
      expand = expansion(add = 365),
      labels = scales::label_date(format = '%Y')
    ) +
    scale_color_brewer(
      label = c(
        legal_any = 'Legal under any',
        legal_only_certain = 'Legal only under certain'
      ),
      type = 'qual'
    ) +
    scale_shape_manual(
      label = c(
        legal_any = 'Legal under any',
        legal_only_certain = 'Legal only under certain'
      ),
      values = 21:22
    ) +
    labs(
      x = 'Year',
      y = 'Percent of respondents'
    )
```

Thus, while the viewpoint that abortion should be legal in all circumstances is a minority viewpoint among Americans, this viewpoint now dominates the Democrats. Since there are only two political parties in the United States that regularly win elections, the legal-abortion-in-all-cases platform of the abortion lobby has an outsized impact on American politics because it controls one of the two main political parties.


## Recommendations for Majority Voter Strategy

Democrats as a party have been captured by the minority viewpoint that abortion should be legal in all circumstances. As we have seen, the majority of Americans want abortion to be legal only in certain circumstances, and the majority of Americans think that abortion should be generally illegal in the second and third trimesters. The abortion lobby, on the other hand, has legalized abortion for any reason into the third trimester with _Roe v. Wade_ and with Ohio's _Issue 1._

If Republicans can successfully campaign for the majority voter, then abortion policy will be an issue of strength for them. However, Democrats have the advantage of being able to excite a very passionate minority, which can cause an increase in voter turnout and be successful in elections by way of the polar voter strategy.

Next year during the 2024 elections, Americans will vote for a President of the United States. Such elections usually have higher voter turnout, which may decrease the effectiveness of the polar voter strategy and increase the effectiveness of the majority voter strategy.

Regardless of whether any given election is a Presidential election or not, candidates using the majority voter strategy must adhere to several strategies in order for abortion policy to be an asset, not a detriment, to their campaigns.


### Reject an All-or-Nothing Framing

A tactic that the abortion lobby has been successful in using is to frame the abortion issue as an all-or-nothing question between abortion being illegal in all circumstances or abortion being legal in all circumstances. As Gallup polling indicates, a minority of Americans think that abortion should be illegal _in all circumstances._ Thus, if the way the issue is framed to voters is this dichotomy, then abortion advocates can cobble together a majority to vote for abortion being legal in all circumstances. However, a minority of Americans _actually want_ abortion to be legal in all circumstances.

Politicians campaigning for the majority voter must be very careful, therefore, not to allow the issue to be framed as whether abortion should be legal or be banned. Instead, they must frame abortion policy as being a question of _under what circumstances_ abortion is to be legal and _under what circumstances_ it is to be illegal.


### Find the Majority Voter

This leads to the next important point. It is imperative for politicians campaigning for the majority voter on abortion policy to _discover_ under what circumstances a majority in their electorate want abortion to be illegal. Since abortion policy is made by the legislatures of the several states in the United States, this must be done at the state level.

This is admittedly more challenging than the all-or-nothing framing that is working for the abortion lobby, but it can be accomplished simply by taking a survey of a representative random sample of registered voters in a given state. Furthermore, polling at the national level can give a political campaign hints on where to begin. Since a majority of Americans think that second and third trimester abortions should be generally illegal, it is likely that the electorate in many states want some kind of gestational age limit on abortion. It is therefore a matter of polling to find out the specific gestational age limit that is desired by majority of voters in a given state.

Setting abortion policy to the majority will was a process that was occurring in the 1960s in the state legislatures, but was overridden by the _Roe v. Wade_ and _Doe v. Bolton_ decisions. For instance, several states had allowed for exceptions to their criminal abortion statutes for cases such as rape or severe fetal deformity, while rejecting attempts to legalize abortion as a method of fertility control. [] It remains to be seen what are the circumstances under which majorities in various states want abortion to be prohibited, but this can be discovered.

In addition to outright prohibition, there are likely many _regulations_ of abortion that a majority of voters in a given state approve.[^regulations] For instance, Gallup polling suggests that a policy requiring abortion providers to provide information about alternatives to abortion has historically been quite popular at the national level in the United States. There are probably numerous regulations of various kinds that would be found to be popular in any given state.

[^regulations]: In addition to legalizing abortion until fetal viability, _Roe v. Wade_ overturned any _regulations_ of abortions done in the first trimester. 

```{r}

```


### Create Simple Messaging

The abortion lobby has a stupidly simple message that they are promulgating that abortion is a "right" and that nefarious evildoers want to "take away your rights." This is an easy message on which to campaign. On the other hand, the Combined Argument and Explanation Against _Issue 1_ included in the Ohio ballots was scattered and unfocused. Campaigning for the majority voter on abortion policy points the way to a much more streamlined message.

This message should be something to the effect of, "Elites used unelected judges who abused their Constitutional authority in order to override the will of the American people and overturn laws in all fifty states pertaining to abortion policy. More recently, the Supreme Court returned matters of abortion policy back to American people and their duly elected representatives. I, as a candidate for ______ office, will fulfill my duty to represent the people of the state of ______. The majority of voters in the state of ______ want..."

Knowledge of the circumstances under which the the majority of voters in a given state want abortion to be illegal and of regulations approved by a majority of voters, as discussed in the previous section, could then be used to substantiate this message.


## Consideration of Polar Voter Strategy

Some politicians opposed to abortion argue that election losses such as those experienced in Virginia are due to use of the majority voter strategy, which has led to lower voter turnout among voters opposed to abortion. There is logic to this. It is highly likely that the abortion lobby, where it has won, has won based on voter turnout.

The abortion lobby has not represented the majority of Americans, but has had political successes. Indeed, rather than following public opinion, the abortion lobby appears to have changed public opinion in its favor by first imposing abortion policy against the will of the majority of Americans by way of _Roe_ and _Doe_ and then worrying about public opinion afterwards.

As previously mentioned, elections in 2024 will be during a Presidential election, which typically has much higher turnout. Therefore, in the short term, strategies which aim to increase turnout at the expense of attracting majority voters may be not effective next year.

In the long term, the main issue with using the polar voter strategy _in opposition_ to abortion is its strong potential to backfire. Indeed, the fact that the abortion lobby has had success using the polar voter strategy is indicative _both_ of how it can be used successfully _and_ that it is currently being used successfully by the abortion lobby. Framing abortion policy as an all-or-nothing question is a tactic with which the abortion lobby has had demonstrable success. Appealing to enthusiastic minorities opposed to abortion risks creating an artificial majority for the abortion lobby's platform of legal abortion in all circumstances.

The polar voter strategy trades in broader appeal among the electorate for increases in turnout among passionate voters. Whether or not this trade is a net positive for a candidate is heavily contingent on the views of voters in the candidate's electorate and the propensity of voters to vote. Thus, it is important for the candidates opposed to abortion to understand the views on abortion policy of the voters in the candidates' electorate, regardless if the majority voter strategy or the polar voter strategy is ultimately used.


## Do Not Be Scared of Abortion Policy

Specific strategy aside, candidates for political office who hope to succeed should run their campaigns with confidence. I have noticed several Republican politicians approach the abortion issue with trepidation and almost apologetically. This is a bad appearance for any candidate for public office. People can sense when others lack confidence, and this lack of confidence does not inspire confidence in others.

One of the advantages of the majority voter strategy is that a candidate is armed with the knowledge that the candidate's platform is aligned with the majority of voters in the candidate's state. Candidates using the polar voter strategy must express personal conviction genuinely and appeal to an excited minority with this conviction. Either way, a candidate hemming and hawing about one's platform on abortion does not flatter a candidate.


## Conclusion

A majority of Democrats want abortion to be legal in all circumstances, which is a view shared by only a minority of Americans. The majority of Americans want abortion to be legal only in certain circumstances and want abortion to be generally illegal in the second and third trimester. Any candidate for public office who can successfully campaign for the majority voter by rejecting an all-or-nothing framing, by identifying the abortion policy desired by a majority of voters in the candidate's electorate, and by messaging this policy in a straightforward and confident manner can exploit this weakness in Democrats' platform.

It may be the case that, in certain constituencies, candidates opposed to abortion can succeed by using a strategy of campaigning for the polar voter, in which a passionate minority wins an election by way of increasing voter turnout. However, this has a strong potential to backfire, given the recent successes of the abortion lobby itself using such a strategy.

Regardless of the particular strategy used to win elections, it is important to remember that politics is only one aspect of human civilization and only one aspect of opposition to abortion. This is ultimately not a struggle over the laws of any one country in any one time and place, but a struggle for the hearts and minds of humanity. Infanticide was not pushed to the extreme margins of society by laws alone. There is a danger in getting tunnel vision around the politics of abortion. Such narrow focus can cause us to ignore better opportunities to use limited resources that would have made more progress toward putting abortion into the history books.

Nonetheless, politics does have _some_ role to play in this process, and unlike other fields, politics only works for your social goals if you win. Therefore, those opposed to abortion who are actively engaged in the political sphere -- politicians, lobbyists, lawyers -- must be very careful how they approach political strategy.

Finally, one of the more overlooked functions of the role of those engaged in the political realm who do win is in funding those aspects of society that prevent abortions and that are in need of much more funding, the most prominent being childcare support and prevention of unintended pregnancy. This is particularly relevant for those elected to legislative bodies in the United States, because in the United States legislatures control the budget and appropriate funds. 

"Childcare" should be construed broadly can mean anything from day care to adoption, depending on the case. Better support for parents, especially those with unexpected pregnancies, is a vital need both in the short term for any individual child and in the long term for changing attitudes toward abortion. Exactly how much of this should be market-based versus state-based and whether policies should be construed as safety nets or entitlements are details that should be informed empirically, but the overall goal of supporting parents, especially parents in need, is clear.

The birth control movement in the United States was founded in direct opposition to abortion as a method of fertility control. However, birth control organizations such as Planned Parenthood have been stolen by the very abortion industry [their founders were attempting to put out of business](/margaret-sanger-on-abortion-in-her-own-words.html). There have been some attempts to stop giving government funding to Planned Parenthood. We need an organization to fund instead, one dedicated to preventing unintended pregnancy exclusively through nonviolent means. The prevention of two unintended pregnancies prevents one abortion on average, and _every_ unintended pregnancy prevented reduces the motivation leveraged by the abortion lobby in order to generate support for abortion.


## Citations

::: {#refs}
:::


## Footnotes
