---
author: "Joshua Born"
bibliography: "../bibliographies/professional-class-cultural-hegemony.bib"
csl: "../bibliographies/chicago-author-date.csl"
date: 2024-10-17
draft: false
include-before: |
  ## Table of Contents
output:
  blogdown::html_page:
    number_sections: false
    toc: true
    toc_depth: 5
summary: "The number of women of an age associated with childbearing in the United States from 1940 to 2019 is quantified and discussed."
tags: ["analysis", "reproductive responsibility", "third-person"]
title: "Women of Reproductive Age in the United States, 1940-2019"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(here)
library(showtext)

font_add_google('Open Sans', 'open')
showtext_auto()
showtext_opts(dpi = 200)
theme_update(text = element_text(family = 'open'))

knitr::opts_chunk$set(
  cache = TRUE,
  dpi = 300,
  echo = FALSE,
  fig.align = 'center',
  fig.height = 3.75,
  fig.retina = 1,
  fig.width = 5
)
```


## Introduction

The number of women of reproductive age is an often-used quantity in fertility studies, most commonly as a denominator in calculating rates and ratios. "Reproductive age" is typically taken in fertility studies to be between 15 years and 44 years of age.


## Data Source

Data were taken from the United States Census Bureau's Population Estimates Program (PEP).

The United States has a decennial[^decennial] census that is mandated by the U.S. Constitution and is taken as being the authoritative count of the population of the United States.

To fill in the population estimates for years in which a decennial census was not taken, the PEP calculates postcensal estimates and intercensal estimates. Postcensal estimates take the most recent decennial census as a starting point and adjust population estimates based on change during the time after the decennial census. Intercensal estimates are generally more accurate than postcensal estimates because they are based on both an earlier decennial census and a later decennial census, calculating population estimates in between the two.

[^decennial]: The census is "decennial" in that it is done once every decade, on the turn of the decade, in 1970, 1980, _etc._

The 2010-2020 intercensal estimates are not yet available. Therefore, the below estimates of women of reproductive age in the United States are based on intercensal estimates for the period of 1940 to 2009 and postcensal estimates for 2010 to 2019.

Furthermore, the only estimates available from the PEP for the time period before 1980 are estimates of both the resident population of the United States _and_ members of the United States armed forces living abroad. Members of the armed forces living abroad are not included in the target population of the National Survey of Family Growth, which is the main instrument used for analysis here, and therefore should not be included in the numbers used for denominators.

However, these are the only data available from the PEP for before 1980. For 1980 and after, the PEP provides estimates for the resident population of the United States not including members of the armed forces living abroad, and so these are used for 1980 and after. Fortunately, the inclusion of members of the armed forces living abroad in the estimates for before 1980 do not appear to result in a substantial discontinuity.

Data for the period 1940-1979 were taken from Excel files available [here](https://www.census.gov/data/tables/time-series/demo/popest/pre-1980-national.html). Data for the period 1980-1989 were taken from fixed-width format text files available [here](https://www.census.gov/data/datasets/time-series/demo/popest/1980s-national.html). Data for the period of 1990-2019 were taken from the [Census API](https://www.census.gov/data/developers.html), using the "Resident Population by Single Year of Age and Sex" from the [1990-2000 Intercensals](https://www.census.gov/data/developers/data-sets/popest-popproj/popest.1990-2000_Intercensals.html#list-tab-2014455046), the "Demographic Characteristics Estimates by Single Year of Age" from the [2000-2010 Intercensals](https://www.census.gov/data/developers/data-sets/popest-popproj/popest.2000-2010_Intercensals.html#list-tab-2014455046), and the "Demographic Characteristics Estimates by Single Year of Age and Sex" from the [Vintage 2019 Postcensals](https://www.census.gov/data/developers/data-sets/popest-popproj/popest.Vintage_2019.html#list-tab-2014455046).


## Plots

```{r load-data}
women_of_reproductive_age <- readRDS(here('data/Census/women_of_reproductive_age_1940-2019.Rds'))
```

```{r total-women-of-reproductive-age-plot, dependson="load-data", fig.cap="Total number of women aged 15 years to 44 years in the United States, based on estimates from the U.S. Census' Population Estimates Program. Estimates left of the dashed line include both the resident population of the United States and members of the armed forces overseas, whereas estimates right of the dashed line include only the resident population of the United States."}
women_of_reproductive_age |>
  summarize(Population.Female = sum(Population.Female), .groups = "drop") |>
  ggplot(
    mapping = aes(
      x = Year,
      y = Population.Female
    )
  ) +
    geom_vline(
      xintercept = 1979.5,
      linetype = 2,
      linewidth = 0.25
    ) + 
    geom_line(
      alpha = 0.4,
      color = 'blue',
      linewidth = 1.25
    ) +
    geom_point(
      color = 'blue',
      fill = 'white',
      size = 1.25,
      shape = 21
    ) +
    scale_shape_manual(
      values = 21
    ) +
    scale_y_continuous(,
      breaks = \(x) seq(0, x[2]+1e6, 2e6),
      labels = scales::label_number(suffix = "M", scale = 1e-6),
    ) +
    scale_x_continuous(
      breaks = \(x) seq(x[1]+1, x[2]-1, 2),
      expand = expansion(add = 1),
      minor_breaks = \(x) seq(x[1]+1, x[2]-1, 1)
    ) +
    labs(
      x = 'Year',
      y = 'Female Population of United States of Reproductive Age (15-44 years)'
    ) +
    theme_classic() +
    theme(
      axis.line.x.bottom = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.line.y.left = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.text.x = element_text(
        angle = 45,
        color = 'black',
        hjust = 0.5,
        vjust = 0.65
      ),
      axis.text.y = element_text(
        color = 'black',
        margin = margin(r = 0)
      ),
      axis.ticks.y = element_blank(),
      axis.ticks.x = element_line(
        color = 'black',
        linewidth = 0.2
      ),
      axis.title.x = element_text(
        margin = margin(t = 2)
      ),
      axis.title.y = element_text(
        margin = margin(r = 10)
      ),
      panel.grid.major = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor.x = element_blank(),
      plot.margin = margin(t = 0)
  )
```

Figure \@ref(fig:total-women-of-reproductive-age-plot) shows that the number of women of reproductive age in the United States has been increasing throughout the analysis period. Thus, even if rates of live birth, induced abortion, _etc.,_ were to remain fixed, it would be expected that the total numbers of births, abortions, _etc.,_ would increase.

The increase in the number of women of reproductive age is roughly "S" shaped, with a period of more rapid increase in between 1960 and 1990 and a plateau of population increase in the 2000s.


```{r women-of-reproductive-age-by-age-group-plot, dependson="load-data", warning=FALSE, fig.height=4, fig.cap="Number of women aged 15 years to 44 years in the United States by age group, based on estimates from the U.S. Census' Population Estimates Program. Estimates left of the dashed line include both the resident population of the United States and members of the armed forces overseas, whereas estimates right of the dashed line include only the resident population of the United States."}
women_of_reproductive_age |>
  ggplot(
    mapping = aes(
      x = Year,
      y = Population.Female,
      color = Age.Group,
      shape = Age.Group
    )
  ) +
    geom_vline(
      xintercept = 1979.5,
      linetype = 2,
      linewidth = 0.25
    ) +
    geom_line(
      alpha = 0.4,
      linewidth = 0.75
    ) +
    geom_point(
      fill = 'white',
      size = 0.75
    ) +
    scale_y_continuous(,
      breaks = \(x) seq(0, x[2]+1e6, 5e5),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_number(suffix = "M", scale = 1e-6),
      limits = c(4e6, NA),
      minor_breaks = NULL
    ) +
    scale_x_continuous(
      breaks = \(x) seq(x[1]+1, x[2]-1, 2),
      expand = expansion(add = 1),
      minor_breaks = \(x) seq(x[1]+1, x[2]-1, 1)
    ) +
    scale_color_brewer(
      palette = 2,
      type = "qual"
    ) +
    guides(color = guide_legend(nrow = 1)) +
    labs(
      x = 'Year',
      y = 'Female Population of United States by Age Group'
    ) +
    theme_classic() +
    theme(
      axis.line.x.bottom = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.line.y.left = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.text.x = element_text(
        angle = 45,
        color = 'black',
        hjust = 0.5,
        vjust = 0.65
      ),
      axis.text.y = element_text(
        color = 'black',
        margin = margin(r = 0)
      ),
      axis.ticks.y = element_blank(),
      axis.ticks.x = element_line(
        color = 'black',
        linewidth = 0.2
      ),
      axis.title.x = element_text(
        margin = margin(t = 2)
      ),
      axis.title.y = element_text(
        margin = margin(r = 10)
      ),
      legend.box.margin = margin(t = 0, r = 0, b = -10, l = 0, unit = "pt"),
      legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
      legend.position = 'top',
      legend.title = element_blank(),
      panel.grid.major = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor.x = element_blank(),
      plot.margin = margin(t = 0)
  )
```

The number of women of reproductive age in the United States is broken down by age group in Figure \@ref(fig:women-of-reproductive-age-by-age-group-plot). There appear to be several cohorts in which population size reached a local maximum. The most prominent are women who were aged 15-19 years in 1976 and thus born some time in 1957 to 1961. The local maximum from this cohort is reflected through all six age groups in successive years after 1976.

Additionally, another local maximum appears among women aged 15 to 19 in 2008, consisting of women born in the period between 1989 and 1993. Within the analysis period, this is reflected again among women aged 20 to 24 and among women aged 25 to 29.

The patterns in the change of the teenage population are of interest for analysis of reproductive responsibility because the teenage population is particularly susceptible to unintended pregnancy due to a combination of high fecundity and behavioral immaturity. Thus, if an apparent decline in unintended pregnancies in the 1980 to 1992 time period were discovered, the possibility would still remain that this reflects only a decline in the number of teenagers in this time period.


## Editorial Note

Fertility surveys and analyses of reproduction often focus on women. On the one hand, this is expedient. The female population of reproductive age is the limiting factor in fertility: ten women and one man could conceivably have ten children in one year, but ten men and one woman cannot. Furthermore, because of the more intimate relationship women have with pregnancy than do men, women can perhaps be expected to have more accurate recall and reporting of events pertaining to reproduction than men.

On the other hand, the focus on women in analyses of reproduction has the unfortunate effect of framing reproduction as a "women's issue" and thus abdicating male reproductive responsibility. This is particularly unfortunate because the United States is going through a period in which more and more children are growing up fatherless, to the detriment of children's outcomes. [@kearney_2023]

Furthermore, every unintended pregnancy is not just the result of the knowledge, attitudes, and behavior of a woman, but also the knowledge, attitudes, and behavior a man. Expectations of reproductive responsibility from men should be equal to expectations from women. This is the case, if not for moral reasons, then at least for practical ones.

Unfortunately, most of the data that are available for the analysis of reproductive responsibility are focused on women. Where there is corresponding data available for men, analyses will be repeated for men. Otherwise, this editorial note will be repeated as a footnote in subsequent analyses.


## Citation

::: {#refs}
:::


## Footnote
