---
title: "The Fallacy of Generic Thinking"
date: 2022-02-16
draft: true
tags: ["reasoning about evidence", "fallacies", "statistics", "psychology", "philosophy", "third-person"]
bibliography: "../bibliographies/fallacy-of-generic-thinking.bib"
csl: "../bibliographies/chicago-author-date.csl"
link-citations: true
implicit-figures: true
output:
  blogdown::html_page:
    md_extensions: ["+footnotes"]
    toc: true
    toc_depth: 5
include-before: |
  <p>Introductory paragraph goes here.</p>
  <!--more-->
  ## Table of Contents
---

```{r setup, echo=FALSE, message=FALSE}
library(data.table)
library(knitr)

opts_chunk$set(
  echo = FALSE
)
```


## Introduction

What unites the many types of [fallacies](/tags/fallacies.html) that this blog discusses is that they would appeal to a mind afflicted with what can be termed "generic thinking." For instance, the article "[How to Scrutinize a Statistic](/how-to-scrutinize-a-statistic.html)" discussed both the need to evaluate [practical significance](/how-to-scrutinize-a-statistic.html#practical-significance) and how this is typically done by quantifying an [effect size](/how-to-scrutinize-a-statistic.html#effect-size). The absence of summarizing an effect size was described as a fallacy of "[generic comparisons](http://localhost:4321/how-to-scrutinize-a-statistic.html#generic-comparisons)."

The phrase "generic comparison" is actually a neologism invented by this blog, but the word "generic" was used deliberately to allude to what is termed "generic generalization," a phrase established in the academic literature.

Generic thinking comprises both generic generalizations and generic comparisons and encompasses all beliefs about an inherently quantitative subject, but that are held without any quantification. This often occurs when someone is thinking about large populations of individuals. For instance, social issues occur, by their very nature, in large populations of individual human beings. Thus, all unquantified social criticism is generic thinking.


## Generic Generalization

In a generic generalization, a property is predicated of a subject that denotes a population of individuals, but no quantification is given.

"All swans are white," "some swans are white," "no swans are white," "approximately 74.2% of swans are white" are all beliefs about a subject &ndash; denoted by "swans" &ndash; that includes a large number of individuals. All of these statements include some kind of quantification, namely "all," "some," "no," or "approximately 74.2%." However, the statement "swans are white" is a generic generalization because it is unquantified.

A body of scholarship pertaining to generic generalizations exists in both academic philosophy and academic psychology. Generic generalizations have been studied by philosophers because they can seem like a logic puzzle. For instance, statements like "dogs have four legs," "ducks lay eggs," "mosquitoes carry the West Nile virus" seem like they have the spirit of truth to them.[^examples]  However, they are all false for a portion of the population they discuss.

[^examples]: "Dogs have four legs" and "ducks lay eggs" are examples used in "Generic Generalizations" in the _Stanford Encyclopedia of Philosophy_ [@leslie_generic_2016] and the West Nile virus example is used in "The Original Sin of Cognition" [@leslie_original_2017].

While it is true that _most_ "dogs have four legs," there are some dogs happily scampering around on three legs. Only female ducks who have reached sexual maturity even have the ability to lay eggs, so "ducks lay eggs" is only true for less than half the population it describes. Mercifully, a very small proportion of mosquitoes actually carry the West Nile virus.

![
A three-legged dog, the existence of which is a false case of the generic generalization "dogs have four legs."<br/>
["Three Legged Dog"](https://www.flickr.com/photos/23743320@N03/7230816680)
by
[Sangudo](https://www.flickr.com/photos/sangudo/)
is licensed under
[CC BY-NC-ND 2.0](https://creativecommons.org/licenses/by-nc-nd/2.0/)
and is not modified.
](/images/three-legged-dog.jpg)


### Resistance to Falsification

Much of the philosophizing around generic generalizations has been an attempt to find some way of interpreting generic generalizations that would allow the ones that seem plausible to be true and the ones that are patently false to be false. [@leslie_generic_2016] This highlights precisely the main issue with generic generalizations: they seem to defy truth conditions.

In particular, generic generalizations are resistant to falsification.  Quantified beliefs have conditions under which the holder of the belief can learn that the belief is false, if it is indeed false. If someone who asserts "all swans are white" encounters a single black swan, then the individual knows the belief is false. Similarly, presenting just one white swan to someone who asserts "all swans are black" demonstrates to this person that the belief is false.

While it is easy to show the assertion "some swans are white" is true since observation of one white swan confirms the belief, falsification is more difficult. To demonstrate the belief is false without any uncertainty requires observations of all the swans in the world. However, this is just a practical matter. In principle, it is known how to demonstrate the belief is false, even if demonstrating it is false would require more work than most are willing to commit.

Statements such as "approximately 72.4\% of swans are white" can be demonstrated false, given a specified [hypothesis test level](/how-to-scrutinize-a-statistic.html#fallacies). This is slightly more complicated by the fact that for a continuous variable any such hypothesis can be arbitrarily rejected given a large enough sample size and enough statistical power, so exact statistical thresholds would have to be stipulated. Once these are all specified, however, such hypotheses can be evaluated as to their truth or falsity according to standard methods.

Thus, quantified beliefs that are false can be demonstrated as false. The same is not the case for generic generalizations. If someone asserts "dogs lay eggs," how many viviparous dogs must be observed before this belief is demonstrated false? Similarly, how many two-legged ducks must be brought to someone who asserts "ducks have four legs" before the holder of the belief must relent? The stubborn can merely retort "not _all_ dogs lay eggs" or "I didn't mean _all_ ducks have two legs." Generic generalizations thus create a threshold that can always be conveniently moved to justify whatever arbitrary belief is desired.

Such resistance to falsification is a nontrivial matter because bigoted attitudes are apt to be construed in terms of generic generalizations. For instance, Leslie [-@leslie_original_2017] uses "Muslims are terrorists" and "Blacks are rapists" as examples of stereotypes formulated as generic generalizations. While these are negative stereotypes, positive stereotypes such as "men are good at math" or "women are empathic" are also often construed as generic generalizations. Thus, generic generalizations are tools for all sorts of [category-based prejudice](/category-based-prejudice.html).

To those concerned about bigotry, this resistance to falsification is a troubling property of generic generalizations because it means that once a negative stereotype has been formulated as a generic generalization in someone's mind, it is that much more difficult to convince the holder of the negative stereotype to change the belief.

If someone were to assert "most Muslims are terrorists" or "80% of Blacks are rapists," these beliefs could be demonstrated false with the appropriate empirical observations and analyses. When someone asserts the generic generalizations "Muslims are terrorists" or "Blacks are rapists," it does not matter _how many_ Muslims are not terrorists or Blacks are not rapists, the generic generalization is resistant to falsification exactly because it is unquantified.


### Uninformativeness

In addition to being resistant to falsification, generic generalizations are not informative because, given a large enough population, an arbitrary number of [confirmatory examples](selection-bias-and-the-fallacy-of-listing-examples.html) can always be found.

For instance, generic generalizations such as "humans are terrorists" or "humans are rapists" seem odd. However, they have a similar amount of actual information as the more specific generic generalizations "Muslims are terrorists" and "Blacks are rapists."

Such generic generalizations about "humans," "Muslims," or "Blacks" assert that, among some population of billions of individuals, there are some unspecified number of individuals who are terrorists or who are rapists. However, this is true of _any_ large enough population of human beings.

The generic generalizations "humans are terrorists" or "humans are rapists" are conspicuous in just how vague and uninformative they are. However, this is precisely the same issue with all generic generalizations. The "humans" versions of these generic generalizations are merely more obvious in illustrating this defect.

Given a large population of humans, there will be among them some unspecified number of terrorists, rapists, charitable givers, volunteers, slalom skiers, stamp collectors, or anything else for that matter. This is a reasonable expectation _a priori_ before any observations are made. Indeed, if a claim were made that a population of millions of human beings did not contain a single rapist, charitable giver, or stamp collector, such a claim would be treated incredulously. Generic generalizations are therefore simply not informative.


### Empirical Study

The fallacy of generic generalization is not just a philosophical concern, thanks in large part to the work of developmental psychologist Susan Gelman, who discovered that people become sensitive to generic generalizations at a young age. [@gelman_childrens_2002]

Psychological research in this area has gone on to demonstrate empirically that the fallacious use of generic generalizations does exist. For instance, Cimpian, Brandone, and Gelman [-@cimpian_generic_2010] found that test subjects accepted generic generalizations at low levels of prevalence, but interpreted them as if they implied high levels of prevalence. Thus, generic generalizations were evaluated for truth or falsity closer to "some" quantification, but interpreted as if they were closer to "all" quantification.

**_[Discuss Cimpian et al here in more detail, at least to the level of Anecdotes Are Not Evidence.]_**


### Bare Plurals and Elision

Because generic generalizations are resistant to falsification, uninformative, and have been found empirically to lead to flawed thinking, generic generalizations as mere fallacies with no merit. Therefore, this article recommends that everyone strive to remove generic generalizations from their thinking.

This is a value judgment that implies that the philosophical program of attempting to salvage generic generalizations is not worthwhile. One might protest, what about the generic generalizations that feel like there is a ring of truth to them, such as "dogs have four legs," "ducks lay eggs," and "mosquitoes carry the West Nile virus?"

Such examples in this article have all been what are called "bare plurals," in which the subject of a sentence is a plural noun that is not proceeded by any modifier word. However, not all uses of bare plurals are truly generic generalizations.

This is due to the messy nature of natural languages. In particular, human beings are prone to make their communication more concise by elision. Sometimes just a few sounds are left out as in the case of a contraction, but whole words, phrases, and clauses can be elided to be inferred from context.

Commonly, a bare plural is used when "all" quantification is implied. For instance, in this very article, it is written that "generic generalizations are resistant to falsification." This is not, however, a generic generalization about generic generalizations. From context, it should be gleaned that "all" quantification is implied, as in "all generic generalizations are resistant to falsification."

If one went about one's day putting "all" in front of each and every universally quantified statement, it would sound strange to native English speakers. In many contexts such as this one, using the word "all" might be used for emphasis, but it is not necessary for the English language to function.

Once one accepts that natural languages function with such elision, then whether or not any given bare plural is a generic generalization becomes a _question,_ not an automatic assumption.

When someone says "ducks lay eggs," there is typically social context surrounding this statement from which a more complicated thought can be inferred, such as "the system of reproduction used by ducks is oviparous."

![
A child with an adult encountering ducks swimmign around a pond.<br/>
[Image](https://pixabay.com/photos/people-adult-autumn-back-boy-2942837/?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=2942837)
by
[freestocks-photos](https://pixabay.com/users/freestocks-photos-7014431/?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=2942837)
from
[Pixabay](https://pixabay.com/?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=2942837).
](/images/duck-pond.jpg)

For instance, a child might ask if a duck found in a pond will give birth to little chicks that can swim about, to which an adult might reply, "no, ducks lay eggs." Given context, the longer statement about reproductive systems is implied, but rarely do English speakers talk in this way, and instead this is shortened to "ducks lay eggs."

Furthermore, one could assert "a dog without genetic aneuploidy and absent any physical trauma, allowed to develop in a salubrious environment including nutritional needs, grows into an organism with four legs." That is a lot of words, which can be elided to the simple statement "dogs have four legs" given social context.

Similarly, those who claim to have gotten West Nile virus from an ant bite might be curtly told, "no, mosquitoes carry the West Nile virus." It is likely in this case the proposition "mosquitoes are the only species that function as a vector for West Nile virus transmission in humans" has simply been shortened to "mosquitoes carry the West Nile virus."


### How to Avoid the Fallacy

One might protest that perhaps all generic generalizations could be explained as instances of this kind of elision. Such a protest is consistent with this article's recommendation for how to avoid the fallacy of generic generalizations. The remedy is straightforward: whenever a generic generalization is noticed, it should be replaced by a quantified generalization.

If there is indeed elided quantification lurking beneath a supposed generic generalization, then making the quantification explicit demonstrates that the belief is not truly a generic generalization. If no quantification can be produced, then the belief is indeed a generic generalization, and the holder of the belief should reflect upon the flaws in this way of thinking.


## Generic Comparison

In a generic comparison, it is asserted that one group has "more" or "less" of some attribute than another group, but no further quantification is specified.

As discussed previously, the phrase "generic comparison" is not as established as the phrase "generic generalization." Indeed, while generic generalizations are studied by philosophers and psychologists, the phrase "generic comparisons" encompasses criticisms that are sometimes discussed by empirical researchers and applied statisticians, but are not as widely discussed outside of this niche. However, these criticisms are just as relevant, if not more so, whenever thinking about issues involving large populations of individuals, such as social issues.

Unlike a generic generalization, which is generic in precisely one way, there are two ways in which a generic comparison can be generic.

A generic generalization fails to describe the proportion of the population for which its predicate is true. When "swans are white" is asserted, it is not clear if this means that 2% or 50% or 99% of swans are white. Similarly, if it is settled that, by any reasonable measure, one group has "more" of something than the other, a generic comparison leaves unspecified how much "more" one group has than the other.

Additionally, there are many ways to measure the distribution of a variable in a population, and different ways might lead to different directions of "more" and "less." Thus, generic comparisons can sometimes be ambiguous about which direction the "more" should go.

This section first addresses the issue of ambiguous direction, which is peculiar to generic comparisons, before discussing the issues stemming from unspecified magnitude, which are similar to the issues with generic generalizations.


### Ambiguous Direction


What makes comparisons between groups more complicated than comparing individuals[^observational] is that comparison between groups is a comparison of distributions, rather than point values. For instance, when comparing the height of two people, who can be labeled "A" and "B," there will be a single point value for the height of person A, e.g., 68 inches, and a single point value for the height of person B, e.g., 75 inches. In this case, it is true that person A is shorter than person B, and false that person A is taller than person B.

[^observational]: In statistical jargon, here "group" is taken to mean a domain and "individual" is taken to mean an observational unit.

On the other hand, if "A" and "B" represent groups of thousands of people, there is not a single point value that captures all the variation in the thousands of heights of people in group A or a single point value that can represent the thousands of heights of people in group B. In this case, the two _distributions_ must be compared.

Unless the two distributions are completely disjoint, some people in group A will be taller than some people in group B, and some people in group B will be taller than some people in group A. It is not simply true or false that group A "is taller" than group B or group A "is shorter" than group B.

One could calculate and compare the arithmetic means of the heights in the two groups, but arithmetic mean is _just_ one measure of the central tendency of a distribution. There are other measures of central tendency, such as the median, which also have their merits. More importantly, there are other properties of a distribution that are just as relevant, such as the dispersion or spread, the skewness, and to a lesser extent, the kurtosis.

Even with just the two basic statistics of mean and median, it is possible for the mean of group A to be greater than the mean of group B, but the median of group A to be _less than_ the median of group B. Simplifying the heights example to involve only 30 people instead of thousands of people in each group leads to Table 1.


```{r, include=FALSE}
n <- 30

set.seed(20220216)

group_A <- ifelse(
  rbinom(n, size = 1, prob = 0.3) == 1,
  rnorm(n, 67, 1),
  rnorm(n, 56, 2)
)

group_B <- rnorm(n, 58, 2)

dt <- rbind(
  data.table(
    Group = 'A',
    `Height (in)` = sort(round(group_A, 1))
  ),
  data.table(
    Group = 'B',
    `Height (in)` = sort(round(group_B, 1))
  )
)

dt[, Person := .I]

setcolorder(dt, 'Person')
```

<p class="caption">Table 1: The randomly generated heights of two groups of 30 people each.</p>

<div class="three-columns">
```{r}
kable(dt[1:15], align = 'ccc')
kable(dt[16:30], align = 'ccc')
kable(dt[31:45], align = 'ccc')
kable(dt[46:60], align = 'ccc')
```
</div>

```{r, include=FALSE}
mean_A <- round(mean(group_A), 1)
median_A <- round(median(group_A), 1)
mean_B <- round(mean(group_B), 1)
median_B <- round(median(group_B), 1)
```

A quick calculation shows that the mean height of people in group A is `r mean_A` inches, which is greater than the mean height of people in group B of `r mean_B` inches. However, the median height of people in group A is `r median_A` inches, which is actually less than the median height of people in group B of `r median_B` inches.[^outliers]

[^outliers]: This is simply due to the fact that medians are more resistant to outliers than means, and 30% of group A consists of exceptionally tall people.

This ambiguity can occur even between two of the most basic statistics used to describe a distribution: mean and median. As more and more statistics are used, there are more and more opportunities for such ambiguities to occur. A similar phenomenon, but with means and totals, was described in this blog's discussion of [missing denominators](/how-to-scrutinize-a-statistic.html#missing-denominators).

The important takeaway is that it always flawed to think about a group having "more" or "less" of something than another group. Assertions such as these are outside the realm of truth or falsity. Rather, it is correct to think about whether a _specific statistic_ &ndash; such as a mean, median, total, proportion, rate, ratio, etc. &ndash; for one population is greater or lesser than that statistic for another population. Assertions about differences in statistics between two groups _are_ true or false.


### Unquantified Magnitude

For any given variable, it is extremely unlikely that it would have the _exact same_ distribution in two different populations. Thus, when comparing two groups, any variable is going to have a greater mean (or other statistic) for one group than for the other.

Because nearly _every_ variable has at least a trivially different distribution between two groups, simply asserting that a mean (or other statistic) is greater for one group than the other is largely uninformative. Given enough precision, no two means will be equal.

Thus, generic generalizations are resistant to falsification. If someone asserts that group A makes more money on average than group B, a study which finds that members of group A have a mean annual income of \$46,300 and members of group B have a mean annual income of \$46,300 does not conclusively demonstrate that this assertion is false. The person making the assertion could simply demand more precision with the expectation that the means, calculated to the dollar, would be different, such as with mean incomes of \$46,382 and \$46,311.

Even if the mean annual incomes of the two groups were found to be \$46,382 and \$46,382, there might be a difference if precision were increased to the cent. Indeed, since a mean is a ratio between a sum and a count, this could continue _ad nauseum_ to an arbitrary level of precision.[^significance]

[^significance]: This is a simplification for illustrative purposes. This is not how statistical analysis works in practice. In reality, such an encounter would turn on how much power a study has and whether a difference is [statistically significant](/how-to-scrutinize-a-statistic.html#statistical-significance).

While generic comparisons are in this way resistant to falsification, they are not quite as resistant to falsification as generic generalizations. With a generic generalization, as long as at its predicate is true of least one member of its subject population, it resists falsification. With a generic comparison, assertions have a direction, i.e., the "more" or "less" component. If this were asserted to be in the wrong direction, then the generic comparison could be readily falsified.[^power]

[^power]: Real-world falsification is, of course, still contingent on a study having enough statistical power to detect the difference, but in principle, the falsification can be done.

As long as the direction of the generic comparison is correct, however, the generic comparison is resistant to falsification. Like generic generalizations, generic comparisons that are resistant to falsification are also highly uninformative. For instance, asserting that group A makes more money on average than group B is uninformative. It could be that the mean annual income in group A is \$32,950 versus a mean income in group B of \$28,552. It could, however, be that the mean annual income of group A is \$157,800 versus a mean income in group B of \$28,552.

Because these kind of generic comparisons are uninformative, they can be used to reach specious conclusions in flawed analysis, a practice all too commonly seen in social or behavioral research. The general scheme is to claim that the mean (or other statistic) of variable _X_ is greater for group A than for group B because the mean (or other statistic) of variable _Y_ is greater for group A than for group B. For instance, someone might claim that group A has a higher mean income than group B because the mean hours worked per week is higher for group A than for group B.

It _might_ very well be the case that the differences in the number of hours worked per week between the two groups explains a substantial portion of the difference in average income between the two groups, but the same could be said for _any other variable_ because, for any variable, it is extremely unlikely that the mean of the variable would be exactly equal between the two groups. Thus, by the exact same logic, _anything_ can explain the difference in mean income between group A or group B.

Quantification of the difference in mean hours worked per week between the two groups and of the difference in mean annual income between the two groups might cast doubt on such an explanation. For instance, if it is discovered that the mean hours worked per week is 42.7 for group A and 41.3 for group B, but the mean annual income is \$74,920 for group A and \$39,500 for group B, such a small difference in mean hours worked per week seems unlikely to explain such a large difference in mean income.

Therefore, quantification might be withheld by deceptive analyses intended to make an explanation seem more plausible than it actually is. Alternatively, it might be withheld by flawed, lazy analyses that do not know any better or do not invest enough time or resources in the investigation.

Even with quantification, simply comparing differences in mean (or other statistic) between groups is itself insufficient for judging how much of the differences in one variable is explained by differences in another variable.[^variation] However, quantification of means (or other statistics) is at least a step in the right direction and away from generic thinking.

[^variation]: To determine how much of variation one variable is explained by another variable, both variables should be measured in the same observational units, and techniques as linear models or analysis of variance (ANOVA) should be used. "Observational unit" is jargon for the individuals that compose the groups, such as, for example, an individual person.


### How to Avoid the Fallacy

The answer to the fallacy of generic comparisons is similar to the answer to the fallacy of generic generalizations: all that is needed is quantification. However, quantifying a generic comparison is much more nuanced than quantifying a generic generalization.

When quantifying a generic generalization, typically all that is needed is a proportion, in particular, the proportion of the subject population for which the predicate of the generalization is true. For instance, specifying "72.4\%" in "approximately 72.4% of swans are white" is relatively straightforward.

Various ways to provide a summary differences between groups were given in the discussion of [effect size](/how-to-scrutinize-a-statistic.html#effect-size). While an effect size is a helpful initial summary of a difference between groups, it is just that, an initial summary and not a complete description of the difference. How to quantify differences between groups is a deep enough topic to merit its own treatment, but is beyond the scope of this article.


## Further Investigation

Generic thinking occupies a central position in this blog's discussion of flaws in reasoning about evidence, and there are many opportunities for further investigation. Many of the [fallacies](/tags/fallacies.html) discussed in their own articles either are specific kinds of generic thinking, or are supported by or derived from generic thinking. In order to present an alternative to generic comparisons, there is an opportunity for further investigation into the subtleties of comparing differences between groups.

There is much in the field of psychology pertaining to generic thinking. A review of the psychological literature around the study of generic generalizations is an opportunity for further investigation. Generic thinking is believed to be related to the psychology of perceptions of social groups as entities themselves, so a review of the psychology of entitativity is another opportunity. This, in turn, may be related to the phenomenon of psychological essentialism, which presents yet another opportunity.


## Conclusion

Generic thinking is an extremely flawed, but all too common, way of viewing the world. Purging generic thinking from one's mind may seem like an aspirational goal. It might be unreasonable to expect anyone to be entirely free of generic thinking, but it is also unreasonable to expect any society to be entirely devoid of house fires, and this does not negate the fact that the desired number of house fires is none.

In particular, care should be taken when generic thinking rears its ugly head in social criticism.

First, this is because social criticism is one of the sorts of thinking when people are &ndash; supposedly &ndash; at their most serious, and while generic thinking in daily banter may be of little consequence, social criticism deserves better for its seriousness.

Second, as touched on earlier, social criticism is inherently quantitative; what separates social criticism from human interest stories is that while human interest might discuss the affairs of one or a handful of people, social criticism directly discusses the state of affairs among millions if not billions of individual people.

Third, generic thinking in social criticism serves little purpose but prejudice, and guarding against generic thinking is thus a way to guard against bigotry.

Quantification is proposed here as a remedy both to generic generalizations and to generic comparisons. While quantification does away with literal generic thinking, it is only a concrete mechanism that presents itself whenever generic thinking is challenged.

Underlying generic thinking is a blunt, inflexible worldview that puts vast numbers of individuals into categories and thinks it useful to hurl around broad, vague ideas about these categories and all the many individuals in them.

Opposed to this are worldviews that realize any population of vast numbers of individuals exhibits _variation,_ and that to understand large populations _is_ to understand this variation. Quantification is merely the tool that such worldviews use to describe variation. Thus, much as quantification is the answer to the _letter_ of generic thinking, to counter the _spirit_ of generic thinking is to expect, seek, understand, and describe variation in the world.


## Citations

::: {#refs}
:::


## Footnotes
