---
title: "The Fallacy of Generic Thinking"
date: 2022-02-16
draft: true
tags: ["reasoning about evidence", "fallacies", "statistics", "psychology", "philosophy", "third-person"]
bibliography: "../bibliographies/fallacy-of-generic-thinking.bib"
csl: "../bibliographies/chicago-author-date.csl"
link-citations: true
implicit-figures: true
output:
  blogdown::html_page:
    md_extensions: ["+footnotes"]
    toc: true
    toc_depth: 5
include-before: |
  <p>Introductory paragraph goes here.</p>
  <!--more-->
  ## Table of Contents
---

```{r setup, echo=FALSE, message=FALSE}
library(data.table)
library(knitr)

opts_chunk$set(
  echo = FALSE
)
```


## Introduction

What unites the many types of [fallacies](/tags/fallacies.html) that this blog discusses is that they would appeal to a mind afflicted with what can be termed "generic thinking." For instance, the article "[How to Scrutinize a Statistic](/how-to-scrutinize-a-statistic.html)" discussed both the need to evaluate [practical significance](/how-to-scrutinize-a-statistic.html#practical-significance) and how this is typically done by quantifying an [effect size](/how-to-scrutinize-a-statistic.html#effect-size). The absence of summarizing an effect size was described as a fallacy of "[generic comparisons](http://localhost:4321/how-to-scrutinize-a-statistic.html#generic-comparisons)."

The phrase "generic comparison" is actually a neologism invented by this blog, but the word "generic" was used deliberately to allude to what is termed "generic generalization," a phrase established in the academic literature.

Generic thinking comprises both generic generalization and generic comparison and encompasses all beliefs about an inherently quantitative subject, but that are held without any quantification. This often occurs when someone is thinking about large populations of individuals. For instance, social issues occur, by their very nature, in large populations of individual human beings. Thus, all unquantified social criticism is generic thinking.


## Generic Generalization

In both academic philosophy and academic psychology, there exists a body of scholarship pertaining to generic generalization. In a generic generalization, a property is predicated of a subject that denotes a population of individuals, but no quantification is given.

"All swans are white," "some swans are white," "no swans are white," "approximately 74.2% of swans are white" are all beliefs about a subject &ndash; denoted by "swans" &ndash; that includes a large number of individuals. All of these statements include some kind of quantification, namely "all," "some," "no," or "approximately 74.2%." However, the statement "swans are white" is a generic generalization because it is unquantified.

Generic generalizations have been studied by philosophers because they can seem like a logic puzzle. For instance, statements like "dogs have four legs," "ducks lay eggs," "mosquitoes carry the West Nile virus" seem like they have the spirit of truth to them.[^examples]  However, they are all false for a portion of the population they discuss.

[^examples]: "Dogs have four legs" and "ducks lay eggs" are examples used in "Generic Generalizations" in the _Stanford Encyclopedia of Philosophy_ [@leslie_generic_2016] and the West Nile virus example is used in "The Original Sin of Cognition" [@leslie_original_2017].

While most dogs have four legs, there are some dogs happily scampering around on three legs. Only female ducks who have reached sexual maturity even have the ability to lay eggs, so "ducks lay eggs" is only true for less than half the population it describes. A mercifully very small proportion of mosquitoes actually carry the West Nile virus.

![
A three-legged dog, the existence of which is a false case of the generic generalization "dogs have four legs."<br/>
["Three Legged Dog"](https://www.flickr.com/photos/23743320@N03/7230816680)
by
[Sangudo](https://www.flickr.com/photos/sangudo/)
is licensed under
[CC BY-NC-ND 2.0](https://creativecommons.org/licenses/by-nc-nd/2.0/)
and is not modified.
](/images/three-legged-dog.jpg)


### Resistance to Falsification

Much of the philosophizing around generic generalizations has been an attempt to find some way of interpreting generic generalizations that would allow the ones that seem plausible to be true and the ones that are patently false to be false. [@leslie_generic_2016] This highlights precisely the main issue with generic generalizations: they seem to defy truth conditions.

In particular, generic generalizations are resistant to falsification.  Quantified beliefs have conditions under which the holder of the belief can learn that the belief is false, if it is indeed false. If someone who asserts "all swans are white" encounters a single black swan, then the individual knows the belief is false. Similarly, presenting just one white swan to someone who asserts "all swans are black" demonstrates to this person that the belief is false.

While it is easy to show the assertion "some swans are white" is true since observation of one white swan confirms the belief, falsification is more difficult. To demonstrate the belief is false without any uncertainty requires observations of all the swans in the world. However, this is just a practical matter. In principle, it is known how to demonstrate the belief is false, even if demonstrating it is false would require more work than most are willing to commit.

Statements such as "approximately 72.4\% of swans are white" can be demonstrated false, given a specified [hypothesis test level](/how-to-scrutinize-a-statistic.html#fallacies). This is slightly more complicated by the fact that for a continuous variable any such hypothesis can be arbitrarily rejected given a large enough sample size and enough statistical power, so exact statistical thresholds would have to be stipulated. Once these are all specified, however, such hypotheses can be evaluated as to their truth or falsity according to standard methods.

Thus, quantified beliefs that are false can be demonstrated as false. The same is not the case for generic generalizations. If someone asserts "dogs lay eggs," how many viviparous dogs must be observed before this belief is demonstrated false? Similarly, how many two-legged ducks must be brought to someone who asserts "ducks have four legs" before the holder of the belief must relent? The stubborn can merely retort "not _all_ dogs lay eggs" or "I didn't mean _all_ ducks have two legs." Generic generalizations thus create a threshold that can always be conveniently moved to justify whatever arbitrary belief is desired.

Such resistance to falsification is a nontrivial matter because many bigoted attitudes are construed in terms of generic generalization. For instance, Leslie [-@leslie_original_2017] uses "Muslims are terrorists" and "Blacks are rapists" as examples of negative stereotypes formulated as generic generalizations.

To those concerned about bigotry, this resistance to falsification is a troubling property of generic generalizations because it means that once a negative stereotype has been formulated as a generic generalization in someone's mind, it is that much more difficult to convince the holder of the negative stereotype to change the belief.

If someone were to assert "most Muslims are terrorists" or "80% of Blacks are rapists," these beliefs could be demonstrated false with the appropriate empirical observations and analyses. When someone asserts the generic generalizations "Muslims are terrorists" or "Blacks are rapists," it does not matter _how many_ Muslims are not terrorists or Blacks are not rapists, the generic generalization is resistant to falsification exactly because it is unquantified.


### Uninformativeness

In addition to being resistant to falsification, generic generalizations are not informative because, given a large enough population, an arbitrary number of [confirmatory examples](selection-bias-and-the-fallacy-of-listing-examples.html) can always be found.

For instance, generic generalizations such as "humans are terrorists" or "humans are rapists" seem odd. However, they have a similar amount of actual information as the more specific generic generalizations "Muslims are terrorists" and "Blacks are rapists."

Such generic generalizations about "humans," "Muslims," or "Blacks" assert that, among some population of billions of individuals, there are some unspecified number of individuals who are terrorists or who are rapists. However, this is true of _any_ large enough population of human beings.

The generic generalizations "humans are terrorists" or "humans are rapists" are conspicuous in just how vague and uninformative they are. However, this is precisely the same issue with all generic generalizations. The "humans" versions of these generic generalizations are merely more obvious in illustrating this defect.

Given a large population of humans, there will be among them some unspecified number of terrorists, rapists, charitable givers, volunteers, slalom skiers, stamp collectors, or anything else for that matter. This is a reasonable expectation _a priori_ before any observations are made. Indeed, if a claim were made that a population of millions of human beings did not contain a single rapist, charitable giver, or stamp collector, such a claim would be treated incredulously. Generic generalizations are therefore simply not informative.


### Empirical Study

The fallacy of generic generalization is not just a philosophical concern, thanks in large part to the work of developmental psychologist Susan Gelman, who discovered that people become sensitive to generic generalizations at a young age. [@gelman_childrens_2002]

Psychological research in this area has gone on to demonstrate empirically that the fallacious use of generic generalizations does exist. For instance, Cimpian, Brandone, and Gelman [-@cimpian_generic_2010] found that test subjects accepted generic generalizations at low levels of prevalence, but interpreted them as if they implied high levels of prevalence. Thus, generic generalizations were evaluated for truth or falsity closer to "some" quantification, but interpreted as if they were closer to "all" quantification.

**_[Discuss Cimpian et al here in more detail, at least to the level of Anecdotes Are Not Evidence.]_**


### Bald Plurals and Elision

Because generic generalizations are resistant to falsification, uninformative, and have been found empirically to lead to flawed thinking, this article dismisses generic generalizations as mere fallacies with no merit. Therefore, this article recommends that everyone strive to remove generic generalizations from their thinking.

This is a value judgment that implies that the philosophical program of attempting to salvage generic generalizations is not worthwhile. One might protest, what about the generic generalizations that feel like there is a ring of truth to them, such as "dogs have four legs," "ducks lay eggs," and "mosquitoes carry the West Nile virus?"

Such examples in this article have all been what are called "bald plurals," in which the subject of a sentence is a plural noun that is not proceeded by any modifier word. However, not all uses of bald plurals are truly generic generalizations.

This is due to the messy nature of natural languages. In particular, human beings are prone to make their communication more concise by elision. Sometimes just a few sounds are left out as in the case of a contraction, but whole words, phrases, and clauses can be elided to be inferred from context.

Commonly, a bald plural is used when "all" quantification is implied. For instance, in this very article, it is written that "generic generalizations are resistant to falsification." This is not, however, a generic generalization about generic generalizations. From context, it should be gleaned that "all" quantification is implied, as in "all generic generalizations are resistant to falsification."

If one went about one's day putting "all" in front of each and every universally quantified statement, it would sound strange to native English speakers. In many contexts such as this one, using the word "all" might be used for emphasis, but it is not necessary for the English language to function.

Once one accepts that natural languages function with such elision, then whether or not any given bald plural is a generic generalization becomes a _question,_ not an automatic assumption.

When someone says "ducks lay eggs," there is typically social context surrounding this statement from which a more complicated thought can be inferred, such as "the system of reproduction used by ducks is oviparous."

![
A child with an adult encountering ducks swimmign around a pond.<br/>
[Image](https://pixabay.com/photos/people-adult-autumn-back-boy-2942837/?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=2942837)
by
[freestocks-photos](https://pixabay.com/users/freestocks-photos-7014431/?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=2942837)
from
[Pixabay](https://pixabay.com/?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=2942837).
](/images/duck-pond.jpg)

For instance, a child might ask if a duck found in a pond will give birth to little chicks that can swim about, to which an adult might reply, "no, ducks lay eggs." Given context, the longer statement about reproductive systems is implied, but rarely do English speakers talk in this way, and instead this is shortened to "ducks lay eggs."

Furthermore, one could assert "a dog without genetic aneuploidy and absent any physical trauma, allowed to develop in a salubrious environment including nutritional needs, grows into an organism with four legs." That is a lot of words, which can be elided to the simple statement "dogs have four legs" given social context.

Similarly, those who claim to have gotten West Nile virus from an ant bite might be curtly told, "no, mosquitoes carry the West Nile virus." It is likely in this case the proposition "mosquitoes are the only species that function as a vector for West Nile virus transmission in humans" has simply been shortened to "mosquitoes carry the West Nile virus."


### How to Avoid the Fallacy

One might protest that perhaps all generic generalizations could be explained as instances of this kind of elision. Such a protest is consistent with this article's recommendation for how to avoid the fallacy of generic generalizations. The remedy is straightforward: whenever a generic generalization is noticed, it should be replaced by a quantified generalization.

If there is indeed elided quantification lurking beneath a supposed generic generalization, then making the quantification explicit demonstrates that the belief is not truly a generic generalization. If no quantification can be produced, then the belief is indeed a generic generalization, and the holder of the belief should reflect upon the flaws in this way of thinking.


## Generic Comparison

In a generic comparison, it is asserted that one group has "more" or "less" of some attribute than another group, but no further quantification is specified. Unlike a generic generalization, which is generic in precisely one way, there are two ways in which a generic comparison can be generic.

A generic generalization fails to describe the proportion of the population for which its predicate is true. When "swans are white" is asserted, it is not clear if this means that 2% or 50% or 99% of swans are white. Similarly, if it is settled that, by any reasonable measure, one group has "more" of something than the other, a generic comparison leaves unspecified how much "more" one group has of something than the other.

Additionally, with generic comparisons, because the judgement of "more" or "less" is left vague, it can be ambiguous which direction the "more" should go and which direction the "less" should go. There are many ways to measure the distribution of a variable in a population, and different ways might lead to different directions of "more" and "less."

This section first addresses the issue of ambiguous direction, which is peculiar to generic comparison, before discussing the issues stemming from unspecified magnitude, which are similar to the issues with generic generalization.


### Ambiguous Direction


What makes group comparisons more complicated than comparing individuals[^observational] is that comparison between groups is a comparison of distributions, rather than point values. For instance, when comparing the height of two people, who can be labeled "A" and "B," there will be a single point value for the height of person A, e.g., 68 inches, and a single point value for the height of person B, e.g., 75 inches. In this case, it is true that person A is shorter than person B and false that person A is taller than person B.

[^observational]: In statistical jargon, here "group" is taken to mean a domain and "individual" is taken to mean an observational unit.

On the other hand, if "A" and "B" represent groups of thousands of people, there is not a single point value that captures all the variation in the thousands of heights of people in group A or a single point value that can represent the thousands of heights of people in group B. In this case, the two _distributions_ must be compared.

Unless the two distributions are completely disjoint, some people in group A will be taller than some people in group B, and some people in group B will be taller than some people in group A. It is not simply true or false that group A "is taller" than group B or group A "is shorter" than group B.

One could calculate and compare the arithmetic means of the heights in the two groups, but arithmetic mean is _just_ one measure of the central tendency of a distribution. There are other measures of central tendency, such as the median, which also have their merits. More importantly, there are other properties of a distribution that are just as relevant, such as the dispersion or spread, the skewness, and to a lesser extent, the kurtosis.

Even with just the two basic statistics of mean and median, it is possible for the mean of group A to be greater than the mean of group B, but the median of group A to be _less than_ the median of group B. Simplifying the heights example to involve only 30 people instead of thousands of people in each group leads to Table 1.


```{r, include=FALSE}
n <- 30

set.seed(20220216)

group_A <- ifelse(
  rbinom(n, size = 1, prob = 0.3) == 1,
  rnorm(n, 67, 1),
  rnorm(n, 56, 2)
)

group_B <- rnorm(n, 58, 2)

dt <- rbind(
  data.table(
    Group = 'A',
    `Height (in)` = sort(round(group_A, 1))
  ),
  data.table(
    Group = 'B',
    `Height (in)` = sort(round(group_B, 1))
  )
)

dt[, Person := .I]

setcolorder(dt, 'Person')
```

<p class="caption">Table 1: The randomly generated heights of two groups of 30 people each.</p>

<div class="three-columns">
```{r}
kable(dt[1:15], align = 'ccc')
kable(dt[16:30], align = 'ccc')
kable(dt[31:45], align = 'ccc')
kable(dt[46:60], align = 'ccc')
```
</div>

```{r, include=FALSE}
mean_A <- round(mean(group_A), 1)
median_A <- round(median(group_A), 1)
mean_B <- round(mean(group_B), 1)
median_B <- round(median(group_B), 1)
```

A quick calculation shows that the mean height of people in group A is `r mean_A` inches, which is greater than the mean height of people in group B of `r mean_B` inches. However, the median height of people in group A is `r median_A` inches, which is actually less than the median height of people in group B of `r median_B` inches.[^outliers]

[^outliers]: This is simply due to the fact that medians are more resistant to outliers than means, and 30% of group A consists of exceptionally tall people.

This ambiguity can occur even between two of the most basic statistics used to describe a distribution: mean and median. As more and more statistics are used, there are more and more opportunities for such ambiguities to occur. A similar phenomenon, but with means and totals, was described in this blog's discussion of [missing denominators](/how-to-scrutinize-a-statistic.html#missing-denominators).

The important takeaway is that it always flawed to think about a group having "more" or "less" of something than another group. Assertions such as these are outside the realm of truth or falsity. Rather, it is correct to think about whether a _specific statistic_ for one population is greater or lesser than that statistic for another population. Assertions about differences in statistics between two groups _are_ true or false.


### Unquantified Magnitude

For any given variable, it is extremely unlikely that it would have the _exact same_ distribution in two different populations. Thus, between two groups, one group is going to have "more" of something than the other. More properly, when comparing two groups, any variable is going to have a greater mean, median, total, proportion, etc., for one group than for the other.

Because _every_ variable is at least trivially different between two groups, simply asserting that a mean or other statistic is greater for one group than the other is largely uninformative. Given enough precision, no two means will be equal. This is a phenomenon described in a well-known quote by statistican John Tukey.

_**[Put Tukey quote here after reading paper.]**_

Thus, generic generalizations are resistant to falsification. If someone asserts that group A makes more money on average than group B, a study which finds that members of group A have a mean income of \$46,300 and members of group B have a mean income of \$46,300 does not conclusively demonstrate that this assertion is false. The person making the assertion could simply demand more precision with the expectation that the means, taken to the dollar, would be different, such as with mean incomes of \$46,832 and \$46,311.

Even if the mean incomes of the two groups were found to be \$46,712 and \$46,712, the person making the assertion could speculate that there would be a difference if precision were increased to the cent. Indeed, since a mean is a ratio between a sum and a count, this could continue _ad nauseum_ to an arbitrary level of precision.[^significance]

[^significance]: This is a simplification for illustrative purposes. This is not how statistical analysis works in practice. In reality, such an encounter would turn on how much power a study has and whether a difference is [statistically significant](/how-to-scrutinize-a-statistic.html#statistical-significance).

While generic comparisons are in this way resistant to falsification, they are not quite as resistant to falsification as generic generalizations. With a generic generalization, as long as at its predicate is true of least one member of its subject population, it resists falsification. With a generic comparison, assertions have a direction, i.e., the "more" or "less" component. If this is asserted to be in the wrong direction, then the generic comparison can be readily falsified.[^power]

[^power]: Real-world falsification is, of course, still contingent on a study having enough statistical power to detect the difference, but in principle, the falsification can be done.

As long as the direction of the generic comparison is correct, however, the generic comparison is resistant to falsification. As was seen with generic generalizations, generic comparisons that are resistant to falsification are also highly uninformative. If for every variable, and for every pairing of any two groups, the mean of the variable is greater for one group than for the other, assertions that this is the case are not saying much at all. The same can be said for an arbitrary number of other comparisons.


* _**Fallacious use in empirical research (group A has "more" of X than group B, and this explains why group A has "more" of Y than group B)**_
* _**"Generic comparisons" more of a statistician niche**_
* _**Add practical significance discussion to generic generalizations?**_


### How to Avoid the Fallacy

[...]

A first answer to this is a summary of [effect size](/how-to-scrutinize-a-statistic.html#effect-size). While an effect size is a helpful initial summary of a difference between groups, it is just that, an initial summary and not a complete description of the difference.

How to compare a variable between two groups _well_ is a deep enough topic to merit its own treatment in a future article of this blog; this article focuses on the fallacious comparison of a variable between two groups.

[...]

## Conclusion

* Follow-ups
    * Review of psychological literature on generic generalizations
    * " " " " " entitativity
    * " " " " " essentialism
    * How-to on comparing response variable between two or more groups
* Particular emphasis should be placed on policing generic thinking with regard to social criticism
    * because serious
    * because inherently quantitative
    * guards against prejudice

Purging generic thinking from one's mind is of course something of an aspirational goal. It is unreasonable to expect anyone's thinking to be entirely free of generic generalizations, but it is also unreasonable to expect any society to be entirely devoid of house fires, and this does not negate the fact that the desired number of house fires is not crime at all.


## Citations

::: {#refs}
:::


## Footnotes
