---
title: "Current Contraceptive Usage, with One Master Categorical Variable"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(here)
library(knitr)
library(survey)

i_am('EDA/005_categorized_current_contraceptive_usage.Rmd')

source(here('R/survey_helpers.R'))
source(here('R/import.R'))

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r helper-functions}
tablena <- function(x) {
  table(x, useNA = 'always')
}

in_millions <- function(dt, round_to = 3, do_round = TRUE) {
  dt <- as.data.table(dt)
  cols <- c('estimate', 'se', '2.5 %', '97.5 %')
  sub_dt <- dt[, cols, with = FALSE] / 10^6
  if (do_round) {
    sub_dt <- round(sub_dt, round_to)
  }
  dt[, (cols) := sub_dt]
  dt
}

rounded <- function(dt, round_to = 3) {
  cols <- c('estimate', 'se', '2.5 %', '97.5 %')
  sub_dt <- round(dt[, cols, with = FALSE], round_to)
  dt[, (cols) := sub_dt]
  dt
}

has_level <- function(x, ids) {
  is.element(as.integer(x), ids)
}

standard_millions_plot <- function(data) {
  ggplot(
    data = data,
    mapping = aes(
      y = reorder(level, estimate)
    )
  ) +
  geom_point(
    aes(
      x = estimate
    ),
    size = 2
  ) +
  geom_segment(
    aes(
      yend = reorder(level, estimate),
      x = `2.5 %`,
      xend = `97.5 %`
    ),
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    size = 0.5
  ) +
  scale_x_continuous(
    labels = scales::label_number(suffix = 'M', accuracy = 0.1),
  ) +
  scale_y_discrete(
    labels = scales::wrap_format(50)
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_blank()
  )
}

standard_proportions_plot <- function(data) {
  ggplot(
  data = data,
  mapping = aes(
    y = reorder(level, estimate)
  )
) +
  geom_point(
    aes(
      x = estimate
    ),
    size = 2
  ) +
  geom_segment(
    aes(
      yend = reorder(level, estimate),
      x = `2.5 %`,
      xend = `97.5 %`
    ),
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    size = 0.5
  ) +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    expand = c(0, 0),
    labels = scales::label_percent(1),
    limits = c(0, 1)
  ) +
  scale_y_discrete(
    labels = scales::wrap_format(50)
  ) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    legend.position = 'none'
  )
}

env <- new.env()

env$estimates <- data.table()

collect_estimate <- function(est, title, domain = '', type = '', round_to = 3, plot = FALSE) {
  this_title <- title
  this_domain <- domain
  this_type <- type

  if (type == 'total') {
    dt <- in_millions(est, round_to)
  } else if (type == 'proportion') {
    dt <- rounded(est, round_to)
  } else {
    dt <- as.data.table(est)
  }
  env$estimates <- rbindlist(list(
    env$estimates[
      !(
        title == this_title &
          domain == this_domain &
          type == this_type
      )
    ],
    dt[, .(
      domain = this_domain,
      title = this_title,
      type = this_type,
      level,
      estimate,
      se,
      `2.5 %`,
      `97.5 %`
    )]
  ))

  env$estimates[
    title == this_title &
      domain == this_domain &
      type == this_type
  ]
}
```


```{r fem-load-data}
fem_2017_2019 <- load_NSFG_data('2017_2019', 'FemResp')
fem_2017_2019_formats <- fem_2017_2019$Formats
```

```{r fem-process-data, dependson=c('helper-functions', 'fem-load-data')}
fem_2017_2019_data <- copy(fem_2017_2019$Data[, .(
  CASEID,
  CONSTAT1,
  METH3M1,
  METH3M2,
  METH3M3,
  P1YRELP = factorize(P1YRELP, 'P1YRELP', fem_2017_2019_formats),
  P1YRELP = factorize(P1YRELP, 'P1YRELP', fem_2017_2019_formats),
  POSIBLMN = factorize(POSIBLMN, 'Y1N5RDF', fem_2017_2019_formats),
  POSIBLPG = factorize(POSIBLPG, 'Y1N5RDF', fem_2017_2019_formats),
  PSURGSTR = factorize(PSURGSTR, 'Y1N5C', fem_2017_2019_formats),
  RCURPREG = factorize(RCURPREG, 'RCURPREG', fem_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'Y1N2RECF', fem_2017_2019_formats),
  TUBS = factorize(TUBS, 'Y1N5C', fem_2017_2019_formats),
  WGT2017_2019,
  WYNOLSTP = factorize(WYNOLSTP, 'Y1N5RDF', fem_2017_2019_formats),
  WYNOTUSE = factorize(WYNOTUSE, 'Y1N5RDF', fem_2017_2019_formats)
)])

# remove sterility not from contraceptive procedure from METH3M series
fem_2017_2019_data[
  METH3M3 %in% 20:21,
  `:=`(
    METH3M3 = NA
  )
]
fem_2017_2019_data[
  METH3M2 %in% 20:21,
  `:=`(
    METH3M2 = METH3M3,
    METH3M3 = NA
  )
]
fem_2017_2019_data[
  METH3M1 %in% 20:21,
  `:=`(
    METH3M1 = METH3M2,
    METH3M2 = METH3M3,
    METH3M3 = NA
  )
]

# make METH3M series variables into factors
fem_2017_2019_data[, `:=`(
  METH3M1 = factorize(METH3M1, 'METH3MF', fem_2017_2019_formats),
  METH3M2 = factorize(METH3M2, 'METH3MF', fem_2017_2019_formats),
  METH3M3 = factorize(METH3M3, 'METH3MF', fem_2017_2019_formats)
)]

# create boolean variables for whether each specific method is currently used
for (x in c(1:19, 22:24)) {
  set(
    fem_2017_2019_data,
    j = sprintf('using3m_meth%d', x),
    value = has_level(fem_2017_2019_data$METH3M1, x) |
      has_level(fem_2017_2019_data$METH3M2, x) |
      has_level(fem_2017_2019_data$METH3M3, x)
  )
}

# count any respondents who are sexually active and whose last intercourse was with husband or cohabitating partner who has an active vasectomy as using vasectomy as a contraceptive method
fem_2017_2019_data[, using3m_meth3 := fifelse(
  SEX3MO == 'YES' & PSURGSTR == 'YES' & (
    P1YRELP == 'Current husband or (current husband from whom she is separated)' |
    P1YRELP == 'Current cohabiting partner'
  ),
  TRUE,
  using3m_meth3
)]

# count any respondents with current tubal ligations who are sexually active as currently using tubal ligation as a contraceptive method
fem_2017_2019_data[, using3m_meth4 := fifelse(
  SEX3MO == 'YES' & TUBS == 'YES',
  TRUE,
  using3m_meth4
)]

# create variable counting number of current methods
fem_2017_2019_data[
  ,
  meth3m_count := rowSums(.SD),
  .SDcols = patterns('^using3m_meth')
]

# set master categorization for contraceptive use
fem_2017_2019_data[,
  curr_use := factor(fcase(
    meth3m_count > 0,
      'Used at least one contraceptive method at last intercourse',
    SEX3MO != 'YES',
      'Did not have sexual intercourse with male in past 3 months',
    RCURPREG == 'YES (CURRENTLY PREGNANT)',
      'Currently pregnant',
    (WYNOTUSE == 'Yes' | WYNOLSTP == 'Yes'),
      'Currently trying to conceive',
    POSIBLPG == 'No',
      'Physically unable to conceive',
    POSIBLMN == 'No' & (
      P1YRELP == 'Current cohabiting partner' |
      P1YRELP == 'Current husband or (current husband from whom she is separated)'
    ),
      'Last intercourse was with partner who is physically unable to conceive',
    default = 'Otherwise did not use contraception at last intercourse'
  ))
]

# set master categorization for contraceptive use and try to estimate postpartum
fem_2017_2019_data[,
  curr_use2 := factor(fcase(
    meth3m_count > 0,
      'Used at least one contraceptive method at last intercourse',
    SEX3MO != 'YES',
      'Did not have sexual intercourse with male in past 3 months',
    RCURPREG == 'YES (CURRENTLY PREGNANT)',
      'Currently pregnant',
    CONSTAT1 == 32,
      'Postpartum',
    (WYNOTUSE == 'Yes' | WYNOLSTP == 'Yes'),
      'Currently trying to conceive',
    POSIBLPG == 'No',
      'Physically unable to conceive',
    POSIBLMN == 'No' & (
      P1YRELP == 'Current cohabiting partner' |
      P1YRELP == 'Current husband or (current husband from whom she is separated)'
    ),
      'Last intercourse was with partner who is physically unable to conceive',
    default = 'Otherwise did not use contraception at last intercourse'
  ))
]

# create survey design object
fem_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = fem_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```


```{r male-load-data}
male_2017_2019 <- load_NSFG_data('2017_2019', 'Male')
male_2017_2019_formats <- male_2017_2019$Formats
```

```{r male-process-data, dependson=c('helper-functions', 'male-load-data')}
male_2017_2019_data <- copy(male_2017_2019$Data[, .(
  CWPPOSS = factorize(CWPPOSS, 'Y1N5RDF', male_2017_2019_formats),
  CWPPRGNW = factorize(CWPPRGNW, 'Y1N5RDF', male_2017_2019_formats),
  CWPTRYPG = factorize(CWPTRYPG, 'Y1N5RDF', male_2017_2019_formats),
  FATHPOSS = factorize(FATHPOSS, 'Y1N5RDF', male_2017_2019_formats),
  METH3M1 = factorize(METH3M1, 'METH3MF', male_2017_2019_formats),
  METH3M2 = factorize(METH3M2, 'METH3MF', male_2017_2019_formats),
  METH3M3 = factorize(METH3M3, 'METH3MF', male_2017_2019_formats),
  METH3M4 = factorize(METH3M4, 'METH3MF', male_2017_2019_formats),
  P1COHABIT = factorize(P1COHABIT, 'Y1N5RDF', male_2017_2019_formats),
  P1CURRWIFE = factorize(P1CURRWIFE, 'Y1N5RDF', male_2017_2019_formats),
  PSURGSTR = factorize(PSURGSTR, 'N0Y1CF', male_2017_2019_formats),
  RSURGSTR  = factorize(RSURGSTR, 'N0Y1RDF', male_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'SEX3MO', male_2017_2019_formats),
  WGT2017_2019
)])

# create boolean variables for whether each specific method is currently used
for (x in 1:13) {
  set(
    male_2017_2019_data,
    j = sprintf('using3m_meth%d', x),
    value = has_level(male_2017_2019_data$METH3M1, x) |
      has_level(male_2017_2019_data$METH3M2, x) |
      has_level(male_2017_2019_data$METH3M3, x) |
      has_level(male_2017_2019_data$METH3M4, x)
  )
}

# count any respondents who are sexually active and who have an active vasectomy as using vasectomy as a contraceptive method
male_2017_2019_data[, using3m_meth3 := fifelse(
  SEX3MO == 'YES, HAD INTERCOURSE' & RSURGSTR == 'YES',
  TRUE,
  using3m_meth3
)]

# count any respondents who are sexually active and who are married or cohabitating and whose last intercourse was with wife or cohabitating partner who has active tubal ligation
male_2017_2019_data[, using3m_meth5 := fifelse(
  SEX3MO == 'YES, HAD INTERCOURSE' &
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') &
    PSURGSTR == 'YES',
  TRUE,
  using3m_meth5
)]

# create variable counting number of current methods
male_2017_2019_data[
  ,
  meth3m_count := rowSums(.SD),
  .SDcols = patterns('^using3m_meth')
]

# set master categorization for contraceptive use
male_2017_2019_data[,
  curr_use := factor(fcase(
    meth3m_count > 0,
      'Used at least one contraceptive method at last intercourse',
    SEX3MO != 'YES, HAD INTERCOURSE',
      'Did not have sexual intercourse with female in past 3 months',
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPPRGNW == 'Yes',
      'Last intercourse was with partner who is pregnant',
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPTRYPG == 'Yes',
      'Last intercourse was with partner who is trying to conceive',
    FATHPOSS == 'No',
      'Physically unable to conceive',
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPPOSS == 'No',
      'Last intercourse was with partner who is physically unable to concieve',
    default = 'Otherwise did not use contraception at last intercourse'
  ))
]

# create survey design object
male_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = male_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```

# Estimates of Master Contraceptive Use or Nonuse

## Females

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~curr_use,
    fem_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by current contraceptive status',
)[, .(level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


## Males

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~curr_use,
    male_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'males aged 15-49 living in households in the United States',
  title = 'by current contraceptive status',
)[, .(level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


# Estimation of Postpartum Females

This is calculated for `CONSTAT1` by comparing `CMINTVW` with `CMLSTPRG`. However, `CMLSTPRG` in the public use files does not contain month information, and the public use files do not contain the day of the interview, so this cannot be calculated from the public use files. Because of this, try to estimate postpartum females using `CONSTAT1`.

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~curr_use2,
    fem_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by current contraceptive status, including postpartum',
)[, .(level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

The estimate of postpartum females has a statistically significant difference from zero and does decrease the estimate of those otherwise not using contraception, so it should be included in the estimates, at least for females.

Unfortunately, there is not an equivalent recode for males in the public use data files, so males whose last intercourse was with a postpartum partner cannot be estimated.
