---
title: "Current Contraceptive Usage, with One Master Categorical Variable"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(here)
library(knitr)
library(survey)

i_am('EDA/005_comparison_with_CONSTAT.Rmd')

source(here('R/survey_helpers.R'))
source(here('R/import.R'))

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r helper-functions}
tablena <- function(x) {
  table(x, useNA = 'always')
}

in_millions <- function(dt, round_to = 3, do_round = TRUE) {
  dt <- as.data.table(dt)
  cols <- c('estimate', 'se', '2.5 %', '97.5 %')
  sub_dt <- dt[, cols, with = FALSE] / 10^6
  if (do_round) {
    sub_dt <- round(sub_dt, round_to)
  }
  dt[, (cols) := sub_dt]
  dt
}

rounded <- function(dt, round_to = 3) {
  cols <- c('estimate', 'se', '2.5 %', '97.5 %')
  sub_dt <- round(dt[, cols, with = FALSE], round_to)
  dt[, (cols) := sub_dt]
  dt
}

has_level <- function(x, ids) {
  is.element(as.integer(x), ids)
}

standard_millions_plot <- function(data) {
  ggplot(
    data = data,
    mapping = aes(
      y = reorder(level, estimate)
    )
  ) +
  geom_point(
    aes(
      x = estimate
    ),
    size = 2
  ) +
  geom_segment(
    aes(
      yend = reorder(level, estimate),
      x = `2.5 %`,
      xend = `97.5 %`
    ),
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    size = 0.5
  ) +
  scale_x_continuous(
    labels = scales::label_number(suffix = 'M', accuracy = 0.1),
  ) +
  scale_y_discrete(
    labels = scales::wrap_format(50)
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_blank()
  )
}

standard_proportions_plot <- function(data) {
  ggplot(
  data = data,
  mapping = aes(
    y = reorder(level, estimate)
  )
) +
  geom_point(
    aes(
      x = estimate
    ),
    size = 2
  ) +
  geom_segment(
    aes(
      yend = reorder(level, estimate),
      x = `2.5 %`,
      xend = `97.5 %`
    ),
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    size = 0.5
  ) +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    expand = c(0, 0),
    labels = scales::label_percent(1),
    limits = c(0, 1)
  ) +
  scale_y_discrete(
    labels = scales::wrap_format(50)
  ) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    legend.position = 'none'
  )
}

env <- new.env()

env$estimates <- data.table()

collect_estimate <- function(est, title, domain = '', type = '', round_to = 3, plot = FALSE) {
  this_title <- title
  this_domain <- domain
  this_type <- type

  if (type == 'total') {
    dt <- in_millions(est, round_to)
  } else if (type == 'proportion') {
    dt <- rounded(est, round_to)
  } else {
    dt <- as.data.table(est)
  }
  env$estimates <- rbindlist(list(
    env$estimates[
      !(
        title == this_title &
          domain == this_domain &
          type == this_type
      )
    ],
    dt[, .(
      domain = this_domain,
      title = this_title,
      type = this_type,
      level,
      estimate,
      se,
      `2.5 %`,
      `97.5 %`
    )]
  ))

  env$estimates[
    title == this_title &
      domain == this_domain &
      type == this_type
  ]
}
```


```{r fem-load-data}
fem_2017_2019 <- load_NSFG_data('2017_2019', 'FemResp')
fem_2017_2019_formats <- fem_2017_2019$Formats
```

```{r fem-process-data, dependson=c('helper-functions', 'fem-load-data')}
fem_2017_2019_data <- copy(fem_2017_2019$Data[, .(
  CASEID,
  CONSTAT1 = factorize(CONSTAT1, 'CONSTATF', fem_2017_2019_formats),
  CONSTAT2 = factorize(CONSTAT2, 'CONSTATF', fem_2017_2019_formats),
  CONSTAT3 = factorize(CONSTAT3, 'CONSTATF', fem_2017_2019_formats),
  CONSTAT4 = factorize(CONSTAT4, 'CONSTATF', fem_2017_2019_formats),
  FECUND,
  METH3M1,
  METH3M2,
  METH3M3,
  P1YRELP = factorize(P1YRELP, 'P1YRELP', fem_2017_2019_formats),
  P1YRELP = factorize(P1YRELP, 'P1YRELP', fem_2017_2019_formats),
  POSIBLMN = factorize(POSIBLMN, 'Y1N5RDF', fem_2017_2019_formats),
  POSIBLPG = factorize(POSIBLPG, 'Y1N5RDF', fem_2017_2019_formats),
  PSURGSTR = factorize(PSURGSTR, 'Y1N5C', fem_2017_2019_formats),
  RCURPREG = factorize(RCURPREG, 'RCURPREG', fem_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'Y1N2RECF', fem_2017_2019_formats),
  STRLOPER,
  TUBS = factorize(TUBS, 'Y1N5C', fem_2017_2019_formats),
  WGT2017_2019,
  WYNOLSTP = factorize(WYNOLSTP, 'Y1N5RDF', fem_2017_2019_formats),
  WYNOTUSE = factorize(WYNOTUSE, 'Y1N5RDF', fem_2017_2019_formats)
)])

# remove sterility not from contraceptive procedure from METH3M series
fem_2017_2019_data[
  METH3M3 %in% 20:21,
  `:=`(
    METH3M3 = NA
  )
]
fem_2017_2019_data[
  METH3M2 %in% 20:21,
  `:=`(
    METH3M2 = METH3M3,
    METH3M3 = NA
  )
]
fem_2017_2019_data[
  METH3M1 %in% 20:21,
  `:=`(
    METH3M1 = METH3M2,
    METH3M2 = METH3M3,
    METH3M3 = NA
  )
]

# make METH3M series variables into factors
fem_2017_2019_data[, `:=`(
  METH3M1 = factorize(METH3M1, 'METH3MF', fem_2017_2019_formats),
  METH3M2 = factorize(METH3M2, 'METH3MF', fem_2017_2019_formats),
  METH3M3 = factorize(METH3M3, 'METH3MF', fem_2017_2019_formats)
)]

# create boolean variables for whether each specific method is currently used
for (x in c(1:19, 22:24)) {
  set(
    fem_2017_2019_data,
    j = sprintf('using3m_meth%d', x),
    value = has_level(fem_2017_2019_data$METH3M1, x) |
      has_level(fem_2017_2019_data$METH3M2, x) |
      has_level(fem_2017_2019_data$METH3M3, x)
  )
}

# count any respondents who are sexually active and whose last intercourse was with husband or cohabitating partner who has an active vasectomy as using vasectomy as a contraceptive method
fem_2017_2019_data[, using3m_meth3 := fifelse(
  SEX3MO == 'YES' & PSURGSTR == 'YES' & (
    P1YRELP == 'Current husband or (current husband from whom she is separated)' |
    P1YRELP == 'Current cohabiting partner'
  ),
  TRUE,
  using3m_meth3
)]

# count any respondents with current tubal ligations who are sexually active as currently using tubal ligation as a contraceptive method
fem_2017_2019_data[, using3m_meth4 := fifelse(
  SEX3MO == 'YES' & TUBS == 'YES',
  TRUE,
  using3m_meth4
)]

# create variable counting number of current methods
fem_2017_2019_data[
  ,
  meth3m_count := rowSums(.SD),
  .SDcols = patterns('^using3m_meth')
]

# set master categorization for contraceptive use
fem_2017_2019_data[,
  curr_use := factor(fcase(
    meth3m_count > 0,
      'Used at least one contraceptive method at last intercourse',
    SEX3MO != 'YES',
      'Did not have sexual intercourse with male in past 3 months',
    RCURPREG == 'YES (CURRENTLY PREGNANT)',
      'Currently pregnant',
    (WYNOTUSE == 'Yes' | WYNOLSTP == 'Yes'),
      'Currently trying to conceive',
    POSIBLPG == 'No',
      'Physically unable to conceive',
    POSIBLMN == 'No' & (
      P1YRELP == 'Current cohabiting partner' |
      P1YRELP == 'Current husband or (current husband from whom she is separated)'
    ),
      'Last intercourse was with partner who is physically unable to conceive',
    default = 'Otherwise did not use contraception at last intercourse'
  ))
]

# set master categorization for contraceptive use and try to estimate postpartum
fem_2017_2019_data[,
  curr_use2 := factor(fcase(
    meth3m_count > 0,
      'Used at least one contraceptive method at last intercourse',
    SEX3MO != 'YES',
      'Did not have sexual intercourse with male in past 3 months',
    RCURPREG == 'YES (CURRENTLY PREGNANT)',
      'Currently pregnant',
    CONSTAT1 == 'Postpartum',
      'Postpartum',
    (WYNOTUSE == 'Yes' | WYNOLSTP == 'Yes'),
      'Currently trying to conceive',
    POSIBLPG == 'No',
      'Physically unable to conceive',
    POSIBLMN == 'No' & (
      P1YRELP == 'Current cohabiting partner' |
      P1YRELP == 'Current husband or (current husband from whom she is separated)'
    ),
      'Last intercourse was with partner who is physically unable to conceive',
    default = 'Otherwise did not use contraception at last intercourse'
  ))
]

# create survey design object
fem_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = fem_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```


```{r male-load-data}
male_2017_2019 <- load_NSFG_data('2017_2019', 'Male')
male_2017_2019_formats <- male_2017_2019$Formats
```

```{r male-process-data, dependson=c('helper-functions', 'male-load-data')}
male_2017_2019_data <- copy(male_2017_2019$Data[, .(
  CWPPOSS = factorize(CWPPOSS, 'Y1N5RDF', male_2017_2019_formats),
  CWPPRGNW = factorize(CWPPRGNW, 'Y1N5RDF', male_2017_2019_formats),
  CWPTRYPG = factorize(CWPTRYPG, 'Y1N5RDF', male_2017_2019_formats),
  FATHPOSS = factorize(FATHPOSS, 'Y1N5RDF', male_2017_2019_formats),
  METH3M1 = factorize(METH3M1, 'METH3MF', male_2017_2019_formats),
  METH3M2 = factorize(METH3M2, 'METH3MF', male_2017_2019_formats),
  METH3M3 = factorize(METH3M3, 'METH3MF', male_2017_2019_formats),
  METH3M4 = factorize(METH3M4, 'METH3MF', male_2017_2019_formats),
  P1COHABIT = factorize(P1COHABIT, 'Y1N5RDF', male_2017_2019_formats),
  P1CURRWIFE = factorize(P1CURRWIFE, 'Y1N5RDF', male_2017_2019_formats),
  PSURGSTR = factorize(PSURGSTR, 'N0Y1CF', male_2017_2019_formats),
  RSURGSTR  = factorize(RSURGSTR, 'N0Y1RDF', male_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'SEX3MO', male_2017_2019_formats),
  WGT2017_2019
)])

# create boolean variables for whether each specific method is currently used
for (x in 1:13) {
  set(
    male_2017_2019_data,
    j = sprintf('using3m_meth%d', x),
    value = has_level(male_2017_2019_data$METH3M1, x) |
      has_level(male_2017_2019_data$METH3M2, x) |
      has_level(male_2017_2019_data$METH3M3, x) |
      has_level(male_2017_2019_data$METH3M4, x)
  )
}

# count any respondents who are sexually active and who have an active vasectomy as using vasectomy as a contraceptive method
male_2017_2019_data[, using3m_meth3 := fifelse(
  SEX3MO == 'YES, HAD INTERCOURSE' & RSURGSTR == 'YES',
  TRUE,
  using3m_meth3
)]

# count any respondents who are sexually active and who are married or cohabitating and whose last intercourse was with wife or cohabitating partner who has active tubal ligation
male_2017_2019_data[, using3m_meth5 := fifelse(
  SEX3MO == 'YES, HAD INTERCOURSE' &
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') &
    PSURGSTR == 'YES',
  TRUE,
  using3m_meth5
)]

# create variable counting number of current methods
male_2017_2019_data[
  ,
  meth3m_count := rowSums(.SD),
  .SDcols = patterns('^using3m_meth')
]

# set master categorization for contraceptive use
male_2017_2019_data[,
  curr_use := factor(fcase(
    meth3m_count > 0,
      'Used at least one contraceptive method at last intercourse',
    SEX3MO != 'YES, HAD INTERCOURSE',
      'Did not have sexual intercourse with female in past 3 months',
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPPRGNW == 'Yes',
      'Last intercourse was with partner who is pregnant',
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPTRYPG == 'Yes',
      'Last intercourse was with partner who is trying to conceive',
    FATHPOSS == 'No',
      'Physically unable to conceive',
    (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPPOSS == 'No',
      'Last intercourse was with partner who is physically unable to concieve',
    default = 'Otherwise did not use contraception at last intercourse'
  ))
]

# create survey design object
male_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = male_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```

# Own Estimates of Contraceptive Use or Nonuse

## Females

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~curr_use,
    fem_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by current contraceptive status',
)[, .(level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


## Males

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~curr_use,
    male_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'males aged 15-49 living in households in the United States',
  title = 'by current contraceptive status',
)[, .(level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```



# `CONSTAT`

## Levels

```{r, dependson=c('fem-process-data')}
as.data.table(levels(fem_2017_2019_data$CONSTAT1))[
  , .(
    ID = .I,
    level = V1
  )
] |> kable()
```

## Estimates

```{r, dependson=c('fem-process-data')}
svyci(
  ~CONSTAT1,
  fem_2017_2019_svy,
  svytotal,
  ordered = FALSE
) |> in_millions() |> kable()
```

## Using Contraception

```{r, dependson=c('fem-process-data')}
svyci(
  ~I(as.integer(CONSTAT1) %in% 1:22),
  fem_2017_2019_svy,
  svytotal,
  ordered = FALSE
) |> in_millions() |> kable()
```

## Not Sexually Active

```{r, dependson=c('fem-process-data')}
svyci(
  ~I(as.integer(CONSTAT1) %in% 31:32),
  fem_2017_2019_svy,
  svytotal,
  ordered = FALSE
)[level == TRUE] |> in_millions() |> kable()
```


# Discrepancies

## Not Sexually Active

There is a big discrepancy between the number of sexually inactive females reported here (21.361) and what is estimated by `CONSTAT1` (12.236) million.

Here only females who have had sexual intercourse in the previous 3 months are counted as using contraception. By `CONSTAT1` logic, any females who have had contraceptive sterilizing operations are assigned a `CONSTAT1` value of `01` "Female sterilization."

```{r, dependson=c('fem-process-data')}
svyci(
  ~I(
    SEX3MO != 'YES' &
    FECUND == 1 &
    STRLOPER %in% 1:2
  ),
  fem_2017_2019_svy,
  svytotal
) |> in_millions() |> kable()
```

This accounts for 2 million females of the difference in the estimates.

Likewise, all pregnant females are given a `CONSTAT1` of `30`, regardless of whether they are sexually active.


```{r, dependson=c('fem-process-data')}
svyci(
  ~I(
    SEX3MO != 'YES' &
    RCURPREG == 'YES (CURRENTLY PREGNANT)'
  ),
  fem_2017_2019_svy,
  svytotal
) |> in_millions() |> kable()
```

This only accounts for aabout 0.195 million females of the difference in the estimates.

```{r, dependson=c('fem-process-data')}
svyci(
  ~I(
    SEX3MO != 'YES' &
    !is.element(CONSTAT1, 40:41)
  ),
  fem_2017_2019_svy,
  svytotal
) |> in_millions() |> kable()
```

More generally, though, all of the females who have not had sexual intercourse in the previous 3 months, but who are given a `CONSTAT1` value not `40` or `41` amount to about 9.208 million females. The difference in the estimates is $21.361 - 12.236$ or $9.125$ million females. Thus, this accounts for the difference in the estimates.


# Estimation of Postpartum Females

This is calculated for `CONSTAT1` by comparing `CMINTVW` with `CMLSTPRG`. However, `CMLSTPRG` in the public use files does not contain month information, and the public use files do not contain the day of the interview, so this cannot be calculated from the public use files. Because of this, try to estimate postpartum females using `CONSTAT1`.

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~curr_use2,
    fem_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by current contraceptive status, including postpartum',
)[, .(level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

The estimate of postpartum females has a statistically significant difference from zero and does decrease the estimate of those otherwise not using contraception, so it should be included in the estimates, at least for females.

Unfortunately, there is not an equivalent recode for males in the public use data files, so males whose last intercourse was with a postpartum partner cannot be estimated.



# To Do

* Compare estimates for females against `CONSTAT1`; see recode specification
    * Pill is more for `CONSTAT1` 10.139 vs 7.947
    * Implant 1.459 vs 1.161
    * Injectable 1.423 vs 1.041
    * Nonsurgical sterile female 1.861 vs 0.369
    * Other nonuser had intercourse in the 3 months prior to interview 5.081 vs 3.917
* Go over outline and adjust


# Outline

## Frames for analysis

### By Table

* Female persons
* Male persons

### By Columns

* "Current" -- at last intercourse with male in past 3 months
* "Recent" -- over previous year
* "Ever" -- over lifetime

## Questions for Users

* Which contraceptives?
* How many?
* For reasons other than contraception?
* For "recent," how consistent?
* For "recent" and for multiple methods users, simultaneously or alternatively?
* For "ever" and who have stopped using a method, why did they stop?

## Questions for Nonusers

* Not sexually active with male?
    * Ever or just now?
    * Exclusively same-sex partner(s)?
* Currently pregnant?
* Noncontraceptive surgical sterilization
* Physically unable to have child
* Cohabitating partner or husband surgical sterilization
* Cohabitating partner or husband physically unable to have child
* Trying to get pregnant?
* For those fertile, heterosexually active, not using contraception, and not desiring pregnancy, why no contraception usage?

## Covariates/Subdomains

* Want children in future, but not just not now
* By sex education
* By income
* By religion
    * Roman Catholics are a subdomain of interest.
* Second phase vs. first-phase (to investigate nonresponse bias)

## Other

* Try combining with 2015-2017 data for more precise estimates?