---
title: "Persons Who Have Not Had Intercourse with Someone of the Opposite Sex in Previous 3 Months"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(here)
library(knitr)
library(survey)

i_am('EDA/008_not_sexually_active.Rmd')

source(here('R/EDA_helpers.R'))
source(here('R/import.R'))
source(here('R/survey_helpers.R'))

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r fem-load-data}
fem_2017_2019 <- load_NSFG_data('2017_2019', 'FemResp')
fem_2017_2019_formats <- fem_2017_2019$Formats
```

```{r fem-process-data, dependson=c('fem-load-data')}
fem_2017_2019_data <- copy(fem_2017_2019$Data[, .(
  ATTRACT = factorize(ATTRACT, 'ATTRACT', fem_2017_2019_formats),
  CASEID,
  HADSEX = factorize(HADSEX, 'HADSEX', fem_2017_2019_formats),
  MFLASTP = factorize(MFLASTP, 'MALFEMNAF', fem_2017_2019_formats),
  SAMESEXANY = factorize(SAMESEXANY, 'Y1N5NAC', fem_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'Y1N2RECF', fem_2017_2019_formats),
  WGT2017_2019,
  YNOSEX = factorize(YNOSEX, 'YNOSEX', fem_2017_2019_formats)
)])

# create two-level of SEX3MO
fem_2017_2019_data[, sex_active := factor(fifelse(
  SEX3MO == 'YES',
  'Had sexual intercourse with male in previous 3 months',
  'Did not have sexual intercourse with male in previous 3 months'
))]

# remove empty 'Not Applicable' level from ATTRACT
fem_2017_2019_data[, ATTRACT := factor(ATTRACT)]

# create survey design object
fem_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = fem_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```

```{r male-load-data}
male_2017_2019 <- load_NSFG_data('2017_2019', 'Male')
male_2017_2019_formats <- male_2017_2019$Formats
```

```{r male-process-data, dependson=c('male-load-data')}
male_2017_2019_data <- copy(male_2017_2019$Data[, .(
  ATTRACT = factorize(ATTRACT, 'ATTRACT', male_2017_2019_formats),
  CASEID,
  HADSEX = factorize(HADSEX, 'HADSEX', male_2017_2019_formats),
  MFLASTP = factorize(MFLASTP, 'MALFEMNAF', male_2017_2019_formats),
  SAMESEXANY = factorize(SAMESEXANY, 'Y1N5NAC', male_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'SEX3MO', male_2017_2019_formats),
  WGT2017_2019,
  YNOSEX = factorize(YNOSEX, 'YNOSEX', male_2017_2019_formats)
)])

# create two-level of SEX3MO
male_2017_2019_data[, sex_active := factor(fifelse(
  SEX3MO == 'YES, HAD INTERCOURSE',
  'Had sexual intercourse with female in previous 3 months',
  'Did not have sexual intercourse with female in previous 3 months'
))]

# remove empty 'Not Applicable' level from ATTRACT
male_2017_2019_data[, ATTRACT := factor(ATTRACT)]

# create survey design object
male_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = male_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```


# Females

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(SEX3MO != 'YES'),
    fem_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'who did not have sexual intercourse with male in past 3 months'
) |> kable()
```


## Never Sexually Active

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(HADSEX == 'NO, R NEVER HAD INTERCOURSE'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'females who did not have sexual intercourse with male in past 3 months',
  title = 'who have never had sexual intercourse with male'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(HADSEX == 'NO, R NEVER HAD INTERCOURSE'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'females who did not have sexual intercourse with male in past 3 months',
  title = 'who have never had sexual intercourse with male'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~YNOSEX,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by reason for not having had sexual intercourse with male'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~YNOSEX,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by reason for not having had sexual intercourse with male'
) |> kable()
```


## Attraction to Sexes

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO == 'YES'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have had intercourse with male in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO == 'NO'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have had sexual intercourse with male, but not in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      YNOSEX == 'Other'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male for reason "Other"',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      YNOSEX == 'Other'
    ),
    svytotal,
    ordered = FALSE
  ),
  type = 'total',
  domain = 'females who have never had sexual intercourse with male for reason "Other"',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svytotal,
    ordered = FALSE
  ),
  type = 'total',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

### Chi-squared Test

```{r, dependson=c('fem-process-data')}
svytable(~ATTRACT+sex_active, fem_2017_2019_svy) |> kable()
```

```{r, dependson=c('fem-process-data')}
env$estimates[
  type == 'proportion' &
    title == 'by strength of attraction to sexes' &
    (
      domain == 'females who have not had intercourse with a male in past 3 months' |
      domain == 'females who have had intercourse with male in past 3 months'
    ),
  .(
    domain, level, estimate, se, `2.5 %`, `97.5 %`
  )
] |> kable()
```

```{r, dependson=c('fem-process-data')}
svychisq(~ATTRACT+sex_active, fem_2017_2019_svy)
```



## Sexual Experience with Female Partner

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~SAMESEXANY,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by whether had sexual encounter with female'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~SAMESEXANY,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by whether had sexual encounter with female'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(MFLASTP == 'Female'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by whether last sexual experience was with female partner'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(MFLASTP == 'Female'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by whether last sexual experience was with female partner'
) |> kable()
```



# Males

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(SEX3MO != 'YES, HAD INTERCOURSE'),
    male_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'males aged 15-49 living in households in the United States',
  title = 'who did not have sexual intercourse with female in past 3 months'
) |> kable()
```


## Never Sexually Active

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(HADSEX == 'NO, R NEVER HAD INTERCOURSE'),
    subset(
      male_2017_2019_svy,
      SEX3MO != 'YES, HAD INTERCOURSE'
    ),
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'males who did not have sexual intercourse with female in past 3 months',
  title = 'who never had sexual intercourse with female'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(HADSEX == 'NO, R NEVER HAD INTERCOURSE'),
    subset(
      male_2017_2019_svy,
      SEX3MO != 'YES, HAD INTERCOURSE'
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'males who did not have sexual intercourse with female in past 3 months',
  title = 'who never had sexual intercourse with female'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~YNOSEX,
    subset(
      male_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'males who have never had sexual intercourse with female',
  title = 'by reason for not having had sexual intercourse with female'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      male_2017_2019_svy,
      YNOSEX == 'Other'
    ),
    svytotal,
    ordered = FALSE
  ),
  type = 'total',
  domain = 'males who have never had sexual intercourse with female for reason "Other"',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      male_2017_2019_svy,
      YNOSEX == 'Other'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'males who have never had sexual intercourse with female for reason "Other"',
  title = 'by strength of attraction to sexes'
) |> kable()
```


## Same-Sex Attraction

```{r, dependson=c('male-process-data')}
svytable(~ATTRACT+sex_active, male_2017_2019_svy) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      male_2017_2019_svy,
      SEX3MO != 'YES, HAD INTERCOURSE'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'males who did not have sexual intercourse with female in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      male_2017_2019_svy,
      SEX3MO == 'YES, HAD INTERCOURSE'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'males have had sexual intercourse with female in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('male-process-data')}
svychisq(~ATTRACT+sex_active, male_2017_2019_svy)
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      male_2017_2019_svy,
      SEX3MO != 'YES, HAD INTERCOURSE'
    ),
    svytotal,
    ordered = FALSE
  ),
  type = 'total',
  domain = 'males who did not have sexual intercourse with female in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```


## Sexual Experience with Male Partner

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~SAMESEXANY,
    subset(
      male_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'males who have never had sexual intercourse with female',
  title = 'by whether had sexual encounter with male'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~SAMESEXANY,
    subset(
      male_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'males who have never had sexual intercourse with female',
  title = 'by whether had sexual encounter with male'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(MFLASTP == 'Male'),
    subset(
      male_2017_2019_svy,
      SEX3MO != 'YES, HAD INTERCOURSE'
    ),
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'males who have not had intercourse with a female in past 3 months',
  title = 'by whether last sexual experience was with male partner'
) |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(MFLASTP == 'Male'),
    subset(
      male_2017_2019_svy,
      SEX3MO != 'YES, HAD INTERCOURSE'
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'males who have not had intercourse with a female in past 3 months',
  title = 'by whether last sexual experience was with male partner'
) |> kable()
```



# Things to Investigate

* Among those who otherwise did not use contraception at last intercourse, was this due to complete nonuse or inconsistent use? That is, how many used a method of contraception with last partner during the previous 12 months?
* Summarize contraceptive use in previous 12 months by those not sexually active with opposite sex, by those sexually active and did not use contraception at all, and by contraceptive method among those who used at least one method of contraception
* Among those sexually active with the opposite sex in the previous 12 months and who used at least one method of contraception, how frequently was any method of contraception used?
* Among those who used condoms in previous 12 months, how frequently were condoms used, in the past 12 months and past 4 weeks?
* Among females who used the pill in the previous 12 months, how many pills were missed?
* Among those at risk for unintended pregnancy, why not using contraception (Fem: EH-2, Male: None)
* Why respondents believe themselves to be sterile (Fem: DE-2 REASIMPP & EH-2 WHYNOTPG, Male: None)
* Ever tried over lifetime and why quit (Fem: EA, DC; Male: BB)
    * Include tubal ligation and vasectomy reversal
* Replicate analysis in SAS
    * Attempt single-factor categorical data analysis, e.g., whether totals of otherwise not using contraception and currently trying to conceive are different
* Combining with 2015-2017 data for more precise estimates
* Nonresponse bias by comparing second phase vs. first-phase


## Variables for Recent Contraceptive Use

* Each month for 3 previous years
    * Females
        * EC MONSX whether intercourse occurred in given month
        * ED METHX each method used in each month
        * ED SIMSEQX whether methods used together or at different times
* Past 12 months
    * Females
        * EL for frequency of condom or any contraceptive use and missed pills (also past 4 weeks)
    * Males
        * BC overall frequency of sexual intercourse and condom use (also past 4 weeks)
        * CF methods used with current wife or cohabitating partner and "most often" used method
        * DG methods used with most recent 3 partners, "main" method, and frequency of condom and any method use


# Outline

* Last
    * Use by Method(s)
    * Nonuse by Pregnancy, Intent to Conceive, Sterility, etc
* Not Sexually Active with Opposite Sex
* Recently
    * Consistency
    * Use for Other Reasons (current or previous month)
    * Reasons for Nonuse (current month)
* Use Ever and Discontinuation
