---
title: "Persons Who Have Not Had Intercourse with Someone of the Opposite Sex in Previous 3 Months"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(here)
library(knitr)
library(survey)

i_am('EDA/008_not_sexually_active.Rmd')

source(here('R/survey_helpers.R'))
source(here('R/import.R'))

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r helper-functions}
tablena <- function(x) {
  table(x, useNA = 'always')
}

in_millions <- function(dt, round_to = 3, do_round = TRUE) {
  dt <- as.data.table(dt)
  cols <- c('estimate', 'se', '2.5 %', '97.5 %')
  sub_dt <- dt[, cols, with = FALSE] / 10^6
  if (do_round) {
    sub_dt <- round(sub_dt, round_to)
  }
  dt[, (cols) := sub_dt]
  dt
}

rounded <- function(dt, round_to = 3) {
  cols <- c('estimate', 'se', '2.5 %', '97.5 %')
  sub_dt <- round(dt[, cols, with = FALSE], round_to)
  dt[, (cols) := sub_dt]
  dt
}

has_level <- function(x, ids) {
  is.element(as.integer(x), ids)
}

standard_millions_plot <- function(data) {
  ggplot(
    data = data,
    mapping = aes(
      y = reorder(level, estimate)
    )
  ) +
  geom_point(
    aes(
      x = estimate
    ),
    size = 2
  ) +
  geom_segment(
    aes(
      yend = reorder(level, estimate),
      x = `2.5 %`,
      xend = `97.5 %`
    ),
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    size = 0.5
  ) +
  scale_x_continuous(
    labels = scales::label_number(suffix = 'M', accuracy = 0.1),
  ) +
  scale_y_discrete(
    labels = scales::wrap_format(50)
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_blank()
  )
}

standard_proportions_plot <- function(data) {
  ggplot(
  data = data,
  mapping = aes(
    y = reorder(level, estimate)
  )
) +
  geom_point(
    aes(
      x = estimate
    ),
    size = 2
  ) +
  geom_segment(
    aes(
      yend = reorder(level, estimate),
      x = `2.5 %`,
      xend = `97.5 %`
    ),
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    size = 0.5
  ) +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    expand = c(0, 0),
    labels = scales::label_percent(1),
    limits = c(0, 1)
  ) +
  scale_y_discrete(
    labels = scales::wrap_format(50)
  ) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    legend.position = 'none'
  )
}

env <- new.env()

env$estimates <- data.table()

collect_estimate <- function(est, title, domain = '', type = '', round_to = 3, plot = FALSE) {
  this_title <- title
  this_domain <- domain
  this_type <- type

  if (type == 'total') {
    dt <- in_millions(est, round_to)
  } else if (type == 'proportion') {
    dt <- rounded(est, round_to)
  } else {
    dt <- as.data.table(est)
  }
  env$estimates <- rbindlist(list(
    env$estimates[
      !(
        title == this_title &
          domain == this_domain &
          type == this_type
      )
    ],
    dt[, .(
      domain = this_domain,
      title = this_title,
      type = this_type,
      level,
      estimate,
      se,
      `2.5 %`,
      `97.5 %`
    )]
  ))

  env$estimates[
    title == this_title &
      domain == this_domain &
      type == this_type
  ]
}
```


```{r fem-load-data}
fem_2017_2019 <- load_NSFG_data('2017_2019', 'FemResp')
fem_2017_2019_formats <- fem_2017_2019$Formats
```

```{r fem-process-data, dependson=c('helper-functions', 'fem-load-data')}
fem_2017_2019_data <- copy(fem_2017_2019$Data[, .(
  ATTRACT = factorize(ATTRACT, 'ATTRACT', fem_2017_2019_formats),
  CASEID,
  HADSEX = factorize(HADSEX, 'HADSEX', fem_2017_2019_formats),
  MFLASTP = factorize(MFLASTP, 'MALFEMNAF', fem_2017_2019_formats),
  SAMESEXANY = factorize(SAMESEXANY, 'Y1N5NAC', fem_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'Y1N2RECF', fem_2017_2019_formats),
  WGT2017_2019,
  YNOSEX = factorize(YNOSEX, 'YNOSEX', fem_2017_2019_formats)
)])

# create two-level of SEX3MO
fem_2017_2019_data[, sex_active := factor(fifelse(
  SEX3MO == 'YES',
  'Had sexual intercourse with male in previous 3 months',
  'Did not have sexual intercourse with male in previous 3 months'
))]

# remove empty 'Not Applicable' level from ATTRACT
fem_2017_2019_data[, ATTRACT := factor(ATTRACT)]

# create survey design object
fem_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = fem_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```

```{r male-load-data}
male_2017_2019 <- load_NSFG_data('2017_2019', 'Male')
male_2017_2019_formats <- male_2017_2019$Formats
```

```{r male-process-data, dependson=c('helper-functions', 'male-load-data')}
male_2017_2019_data <- copy(male_2017_2019$Data[, .(
  CASEID,
  SECU,
  SEST,
  WGT2017_2019
)])

# create survey design object
male_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = male_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```


# Females

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(SEX3MO != 'YES'),
    fem_2017_2019_svy,
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'who did not have sexual intercourse with male in past 3 months'
) |> kable()
```


## Never Sexually Active

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(HADSEX == 'NO, R NEVER HAD INTERCOURSE'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'females who did not have sexual intercourse with male in past 3 months',
  title = 'who have never had sexual intercourse with male'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(HADSEX == 'NO, R NEVER HAD INTERCOURSE'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'females who did not have sexual intercourse with male in past 3 months',
  title = 'who have never had sexual intercourse with male'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~YNOSEX,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by reason for not having had sexual intercourse with male'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~YNOSEX,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by reason for not having had sexual intercourse with male'
) |> kable()
```


## Attraction to Sexes

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO == 'YES'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have had intercourse with male in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO == 'NO'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have had sexual intercourse with male, but not in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      YNOSEX == 'Other'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male for reason "Other"',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      YNOSEX == 'Other'
    ),
    svytotal,
    ordered = FALSE
  ),
  type = 'total',
  domain = 'females who have never had sexual intercourse with male for reason "Other"',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svytotal,
    ordered = FALSE
  ),
  type = 'total',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~ATTRACT,
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svymean,
    ordered = FALSE
  ),
  type = 'proportion',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by strength of attraction to sexes'
) |> kable()
```

### Chi-squared Test

```{r, dependson=c('fem-process-data')}
svytable(~ATTRACT+sex_active, fem_2017_2019_svy) |> kable()
```

```{r, dependson=c('fem-process-data')}
env$estimates[
  type == 'proportion' &
    title == 'by strength of attraction to sexes' &
    (
      domain == 'females who have not had intercourse with a male in past 3 months' |
      domain == 'females who have had intercourse with male in past 3 months'
    ),
  .(
    domain, level, estimate, se, `2.5 %`, `97.5 %`
  )
] |> kable()
```

```{r, dependson=c('fem-process-data')}
svychisq(~ATTRACT+sex_active, fem_2017_2019_svy)
```



## Sexual Experience with Female Partner

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~SAMESEXANY,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by whether had sexual encounter with female'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~SAMESEXANY,
    subset(
      fem_2017_2019_svy,
      HADSEX == 'NO, R NEVER HAD INTERCOURSE'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who have never had sexual intercourse with male',
  title = 'by whether had sexual encounter with female'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(MFLASTP == 'Female'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by whether last sexual experience was with female partner'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(MFLASTP == 'Female'),
    subset(
      fem_2017_2019_svy,
      SEX3MO != 'YES'
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'females who have not had intercourse with a male in past 3 months',
  title = 'by whether last sexual experience was with female partner'
) |> kable()
```



# Things to Investigate

* Not sexually active with opposite sex
    * Males
* Among those at risk for unintended pregnancy, why not using contraception
* Why respondents believe themselves to be sterile
* Ever tried over lifetime and why quit
* Time periods of contraceptive use for unintended pregnancy detection and efficacy/failure rate estimation
* Consistency
* Replicate analysis in SAS
    * Attempt single-factor categorical data analysis, e.g., whether totals of otherwise not using contraception and currently trying to conceive are different
* Combining with 2015-2017 data for more precise estimates
* Nonresponse bias by comparing second phase vs. first-phase

## Covariates/Subdomains

* Want children in future, but not just not now
* By sex education
* By income
* By religiosity
