---
title: "Why Not Use Contraception?"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(knitr)
library(survey)

i_am('EDA/011_why_not_use.Rmd')

source(here('R/EDA_helpers.R'))
source(here('R/data_table_helpers.R'))
source(here('R/factor_helpers.R'))
source(here('R/import.R'))
source(here('R/survey_helpers.R'))

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r fem-load-data}
fem_2017_2019 <- load_NSFG_data('2017_2019', 'FemResp')
fem_2017_2019_formats <- fem_2017_2019$Formats

fem_2017_2019_formats <- rbindlist(list(
  fem_2017_2019_formats,
  data.table(
    format_name = 'WHYNOTPGF',
    factor_value = 13,
    factor_label = 'No menstrual cycle or menstrual irregularity'
  )
))

```

```{r fem-process-data, dependson=c('fem-load-data')}
fem_2017_2019_data <- copy(fem_2017_2019$Data[, .(
  CASEID,
  CMINTVW,
  CURRMETH1,
  CURRPREG,
  HPPREGQ = factorize(HPPREGQ, 'HPPREGQ', fem_2017_2019_formats),
  MAINNOUSE  = factorize(MAINNOUSE, 'MAINNOUSE', fem_2017_2019_formats),
  RSTRSTAT,
  SECU,
  SEST,
  WGT2017_2019,
  WHYNOTPG1 = factorize(WHYNOTPG1, 'WHYNOTPGF', fem_2017_2019_formats),
  WHYNOTPG2 = factorize(WHYNOTPG2, 'WHYNOTPGF', fem_2017_2019_formats),
  WHYNOUSING1 = factorize(WHYNOUSING1, 'WHYNOUSNG', fem_2017_2019_formats),
  WHYNOUSING2 = factorize(WHYNOUSING2, 'WHYNOUSNG', fem_2017_2019_formats),
  WHYNOUSING3 = factorize(WHYNOUSING3, 'WHYNOUSNG', fem_2017_2019_formats),
  WHYNOUSING4 = factorize(WHYNOUSING4, 'WHYNOUSNG', fem_2017_2019_formats),
  WHYNOUSING5 = factorize(WHYNOUSING5, 'WHYNOUSNG', fem_2017_2019_formats),
  WYNOTUSE = factorize(WYNOTUSE, 'Y1N5RDF', fem_2017_2019_formats)
)])


# create pivot table for MONSX series variables
fem_2017_2019_MONSX <- create_pivot_table(
  'CMMONSX',
  2:48,
  fem_2017_2019$Data,
  c('CASEID', 'CMINTVW', 'CMLSTYR')
)[
  create_pivot_table(
    'MONSX',
    2:48,
    fem_2017_2019$Data,
    'CASEID'
  ),
  on = c(CASEID = 'CASEID', CMMONSXID = 'MONSXID')
]


# create boolean variable for whether MONSX series variables indicate respondent had intercourse during month of interview
fem_2017_2019_data <- fem_2017_2019_MONSX[
  CMINTVW == CMMONSX,
  .(
    CASEID,
    MONSX,
    CMMONSX,
    had_sex_this_month = MONSX == 1
  )
][
  fem_2017_2019_data,
  on = 'CASEID'
]


# create pivot table for METHX series variables
fem_2017_2019_METHX <- create_pivot_table(
  'METHX',
  1:192,
  fem_2017_2019$Data,
  'CASEID'
)

fem_2017_2019_METHX[,
  CMMHCALXID := ceiling(METHXID / 4)
]

fem_2017_2019_METHX <- fem_2017_2019_METHX[
  create_pivot_table(
    'CMMHCALX',
    1:48,
    fem_2017_2019$Data,
    c('CASEID', 'CMINTVW', 'CMLSTYR')
  ),
  on = c('CASEID', 'CMMHCALXID')
]


# create boolean variable for whether METHX series variables indicate respondent used at least one method of contraception during month of interview
fem_2017_2019_data <- fem_2017_2019_METHX[
  CMINTVW == CMMHCALX,
  .(
    sterilized = any(METHX == 6, na.rm = TRUE)
  ),
  by = CASEID
][
  fem_2017_2019_data,
  on = 'CASEID'
]

fem_2017_2019_data[, sterilized := na_to_false(sterilized)]


# create categorical variable for contraceptive nonuse status
fem_2017_2019_data[, nonuse_status := factor(fcase(
  WYNOTUSE == 'Not Applicable',
    'Outside of universe: currently pregnant, did not have sex in month of interview, sterilized or sterile, or used a contraceptive method in month of interview',
  WYNOTUSE == 'Yes',
    'Trying to get pregnant',
  HPPREGQ == 'Yes',
    'Partner wants respondent to get pregnant',
  WYNOTUSE == "Don't know" | HPPREGQ == "Don't know",
    "Don't know",
  WYNOTUSE == 'Refused' | HPPREGQ == 'Refused',
    'Refused',
  HPPREGQ == '(If volunteered:) no current partner',
    'No current partner',
  WHYNOUSING1 != 'Not Applicable',
    'Other',
  default = 'Slipped through!'
))]


# create boolean variables for each nonuse reason
for (x in c(1:9, 98:99)) {
  set(
    fem_2017_2019_data,
    j = sprintf('nonuse_reason%d', x),
    value = has_level(fem_2017_2019_data$WHYNOUSING1, x) |
      has_level(fem_2017_2019_data$WHYNOUSING2, x) |
      has_level(fem_2017_2019_data$WHYNOUSING3, x) |
      has_level(fem_2017_2019_data$WHYNOUSING4, x) |
      has_level(fem_2017_2019_data$WHYNOUSING5, x)
  )
}


# create boolean variables for each WHYNOTPG1 reason
for (x in c(1:13, 97:99)) {
  set(
    fem_2017_2019_data,
    j = sprintf('impossible_reason%d', x),
    value = has_level(fem_2017_2019_data$WHYNOTPG1, x) |
      has_level(fem_2017_2019_data$WHYNOTPG2, x)
  )
}


# create variable counting number of nonuse reasons
fem_2017_2019_data[
  ,
  nonuse_reason_count := rowSums(.SD),
  .SDcols = patterns('^nonuse_reason')
]


# create categorical variable combining WHYNOUSING1 and MAINNOUSE
fem_2017_2019_data[, nonuse_main_reason := factor(fcase(
  is.na(nonuse_reason_count) | nonuse_reason_count == 0,
    'Not Applicable',
  nonuse_reason_count == 1,
    as.character(WHYNOUSING1),
  nonuse_reason_count > 1,
    as.character(MAINNOUSE),
  default = 'Slipped through'
))]


# create survey design object
fem_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = fem_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```

```{r, dependson='fem-process-data'}
fem_2017_2019_data[
  CURRPREG == 5 &
    had_sex_this_month &
    !is.element(RSTRSTAT, 1:2) &
    !sterilized &
    CURRMETH1 == 1,
  table(WYNOTUSE)
]

fem_2017_2019_data[, tablena(WYNOTUSE)]
```

It appears the universe for `WYNOTUSE` is once again not as expected from the codebook.


These are the rows for which there is a `WYNOTUSE` value, but aren't in the supposed universe for `WYNOTUSE`:

```{r, dependson='fem-process-data'}
fem_2017_2019_data[
  WYNOTUSE != 'Not Applicable',
  .(
    CASEID,
    CMINTVW,
    CURRPREG,
    MONSX,
    CMMONSX,
    had_sex_this_month,
    RSTRSTAT,
    sterilized,
    CURRMETH1,
    universe = (
      CURRPREG == 5 &
      had_sex_this_month &
      !is.element(RSTRSTAT, 1:2) &
      !sterilized &
      CURRMETH1 == 1
    )
  )
][!(universe)] |> kable()
```

These rows have the correct values for `CURRPREG`, `CURRMETH1`, and whether sterilized based on `METHX` series. However, some of these respondents are have "No" or "Don't know" for whether had intercourse this month based on the `MONSX` series, and some of these respondents are determined to be nonsurgically sterile based on `RSTRSTAT`.

These are the rows that are in the supposed universe for `WYNOTUSE`, but do not have a `WYNOTUSE` value:

```{r, dependson='fem-process-data'}
fem_2017_2019_data[
  WYNOTUSE == 'Not Applicable',
  .(
    CASEID,
    CMINTVW,
    CURRPREG,
    MONSX,
    CMMONSX,
    had_sex_this_month,
    RSTRSTAT,
    sterilized,
    CURRMETH1,
    universe = (
      CURRPREG == 5 &
      had_sex_this_month &
      !is.element(RSTRSTAT, 1:2) &
      !sterilized &
      CURRMETH1 == 1
    )
  )
][(universe)] |> kable()
```

I'm going to proceed with this inconsistency for now, pretending that the universe for `WYNOTUSE` and related variables is as specified in the codebook.

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(WYNOTUSE != 'Not Applicable'),
    fem_2017_2019_svy,
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'who are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
)[, .(domain, title, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(WYNOTUSE),
    subset(
      fem_2017_2019_svy,
      WYNOTUSE != 'Not Applicable'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by whether trying to become pregnant',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(WYNOTUSE),
    subset(
      fem_2017_2019_svy,
      WYNOTUSE != 'Not Applicable'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by whether trying to become pregnant',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(HPPREGQ),
    subset(
      fem_2017_2019_svy,
      WYNOTUSE != 'Yes' & WYNOTUSE != 'Not Applicable'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who are not trying to become pregnant, not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by whether partner wants them to become pregnant',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(HPPREGQ),
    subset(
      fem_2017_2019_svy,
      WYNOTUSE != 'Yes' & WYNOTUSE != 'Not Applicable'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who are not trying to become pregnant, not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by whether partner wants them to become pregnant',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(nonuse_status),
    fem_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by contraceptive nonuse status',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-load-data')}
fem_2017_2019$Data[
  !is.na(WYNOTUSE) & WYNOTUSE %in% c(5, 8, 9) &
    !is.na(HPPREGQ) & HPPREGQ %in% c(5, 8, 9),
  tablena(WHYNOUSING1)
]

fem_2017_2019$Data[
  !is.na(WHYNOUSING1),
  tablena(
    !is.na(WYNOTUSE) & WYNOTUSE %in% c(5, 8, 9) &
    !is.na(HPPREGQ) & HPPREGQ %in% c(5, 8, 9)
  )
]
```

```{r, dependson='fem-load-data'}
dcast(
  fem_2017_2019_formats[
    format_name %in% c(
      'WHYNOUSNG',
      'MAINNOUSE'
    )
  ],
  factor_value ~ format_name,
  value.var = 'factor_label'
) |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:9, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~nonuse_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          nonuse_status == 'Other'
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'WHYNOUSNG', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by reason for not using contraception, with multiple mentions counted multiple times',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:9, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~nonuse_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          nonuse_status == 'Other'
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'WHYNOUSNG', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by reason for not using contraception, with multiple mentions counted multiple times',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(WHYNOUSING1),
    subset(
      fem_2017_2019_svy,
      nonuse_status == 'Other' &
        nonuse_reason_count == 1
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who only one reason for not using contraception, who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by reason for not using contraception',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(WHYNOUSING1),
    subset(
      fem_2017_2019_svy,
      nonuse_status == 'Other' &
        nonuse_reason_count == 1
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who have only one reason for not using contraception, who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by reason for not using contraception',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  nonuse_reason_count > 1,
  tablena(MAINNOUSE)
]
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  MAINNOUSE != 'Not Applicable',
  tablena(nonuse_reason_count)
]
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(MAINNOUSE),
    subset(
      fem_2017_2019_svy,
      nonuse_status == 'Other' &
        nonuse_reason_count > 1
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who have more than one reason for not using contraception, who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by main reason for not using contraception',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(MAINNOUSE),
    subset(
      fem_2017_2019_svy,
      nonuse_status == 'Other' &
        nonuse_reason_count > 1
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who have more than one reason for not using contraception, who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by main reason for not using contraception',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(nonuse_main_reason),
    subset(
      fem_2017_2019_svy,
      nonuse_status == 'Other'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by main or only reason for not using contraception',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(nonuse_main_reason),
    subset(
      fem_2017_2019_svy,
      nonuse_status == 'Other' &
        nonuse_reason_count > 1
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females who are not trying to become pregnant, whose partners do not want them become pregnant, are not currently pregnant, had intercourse in the month of interview, are neither sterilized nor sterile, and did not use a contraceptive in month of interview',
  title = 'by main or only reason for not using contraception',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


# Reasons Respondent Does Not Believe She Can Pregnant

## Universe Check of `WHYNOTPG1`

```{r, dependson=c('fem-process-data'), include=FALSE}
fem_2017_2019$Data[
  fem_2017_2019_data[, (nonuse_reason2)],
  tablena(WHYNOTPG1)
] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  (nonuse_reason2),
  tablena(WHYNOTPG1)
] |> kable()
```

## Format Encoding Issue

```{r, dependson=c('fem-process-data')}
fem_2017_2019_formats[format_name == 'WHYNOTPGF'] |> kable()
```

There are 6 rows that have the value `13` for `WHYNOTPG1`, which is not defined in the `WHYNOTPGF` format. Because of this, they get loaded as `NA` values when `WHYNOTPG1` is factorized.

According to the codebook, `13` should be mapped to "No menstrual cycle or menstrual irregularity."

```{r, dependson=c('fem-process-data')}
fem_2017_2019_formats[factor_label == 'No menstrual cycle or menstrual irregularity'] |> kable()
```

This value appears to be missing entirely.

Manually adding the row to the `fem_2017_2019_formats` in the fem-load-data block fixes this issue.

## Reverse Universe Check of `WHYNOTPG1`

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  WHYNOTPG1 != 'Not Applicable',
  tablena(nonuse_reason2)
] |> kable()
```

The `WHYNOTPG1` is specified if and only if `nonuse_reason2` is true, as expected.

## Estimates

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:13, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~impossible_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          nonuse_reason2
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'WHYNOTPGF', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who mentioned "do not think you can get pregnant" as reason for not using contraception',
  title = 'by reason respondent thinks cannot get pregnant',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:13, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~impossible_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          nonuse_reason2
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'WHYNOTPGF', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who mentioned "do not think you can get pregnant" as reason for not using contraception',
  title = 'by reason respondent thinks cannot get pregnant',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```




# Things to Investigate

* Add table of contents to this EDA file
* Why respondents believe themselves to be sterile (Fem: DE-2 REASIMPP, Male: None)
* Ever tried over lifetime and why quit (Fem: EA, DC; Male: BB)
    * Include tubal ligation and vasectomy reversal
* Collect estimates into coherent report
* Review situations in which universes do not match specifications in codebooks and follow up with NSFG administrators
* Compare own report with published reports
* Replicate analysis in SAS
    * Attempt single-factor categorical data analysis, e.g., whether totals of otherwise not using contraception and currently trying to conceive are different
