---
title: "Reason for Physical Infertility"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(knitr)
library(survey)

i_am('EDA/012_reason_physically_infertile.Rmd')

source(here('R/EDA_helpers.R'))
source(here('R/factor_helpers.R'))
source(here('R/import.R'))
source(here('R/survey_helpers.R'))

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r fem-load-data}
fem_2017_2019 <- load_NSFG_data('2017_2019', 'FemResp')
fem_2017_2019_formats <- fem_2017_2019$Formats
```

```{r fem-process-data, dependson=c('fem-load-data')}
fem_2017_2019_data <- copy(fem_2017_2019$Data[, .(
  CASEID,
  CONSTAT1 = factorize(CONSTAT1, 'CONSTATF', fem_2017_2019_formats),
  LSEXPREG = factorize(LSEXPREG, 'Y1N2RECF', fem_2017_2019_formats),
  METH3M1,
  METH3M2,
  METH3M3,
  P1YRELP = factorize(P1YRELP, 'P1YRELP', fem_2017_2019_formats),
  POSIBLMN = factorize(POSIBLMN, 'Y1N5RDF', fem_2017_2019_formats),
  POSIBLPG = factorize(POSIBLPG, 'Y1N5RDF', fem_2017_2019_formats),
  PSURGSTR = factorize(PSURGSTR, 'Y1N5C', fem_2017_2019_formats),
  REASIMPP = factorize(REASIMPP, 'REASP', fem_2017_2019_formats),
  REASIMPR = factorize(REASIMPR, 'REASR', fem_2017_2019_formats),
  SECU,
  SEST,
  SEX3MO = factorize(SEX3MO, 'Y1N2RECF', fem_2017_2019_formats),
  TUBS = factorize(TUBS, 'Y1N5C', fem_2017_2019_formats),
  WGT2017_2019,
  WYNOLSTP = factorize(WYNOLSTP, 'Y1N5RDF', fem_2017_2019_formats),
  WYNOTUSE = factorize(WYNOTUSE, 'Y1N5RDF', fem_2017_2019_formats)
)])

# remove sterility not from contraceptive procedure from METH3M series
fem_2017_2019_data[
  METH3M3 %in% 20:21,
  `:=`(
    METH3M3 = NA
  )
]
fem_2017_2019_data[
  METH3M2 %in% 20:21,
  `:=`(
    METH3M2 = METH3M3,
    METH3M3 = NA
  )
]
fem_2017_2019_data[
  METH3M1 %in% 20:21,
  `:=`(
    METH3M1 = METH3M2,
    METH3M2 = METH3M3,
    METH3M3 = NA
  )
]

# make METH3M series variables into factors
fem_2017_2019_data[, `:=`(
  METH3M1 = factorize(METH3M1, 'METH3MF', fem_2017_2019_formats),
  METH3M2 = factorize(METH3M2, 'METH3MF', fem_2017_2019_formats),
  METH3M3 = factorize(METH3M3, 'METH3MF', fem_2017_2019_formats)
)]

# create boolean variables for whether each specific method is currently used
for (x in c(1:19, 22:24)) {
  set(
    fem_2017_2019_data,
    j = sprintf('using3m_meth%d', x),
    value = has_level(fem_2017_2019_data$METH3M1, x) |
      has_level(fem_2017_2019_data$METH3M2, x) |
      has_level(fem_2017_2019_data$METH3M3, x)
  )
}

# count any respondents who are sexually active and whose last intercourse was with husband or cohabitating partner who has an active vasectomy as using vasectomy as a contraceptive method
fem_2017_2019_data[, using3m_meth3 := fifelse(
  SEX3MO == 'YES' & PSURGSTR == 'YES' & (
    P1YRELP == 'Current husband or (current husband from whom she is separated)' |
    P1YRELP == 'Current cohabiting partner'
  ),
  TRUE,
  using3m_meth3
)]

# count any respondents with current tubal ligations who are sexually active as currently using tubal ligation as a contraceptive method
fem_2017_2019_data[, using3m_meth4 := fifelse(
  SEX3MO == 'YES' & TUBS == 'YES',
  TRUE,
  using3m_meth4
)]

# create variable counting number of current methods
fem_2017_2019_data[
  ,
  meth3m_count := rowSums(.SD),
  .SDcols = patterns('^using3m_meth')
]

# set master categorization for contraceptive use
fem_2017_2019_data[,
  curr_use := factor(fcase(
    meth3m_count > 0,
      'Used at least one contraceptive method at last intercourse',
    SEX3MO != 'YES',
      'Did not have sexual intercourse with male in past 3 months',
    LSEXPREG == 'YES',
      'Pregnant in month of last intercourse',
    CONSTAT1 == 'Postpartum',
      'Postpartum',
    (WYNOTUSE == 'Yes' | WYNOLSTP == 'Yes'),
      'Currently trying to conceive',
    POSIBLPG == 'No',
      'Physically unable to conceive',
    POSIBLMN == 'No' & (
      P1YRELP == 'Current cohabiting partner' |
      P1YRELP == 'Current husband or (current husband from whom she is separated)'
    ),
      'Last intercourse was with partner who is physically unable to conceive',
    default = 'Otherwise did not use contraception at last intercourse'
  ))
]

# create survey design object
fem_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = fem_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```


# Universe of `POSIBLPG`

```{r, dependson='fem-load-data'}
fem_2017_2019$Data[
  RSURGSTR == 5 & CURRPREG == 5,
  tablena(POSIBLPG)
] |> kable()
```

```{r, dependson='fem-load-data'}
fem_2017_2019$Data[
  !is.na(POSIBLPG),
  table(RSURGSTR, CURRPREG)
]
```

For the most part, the actual universe of `POSIBLPG` matches what is in the codebook. There is one respondent that has a `POSIBLPG` response, but is also surgically sterile according to `RSURGSTR`. This is the one and only exception.


# Universe of `REASIMPR` in Domain of Interest

```{r, dependson='fem-process-data'}
fem_2017_2019$Data[
  fem_2017_2019_data[, curr_use == 'Physically unable to conceive'],
  tablena(REASIMPR)
] |> kable()
```

Since the 'Physically unable to conceive' status is set based on `POSIBLPG`, it should not be surprising that `REASIMPR` is fully specified for the respondents with this status.


# Universe of `REASIMPP` in Domain of Interest

```{r, dependson='fem-process-data'}
fem_2017_2019$Data[
  fem_2017_2019_data[
    , curr_use == 'Last intercourse was with partner who is physically unable to conceive'
  ],
  tablena(REASIMPP)
] |> kable()
```

Likewise, `REASIMPP` is fully specified for the "Last intercourse was with partner who is physically unable to conceive" status.


# Reasons Nonsurgically Sterile

## In Domain of Interest

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPR),
    subset(
      fem_2017_2019_svy,
      curr_use == 'Physically unable to conceive'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females with current contraceptive status of "Physically unable to conceive"',
  title = 'by reason physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPR),
    subset(
      fem_2017_2019_svy,
      curr_use == 'Physically unable to conceive'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females with current contraceptive status of "Physically unable to conceive"',
  title = 'by reason physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

## Generally

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPR),
    subset(
      fem_2017_2019_svy,
      POSIBLPG == 'No'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females physically unable to conceive',
  title = 'by reason physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPR),
    subset(
      fem_2017_2019_svy,
      POSIBLPG == 'No'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females physically unable to conceive',
  title = 'by reason physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


# Reasons Partner Nonsurgically Sterile

## In Domain of Interest

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPP),
    subset(
      fem_2017_2019_svy,
      curr_use == 'Last intercourse was with partner who is physically unable to conceive'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females with current contraceptive status "Last intercourse was with partner who is physically unable to conceive"',
  title = 'by reason partner is physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPP),
    subset(
      fem_2017_2019_svy,
      curr_use == 'Last intercourse was with partner who is physically unable to conceive'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females with current contraceptive status "Last intercourse was with partner who is physically unable to conceive"',
  title = 'by reason partner is physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

## Generally

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPP),
    subset(
      fem_2017_2019_svy,
      POSIBLMN == 'No'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females whose partners are physically unable to conceive',
  title = 'by reason partner is physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REASIMPP),
    subset(
      fem_2017_2019_svy,
      POSIBLMN == 'No'
    ),
    svymean
  ),
  type = 'proportion',
  domain = 'females whose partners are physically unable to conceive',
  title = 'by reason partner is physically unable to conceive',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```
