---
title: "Whether Ever Used or Discontinued Specific Methods"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(knitr)
library(survey)

i_am('EDA/013_ever_used_discontinued.Rmd')

source(here('R/helpers/NSFG/formats.R'))
source(here('R/helpers/NSFG/load.R'))
source(here('R/helpers/data_table.R'))
source(here('R/helpers/deprecated.R'))
source(here('R/helpers/general.R'))

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r fem-load-data}
fem_2017_2019 <- load_NSFG_data('2017_2019', 'FemResp')
```

```{r fem-process-data, dependson=c('fem-load-data')}

# create new format encoding for everused_meth series
fem_everused_format <- copy(fem_2017_2019$Formats[
  (format_name == 'METHHXF' & factor_value %in% c(3:18, 20:21, 25:26, 98:99)) |
    (format_name == 'OTHRMETHF' & factor_value == 24)
])

fem_everused_format[, format_name := 'everused_meth']

fem_2017_2019_formats <- rbindlist(list(
  fem_2017_2019$Formats,
  fem_everused_format,
  data.table(
    format_name = 'everused_meth',
    format_value = c(31:33),
    format_label = c(
      'Copper-bearing IUD (such as Copper-T or ParaGard)',
      'Hormonal IUD (such as Mirena, Skyla, Liletta, or Kyleena)',
      'Other or unknown IUD'
    )
  )
))


# copy working data table
fem_2017_2019_data <- copy(fem_2017_2019$Data[, .(
  CASEID,
  CONDOMR,
  DEPOPROV,
  ESSURE,
  EVERTUBS,
  EVIUDTYP1,
  EVIUDTYP2,
  EVIUDTYP3,
  HADSEX = factorize(HADSEX, 'HADSEX', fem_2017_2019_formats),
  METHSTOP01,
  METHSTOP02,
  METHSTOP03,
  METHSTOP04,
  METHSTOP05,
  METHSTOP06,
  METHSTOP07,
  METHSTOP08,
  METHSTOP09,
  METHSTOP10,
  MORNPILL,
  OTHRMETH01,
  OTHRMETH02,
  OTHRMETH03,
  OTHRMETH04,
  OTHRMETH05,
  OTHRMETH06,
  OTHRMETH07,
  PATCH,
  PILLR,
  REASCOND01,
  REASCOND02,
  REASCOND03,
  REASCOND04,
  REASCOND05,
  REASCOND06,
  REASCOND07,
  REASPILL01,
  REASPILL02,
  REASPILL03,
  REASPILL04,
  REASPILL05,
  REASPILL06,
  REASPILL07,
  REASPILL08,
  REASPILL08,
  REASPILL09,
  REASIUD01,
  REASIUD02,
  REASIUD03,
  REASIUD04,
  REASIUD05,
  REASIUD06,
  REASIUD07,
  REASIUD08,
  REVSTUBL = factorize(REVSTUBL, 'Y1N5RDF', fem_2017_2019_formats),
  RHYTHM,
  RING,
  SECU,
  SEST,
  STOPCOND1,
  STOPCOND2,
  STOPIUD1,
  STOPIUD2,
  STOPIUD3,
  STOPIUD4,
  STOPIUD5,
  STOPPILL1,
  STOPPILL2,
  STOPPILL3,
  STOPPILL4,
  STOPPILL5,
  STOPPILL6,
  TEMPSAFE,
  TYPEIUD_1,
  TYPEIUD_2,
  TYPEIUD_3,
  VASECTMY,
  WGT2017_2019,
  WIDRAWAL
)])


# create boolean variables for whether ever used specific contraceptive methods
fem_2017_2019_data[, `:=`(
  everused_meth3  = na_to_false(PILLR == 1),
  everused_meth4  = na_to_false(CONDOMR == 1),
  everused_meth5  = na_to_false(VASECTMY == 1),
  everused_meth6  = na_to_false(EVERTUBS %in% c(1, 4) | ESSURE == 1),
  everused_meth7  = na_to_false(WIDRAWAL == 1),
  everused_meth8  = na_to_false(DEPOPROV == 1),
  everused_meth10 = na_to_false(RHYTHM == 1),
  everused_meth11 = na_to_false(TEMPSAFE == 1),
  everused_meth20 = na_to_false(MORNPILL == 1),
  everused_meth25 = na_to_false(PATCH == 1),
  everused_meth26 = na_to_false(RING == 1),
  everused_meth33 = na_to_false(
    has_level(EVIUDTYP1, c(3, 9)) |
    has_level(EVIUDTYP2, c(3, 9)) |
    has_level(EVIUDTYP3, c(3, 9))
  )
)]

for (x in c(9, 12:18, 21, 24)) {
  fem_2017_2019_data[
    ,
    (sprintf('everused_meth%d', x)) := na_to_false(
      has_level(OTHRMETH01, x) |
      has_level(OTHRMETH02, x) |
      has_level(OTHRMETH03, x) |
      has_level(OTHRMETH04, x) |
      has_level(OTHRMETH05, x) |
      has_level(OTHRMETH06, x) |
      has_level(OTHRMETH07, x)
    )
  ]
}

for (x in c(1:2)) {
  fem_2017_2019_data[
    ,
    (sprintf('everused_meth%d', 30 + x)) := na_to_false(
      has_level(EVIUDTYP1, x) |
      has_level(EVIUDTYP2, x) |
      has_level(EVIUDTYP3, x)
    )
  ]
}


# create boolean variables for whether specific contraceptive methods were discontinued due to dissatisfaction
for (x in c(3:18, 20:21, 24:26)) {
  fem_2017_2019_data[
    ,
    (sprintf('everquit_meth%d', x)) := na_to_false(
      has_level(METHSTOP01, x) |
      has_level(METHSTOP02, x) |
      has_level(METHSTOP03, x) |
      has_level(METHSTOP04, x) |
      has_level(METHSTOP05, x) |
      has_level(METHSTOP06, x) |
      has_level(METHSTOP07, x) |
      has_level(METHSTOP08, x) |
      has_level(METHSTOP09, x) |
      has_level(METHSTOP10, x)
    )
  ]
}

for (x in c(1:2)) {
  fem_2017_2019_data[
    ,
    (sprintf('everquit_meth%d', 30 + x)) := na_to_false(
      has_level(TYPEIUD_1, x) |
      has_level(TYPEIUD_2, x) |
      has_level(TYPEIUD_3, x)
    )
  ]
}

fem_2017_2019_data[
  ,
  everquit_meth33 := na_to_false(
    has_level(TYPEIUD_1, c(3, 9)) |
    has_level(TYPEIUD_2, c(3, 9)) |
    has_level(TYPEIUD_3, c(3, 9))
)]


# create boolean variables for reason a method was discontinued due to dissatisfaction
for (x in c(1:15, 98:99)) {
  fem_2017_2019_data[
    ,
    (sprintf('pillquit_reason%d', x)) := na_to_false(
      has_level(REASPILL01, x) |
      has_level(REASPILL02, x) |
      has_level(REASPILL03, x) |
      has_level(REASPILL04, x) |
      has_level(REASPILL05, x) |
      has_level(REASPILL06, x) |
      has_level(REASPILL07, x) |
      has_level(REASPILL08, x) |
      has_level(REASPILL09, x)
    )
  ]

  fem_2017_2019_data[
    ,
    (sprintf('condquit_reason%d', x)) := na_to_false(
      has_level(REASCOND01, x) |
      has_level(REASCOND02, x) |
      has_level(REASCOND03, x) |
      has_level(REASCOND04, x) |
      has_level(REASCOND05, x) |
      has_level(REASCOND06, x) |
      has_level(REASCOND07, x)
    )
  ]

  fem_2017_2019_data[
    ,
    (sprintf('iudquit_reason%d', x)) := na_to_false(
      has_level(REASIUD01, x) |
      has_level(REASIUD02, x) |
      has_level(REASIUD03, x) |
      has_level(REASIUD04, x) |
      has_level(REASIUD05, x)
    )
  ]
}


# create boolean variables for reason volunteered by respondents for dissatisfaction with pill
for (x in c(1:23, 97:99)) {
  fem_2017_2019_data[
    ,
    (sprintf('specificpill_reason%d', x)) := na_to_false(
      has_level(STOPPILL1, x) |
      has_level(STOPPILL2, x) |
      has_level(STOPPILL3, x) |
      has_level(STOPPILL4, x) |
      has_level(STOPPILL5, x) |
      has_level(STOPPILL6, x)
    )
  ]
}


# create boolean variables for reason volunteered by respondents for dissatisfaction with condoms
for (x in c(1:9, 97:99)) {
  fem_2017_2019_data[
    ,
    (sprintf('specificcond_reason%d', x)) := na_to_false(
      has_level(STOPCOND1, x) |
      has_level(STOPCOND2, x)
    )
  ]
}


# create boolean variables for reason volunteered by respondents for dissatisfaction with an IUD
for (x in c(1:13, 97:99)) {
  fem_2017_2019_data[
    ,
    (sprintf('specificiud_reason%d', x)) := na_to_false(
      has_level(STOPIUD1, x) |
      has_level(STOPIUD2, x) |
      has_level(STOPIUD3, x) |
      has_level(STOPIUD4, x) |
      has_level(STOPIUD5, x)
    )
  ]
}


# create survey design object
fem_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = fem_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```


# Encoding of `OTHRMETH` Series

```{r, dependson='fem-load-data'}
dcast(
  fem_2017_2019_formats[
    format_name %in% c(
      'OTHRMETHF',
      'METHHXF',
      'METHSTOPF'
    )
  ],
  factor_value ~ format_name,
  value.var = 'factor_label'
) |> kable()
```

The `OTHRMETH` series uses the `OTHERMETHF` format, which is mostly a subset of the `METHHXF` format. The only exception is 24 Lunelle.


# Ever Used Method

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(3:18, 20:21, 24:26, 31:33),
    function(x) {
      svyci(
        as.formula(sprintf('~everused_meth%d', x)),
        fem_2017_2019_svy,
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'everused_meth', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by whether ever used each contraception method',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(3:18, 20:21, 24:26, 31:33),
    function(x) {
      svyci(
        as.formula(sprintf('~everused_meth%d', x)),
        fem_2017_2019_svy,
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'everused_meth', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by whether ever used each contraception method',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


# Females Who Have Ever Had Intercourse with a Male

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(HADSEX),
    fem_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by whether ever had intercourse with a male',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(HADSEX),
    fem_2017_2019_svy,
    svymean
  ),
  type = 'proportion',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by whether ever had intercourse with a male',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


# Had Sex and Ever Used Method

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(3:18, 20:21, 24:26, 31:33),
    function(x) {
      svyci(
        as.formula(sprintf('~everused_meth%d', x)),
        subset(
          fem_2017_2019_svy,
          HADSEX == 'YES, R EVER HAD INTERCOURSE'
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'everused_meth', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have had sexual intercourse with a male',
  title = 'by whether ever used each contraception method',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(3:18, 20:21, 24:26, 31:33),
    function(x) {
      svyci(
        as.formula(sprintf('~everused_meth%d', x)),
        subset(
          fem_2017_2019_svy,
          HADSEX == 'YES, R EVER HAD INTERCOURSE'
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'everused_meth', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have had sexual intercourse with a male',
  title = 'by whether ever used each contraception method',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


# Ever Discontinued Due to Dissatisfication

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(3:18, 21, 24:26, 31:33),
    function(x) {
      svyci(
        as.formula(sprintf('~everquit_meth%d', x)),
        fem_2017_2019_svy,
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'everused_meth', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'by whether ever discontinued a contraception method due to dissatisfaction',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(3:18, 21, 24:26, 31:33),
    function(x) {
      svyci(
        as.formula(sprintf('~everquit_meth%d', x)),
        subset(
          fem_2017_2019_svy,
          fem_2017_2019_data[, get(sprintf('everused_meth%d', x))]
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'everused_meth', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who used the specific method of contraception',
  title = 'by whether ever discontinued due to dissatisfaction',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```



# Tubal Ligation Reversal

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~everused_meth6,
    fem_2017_2019_svy,
    svytotal
  )[level == TRUE],
  type = 'total',
  domain = 'females aged 15-49 living in households in the United States',
  title = 'who have ever had a tubal ligation or similar procedure',
)[, .(domain, title, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~factor(REVSTUBL),
    subset(
      fem_2017_2019_svy,
      everused_meth6
    ),
    svytotal
  ),
  type = 'total',
  domain = 'females who have ever had a tubal ligation or similar procedure',
  title = 'by whether ever had a reversal procedure',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  svyci(
    ~I(REVSTUBL == 'Yes'),
    subset(
      fem_2017_2019_svy,
      everused_meth6
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'females who have ever had a tubal ligation or similar procedure',
  title = 'who ever had a reversal procedure',
)[, .(domain, title, level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```



# Vasectomy Reversal

```{r male-load-data, depeondson='setup'}
male_2017_2019 <- load_NSFG_data('2017_2019', 'Male')
male_2017_2019_formats <- male_2017_2019$Formats
```

```{r male-process-data, dependson=c('male-load-data')}
male_2017_2019_data <- copy(male_2017_2019$Data[, .(
  RVRSVAS = factorize(RVRSVAS, 'Y1N5RDF', male_2017_2019_formats),
  SECU,
  SEST,
  TYPEOPER = factorize(TYPEOPER, 'TYPEOPER', male_2017_2019_formats),
  WGT2017_2019
)])


# create survey design object
male_2017_2019_svy <- svydesign(
  ids = ~SECU,
  strata = ~SEST,
  data = male_2017_2019_data,
  nest = TRUE,
  weights = ~WGT2017_2019
)
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(TYPEOPER == 'Vasectomy'),
    male_2017_2019_svy,
    svytotal
  ),
  type = 'total',
  domain = 'males aged 15-49 living in households in the United States',
  title = 'who have ever had a vasectomy',
)[, .(domain, title, type, level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(TYPEOPER == 'Vasectomy'),
    male_2017_2019_svy,
    svymean
  ),
  type = 'proportion',
  domain = 'males aged 15-49 living in households in the United States',
  title = 'who have ever had a vasectomy',
)[, .(domain, title, type, level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~factor(RVRSVAS),
    subset(
      male_2017_2019_svy,
      TYPEOPER == 'Vasectomy'
    ),
    svytotal
  ),
  type = 'total',
  domain = 'males who have ever had a vasectomy',
  title = 'by whether ever had a reversal procedure',
)[, .(domain, title, type, level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('male-process-data')}
collect_estimate(
  svyci(
    ~I(RVRSVAS == 'Yes'),
    subset(
      male_2017_2019_svy,
      TYPEOPER == 'Vasectomy'
    ),
    svyciprop
  ),
  type = 'proportion',
  domain = 'males who have ever had a vasectomy',
  title = 'who ever had a reversal procedure',
)[, .(domain, title, type, level, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```



# Universe Checks of Reasons for Dissatisfaction Variables

## `REASPILL` Series Universe Check

```{r, dependson=c('fem-process-data')}
fem_2017_2019$Data[
  fem_2017_2019_data$everquit_meth3,
  tablena(REASPILL01)
] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  fem_2017_2019$Data[, !is.na(REASPILL01)],
  tablena(everused_meth3)
] |> kable()
```

The universe of `REASPILL01` is as specified in the codebook.


## `REASMFMT` Encoding

```{r, dependson='fem-load-data'}
fem_2017_2019_formats[format_name == 'REASMFMT'] |> kable()
```


## `STOPPILL` Series Universe Check

```{r, dependson=c('fem-process-data')}
fem_2017_2019$Data[
  fem_2017_2019_data$pillquit_reason3 |
    fem_2017_2019_data$pillquit_reason6 |
    fem_2017_2019_data$pillquit_reason15,
  tablena(STOPPILL1)
] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  fem_2017_2019$Data[, !is.na(STOPPILL1)],
  .(
    .N,
    true = sum(
      pillquit_reason3 |
        pillquit_reason6 |
        pillquit_reason15
    ),
    false = sum(!(
      pillquit_reason3 |
        pillquit_reason6 |
        pillquit_reason15
    ))
  )
] |> kable()
```

The universe of `STOPPILL1` is as specified in the codebook.


## `REASCOND` Universe Check

```{r, dependson=c('fem-process-data')}
fem_2017_2019$Data[
  fem_2017_2019_data$everquit_meth4,
  tablena(REASCOND01)
] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  fem_2017_2019$Data[, !is.na(REASCOND01)],
  tablena(everused_meth4)
] |> kable()
```

The universe of `REASCOND01` is as specified in the codebook.


## `STOPCOND` Series Universe Check

```{r, dependson=c('fem-process-data')}
fem_2017_2019$Data[
  fem_2017_2019_data$condquit_reason3 |
    fem_2017_2019_data$condquit_reason6 |
    fem_2017_2019_data$condquit_reason15,
  tablena(STOPCOND1)
] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  fem_2017_2019$Data[, !is.na(STOPCOND1)],
  .(
    .N,
    true = sum(
      condquit_reason3 |
        condquit_reason6 |
        condquit_reason15
    ),
    false = sum(!(
      condquit_reason3 |
        condquit_reason6 |
        condquit_reason15
    ))
  )
] |> kable()
```

The universe of `STOPCOND1` is as specified in the codebook.


## `REASIUD` Universe Check

```{r, dependson=c('fem-process-data')}
fem_2017_2019$Data[
  fem_2017_2019_data$everquit_meth31 |
    fem_2017_2019_data$everquit_meth32 |
    fem_2017_2019_data$everquit_meth33,
  tablena(REASIUD01)
] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  is.na(REASIUD01) & (
    everquit_meth31 |
      everquit_meth32 |
      everquit_meth33
  ),
  .(
    CASEID,
    everquit_meth31,
    everquit_meth32,
    everquit_meth33,
    METHSTOP01,
    METHSTOP02,
    METHSTOP03,
    METHSTOP04,
    METHSTOP05,
    METHSTOP06,
    METHSTOP07,
    METHSTOP08,
    METHSTOP09,
    METHSTOP10,
    TYPEIUD_1,
    TYPEIUD_2,
    TYPEIUD_3
  )
]
```

There is a logical inconsistency with cases 86949 and 88045. They do not have a value of 19 for any `METHSTOP` variable, but yet have answers for `TYPEIUD_1`. This does not make sense. If a respondent has not indicated they have stopped using an IUD due to dissatisfaction, then they should not be asked what type of IUD they stopped using due to dissatisfaction.

This is getting frustrating. Going forward, I'm not going to try to resolve such inconsistencies. Instead I'm going to create a "No answer provided in public use data files" category to handle these issues, and then explain why these answers exist in the report.

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  fem_2017_2019$Data[, !is.na(REASIUD01)],
  .(
    .N,
    true = sum(
      everquit_meth31 |
        everquit_meth32 |
        everquit_meth33
    ),
    false = sum(!(
      everquit_meth31 |
        everquit_meth32 |
        everquit_meth33
    ))
  )
] |> kable()
```


## `STOPIUD` Series Universe Check

```{r, dependson=c('fem-process-data')}
fem_2017_2019$Data[
  fem_2017_2019_data$iudquit_reason3 |
    fem_2017_2019_data$iudquit_reason6 |
    fem_2017_2019_data$iudquit_reason15,
  tablena(STOPIUD1)
] |> kable()
```

```{r, dependson=c('fem-process-data')}
fem_2017_2019$Data[
  (
    fem_2017_2019_data$iudquit_reason3 |
    fem_2017_2019_data$iudquit_reason6 |
    fem_2017_2019_data$iudquit_reason15
  ) & is.na(STOPIUD1),
  .(
    CASEID,
    REASIUD01,
    REASIUD02,
    REASIUD03,
    REASIUD04,
    REASIUD05,
    STOPIUD1
  )
] |> kable()
```

Case 83425 does not have a `STOPIUD` value, despite 15 being its `REASIUD01` value.

```{r, dependson=c('fem-process-data')}
fem_2017_2019_data[
  fem_2017_2019$Data[, !is.na(STOPIUD1)],
  .(
    .N,
    true = sum(
      iudquit_reason3 |
        iudquit_reason6 |
        iudquit_reason15
    ),
    false = sum(!(
      iudquit_reason3 |
        iudquit_reason6 |
        iudquit_reason15
    ))
  )
] |> kable()
```



# `STOPPILL` Encoding

```{r, dependson='fem-load-data'}
fem_2017_2019_formats[format_name == 'STOPPILLFMT'] |> kable()
```


# Reason(s) for Dissatisfaction leading to Discontinuation of the Pill

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~pillquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth3
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using the pill due to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with the pill',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~pillquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth3
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using the pill due to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with the pill',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:23, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificpill_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          pillquit_reason3 | pillquit_reason6 | pillquit_reason15
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPPILLFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using the pill because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction with the pill',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:23, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificpill_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          pillquit_reason3 | pillquit_reason6 | pillquit_reason15
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPPILLFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using the pill because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction with the pill',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```



# `STOPCONDFMT` Encoding

```{r, dependson='fem-data-load'}
fem_2017_2019_formats[format_name == 'STOPCONDFMT'] |> kable()
```



# Reason(s) for Dissatisfaction leading to Discontinuation of Condoms

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~condquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth4
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using condoms due to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with condoms',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~condquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth4
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using condoms due to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with condoms',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:9, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificcond_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          condquit_reason3 | condquit_reason6 | condquit_reason15
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPCONDFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using the comdoms because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction with condoms',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:9, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificcond_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          condquit_reason3 | condquit_reason6 | condquit_reason15
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPCONDFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using the comdoms because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction with condoms',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```



# `STOPIUDFMT` Encoding

```{r, dependson='fem-data-load'}
fem_2017_2019_formats[format_name == 'STOPIUDFMT'] |> kable()
```


# Reason(s) for Dissatisfaction Leading to Discontinuation of Hormonal IUDs

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~iudquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth32
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using a hormonal IUD to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~iudquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth32
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using a hormonal IUD to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:13, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificiud_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth32 & (iudquit_reason3 | iudquit_reason6 | iudquit_reason15)
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPIUDFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using an hormonal IUD because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:13, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificiud_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth32 & (iudquit_reason3 | iudquit_reason6 | iudquit_reason15)
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPIUDFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using an hormonal IUD because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```


# Reason(s) for Dissatisfaction Leading to Discontinuation of Copper IUDs

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~iudquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth31
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using a copper IUD to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:15, 98:99),
    function(x) {
      svyci(
        as.formula(sprintf('~iudquit_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth31
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'REASMFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using a copper IUD to dissatisfaction',
  title = 'by reason(s) for dissatisfaction with the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:13, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificiud_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth31 & (iudquit_reason3 | iudquit_reason6 | iudquit_reason15)
        ),
        svytotal
      )[
        level == TRUE
      ][
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPIUDFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'total',
  domain = 'females who have discontinued using a copper IUD because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```

```{r, dependson=c('fem-process-data')}
collect_estimate(
  rbindlist(lapply(
    c(1:13, 97:99),
    function(x) {
      svyci(
        as.formula(sprintf('~specificiud_reason%d', x)),
        subset(
          fem_2017_2019_svy,
          everquit_meth31 & (iudquit_reason3 | iudquit_reason6 | iudquit_reason15)
        ),
        svyciprop
      )[
        , .(level = x, estimate, se, `2.5 %`, `97.5 %`)
      ]
    }
  ))[
    order(-estimate),
    .(
      level = factorize(level, 'STOPIUDFMT', fem_2017_2019_formats),
      estimate, se, `2.5 %`, `97.5 %`
    )
  ],
  type = 'proportion',
  domain = 'females who have discontinued using a copper IUD because "too difficult to use," "side effects," or "other" reason',
  title = 'by reason(s) volunteered for dissatisfaction the IUD',
)[, .(level, type, estimate, se, `2.5 %`, `97.5 %`)] |> kable()
```
