---
title: "Contraception Use among Roman Catholics"
author: "Joshua Born"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(knitr)
library(survey)

i_am('EDA/014_roman_catholics.Rmd')

source(here('R/helpers/NSFG/formats.R'))
source(here('R/helpers/NSFG/load.R'))
source(here('R/helpers/NSFG/survey.R'))
source(here('R/helpers/deprecated.R'))
source(here('R/helpers/huxtable.R'))
source(here('R/helpers/survey.R'))

opts_chunk$set(
  cache = TRUE,
  dpi = 72,
  echo = TRUE,
  fig.align = 'center',
  fig.asp = 0.5,
  fig.retina = 2,
  fig.width = 10
)
```

```{r load-data}
fem1719 <- load_NSFG_data('2017_2019', 'FemResp')
fem1719dt <- copy(fem1719$Data)
male1719 <- load_NSFG_data('2017_2019', 'Male')
male1719dt <- copy(male1719$Data)
```

```{r process-data, dependson='load-data'}
fem1719dt <- copy(fem1719$Data[, .(
  AGE_R,
  CASEID,
  SECU,
  SEST,
  WGT2017_2019
)])

male1719dt <- copy(male1719$Data[, .(
  AGE_R,
  CASEID,
  SECU,
  SEST,
  WGT2017_2019
)])
```


# Effective Sample Sizes

```{r effective-sample-sizes, dependson='process-data'}
fem1719formats <- copy(fem1719$Formats)

set_NSFG_variables_as_factors(
  fem1719$Data,
  fem1719dt,
  fem1719formats,
  variables = c(
    HADSEX = 'HADSEX',
    RELIGION = 'RELIGION',
    SEX3MO = 'Y1N2RECF'
  )
)

male1719formats <- copy(male1719$Formats)

set_NSFG_variables_as_factors(
  male1719$Data,
  male1719dt,
  male1719formats,
  variables = c(
    HADSEX = 'HADSEX',
    RELIGION = 'RELIGION',
    SEX3MO = 'SEX3MO'
  )
)

fem1719dt[, .N, by = RELIGION]

male1719dt[, .N, by = RELIGION]

fem1719dt[, `:=`(
  catholic1 = factor(fifelse(
    RELIGION == 'CATHOLIC',
    'Catholic',
    'Not Catholic'
  )),
  catholic2 = factor(fcase(
    RELIGION == 'CATHOLIC', 'Catholic',
    RELIGION == 'NO RELIGION', 'No Religion',
    default = 'Other Religion'
  ))
)]

male1719dt[, `:=`(
  catholic1 = factor(fifelse(
    RELIGION == 'Catholic',
    'Catholic',
    'Not Catholic'
  )),
  catholic2 = factor(fcase(
    RELIGION == 'Catholic', 'Catholic',
    RELIGION == 'No religion', 'No Religion',
    default = 'Other Religion'
  ))
)]

fem1719dt[, .N, by = catholic1]

fem1719dt[, .N, by = catholic2]

male1719dt[, .N, by = catholic1]

male1719dt[, .N, by = catholic2]

fem1719dt[, .N, by = .(catholic2, AGE_R)][order(AGE_R, catholic2)] |>
  dcast(AGE_R ~ catholic2, value.var = 'N') |>
  kable()

male1719dt[, .N, by = .(catholic2, AGE_R)][order(AGE_R, catholic2)] |>
  dcast(AGE_R ~ catholic2, value.var = 'N') |>
  kable()

fem1719dt[, .N, by = HADSEX]

male1719dt[, .N, by = HADSEX]

fem1719dt[HADSEX == 'YES, R EVER HAD INTERCOURSE', .N, by = SEX3MO]

male1719dt[HADSEX == 'YES, R EVER HAD INTERCOURSE', .N, by = SEX3MO]

fem1719dt[
  SEX3MO == 'YES',
  .N,
  by = .(catholic2, AGE_R)
][order(AGE_R, catholic2)] |>
  dcast(AGE_R ~ catholic2, value.var = 'N') |>
  kable()

male1719dt[
  SEX3MO == 'YES, HAD INTERCOURSE',
  .N,
  by = .(catholic2, AGE_R)
][order(AGE_R, catholic2)] |>
  dcast(AGE_R ~ catholic2, value.var = 'N') |>
  kable()

fem1719dt[,
  age_range := cut(
    AGE_R,
    breaks = c(seq(15, 21, 2), seq(25, 45, 5), 51),
    right = FALSE
  )
]

fem1719dt[
  SEX3MO == 'YES',
  .N,
  by = .(catholic2, age_range)
][order(age_range, catholic2)] |>
  dcast(age_range ~ catholic2, value.var = 'N') |>
  kable()

male1719dt[,
  age_range := cut(
    AGE_R,
    breaks = c(seq(15, 21, 2), seq(25, 45, 5), 51),
    right = FALSE
  )
]

male1719dt[
  SEX3MO == 'YES, HAD INTERCOURSE',
  .N,
  by = .(catholic2, age_range)
][order(age_range, catholic2)] |>
  dcast(age_range ~ catholic2, value.var = 'N') |>
  kable()
```


# Population Sizes

```{r population-sizes, dependson='effective-sample-sizes'}
estimate_totals_and_percentages(
  ~catholic2,
  get_NSFG_survey_design(fem1719dt)
) |> style_totals_and_percentages()

fem1719dt[,
  sexually_active := factor(fifelse(
    SEX3MO == 'YES',
    'Sexually Active',
    'Not Sexually Active'
  ))
]

style_and_combine_totals_and_percentages_horizontally(
  'Catholic',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(fem1719dt),
      catholic2 == 'Catholic'
    )
  ),
  'Other Religion',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(fem1719dt),
      catholic2 == 'Other Religion'
    )
  )
)

style_and_combine_totals_and_percentages_horizontally(
  'Catholic',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(fem1719dt),
      catholic2 == 'Catholic'
    )
  ),
  'No Religion',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(fem1719dt),
      catholic2 == 'No Religion'
    )
  )
)

male1719dt[,
  sexually_active := factor(fifelse(
    SEX3MO == 'YES, HAD INTERCOURSE',
    'Sexually Active',
    'Not Sexually Active'
  ))
]

style_and_combine_totals_and_percentages_horizontally(
  'Catholic',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(male1719dt),
      catholic2 == 'Catholic'
    )
  ),
  'Other Religion',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(male1719dt),
      catholic2 == 'Other Religion'
    )
  )
)

style_and_combine_totals_and_percentages_horizontally(
  'Catholic',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(male1719dt),
      catholic2 == 'Catholic'
    )
  ),
  'No Religion',
  estimate_totals_and_percentages(
    ~sexually_active,
    subset(
      get_NSFG_survey_design(male1719dt),
      catholic2 == 'No Religion'
    )
  )
)

fem1719dt[, person := 1]

this <- confint(svyby(
  ~person,
  ~catholic2+age_range,
  get_NSFG_survey_design(fem1719dt),
  svytotal
)) |> as.data.table(keep.rownames = TRUE)

this[, c('religion', 'age_group') := tstrsplit(rn, '.', fixed = TRUE)]

this[, CI := sprintf(
  "(%.3fM, %.3fM)",
  `2.5 %` / 1e6,
  `97.5 %` / 1e6
)]

dcast(
  this,
  age_group ~ religion,
  value.var = 'CI'
) |> kable()

male1719dt[, person := 1]

this <- confint(svyby(
  ~person,
  ~catholic2+age_range,
  get_NSFG_survey_design(male1719dt),
  svytotal
)) |> as.data.table(keep.rownames = TRUE)

this[, c('religion', 'age_group') := tstrsplit(rn, '.', fixed = TRUE)]

this[, CI := sprintf(
  "(%.3fM, %.3fM)",
  `2.5 %` / 1e6,
  `97.5 %` / 1e6
)]

dcast(
  this,
  age_group ~ religion,
  value.var = 'CI'
) |> kable()
```

I'm a little bit worried about the size of the sampling error compared with the estimates for population size. Maybe it would be worth pulling in more observations, such as including the 2015-2017 data set, which is similar to 2017-2019. Because the analysis is just going to use some basic recode variables, it could use the full 2011-2019 data set, even, but then respondents aged 45-49 would have to be discarded.
