---
title: "Contraceptive Use Intervals in 2002 NSFG (Cycle 6)"
author: "Joshua Born"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(knitr)
library(survey)

i_am('EDA/016_contraceptive_use_intervals_2002.Rmd')

source(here('R/NSFG_helpers.R'))
source(here('R/general_helpers.R'))

opts_chunk$set(
  cache = TRUE,
  dpi = 72,
  echo = TRUE,
  fig.align = 'center',
  fig.asp = 0.5,
  fig.retina = 2,
  fig.width = 10
)
```

```{r load-data}
fem2002 <- load_NSFG_data('2002', 'FemResp')
```

```{r process-data, dependson='load-data'}
fem2002dt <- copy(
  fem2002$Data[, lapply(.SD, function(x) {
    if (class(x) == 'logical') {
      as.integer(x)
    } else {
      x
    }
  })]
)

cal2002 <- melt(
  fem2002dt,
  id.vars = c('CASEID'),
  measure.vars = patterns('^MONSX[[:digit:]]+'),
  value.name = 'MONSX'
)[, .(
    CASEID,
    cm_value = as.integer(sub(
      '^MONSX([[:digit:]]+)',
      '\\1',
      variable
    )),
    MONSX = factor(
      MONSX,
      levels = c(1, 5, 7, 8, 9),
      labels = c('YES', 'NO', 'NOT ASCERTAINED', 'REFUSED', "DON'T KNOW")
    )
  )
][
  melt(
    fem2002dt,
    id.vars = c('CASEID'),
    measure.vars = patterns('^NUMMULT[[:digit:]]*', '^SIMSEQ[[:digit:]]*'),
    value.name = c('NUMMULT', 'SIMSEQ'),
    variable.name = 'month_offset',
    variable.factor = FALSE
  )[, .(
    CASEID,
    cm_value = 1188L + nafill(as.integer(month_offset), fill = 1),
    NUMMULT,
    SIMSEQ = factorize_NSFG_variable(
      SIMSEQ,
      'TE_158F',
      formats_table = fem2002$Formats,
      fill_na = FALSE
    )
  )],
  on = c('CASEID', 'cm_value')
][
  dcast(
    melt(
      fem2002dt,
      id.vars = c('CASEID'),
      measure.vars = patterns('^METHHIST[[:digit:]]*'),
      value.name = 'METHHIST'
    )[, .(
      CASEID,
      cm_value =  1188L + as.integer(sub(
        '^METHHIST([[:digit:]]{2,2})[[:digit:]]',
        '\\1',
        variable
      )),
      mention_num = as.integer(sub(
        '^METHHIST[[:digit:]]{2,2}([[:digit:]])',
        '\\1',
        variable
      )),
      METHHIST = factorize_NSFG_variable(
        METHHIST,
        'TE_157F',
        formats_table = fem2002$Formats,
        fill_na = FALSE
      ),
      varname = 'METHHIST'
    )],
    formula = CASEID + cm_value ~ varname + mention_num,
    value.var = c('METHHIST')
  ),
  on = c('CASEID', 'cm_value')
][
  dcast(
    melt(
      fem2002dt,
      id.vars = c('CASEID'),
      measure.vars = patterns('^MTHUSIMX[[:digit:]]*'),
      value.name = 'MTHUSIMX'
    )[, .(
      CASEID,
      cm_value =  1189L + nafill(
        as.integer(sub(
          '^MTHUSIMX([[:digit:]]{0,2})[[:digit:]]{2}$',
          '\\1',
          variable
        )),
        fill = 0
      ),
      mention_num = as.integer(sub(
        '^MTHUSIMX[[:digit:]]{0,2}([[:digit:]]{2})$',
        '\\1',
        variable
      )),
      MTHUSIMX = factorize_NSFG_variable(
        MTHUSIMX,
        'TE_159F',
        formats_table = fem2002$Formats,
        fill_na = FALSE
      ),
      varname = 'MTHUSIMX'
    )],
    formula = CASEID + cm_value ~ varname + mention_num,
    value.var = 'MTHUSIMX'
  ),
  on = c('CASEID', 'cm_value')
]

setkeyv(cal2002, c('CASEID', 'cm_value'))

cal2002[, cm_label := to_century_month(cm_value)]

setcolorder(cal2002, c('CASEID', 'cm_value', 'cm_label'))
```


## Notes

* Look at the dates for sterilizing operations and override contraceptive use calendars with this information.
* Handle non-contraceptive sterility for both
    * respondent and
    * respondent's partner.
* Going forward doing analysis on different years of the NSFG, having the NSFG helper functions set and get variables is going to be unwieldy. I should remove that and just pass in the relevant tables as function parameters.
