---
title: "Retrieval of Census Counts for Post-Stratification of NSFG"
author: "Joshua Born"
date: "2023-06-18"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(censusapi)
library(dplyr)
library(gt)
library(here)
library(knitr)
library(readr)
library(tidyr)

opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  tidy = TRUE
)
```


## API Key

```{r check-api-key}
Sys.setenv(CENSUS_KEY='7c06499a2432f6947e76ee1cf0876fc4ee43ce98')
Sys.getenv("CENSUS_KEY")
```


## Notes

The "Resident" population universe appears to the only universe available for the older time periods. For consistency, I will use the "Resident" population counts for post-stratification, even though the "Resident" population is larger than the actual target population of the NSFG.


### Vintage 2019

Because the 2010-2020 Intercensal Estimates are not available yet, I will have to use the 2019 Vintage estimates for the 2011-2015 and 2015-2019 NSFGs.

```{r get-monthly-codes-2019, dependson="check-api-key"}
months2019 <- getCensus(
  name = 'pep/natmonthly',
  vars = c(
    'MONTHLY',
    'MONTHLY_DESC'
  ),
  region = 'us:*',
  vintage = '2019'
)

months2019 |>
  select(MONTHLY, MONTHLY_DESC) |>
  filter(MONTHLY %in% c('43', '91')) |>
  gt()
```


### 2000-2010 Intercensals

The `pep/int_natmonthly` data set does not have `HISP`, `RACE`, or `SEX` variables and so cannot be used for post-stratification of the NSFG. Instead, I will have to use the `pep/int_charagegroups`.

```{r get-monthly-codes-2000-2010, dependson="check-api-key"}
census2000_2010 <- getCensus(
  name = 'pep/int_charagegroups',
  vars = c(
    'DATE_',
    'DATE_DESC'
  ),
  region = 'us:1',
  vintage = '2000'
)

census2000_2010 |>
  select(DATE_, DATE_DESC) |>
  gt()
```


### 1990-2000 Intercensals

* The `pep/int_charagegroups` data set for the 1990-2000 Intercensals is released _by county_, and can only be downloaded from the API for a specific state. All of these downloads will have to be summed up to national estimates.
* The Cycle 5 NSFG User's Guide does not specify what variables were used for post-stratification. The same variables as other surveys, i.e., age group, race, and Hispanic origin, will be used here. (Sex will necessarily be used because Cycle 4 only surveyed women.)
* While there is a web page defining the encoding of the 2000-2010 Intercensals, there does not appear to be on for 1990-2000. It might be safe to assume that variable encoding did not change between 1990-2000 and 2000-2010, but that leaves the encoding of `RACE_SEX` to be a mystery.

```{r get-pep-int-charagegroups-1990-2000, dependson="check-api-key"}
pep_int_charagegroups1990_2000_state02 <- getCensus(
  name = 'pep/int_charagegroups',
  vars = c(
    'POP', 'YEAR', 'AGEGRP', 'RACE_SEX', 'HISP', 'SUMLEV'
  ),
  region = 'county:*',
  regionin = 'state:02',
  vintage = '1990'
)

pep_int_charagegroups1990_2000_state02 |>
  select(YEAR) |>
  distinct()

pep_int_charagegroups1990_2000_state02 |>
  select(AGEGRP) |>
  distinct()

pep_int_charagegroups1990_2000_state02 |>
  select(RACE_SEX) |>
  distinct()

pep_int_charagegroups1990_2000_state02 |>
  select(HISP) |>
  distinct()

pep_int_charagegroups1990_2000_state02 |>
  select(SUMLEV) |>
  distinct()
```


### 1980-1990 Intercensals

* It does not appear that the 1980-1990 Intercensal data sets are available through the API. "Estimates of the Population of Counties in the United States by Age, Sex, and Race" is available as a downloadable Microsoft Excel file for each year. See file `pe-02-1986.xls`. It will have to be parsed and summed up to the national estimates.
* The file does not specify what the universe used is, so it is likely the "Resident" population universe.
* The Cycle 4 NSFG was post-stratified by "by age, race, marital status, and parity (the number of live births the woman has had) obtained from the U.S. Bureau of the Census." Marital status and parity are not in the PEP. However, sex, age group, and race are available, these will be used in the calibration.


## Downloads and Processing

```{r define-pep-natmonthly-functions}
decompose_race_ethn_by_age <- function(these_data) {
  nonhispanic_allrace <- these_data |>
    filter(
      HISP == 1 & RACE == 0
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  nonhispanic_black <- these_data |>
    filter(
      HISP == 1 & RACE == 8
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  nonhispanic_nonblack <- inner_join(
    nonhispanic_allrace,
    nonhispanic_black |>
      rename('Black_POP' = 'POP'),
    by = join_by(AGE, SEX, SEX_f, UNIVERSE, UNIVERSE_f),
    relationship = 'one-to-one'
  ) |>
    mutate(POP = POP - Black_POP) |> 
    select(-Black_POP)
  
  hispanic <- these_data |>
    filter(
      HISP == 2 & RACE == 0
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  
  bind_rows(
    mutate(nonhispanic_black, race_ethn = 1),
    mutate(hispanic, race_ethn = 2),
    mutate(nonhispanic_nonblack, race_ethn = 3)
  ) |>
    mutate(
      race_ethn_f = factor(
        race_ethn,
        labels = c(
          'Non-Hispanic Black',
          'Hispanic',
          'Non-Hispanic Non-Black'
        )
      )
    )
}

process_pep_natmonthly <- function(these_data) {
  tibble(these_data) |>
    select(-us, -MONTHLY, -MONTHLY_DESC) |>
    mutate(across(!ends_with(c('_DESC', 'UNIVERSE')), ~ as.integer(.x))) |>
    mutate(
      HISP_f = factor(
        HISP,
        levels = 0:2,
        labels = c(
          'Both Hispanic Origins',
          'Non-Hispanic',
          'Hispanic'
        )
      ),
      RACE_f = factor(
        RACE,
        levels = 0:11,
        labels =  c(
          'All races',
          'White alone',
          'Black alone',
          'American Indian and Alaska Native alone',
          'Asian alone',
          'Native Hawaiian and Other Pacific Islander alone',
          'Two or more races',
          'White alone or in combination',
          'Black alone or in combination',
          'American Indian and Alaska Native alone or in combination',
          'Asian alone or in combination',
          'Native Hawaiian and Other Pacific Islander alone or in combination'
        )
      ),
      SEX_f = factor(
        SEX,
        levels = 0:2,
        labels = c(
          'Both Sexes',
          'Male',
          'Female'
        )
      ),
      UNIVERSE_f = factor(
        UNIVERSE,
        levels = c('H', 'N', 'C', 'R', 'P'),
        labels = c(
          "Household population",
          "Civilian noninstitutionalized population",
          "Civilian population",
          "Resident population",
          "Resident population plus Armed Forces overseas"
        )
      )
    ) |>
    decompose_race_ethn_by_age() |>
    filter(AGE >= 15 & AGE <= 49) |>
    mutate(
      age_group = cut(
        AGE,
        breaks = c(14, 19, 24, 29, 34, 39, 44, 49),
        labels = c('15-19', '20-24', '25-29', '30-34', '35-39', '40-44', '45-49')
      ),
      AGEGROUP = case_match(
        age_group,
        '15-19' ~ 4,
        '20-24' ~ 5,
        '25-29' ~ 6,
        '30-34' ~ 7,
        '35-39' ~ 8,
        '40-44' ~ 9,
        '45-49' ~ 10
      ),
      AGEGROUP_f = factor(
        AGEGROUP,
        levels = 0:31,
        labels = c(
          "All ages",
          "Age 0 to 4 years",
          "Age 5 to 9 years",
          "Age 10 to 14 years",
          "Age 15 to 19 years",
          "Age 20 to 24 years",
          "Age 25 to 29 years",
          "Age 30 to 34 years",
          "Age 35 to 39 years",
          "Age 40 to 44 years",
          "Age 45 to 49 years",
          "Age 50 to 54 years",
          "Age 55 to 59 years",
          "Age 60 to 64 years",
          "Age 65 to 69 years",
          "Age 70 to 74 years",
          "Age 75 to 79 years",
          "Age 80 to 84 years",
          "Age 85 years and older",
          "Under 18 years",
          "5 to 13 years",
          "14 to 17 years",
          "18 to 64 years",
          "18 to 24 years",
          "25 to 44 years",
          "45 to 64 years",
          "65 years and over",
          "85 years and over",
          "16 years and over",
          "18 years and over",
          "15 to 44 years",
          "Median age"
        )
      )
    ) |>
    select(-AGE, -age_group) |>
    group_by(UNIVERSE, UNIVERSE_f, SEX, SEX_f, race_ethn, race_ethn_f, AGEGROUP, AGEGROUP_f) |>
    summarize(pop = sum(POP), .groups = 'drop') |>
    filter(SEX != 0)
}
```


### 2015-2019

```{r download-pep-natmonthly-91}
pep_natmonthly_91 <- getCensus(
    name = 'pep/natmonthly',
    vars = c(
      'MONTHLY_DESC',
      'AGE',
      'HISP',
      'RACE',
      'SEX',
      'POP',
      'UNIVERSE'
    ),
    monthly = 91,
    region = 'us:*',
    vintage = '2019'
  )

pep_natmonthly_91_date <- unique(pull(pep_natmonthly_91, MONTHLY_DESC))
```

Date: `r pep_natmonthly_91_date`

```{r process-census-counts-for-2015-2019, dependson=c("define-pep-natmonthly-functions", "download-pep-natmonthly-91")}
census_for_2015_2019 <- process_pep_natmonthly(pep_natmonthly_91) |>
  mutate(survey = '2015-2019', .before = everything())

census_for_2015_2019 |>
  filter(UNIVERSE == 'R') |>
  select(SEX_f, race_ethn_f, AGEGROUP_f, pop) |>
  gt() |>
  fmt_number(decimals = 0)
```


### 2011-2015

```{r download-pep-natmonthly-43}
pep_natmonthly_43 <- getCensus(
    name = 'pep/natmonthly',
    vars = c(
      'MONTHLY_DESC',
      'AGE',
      'HISP',
      'RACE',
      'SEX',
      'POP',
      'UNIVERSE'
    ),
    monthly = 43,
    region = 'us:*',
    vintage = '2019'
  )

pep_natmonthly_43_date <- unique(pull(pep_natmonthly_43, MONTHLY_DESC))
```

Date: `r pep_natmonthly_43_date`

```{r process-census-counts-for-2011-2015, dependson=c("define-pep-natmonthly-functions", "download-pep-natmonthly-43")}
census_for_2011_2015 <- process_pep_natmonthly(pep_natmonthly_43) |>
  mutate(survey = '2011-2015', .before = everything())

census_for_2011_2015 |>
  filter(UNIVERSE == 'R') |>
  select(SEX_f, race_ethn_f, AGEGROUP_f, pop) |>
  gt() |>
  fmt_number(decimals = 0)
```


### 2006-2010

```{r define-pep-int-charagegroups-functions}
decompose_race_ethn_by_agegroup <- function(these_data) {
  nonhispanic_allrace <- these_data |>
    filter(
      HISP == 1 & RACE == 0
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  nonhispanic_black <- these_data |>
    filter(
      HISP == 1 & RACE == 8
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  nonhispanic_nonblack <- inner_join(
    nonhispanic_allrace,
    nonhispanic_black |>
      rename('Black_POP' = 'POP'),
    by = join_by(AGEGROUP, AGEGROUP_f, SEX, SEX_f, UNIVERSE, UNIVERSE_f),
    relationship = 'one-to-one'
  ) |>
    mutate(POP = POP - Black_POP) |> 
    select(-Black_POP)
  
  hispanic <- these_data |>
    filter(
      HISP == 2 & RACE == 0
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  
  bind_rows(
    mutate(nonhispanic_black, race_ethn = 1),
    mutate(hispanic, race_ethn = 2),
    mutate(nonhispanic_nonblack, race_ethn = 3)
  ) |>
    mutate(
      race_ethn_f = factor(
        race_ethn,
        labels = c(
          'Non-Hispanic Black',
          'Hispanic',
          'Non-Hispanic Non-Black'
        )
      )
    )
}

process_pep_int_charagegroups <- function(these_data) {
  tibble(these_data) |>
    select(-us, -DATE_, -DATE_DESC) |>
    mutate(across(!ends_with(c('_DESC', 'UNIVERSE')), ~ as.integer(.x))) |>
    mutate(
      AGEGROUP_f = factor(
        AGEGROUP,
        levels = 0:31,
        labels = c(
          "All ages",
          "Age 0 to 4 years",
          "Age 5 to 9 years",
          "Age 10 to 14 years",
          "Age 15 to 19 years",
          "Age 20 to 24 years",
          "Age 25 to 29 years",
          "Age 30 to 34 years",
          "Age 35 to 39 years",
          "Age 40 to 44 years",
          "Age 45 to 49 years",
          "Age 50 to 54 years",
          "Age 55 to 59 years",
          "Age 60 to 64 years",
          "Age 65 to 69 years",
          "Age 70 to 74 years",
          "Age 75 to 79 years",
          "Age 80 to 84 years",
          "Age 85 years and older",
          "Under 18 years",
          "5 to 13 years",
          "14 to 17 years",
          "18 to 64 years",
          "18 to 24 years",
          "25 to 44 years",
          "45 to 64 years",
          "65 years and over",
          "85 years and over",
          "16 years and over",
          "18 years and over",
          "15 to 44 years",
          "Median age"
        )
      ),
      HISP_f = factor(
        HISP,
        levels = 0:2,
        labels = c(
          'Both Hispanic Origins',
          'Non-Hispanic',
          'Hispanic'
        )
      ),
      RACE_f = factor(
        RACE,
        levels = 0:11,
        labels =  c(
          'All races',
          'White alone',
          'Black alone',
          'American Indian and Alaska Native alone',
          'Asian alone',
          'Native Hawaiian and Other Pacific Islander alone',
          'Two or more races',
          'White alone or in combination',
          'Black alone or in combination',
          'American Indian and Alaska Native alone or in combination',
          'Asian alone or in combination',
          'Native Hawaiian and Other Pacific Islander alone or in combination'
        )
      ),
      SEX_f = factor(
        SEX,
        levels = 0:2,
        labels = c(
          'Both Sexes',
          'Male',
          'Female'
        )
      ),
      UNIVERSE_f = factor(
        UNIVERSE,
        levels = c('H', 'N', 'C', 'R', 'P'),
        labels = c(
          "Household population",
          "Civilian noninstitutionalized population",
          "Civilian population",
          "Resident population",
          "Resident population plus Armed Forces overseas"
        )
      )
    ) |>
    decompose_race_ethn_by_agegroup() |>
    group_by(UNIVERSE, UNIVERSE_f, SEX, SEX_f, race_ethn, race_ethn_f, AGEGROUP, AGEGROUP_f) |>
    summarize(pop = sum(POP), .groups = 'drop') |>
    filter(SEX != 0 & AGEGROUP %in% 4:9)
}
```


```{r download-pep-int-charagegroups-2000-2010}
pep_int_charagegroups_2000_2010 <- getCensus(
  name = 'pep/int_charagegroups',
  vars = c(
    'UNIVERSE',
    'DATE_',
    'DATE_DESC',
    'AGEGROUP',
    'HISP',
    'POP',
    'RACE',
    'SEX'
  ),
  region = 'us:1',
  vintage = '2000'
)

pep_int_charagegroups_10_date <- pep_int_charagegroups_2000_2010 |>
  filter(DATE_ == 10) |>
  pull(DATE_DESC) |>
  unique()

pep_int_charagegroups_4_date <- pep_int_charagegroups_2000_2010 |>
  filter(DATE_ == 4) |>
  pull(DATE_DESC) |>
  unique()
```

Date: `r pep_int_charagegroups_10_date`

```{r process-census-counts-for-2006-2010, dependson=c("define-pep-int-charagegroups-functions", "download-pep-int-charagegroups-2000-2010")}
census_for_2006_2010 <- pep_int_charagegroups_2000_2010 |>
  filter(DATE_ == 10) |>
  process_pep_int_charagegroups() |>
  mutate(survey = '2006-2010', .before = everything())

census_for_2006_2010 |>
  select(SEX_f, race_ethn_f, AGEGROUP_f, pop) |>
  gt() |>
  fmt_number(decimals = 0)
```


### Cycle 6

Date: `r pep_int_charagegroups_4_date`

```{r process-census-counts-for-cycle-6, dependson=c("define-pep-int-charagegroups-functions", "download-pep-int-charagegroups-2000-2010")}
census_for_cycle_6 <- pep_int_charagegroups_2000_2010 |>
  filter(DATE_ == 4) |>
  process_pep_int_charagegroups() |>
  mutate(survey = 'Cycle 6', .before = everything())

census_for_cycle_6 |>
  select(SEX_f, race_ethn_f, AGEGROUP_f, pop) |>
  gt() |>
  fmt_number(decimals = 0)
```


### Cycle 4

```{r process-census-counts-for-cycle-4}
census_for_cycle_4 <- read_csv(
    here('data/Census/pe-02-1986_processed.csv'),
    show_col_types = FALSE
  ) |>
  select(-YEAR, -FIPS) |>
  group_by(race_sex) |>
  summarize(across(everything(), sum)) |>
  pivot_longer(-race_sex, names_to = 'age_group', values_to = 'pop') |>
  mutate(
    race = if_else(
      grepl('Black', race_sex),
      1,
      2
    ),
    race_f = factor(
      race,
      levels = 1:2,
      labels = c(
        'Black',
        'Non-Black'
      )
    ),
    SEX = case_when(
      grepl('female', race_sex) ~ 2,
      grepl('male', race_sex) ~ 1
    ),
    SEX_f = factor(
      SEX,
      levels = 0:2,
      labels = c(
        'Both Sexes',
        'Male',
        'Female'
      )
    ),
    AGEGROUP = case_match(
      age_group,
      '15_to_19' ~ 4,
      '20_to_24' ~ 5,
      '25_to_29' ~ 6,
      '30_to_34' ~ 7,
      '35_to_39' ~ 8,
      '40_to_44' ~ 9,
      '45_to_49' ~ 10,
      '50_to_54' ~ 11,
      '55_to_59' ~ 12
    ),
    AGEGROUP_f = factor(
      AGEGROUP,
      levels = 0:31,
      labels = c(
        "All ages",
        "Age 0 to 4 years",
        "Age 5 to 9 years",
        "Age 10 to 14 years",
        "Age 15 to 19 years",
        "Age 20 to 24 years",
        "Age 25 to 29 years",
        "Age 30 to 34 years",
        "Age 35 to 39 years",
        "Age 40 to 44 years",
        "Age 45 to 49 years",
        "Age 50 to 54 years",
        "Age 55 to 59 years",
        "Age 60 to 64 years",
        "Age 65 to 69 years",
        "Age 70 to 74 years",
        "Age 75 to 79 years",
        "Age 80 to 84 years",
        "Age 85 years and older",
        "Under 18 years",
        "5 to 13 years",
        "14 to 17 years",
        "18 to 64 years",
        "18 to 24 years",
        "25 to 44 years",
        "45 to 64 years",
        "65 years and over",
        "85 years and over",
        "16 years and over",
        "18 years and over",
        "15 to 44 years",
        "Median age"
      )
    )
  ) |> 
  filter(!is.na(AGEGROUP)) |>
  group_by(SEX, SEX_f, race, race_f, AGEGROUP, AGEGROUP_f) |>
  summarize(pop = sum(pop), .groups = 'drop') |>
  mutate(survey = 'Cycle 4', .before = everything())

census_for_cycle_4 |>
  select(SEX_f, race_f, AGEGROUP_f, pop) |>
  gt() |>
  fmt_number(decimals = 0)
```


## Final, Combined Data Table

```{r combine-census-counts, dependson=c("process-census-counts-for-2015-2019", "process-census-counts-for-2011-2015", "process-census-counts-for-2006-2010", "process-census-counts-for-cycle-6", "process-census-counts-for-cycle-4")}
census_combined <- bind_rows(
    census_for_2015_2019,
    census_for_2011_2015,
    census_for_2006_2010,
    census_for_cycle_6,
    census_for_cycle_4
  ) |>
  mutate(
    survey_f = factor(
      survey,
      ordered = TRUE,
      levels = c(
        '2015-2019',
        '2011-2015',
        '2006-2010',
        'Cycle 6',
        'Cycle 5',
        'Cycle 4'
      )
    ),
    .after = survey
  )

saveRDS(census_combined, here('data/Census/combined_pep.Rds'))
```

