---
title: "Feasibility of Calibration of NSFG to Live Births and Induced Abortions"
author: "Joshua Born"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggsci)
library(gt)
library(gtExtras)
library(haven)
library(here)
library(lubridate)
library(readr)
library(scales)
library(srvyr)
library(tidyr)

knitr::opts_chunk$set(
  cache = TRUE,
  dpi=200,
  echo = FALSE,
  fig.align = 'center',
  fig.height=9.6,
  fig.retina=1.5,
  fig.width=9.6
)
```

```{r load-data, message=FALSE}
guttmacher_raw <- read_csv(
  here('data/Guttmacher/NationalAndStatePregnancy_PublicUse.csv')
)
vital_births <- read_csv(
  here('data/NCHS_-_Births_and_General_Fertility_Rates__United_States.csv'),
  name_repair = 'universal'
)
preg2017_2019 <- read_sas(
  here('data/NSFG/d2017_2019fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2017_2019fempreg.sas7bcat')
)
preg2015_2017 <- read_sas(
  here('data/NSFG/d2015_2017fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2015_2017fempreg.sas7bcat')
)
preg2013_2015 <- read_sas(
  here('data/NSFG/d2013_2015fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2013_2015fempreg.sas7bcat')
)
preg2011_2013 <- read_sas(
  here('data/NSFG/d2011_2013fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2011_2013fempreg.sas7bcat')
)
wgt2011_2019 <- read_sas(
  here('data/NSFG/d2011_2019femwgt.sas7bdat')
)
preg2006_2010 <- read_sas(
  here('data/NSFG/d2006_2010fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2006_2010fempreg.sas7bcat')
)
preg2002_2003 <- read_sas(
  here('data/NSFG/d2002_2003fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2002_2003fempreg.sas7bcat')
)
preg1995_1995 <- read_sas(
  here('data/NSFG/d1995_1995fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d1995_1995fempreg.sas7bcat')
)
preg1985_1987 <- read_sas(
  here('data/NSFG/d1985_1987fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d1985_1987fempreg.sas7bcat')
)
preg1982_1983 <- read_sas(
  here('data/NSFG/d1982_1983fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d1982_1983fempreg.sas7bcat')
)
fem1982_1983 <- read_sas(
  here('data/NSFG/d1982_1983femresp.sas7bdat'),
  catalog_file = here('data/NSFG/d1982_1983femresp.sas7bcat')
)
```


```{r cycle-3-interview-dates, dependson="load-data", include=FALSE}
cm_to_date <- function(x) {
  x_int <- as.integer(x)
  if_else(
    x_int > 9000 | x_int == 0,
    NA,
    ym(paste0(
      floor((x_int - 1) / 12) + 1900,
      factor(
        x_int %% 12,
        levels = 0:11,
        labels = c('Dec', 'Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov')
      )
    ))
  )
}

fem1982_1983 |>
  mutate(interview_date = cm_to_date(fem1982_1983$R7_CM)) |>
  select(interview_date) |>
  group_by(interview_date) |>
  count()
```


```{r calculate-census-counts-by-year, dependson="load-data"}
guttmacher_national <- guttmacher_raw |> 
  filter(state == 'US') |>
  select(-state, -notes) |>
  mutate(abortions_interpolated = year %in%
      c(1983, 1986, 1989, 1990, 1993, 1994, 1997, 1998, 2001, 2002, 2003, 2006,
        2009, 2012, 2015)
  )

census_counts <- bind_rows(
  guttmacher_national |>
    filter(!abortions_interpolated) |> 
    rename(Year = year, Total = abortionstotal) |>
    select(Year, Total) |>
    mutate(Outcome = 'Abortions'),
  vital_births |> 
    filter(Year >= 1973) |>
    rename(Total = Birth.Number) |>
    select(Year, Total) |>
    mutate(Outcome = 'Births')
)
```

```{r plot-census-counts-by-year, dependson="calculate-census-counts-by-year", include=FALSE}
ggplot(
  data = census_counts,
  mapping = aes(
    y = Total,
    x = Year,
    color = Outcome,
    shape = Outcome
  )
) +
  geom_point() +
  scale_y_continuous(
    breaks = \(x) seq(x[1], x[2], 5e5),
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA),
    minor_breaks = \(x) seq(x[1], x[2], 25e4)
  ) +
  scale_x_continuous(
    breaks = \(x) seq(x[1]+1, x[2], 4),
    expand = expansion(add = 1),
    minor_breaks = \(x) seq(x[1]+1, x[2], 2)
  ) + 
  theme_bw() +
  theme(
    legend.position = 'bottom',
    legend.title = element_blank(),
    axis.ticks.y = element_blank()
  )
```

```{r create-survey-objects, dependson="load-data"}
cm_to_year_cycle5 <- function(x) {
  x_int <- as.integer(x)
  ifelse(
    x_int %in% c(NA, 0, 9997, 9998, 9999),
    NA,
    floor((x_int - 1) / 12) + 1900
  )
}

count_abortions_and_births <- function(tbl) {
  mutate(
    tbl,
    Abortions = if_else(OUTCOME == 2, 1, 0),
    Births = if_else(OUTCOME == 1, 1, 0),
  )
}

preg_outcomes2015_2019 <- bind_rows(
  select(preg2015_2017, SECU, SEST, CASEID, DATEND, OUTCOME, AGEPREG),
  select(preg2017_2019, SECU, SEST, CASEID, DATEND, OUTCOME, AGEPREG)
) |>
left_join(
  wgt2011_2019,
  by = join_by(CASEID)
) |>
mutate(
  Year = DATEND
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2015_2019,
  nest = TRUE
)

preg_outcomes2011_2015 <- bind_rows(
  select(preg2011_2013, SECU, SEST, CASEID, DATEND, OUTCOME, AGEPREG) |>
    mutate(
      # Not sure what happened here when the data was exported from SAS, but this adjustment seems to fix it.
      DATEND = as.numeric(DATEND) + 3653
    ), 
  select(preg2013_2015, SECU, SEST, CASEID, DATEND, OUTCOME, AGEPREG),
) |>
mutate(
  AGEPREG = floor(AGEPREG / 100)
) |>
left_join(
  wgt2011_2019,
  by = join_by(CASEID)
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |>
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2011_2015,
  nest = TRUE
)


preg_outcomes2011_2019 <- bind_rows(
  bind_rows(
    select(preg2011_2013, SECU, SEST, CASEID, DATEND, OUTCOME) |>
      mutate(
        # Not sure what happened here when the data was exported from SAS, but this adjustment seems to fix it.
        DATEND = as.numeric(DATEND) + 3653
      ), 
    select(preg2013_2015, SECU, SEST, CASEID, DATEND, OUTCOME),
  ) |> mutate(
    Year = cm_to_year_cycle5(DATEND)
  ),
  bind_rows(
    select(preg2015_2017, SECU, SEST, CASEID, DATEND, OUTCOME),
    select(preg2017_2019, SECU, SEST, CASEID, DATEND, OUTCOME)
  ) |>
    mutate(
      Year = DATEND
    )
) |>
mutate(
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |>
left_join(
  wgt2011_2019,
  by = join_by(CASEID)
) |>
count_abortions_and_births() |>
filter(!is.na(WGT2011_2019)) |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2011_2019,
  nest = TRUE
)


preg_outcomes2011_2013 <- select(
  preg2011_2013, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2011_2013
) |>
mutate(
  DATEND = as.numeric(DATEND) + 3653,
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2011_2013,
  nest = TRUE
)


preg_outcomes2013_2015 <- select(
  preg2013_2015, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2013_2015
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2013_2015,
  nest = TRUE
)


preg_outcomes2015_2017 <- select(
  preg2015_2017, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2015_2017
) |>
mutate(
  Year = DATEND,
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2015_2017,
  nest = TRUE
)


preg_outcomes2017_2019 <- select(
  preg2017_2019, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2017_2019
) |>
mutate(
  Year = DATEND,
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2017_2019,
  nest = TRUE
)


preg_outcomes2006_2010 <- select(
  preg2006_2010, SECU, SEST, CASEID, DATEND, OUTCOME, AGEPREG, WGTQ1Q16
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  )),
  AGEPREG = floor(AGEPREG / 100)
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGTQ1Q16,
  nest = TRUE
)


preg_outcomes2002_2003 <- select(
  preg2002_2003, SECU_P, SEST, CASEID, DATEND, OUTCOME, AGEPREG, FINALWGT
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  )),
  AGEPREG = floor(AGEPREG / 100)
) |>
count_abortions_and_births() |>
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU_P,
  weights = FINALWGT,
  nest = TRUE
)


preg_outcomes1995_1995 <- select(
  preg1995_1995, PANEL, COL_STR, CASEID, DATEND, OUTCOME, AGEPREG, POST_WT
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  )),
  AGEPREG = floor(AGEPREG / 100)
) |>
count_abortions_and_births() |>
as_survey_design(
  strata = COL_STR,
  ids = PANEL,
  weights = POST_WT,
  nest = TRUE
)


cm_to_year_cycle4 <- function(x) {
  x_int <- as.integer(x)
  specials <- c(NA, 0, 99585, 99696, 99999)
  x_modified <- if_else(
    !is.element(x_int, specials) & x_int > 9000,
    x_int - 90000,
    x_int
  )
  ifelse(
    x_modified %in% specials,
    NA,
    floor((x_modified - 1) / 12) + 1900
  )
}

preg_outcomes1985_1987 <- select(
  preg1985_1987, SEQNUM, OUTCOME, AGEPREG, B14MO, W_5, starts_with('REP')
) |>
mutate(
  Year = cm_to_year_cycle4(B14MO),
  W_5 = W_5 / 100, # From Codebook: IMPLIED DECIMAL BETWEEN POSTIONS 370 AND 371
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "MISCARRIAGE" = 3,
    "STILLBIRTH" = 4,
    "CURRENT PREGNANCY" = 5
  )),
  across(
    starts_with('REP'),
    ~ .x / 100 # From Codebook: THERE IS AN IMPLIED DECIMAL BEFORE THE LAST TWO DIGITS OF EACH WEIGHT.
  ),
  AGEPREG = floor(AGEPREG / 100)
) |>
count_abortions_and_births() |>
as_survey_rep(
  variables = c(OUTCOME, AGEPREG, Year, Births, Abortions),
  weights = W_5,
  repweights = starts_with('REP'),
  type = 'BRR'
)


cm_to_year_cycle3 <- function(x) {
  x_int <- as.integer(x)
  specials <- c(NA, 0, 9696, 9797, 9898, 9999)
  x_modified <- if_else(
    !is.element(x_int, specials) & x_int > 9000,
    x_int - 9000,
    x_int
  )
  ifelse(
    x_modified %in% specials,
    NA,
    floor((x_modified - 1) / 12) + 1900
  )
}

preg_outcomes1982_1983 <- left_join(
  select(
    preg1982_1983, CASEID, B22A, OUTCOME
  ),
  select(
    fem1982_1983, CASEID, PSUCODE, PSEUDOST, PAIRNUM, FINALWT
  ),
  by = 'CASEID'
) |>
mutate(
  PAIRNUM = if_else(
    PSEUDOST == 21 & PSUCODE == 62,
    2,
    PAIRNUM
  ),
  FINALWT = FINALWT / 10,
  Year = cm_to_year_cycle3(B22A),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "MISCARRIAGE" = 3,
    "STILLBIRTH" = 4,
    "CURRENT PREGNANCY" = 5
  ))
) |>
count_abortions_and_births() |>
as_survey_design(
  strata = PSEUDOST,
  ids = PAIRNUM,
  weights = FINALWT,
  nest = TRUE
)
```

```{r generate-estimates, dependson=c("create-survey-objects", "calculate-census-counts-by-year")}
estimate_births_and_abortions <- function(census_counts, survey, survey_label, begin, end) {
  these_estimates <- survey |>
    group_by(Year) |> 
    summarize(
      Abortions = survey_total(Abortions, vartype = 'ci'),
      Births = survey_total(Births, vartype = 'ci'),
      AbortionsUnweighted = unweighted(sum(Abortions)),
      BirthsUnweighted = unweighted(sum(Births))
    ) |> 
    filter(!is.na(Year))

  bind_rows(
    these_estimates |>
      select(
        Year, starts_with('Abortions')
      ) |>
      rename(
        Estimate = 'Abortions',
        Estimate_low = 'Abortions_low',
        Estimate_upp = 'Abortions_upp',
        UnweightedCount = 'AbortionsUnweighted'
      ) |>
      mutate(
        Outcome = 'Abortions'
      ),
    these_estimates |>
      select(
        Year, starts_with('Births')
      ) |>
      rename(
        Estimate = 'Births',
        Estimate_low = 'Births_low',
        Estimate_upp = 'Births_upp',
        UnweightedCount = 'BirthsUnweighted'
      ) |>
      mutate(
        Outcome = 'Births'
      )
  ) |>
  mutate(
    Begin = begin,
    End = end,
    SurveyLabel = survey_label
  ) |>
  left_join(
    census_counts,
    by = c('Year', 'Outcome')
  )
}

estimates <- bind_rows(
  estimate_births_and_abortions(census_counts, preg_outcomes2015_2019, '2015-2019', my('September 2015'), my('September 2019')),
  estimate_births_and_abortions(census_counts, preg_outcomes2011_2015, '2011-2015', my('September 2011'), my('September 2015')),
  estimate_births_and_abortions(census_counts, preg_outcomes2011_2019, '2011-2019', my('September 2011'), my('September 2019')),
  estimate_births_and_abortions(census_counts, preg_outcomes2011_2013, '2011-2013', my('September 2011'), my('September 2013')),
  estimate_births_and_abortions(census_counts, preg_outcomes2013_2015, '2013-2015', my('September 2013'), my('September 2015')),
  estimate_births_and_abortions(census_counts, preg_outcomes2015_2017, '2015-2017', my('September 2015'), my('September 2017')),
  estimate_births_and_abortions(census_counts, preg_outcomes2017_2019, '2017-2019', my('September 2017'), my('September 2019')),
  estimate_births_and_abortions(census_counts, preg_outcomes2006_2010, '2006-2010', my('June 2006'),      my('June 2010')),
  estimate_births_and_abortions(census_counts, preg_outcomes2002_2003, 'Cycle 6', my('March 2002'),     my('March 2003')),
  estimate_births_and_abortions(census_counts, preg_outcomes1995_1995, 'Cycle 5', my('January 1995'),   my('October 1995')),
  estimate_births_and_abortions(census_counts, preg_outcomes1985_1987, 'Cycle 4', my('December 1985'),  my('March 1987')),
  estimate_births_and_abortions(census_counts, preg_outcomes1982_1983, 'Cycle 3', my('August 1982'),    my('April 1983'))
)
```

```{r define-plot-function}
plot_CIs_against_census_counts <- function(these_data, survey_label, outcome) {
  if (outcome == 'Births') {
    y_axis_label <- 'Live Births'
    title_string <- paste('Number of live births in the United States per year')
    subtitle_string <- paste('Confidence intervals estimated by the', survey_label, "NSFG (brackets) compared with the NVSS counts (X's)")
  }
  if (outcome == 'Abortions') {
    y_axis_label <- 'Induced Abortions'
    title_string <- paste('Number of induced abortions in the United States per year')
    subtitle_string <- paste('Confidence intervals estimated by the', survey_label, "NSFG (brackets) compared with the Guttmacher APC counts (X's)")
  }

  filter(
      these_data,
      Outcome == outcome,
      SurveyLabel == survey_label
  ) |>
  ggplot() +
  geom_vline(
    aes(
      xintercept = year(Begin)
    ),
    color = 'red',
    linetype = 2
  ) +
  geom_vline(
    aes(
      xintercept = year(End)
    ),
    color = 'red',
    linetype = 2
  ) +
  geom_segment(
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    aes(
      x = Year,
      xend = Year,
      y = Estimate_low,
      yend = Estimate_upp
    ),
    color = 'darkblue',
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Estimate
    ),
    color = 'darkblue',
    fill = 'white',
    shape = 21,
    size = 3,
    stroke = 0.75,
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Total
    ),
    color = 'darkorange',
    shape = 4,
    size = 4,
    stroke = 1,
    na.rm = TRUE
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  scale_x_continuous(
    breaks = \(x) seq(x[1], x[2], 2),
    expand = expansion(add = 1)
  ) +
  labs(
    title = title_string,
    subtitle = subtitle_string,
    y = y_axis_label
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      color = 'black',
      hjust = 0.5,
      vjust = 0.65
    ),
    axis.ticks.y = element_blank(),
    legend.position = 'bottom',
    legend.title = element_blank()
  )
}
```



## Live Births

### Two-year and Eight-Year Data Sets

#### 2011-2019

```{r births-2011-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2011-2019',
  outcome = 'Births'
)
```

While using the combined 2011-2019 data set does result in a slight decrease in estimator variance compared with the 2011-2015 stacked data as reflected in narrower confidence intervals for the number of live births per year, it comes as the expense of a much more bias.


#### 2011-2013

```{r births-2011-2013, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2011-2013',
  outcome = 'Births'
)
```


#### 2013-2015

```{r births-2013-2015, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2013-2015',
  outcome = 'Births'
)
```


#### 2015-2017

```{r births-2015-2017, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2015-2017',
  outcome = 'Births'
)
```


#### 2017-2019

```{r births-2017-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2017-2019',
  outcome = 'Births'
)
```

Using the 2-year data sets of the NSFG data appears to result in more variance of the estimators of live births, and without any decrease in bias of the estimators. Sometimes the 4-year estimates appear to be less biased, and sometimes the 2-year estimates to appear to be less biased.



### Four-Year Data Sets

The four-year data sets of the 2011-2019 NSFG appear to produce the best overall estimators of number of live births per year.


#### 2015-2019

```{r births-2015-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2015-2019',
  outcome = 'Births'
)
```


#### 2011-2015

```{r births-2011-2015, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2011-2015',
  outcome = 'Births'
)
```




#### 2006-2010

```{r births-2006-2010, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2006-2010',
  outcome = 'Births'
)
```


#### 2002-2003 (Cycle 6)

```{r births-2002-2003, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 6',
  outcome = 'Births'
)
```


#### 1995 (Cycle 5)

```{r births-1995, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 5',
  outcome = 'Births'
)
```


#### 1985-1987 (Cycle 4)

```{r births-1985-1987, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 4',
  outcome = 'Births'
)
```


#### 1982-1983 (Cycle 3)

```{r births-1982-1983, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 3',
  outcome = 'Births'
)
```



## Induced Abortions

Less than half of induced abortions are reported on the NSFG.


### 2015-2019

```{r abortions-2015-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2015-2019',
  outcome = 'Abortions'
)
```


### 2011-2015

```{r abortions-2011-2015, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2011-2015',
  outcome = 'Abortions'
)
```


### 2006-2010

```{r abortions-2006-2010, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = '2006-2010',
  outcome = 'Abortions'
)
```


### 2002-2003 (Cycle 6)

```{r abortions-2002-2003, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 6',
  outcome = 'Abortions'
)
```


### 1995 (Cycle 5)

```{r abortions-1995, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 5',
  outcome = 'Abortions'
)
```


### 1985-1987 (Cycle 4)

```{r abortions-1985-1987, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 4',
  outcome = 'Abortions'
)
```


### 1982-1983 (Cycle 3)

```{r abortions-1982-1983, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  survey_label = 'Cycle 3',
  outcome = 'Abortions'
)
```


## Which Surveys' Estimates Cover Census Counts of Live Births

These are years for which the census count of live births is covered by at least one 95% confidence interval for live births based on the NSFG.

```{r list-coverage-by-year, depends="generate-estimates"}
estimates |>
  filter(
    Year <= year(Begin) &
    (year(Begin) < 2011 | !is.element(year(End) - year(Begin), c(2, 8))) &
    Outcome == 'Births' &
    Estimate_low <= Total & Total <= Estimate_upp
  ) |>
  mutate(
    Bias = Total - Estimate,
    CI_Width = Estimate_upp - Estimate_low
  ) |>
  select(Year, Bias, CI_Width, UnweightedCount, SurveyLabel, Begin, End) |>
  arrange(desc(Year)) |>
  gt() |>
  sub_missing(
    missing_text = ''
  ) |>
  fmt_number(
    c('Bias', 'CI_Width'),
    drop_trailing_zeros = TRUE,
    decimals = 0
  ) |>
  fmt_date(
    c('Begin', 'End'),
    date_style = 'yM'
  )
```


## Which Surveys to Use to Estimate Which Years

These are the survey data sets that will be used to estimate each year of pregnancy outcomes.

```{r assign-years}
years_assignment <- tribble(
  ~Year, ~SurveyLabel,
  2014, '2015-2019',
  2013, '2015-2019',
  2012, '2015-2019',
  2011, '2015-2019', # However, 2011-2015 seems better, even while losing October, November, December. Using a partial year sample might be biased in other ways, nonetheless.
  2010, '2011-2015',
  2009, '2011-2015',
  2008, '2011-2015',
  2007, '2011-2015',
  2006, '2011-2015',
  2005, '2006-2010',
  2004, '2006-2010',
  2003, '2006-2010',
  2002, '2006-2010',
  2001, 'Cycle 6',
  2000, 'Cycle 6', # However, 2006-2010 seems better.
  1999, 'Cycle 6', # Again, 2006-2010 seems better.
  1998, 'Cycle 6', # Again, 2006-2010 seems better.
  1997, 'Cycle 6', # Cycle 6 is the only option.
  1996, 'Cycle 6', # Cycle 6 outperforms 2006-2010.
  1995, 'Cycle 6', # Cycle 6 is the only option.
  1994, 'Cycle 5',
  1993, 'Cycle 5',
  1992, 'Cycle 5',
  1991, 'Cycle 5',
  1990, 'Cycle 5', # No estimates cover 1990.
  1989, 'Cycle 5',
  1988, 'Cycle 5',
  1987, 'Cycle 5',
  1986, 'Cycle 5',
  1985, 'Cycle 4',
  1984, 'Cycle 4', # Cycle 4 does not cover in 1984.
  1983, 'Cycle 4', # Cycle 5 outperforms Cycle 4 here.
  1982, 'Cycle 4', # No estimates cover 1982.
  1981, 'Cycle 4',
  1980, 'Cycle 4'
)

years_assignment |>
  rename(Survey = 'SurveyLabel') |>
  gt()
```


### Composite Plot of Live Births

```{r live-births-composite-plot, dependson=c("assign-years", "define-plot-function", "generate-estimates")}
years_assignment |>
  left_join(estimates, by = join_by(Year, SurveyLabel)) |>
  filter(Outcome == 'Births') |>
  mutate(
    Survey = factor(
      SurveyLabel,
      levels = c('Cycle 4', 'Cycle 5', 'Cycle 6', '2006-2010', '2011-2015', '2015-2019'),
      ordered = TRUE
    )
  ) |>
  ggplot() +
  geom_segment(
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    aes(
      x = Year,
      xend = Year,
      y = Estimate_low,
      yend = Estimate_upp,
      color = Survey
    ),
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Estimate,
      color = Survey
    ),
    fill = 'white',
    shape = 21,
    size = 3,
    stroke = 0.75,
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Total
    ),
    shape = 4,
    size = 4,
    stroke = 0.625,
    na.rm = TRUE
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  scale_x_continuous(
    breaks = \(x) seq(x[1], x[2], 2),
    expand = expansion(add = 1)
  ) +
  scale_color_cosmic() +
  guides(colour = guide_legend(nrow = 1)) +
  labs(
    title = 'Number of live births in the United States per year',
    subtitle = "Confidence intervals estimated by the NSFG (brackets) compared with the NVSS counts (X's)",
    y = 'Live Births'
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      color = 'black',
      hjust = 0.5,
      vjust = 0.65
    ),
    axis.ticks.y = element_blank(),
    legend.position = 'top',
    legend.title = element_blank()
  )

```


### Noncoverage of Live Births

These are years for which the NSFG estimated confidence interval for live births does not cover the NVSS census count of live births.

```{r count-CIs-that-cover, dependson=c("years-assignment", "generate-estimates")}
births_estimates <- years_assignment |>
  left_join(
    estimates,
    by = join_by(Year, SurveyLabel)
  ) |>
  filter(Outcome == 'Births') |>
  mutate(
    Covers = Estimate_low <= Total & Total <= Estimate_upp
  )

coverage_percent <- births_estimates |>
  summarize(`Coverage Percent` = mean(Covers)) |>
  pull()

births_estimates |>
  filter(!Covers) |>
  select(-Covers, -Outcome) |>
  gt() |>
  fmt_number(
    -Year,
    drop_trailing_zeros = TRUE,
    decimals = 0
  ) |>
  fmt_date(
    c('Begin', 'End'),
    date_style = 'yM'
  )
```

The proportion of confidence intervals that cover the census counts of live births is `r percent(coverage_percent, accuracy = 0.1)`.


### Composite Plot of Induced Abortions

```{r induced-abortions-composite-plot, dependson=c("assign-years", "define-plot-function", "generate-estimates")}
years_assignment |>
  left_join(estimates, by = join_by(Year, SurveyLabel)) |>
  filter(Outcome == 'Abortions') |>
  mutate(
    Survey = factor(
      SurveyLabel,
      levels = c('Cycle 4', 'Cycle 5', 'Cycle 6', '2006-2010', '2011-2015', '2015-2019'),
      ordered = TRUE
    )
  ) |>
  ggplot() +
  geom_segment(
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    aes(
      x = Year,
      xend = Year,
      y = Estimate_low,
      yend = Estimate_upp,
      color = Survey
    ),
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Estimate,
      color = Survey
    ),
    fill = 'white',
    shape = 21,
    size = 3,
    stroke = 0.75,
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Total
    ),
    shape = 4,
    size = 4,
    stroke = 0.625,
    na.rm = TRUE
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  scale_x_continuous(
    breaks = \(x) seq(x[1], x[2], 2),
    expand = expansion(add = 1)
  ) +
  scale_color_cosmic() +
  guides(colour = guide_legend(nrow = 1)) +
  labs(
    title = 'Number of induced abortions in the United States per year',
    subtitle = "Confidence intervals estimated by the NSFG (brackets) compared with the Guttmacher APC counts (X's)",
    y = 'Induced Abortions'
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      color = 'black',
      hjust = 0.5,
      vjust = 0.65
    ),
    axis.ticks.y = element_blank(),
    legend.position = 'top',
    legend.title = element_blank()
  )

```



## Number of and Weights of Respondents with Induced Abortions

These are the weights that will have a ~3 times increase during calibration.

```{r abortion-outcome-observations, dependson=c("create-survey-objects")}
abortion_outcomes <- bind_rows(
  preg_outcomes2015_2019$variables |>
    select(Year, CASEID, AGEPREG, Abortions, WGT2015_2019) |>
    rename(Weight = 'WGT2015_2019') |>
    mutate(SurveyLabel = '2015-2019', .before = everything()),
  preg_outcomes2011_2015$variables |>
    select(Year, CASEID, AGEPREG, Abortions, WGT2011_2015) |>
    rename(Weight = 'WGT2011_2015') |>
    mutate(SurveyLabel = '2011-2015', .before = everything()),
  preg_outcomes2006_2010$variables |>
    select(Year, CASEID, AGEPREG, Abortions, WGTQ1Q16) |>
    rename(Weight = 'WGTQ1Q16') |>
    mutate(SurveyLabel = '2006-2010', .before = everything()),
  preg_outcomes2002_2003$variables |>
    mutate(CASEID = as.numeric(CASEID)) |>
    select(Year, CASEID, AGEPREG, Abortions, FINALWGT) |>
    rename(Weight = 'FINALWGT') |>
    mutate(SurveyLabel = 'Cycle 6', .before = everything()),
  preg_outcomes1995_1995$variables |>
    select(Year, CASEID, AGEPREG, Abortions, POST_WT) |>
    rename(Weight = 'POST_WT') |>
    mutate(SurveyLabel = 'Cycle 5', .before = everything()),
  preg_outcomes1985_1987$variables |>
    mutate(
      Year = cm_to_year_cycle4(B14MO),
      W_5 = W_5 / 100, # From Codebook: IMPLIED DECIMAL BETWEEN POSTIONS 370 AND 371
      Abortions = if_else(OUTCOME == 2, 1, 0)
    ) |>
    select(
      Year, SEQNUM, AGEPREG, Abortions, W_5
    ) |>
    rename(
      CASEID = 'SEQNUM',
      Weight = 'W_5'
    ) |>
    mutate(SurveyLabel = 'Cycle 4', .before = everything())
) |>
  mutate(
    Survey = factor(
      SurveyLabel,
      levels = c('Cycle 4', 'Cycle 5', 'Cycle 6', '2006-2010', '2011-2015', '2015-2019'),
      ordered = TRUE
    ),
    age_group = cut(
      AGEPREG,
      breaks = c(0, 15, 20, 25, 30, 35, 40, 45, Inf),
      right = FALSE
    )
  )

abortion_outcomes |>
  group_by(Survey, AGEPREG, age_group) |>
  count() |>
  View()

years_assignment |>
  left_join(abortion_outcomes, by = join_by(Year, SurveyLabel)) |>
  filter(Abortions == 1) |>
  group_by(Survey) |>
  summarize(
    NumberRespondents = n_distinct(CASEID),
    MedianWeight = median(Weight),
    MeanWeight = mean(Weight),
    StdDevWeight = sd(Weight),
    MinWeight = min(Weight),
    MaxWeight = max(Weight)
  ) |>
  arrange(desc(Survey)) |>
  gt() |>
  fmt_number(
    ends_with('Weight'),
    decimals = 1
  )

years_assignment |>
  left_join(abortion_outcomes, by = join_by(Year, SurveyLabel)) |>
  filter(Abortions == 1) |>
  group_by(Survey, Year) |>
  summarize(
    NumberRespondents = n_distinct(CASEID),
    MedianWeight = median(Weight),
    MeanWeight = mean(Weight),
    StdDevWeight = sd(Weight),
    MinWeight = min(Weight),
    MaxWeight = max(Weight),
    .groups = 'drop_last'
  ) |>
  arrange(desc(Year)) |>
  gt() |>
  fmt_number(
    ends_with('Weight'),
    decimals = 1
  )

years_assignment |>
  left_join(abortion_outcomes, by = join_by(Year, SurveyLabel)) |>
  filter(Abortions == 1) |>
  group_by(Survey, Year, age_group) |>
  summarize(
    NumberRespondents = n_distinct(CASEID),
    MedianWeight = median(Weight),
    MeanWeight = mean(Weight),
    StdDevWeight = sd(Weight),
    MinWeight = min(Weight),
    MaxWeight = max(Weight),
    .groups = 'drop_last'
  ) |>
  arrange(desc(Year)) |>
  gt() |>
  fmt_number(
    ends_with('Weight'),
    decimals = 1
  )
```


## Number of Induced Abortions in Extremal Age Groups

```{r extremal-age-range-counts, dependson="load-data"}
guttmacher_raw |>
  filter(state == 'US') |>
  mutate(
    abortions1519 = abortions1517 + abortions1819,
    prop_abortionslt15 = abortionslt15 / abortionstotal,
    prop_abortions1519 = abortions1519 / abortionstotal,
    prop_abortions40plus = abortions40plus / abortionstotal
  ) |>
  select(year, abortionslt15, abortions1519, abortions40plus, starts_with('prop_')) |>
  gt() |>
  fmt_percent(starts_with('prop_')) |>
  fmt_number(starts_with('abortions'), decimal = 0)
```
