---
title: "Calibration of 2015-2019 NSFG by Traditional Calibration"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
    css: wide.css
---

```{r setup, include=FALSE}
library(dplyr)
library(gt)
library(haven)
library(here)
library(purrr)
library(readr)
library(srvyr)
library(stringr)
library(tidyr)

knitr::opts_chunk$set(
  cache = TRUE,
  dpi=200,
  echo = TRUE,
  fig.align = 'center',
  fig.height=9.6,
  fig.retina=1.5,
  fig.width=9.6
)
```


## Data Preparation

### Load NSFG Data

```{r load-NSFG-data}
NSFG_2017_2019_preg_raw <- read_sas(
  data_file = here('data/NSFG/d2017_2019fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2017_2019fempreg.sas7bcat')
)
NSFG_2015_2017_preg_raw <- read_sas(
  data_file = here('data/NSFG/d2015_2017fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2015_2017fempreg.sas7bcat')
)
NSFG_2011_2019_wgts <- read_sas(
  data_file = here('data/NSFG/d2011_2019femwgt.sas7bdat')
)
NSFG_2017_2019_fem_raw <- read_sas(
  data_file = here('data/NSFG/d2017_2019femresp.sas7bdat'),
  catalog_file = here('data/NSFG/d2017_2019femresp.sas7bcat')
)
NSFG_2015_2017_fem_raw <- read_sas(
  data_file = here('data/NSFG/d2015_2017femresp.sas7bdat'),
  catalog_file = here('data/NSFG/d2015_2017femresp.sas7bcat')
)
```


### Select NSFG Variables

```{r select-NSFG-variables, dependson="load-NSFG-data"}
NSFG_2015_2019_preg_data <- bind_rows(
  select(
    NSFG_2015_2017_preg_raw,
    SECU, SEST, CASEID,
    DATEND,
    OUTCOME,
    AGEPREG,
    NBRNLV_S,
    FMAROUT5,
    HISPRACE2,
    PRGLNGTH,
    RMAROUT6,
    PARITY,
    POVERTY,
    HIEDUC,
    RELIGION,
    PREGORDR,
    BRNOUT
  ),
  select(
    NSFG_2017_2019_preg_raw,
    SECU, SEST, CASEID,
    DATEND,
    OUTCOME,
    AGEPREG,
    NBRNLV_S,
    FMAROUT5,
    HISPRACE2,
    GEST_OTHR,
    RMAROUT6,
    PARITY,
    POVERTY,
    HIEDUC,
    RELIGION,
    PREGORDR,
    BRNOUT
  )
) |>
  left_join(
    select(NSFG_2011_2019_wgts, CASEID, WGT2015_2019),
    by = join_by(CASEID)
  ) |>
  mutate(
    one = 1,
    OUTCOME = as_factor(labelled(OUTCOME, c(
      "LIVE BIRTH" = 1,
      "INDUCED ABORTION" = 2,
      "STILLBIRTH" = 3,
      "MISCARRIAGE" = 4,
      "ECTOPIC PREGNANCY" = 5,
      "CURRENT PREGNANCY" = 6
    ))),
    age_under_20 = (AGEPREG < 20),
    age_under_15 = (AGEPREG < 15),
    age_15_to_17 = (15 <= AGEPREG & AGEPREG <= 17),
    age_18_to_19 = (18 <= AGEPREG & AGEPREG <= 19),
    age_15_to_19 = (15 <= AGEPREG & AGEPREG <= 19),
    age_20_to_24 = (20 <= AGEPREG & AGEPREG <= 24),
    age_25_to_29 = (25 <= AGEPREG & AGEPREG <= 29),
    age_30_to_34 = (30 <= AGEPREG & AGEPREG <= 34),
    age_35_to_39 = (35 <= AGEPREG & AGEPREG <= 39),
    age_40_and_over = (40 <= AGEPREG),
    age_group = cut(
      AGEPREG,
      breaks = c(0, seq(20, 40, 5), Inf),
      right = FALSE,
      labels = c(
        "Under 20 years",
        "20-24 years",
        "25-29 years",
        "30-34 years",
        "35-39 years",
        "40 years and over"
      )
    ),
    FMAROUT5 = as_factor(labelled(FMAROUT5, c(
      "MARRIED" = 1,
      "DIVORCED" = 2,
      "WINDOWED" = 3,
      "SEPARATED" = 4,
      "NEVER MARRIED" = 5
    ))),
    marital_status = as.factor(if_else(
      FMAROUT5 == 'MARRIED',
      'Married',
      'Unmarried'
    )),
    HISPRACE2 = as_factor(labelled(HISPRACE2, c(
      'Hispanic' = 1,
      'Non-Hispanic White' = 2,
      'Non-Hispanic Black' = 3,
      'Non-Hispanic Other' = 4
    ))),
    gestational_age = case_when(
      GEST_OTHR == 1 | PRGLNGTH %in% 0:13 ~ '1st trimester (13 or less weeks)',
      GEST_OTHR %in% 2:3 | PRGLNGTH %in% 14:95 ~ '2nd or 3rd trimester (14 or more weeks)'
    ),
    relationship_status = case_when(
      RMAROUT6 == 1 ~ 'Married',
      RMAROUT6 == 5 ~ 'Cohabitating, not married',
      RMAROUT6 == 6 ~ 'Never-married, not cohabiting',
      RMAROUT6 %in% 2:4 ~ 'Previously married, not cohabiting'
    ),
    RMAROUT6 = as_factor(labelled(RMAROUT6, c(
      "MARRIED" = 1,
      "DIVORCED" = 2,
      "WINDOWED" = 3,
      "SEPARATED" = 4,
      "COHABITING" = 5,
      "NEVER MARRIED, NOT COHABITING" = 6
    ))),
    birth = if_else(
      OUTCOME == 'LIVE BIRTH',
      1,
      0
    ),
    abortion = if_else(
      OUTCOME == 'INDUCED ABORTION',
      1,
      0
    ),
    poverty_level = cut(
      POVERTY,
      breaks = c(0, 99, 199, Inf),
      include.lowest = TRUE,
      labels = c('<100', '100-199', '>=200')
    ),
    educational_attainment = case_when(
      HIEDUC %in% 5:8 ~ '<high school',
      HIEDUC == 9 ~ 'High school graduate/GED',
      HIEDUC %in% 10:11 ~ 'Some college/associate degree',
      HIEDUC %in% 12:15 ~ 'College graduate',
    ),
    HIEDUC = as_factor(labelled(HIEDUC, c(
      "9TH GRADE OR LESS" = 5,
      "10TH GRADE" = 6,
      "11TH GRADE" = 7,
      "12TH GRADE, NO DIPLOMA (NOR GED)" = 8,
      "HIGH SCHOOL GRADUATE (DIPLOMA OR GED)" = 9,
      "SOME COLLEGE BUT NO DEGREE" = 10,
      "ASSOCIATE DEGREE IN COLLEGE/UNIVERSITY" = 11,
      "BACHELOR'S DEGREE" = 12,
      "MASTER'S DEGREE" = 13,
      "DOCTORATE DEGREE" = 14,
      "PROFESSIONAL DEGREE" = 15
    ))),
    religious_affiliation = case_when(
      RELIGION == 1 ~ 'None',
      RELIGION == 2 ~ 'Roman Catholic',
      RELIGION == 3 ~ 'Protestant',
      RELIGION == 4 ~ 'Other'
    ),
    RELIGION = as_factor(labelled(RELIGION, c(
      "NO RELIGION" = 1,
      "CATHOLIC" = 2,
      "PROTESTANT" = 3,
      "OTHER RELIGIONS" = 4
    ))),
    nativity = case_when(
      BRNOUT == 1 ~ 'Foreign-born',
      BRNOUT == 5 ~ 'U.S.-born',
      !is.element(BRNOUT, c(1, 5)) ~ 'Unknown'
    )
  ) |>
  rename(
    Year = DATEND
  ) |>
  arrange(
    CASEID, PREGORDR
  ) |>
  group_by(
    CASEID
  ) |>
  mutate(
    births = cumsum(birth),
    previous_births = lag(births, default = 0),
    previous_births_topcoded4 = factor(if_else(
      previous_births >= 4,
      '>=4',
      as.character(previous_births)
    )),
    previous_births_topcoded2 = factor(if_else(
      previous_births >= 2,
      '>=2',
      as.character(previous_births)
    )),
    abortions = cumsum(abortion),
    previous_abortions = lag(abortions, default = 0),
    previous_abortions_topcoded3 = factor(if_else(
      previous_abortions >= 3,
      '>=3',
      as.character(previous_abortions)
    )),
    any_previous_abortions = factor(if_else(
      previous_abortions >= 1,
      '>=1',
      'None'
    ))
  ) |>
  ungroup()

NSFG_2015_2019_fem_data <- bind_rows(
  select(NSFG_2017_2019_fem_raw, CASEID, SEST, SECU, AGER, HISPRACE2, ABORTION),
  select(NSFG_2015_2017_fem_raw, CASEID, SEST, SECU, AGER, HISPRACE2, ABORTION)
) |>
  mutate(
    age_group = cut(
      AGER,
      breaks = c(14, 19, 24, 29, 34, 39, 44, 50),
      labels = c('15-19', '20-24', '25-29', '30-34', '35-39', '40-44', '45-49')
    ),
    HISPRACE2 = as_factor(labelled(HISPRACE2, c(
      'Hispanic' = 1,
      'Non-Hispanic White' = 2,
      'Non-Hispanic Black' = 3,
      'Non-Hispanic Other' = 4
    ))),
    count_abortions = coalesce(ABORTION, 0)
  ) |> 
  left_join(
    select(NSFG_2011_2019_wgts, CASEID, WGT2015_2019),
    by = join_by(CASEID)
  )
```


## Estimation Uncalibrated

### Create NSFG Survey Object

```{r create-NSFG-survey-object, dependson="select-NSFG-variables"}
NSFG_2015_2019_preg_srv <- NSFG_2015_2019_preg_data |>
  as_survey_design(
    strata = SEST,
    ids = SECU,
    weights = WGT2015_2019,
    nest = TRUE
  )
```


### Estimate Births Uncalibrated

```{r estimate-births-uncalibrated, dependson="create-NSFG-survey-object"}
NSFG_2015_2019_unc_est_births <- bind_rows(
  NSFG_2015_2019_preg_srv |>
    group_by(Year, OUTCOME, age_group) |>
    summarize(
      Est_Births = survey_total(NBRNLV_S, vartype = 'ci'),
      n = unweighted(sum(one))
    ) |>
    ungroup() |>
    filter(
      Year %in% 2011:2014 & OUTCOME == 'LIVE BIRTH'
    ) |>
    select(-OUTCOME) |>
    rename(Level = age_group) |>
    mutate(Variable = 'Age group', .after = Year),
  NSFG_2015_2019_preg_srv |>
    group_by(Year, OUTCOME, marital_status) |>
    summarize(
      Est_Births = survey_total(NBRNLV_S, vartype = 'ci'),
      n = unweighted(sum(one))
    ) |>
    ungroup() |>
    filter(
      Year %in% 2011:2014 & OUTCOME == 'LIVE BIRTH'
    ) |>
    select(-OUTCOME) |>
    rename(Level = marital_status) |>
    mutate(Variable = 'Marital status', .after = Year),
  NSFG_2015_2019_preg_srv |>
    group_by(Year, OUTCOME, HISPRACE2) |>
    summarize(
      Est_Births = survey_total(NBRNLV_S, vartype = 'ci'),
      n = unweighted(sum(one))
    ) |>
    ungroup() |>
    filter(
      Year %in% 2011:2014 & OUTCOME == 'LIVE BIRTH'
    ) |>
    select(-OUTCOME) |>
    rename(Level = HISPRACE2) |>
    mutate(Variable = 'Race and ethnicity', .after = Year)
)
```


### Estimate Abortions Uncalibrated

```{r estimate-abortions-uncalibrated, dependson="create-NSFG-survey-object"}
estimate_abortions_by_domain <- function(variable, label) {
  external_vector <- c(Level = variable)
  NSFG_2015_2019_preg_srv |>
    group_by(across(all_of(c('Year', 'OUTCOME', variable)))) |>
    summarize(
      n = unweighted(sum(one)),
      Est_Abortions = survey_total(one, vartype = 'ci'),
      .groups = 'drop'
    ) |>
    rename(any_of(external_vector)) |>
    mutate(
      Variable = label,
      .after = Year
    ) |>
    filter(
      Year %in% 2011:2014 & OUTCOME == 'INDUCED ABORTION'
    ) |>
    select(-OUTCOME)
}

NSFG_2015_2019_unc_est_abortions <- bind_rows(
  NSFG_2015_2019_preg_srv |>
    group_by(Year, OUTCOME) |>
    summarize(
      Est_age_under_15 = survey_total(age_under_15, vartype = 'ci'),
      n_age_under_15 = unweighted(sum(age_under_15)),
      Est_age_under_20 = survey_total(age_under_20, vartype = 'ci'),
      n_age_under_20 = unweighted(sum(age_under_20)),
      Est_age_15_to_17 = survey_total(age_15_to_17, vartype = 'ci'),
      n_age_15_to_17 = unweighted(sum(age_15_to_17)),
      Est_age_18_to_19 = survey_total(age_18_to_19, vartype = 'ci'),
      n_age_18_to_19 = unweighted(sum(age_18_to_19)),
      Est_age_15_to_19 = survey_total(age_15_to_19, vartype = 'ci'),
      n_age_15_to_19 = unweighted(sum(age_15_to_19)),
      Est_age_20_to_24 = survey_total(age_20_to_24, vartype = 'ci'),
      n_age_20_to_24 = unweighted(sum(age_20_to_24)),
      Est_age_25_to_29 = survey_total(age_25_to_29, vartype = 'ci'),
      n_age_25_to_29 = unweighted(sum(age_25_to_29)),
      Est_age_30_to_34 = survey_total(age_30_to_34, vartype = 'ci'),
      n_age_30_to_34 = unweighted(sum(age_30_to_34)),
      Est_age_35_to_39 = survey_total(age_35_to_39, vartype = 'ci'),
      n_age_35_to_39 = unweighted(sum(age_35_to_39)),
      Est_age_40_and_over = survey_total(age_40_and_over, vartype = 'ci'),
      n_age_40_and_over = unweighted(sum(age_40_and_over))
    ) |>
    ungroup() |>
    filter(
      Year %in% 2011:2014 & OUTCOME == 'INDUCED ABORTION'
    ) |>
    select(-OUTCOME) |>
    pivot_longer(cols = starts_with(c('Est_', 'n_'))) |>
    mutate(
      Variable = 'Age group',
      Level = sub('^.*_age_', '', name) %>% 
        sub('(_low|_upp)$', '', .) %>%
        sub('_to_', '-', .) %>% 
        gsub('_', ' ', .) %>% 
        sub('([[:digit:]]{2})([^[:digit:]]*$)', '\\1 years\\2', .) %>% 
        str_to_sentence(),
      ColName = case_when(
        grepl('_low', name) ~ 'Est_Abortions_low',
        grepl('_upp', name) ~ 'Est_Abortions_upp',
        grepl('Est_', name) ~ 'Est_Abortions',
        grepl('n_', name) ~ 'n'
      )
    ) |>
    pivot_wider(
      id_cols = c(Year, Variable, Level),
      names_from = ColName
    ),
  estimate_abortions_by_domain('gestational_age', 'Gestational age'),
  estimate_abortions_by_domain('marital_status', 'Relationship status') |>
    filter(Level == 'Unmarried'),
  estimate_abortions_by_domain('relationship_status', 'Relationship status'),
  estimate_abortions_by_domain('previous_births_topcoded4', 'No. of prior births'),
  estimate_abortions_by_domain('previous_births_topcoded2', 'No. of prior births') |>
    filter(Level == '>=2'),
  estimate_abortions_by_domain('poverty_level', 'Family income as % of poverty level'),
  estimate_abortions_by_domain('educational_attainment', 'Educational attainment'),
  estimate_abortions_by_domain('religious_affiliation', 'Religious affiliation'),
  estimate_abortions_by_domain('HISPRACE2', 'Race/ethnicity'),
  estimate_abortions_by_domain('previous_abortions_topcoded3', 'No. of previous induced abortions'),
  estimate_abortions_by_domain('any_previous_abortions', 'No. of previous induced abortions'),
  estimate_abortions_by_domain('nativity', 'Nativity')
)
```


## Calibration

### Set Calibration Variables

```{r set-calibration-variables, dependson="select-NSFG-variables"}
indicators <- model.matrix(~age_group:HISPRACE2-1, data=NSFG_2015_2019_fem_data)

colnames(indicators) <- paste0(
  'ind_',
  gsub(
    '[ :-]',
    '_',
    gsub('_group|HISPRACE2', '', colnames(indicators))
  )
)

NSFG_2015_2019_for_calibration <- bind_cols(
  NSFG_2015_2019_fem_data,
  indicators
) |>
  mutate(
    across(
      starts_with('ind_'),
      \(x) factor(x, levels = 0:1, labels = c('No', 'Yes'))
    )
  ) |>
  left_join(
    NSFG_2015_2019_preg_data |>
      mutate(
        count_abortions_2011_2014 = if_else(
          OUTCOME == 'INDUCED ABORTION' & Year %in% 2011:2014,
          1,
          0
        ),
        count_births_2014 = if_else(
          OUTCOME == 'LIVE BIRTH' & Year == 2014,
          1,
          0
        ),
        count_births_2013 = if_else(
          OUTCOME == 'LIVE BIRTH' & Year == 2013,
          1,
          0
        ),
        count_births_2012 = if_else(
          OUTCOME == 'LIVE BIRTH' & Year == 2012,
          1,
          0
        ),
        count_births_2011 = if_else(
          OUTCOME == 'LIVE BIRTH' & Year == 2011,
          1,
          0
        )
      ) |>
      select(CASEID, starts_with('count_')) |>
      group_by(CASEID) |>
      summarize(across(everything(), sum)),
    by = 'CASEID'
  ) |> 
  mutate(across(starts_with('count_'), \(x) coalesce(x, 0)))

rm(indicators)
```


### Counts of Respondents Per Cell Used for Calibration

```{r count-indicators-for-calibration, dependson="set-calibration-variables"}
NSFG_2015_2019_for_calibration |>
  select(starts_with('ind_')) |>
  map(table) |>
  bind_rows(.id = 'Indicator Variable') |>
  gt()
```

```{r count-counts-for-calibration, dependson="set-calibration-variables"}
NSFG_2015_2019_for_calibration |>
  select(starts_with('count_')) |>
  map(table) |>
  bind_rows(.id = 'Count Variable') |>
  gt()
```


## Test Harness

### Create Births Test Harness

```{r create-births-test-harness, message=FALSE}
test_harness_births <- bind_rows(
  read_tsv(
      here('data/NVSS/2011-2014/age.txt'),
      name_repair = 'universal',
      show_col_types = FALSE
    ) |>
    filter(is.na(Notes)) |>
    mutate(Level = case_when(
      Age.of.Mother.9 %in% c(
        'Under 15 years',
        '15-19 years'
      ) ~ 'Under 20 years',
      Age.of.Mother.9 %in% c(
        '40-44 years',
        '45-49 years',
        '50 years and over'
      ) ~ '40 years and over',
      .default = Age.of.Mother.9
    )) |>
    select(Year, Level, Births) |>
    group_by(Year, Level) |>
    summarize(Births = sum(Births), .groups = 'drop') |>
    mutate(Variable = 'Age group', .after = Year),
  read_tsv(
      here('data/NVSS/2011-2014/marital_status.txt'),
      name_repair = 'universal',
      show_col_types = FALSE
    ) |>
    filter(is.na(Notes)) |>
    rename(Level = Marital.Status) |>
    select(Year, Level, Births) |>
    mutate(Variable = 'Marital status'),
  read_tsv(
      here('data/NVSS/2011-2014/race_ethnicity.txt'),
      name_repair = 'universal',
      show_col_types = FALSE
    ) |>
    rename(
      Hispanic.Origin = Mother.s.Hispanic.Origin,
      Race = Mother.s.Bridged.Race
    ) |>
    filter(
      (is.na(Notes) & Hispanic.Origin != 'Hispanic or Latino') |
        (Notes == 'Total' & Hispanic.Origin == 'Hispanic or Latino')
    ) |>
    select(Notes, Year, Hispanic.Origin, Race, Births) |>
    mutate(
      Level = case_when(
        Hispanic.Origin == 'Hispanic or Latino' ~ 'Hispanic',
        Race == 'White' ~ 'Non-Hispanic White',
        Race == 'Black or African American' ~ 'Non-Hispanic Black',
        .default = 'Non-Hispanic Other'
      )
    ) |>
    group_by(Year, Level) |>
    summarize(Births = sum(Births), .groups = 'drop') |>
    mutate(Variable = 'Race and ethnicity')
) |>
  arrange(-Year, Variable)
```


### Create Abortions Test Harness

```{r load-and-process-guttmacher-APC-data}
guttmacher_APC_national <- read_csv(
    here('data/Guttmacher/NationalAndStatePregnancy_PublicUse.csv'),
    show_col_types = FALSE
  ) |>
  filter(state == 'US') |>
  select(-state, -notes) |>
  mutate(interpolated = year %in%
      c(1983, 1986, 1989, 1990, 1993, 1994, 1997, 1998, 2001, 2002, 2003, 2006,
        2009, 2012, 2015)
  )

guttmacher_APC_2014 <- guttmacher_APC_national |>
  filter(year == 2014) |>
  select(starts_with('abortions'))

guttmacher_APC_2014_total <- pull(guttmacher_APC_2014, abortionstotal)

guttmacher_APC_2012 <- guttmacher_APC_national |>
  filter(year == 2012) |>
  select(starts_with('abortions'))

guttmacher_APC_2012_total <- pull(guttmacher_APC_2012, abortionstotal)

guttmacher_APC_2011 <- guttmacher_APC_national |>
  filter(year == 2011) |>
  select(starts_with('abortions'))

guttmacher_APC_2011_total <- pull(guttmacher_APC_2011, abortionstotal)
```

```{r create-abortions-test-harness, dependson="load-and-process-guttmacher-APC-data"}
guttmacher_APS_2014 <- bind_rows(
  tribble(
    ~name, ~Level, ~PctAPS2014,
    'abortionslt20',   'Under 20 years',     11.9,
    'abortions1517',   '15-17 years',     3.4,
    'abortions1819',   '18-19 years',     8.2,
    'abortions2024',   '20-24 years',     33.6,
    'abortions2529',   '25-29 years',     26.5,
    'abortions3034',   '30-34 years',     15.9,
    'abortions3539',   '35-39 years',     9.1,
    'abortions40plus', '40 years and over',   3.1
  ) |> mutate(
    Variable = 'Age group',
    .before = everything()
  ) |>
    left_join(
      pivot_longer(
        guttmacher_APC_2014,
        cols = everything(),
        values_to = 'TtlAPC2014'
      ),
      by = 'name'
    ) |>
    select(-name),
  tribble(
    ~Level, ~PctAPS2014,
    'Married            ',               14.3,
    'Cohabitating, not married',         31.0,
    'Never-married, not cohabiting',     45.9,
    'Previously married, not cohabiting', 8.8
  ) |>  mutate(
    Variable = 'Relationship status',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctAPS2014,
    'Non-Hispanic White',    38.7,
    'Non-Hispanic Black',    27.6,
    'Hispanic',              24.8,
    'Non-Hispanic Other',    8.8
  ) |>  mutate(
    Variable = 'Race/ethnicity',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctAPS2014,
    'U.S.-born',     83.9,
    'Foreign-born',  16.1
  ) |>  mutate(
    Variable = 'Nativity',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctAPS2014,
    '<high school',                  8.9,
    'High school graduate/GED',      27.0,
    'Some college/associate degree', 40.9,
    'College graduate',              23.1
  ) |>  mutate(
    Variable = 'Educational attainment',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctAPS2014,
    '0',   40.7,
    '1',   26.2,
    '>=2', 33.1
  ) |>  mutate(
    Variable = 'No. of prior births',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctAPS2014,
    '>=1', 44.8
  ) |>  mutate(
    Variable = 'No. of previous induced abortions',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctAPS2014,
    '<100',    49.3,
    '100-199', 25.7,
    '>=200',   25.0
  ) |>  mutate(
    Variable = 'Family income as % of poverty level',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctAPS2014,
    'Mainline Protestant',    17.3,
    'Evangelical Protestant', 12.8,
    'Roman Catholic',         23.7,
    'Other',                  8.2,
    'None',                   38.0
  ) |>  mutate(
    Variable = 'Religious affiliation',
    .before = everything()
  )
) |> mutate(
  Level = str_squish(Level),
  PctAPS2014 = PctAPS2014 / 100,
  TtlAPS2014 = as.integer(round(guttmacher_APC_2014_total * PctAPS2014)),
  .before = 'TtlAPC2014'
)


CDC_2014 <- bind_rows(
  tribble(
    ~name, ~Level, ~PctCDC,
    'abortionslt15',   'Under 15 years',   0.3,
    'abortions1519',   '15-19 years',   10.4,
    'abortions2024',   '20-24 years',   32.2,
    'abortions2529',   '25-29 years',   26.7,
    'abortions3034',   '30-34 years',   17.1,
    'abortions3539',   '35-39 years',   9.7,
    'abortions40plus', '40 years and over', 3.6
  ) |> mutate(
    Variable = 'Age group',
    .before = everything()
  ) |> left_join(
    pivot_longer(
      guttmacher_APC_2014,
      cols = everything(),
      values_to = 'TtlAPC'
    ),
    by = 'name'
  ) |> select(-name),
  tribble(
    ~Level, ~PctCDC,
    '<=8',   64.9,
    '9-13',  26.1,
    '14-15', 3.6,
    '16-17', 2.2,
    '18-20', 1.9,
    '>=21',  1.3
  ) |>  mutate(
    Variable = 'Gestational age',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    'Non-Hispanic White',    38.0,
    'Non-Hispanic Black',    36.0,
    'Non-Hispanic Other',    7.7,
    'Hispanic', 18.3,
  ) |>  mutate(
    Variable = 'Race/ethnicity',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    'Married',    14.5,
    'Unmarried',  85.5
  ) |>  mutate(
    Variable = 'Relationship status',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    '0',   40.4,
    '1',   26.0,
    '2',   19.7,
    '3',   8.7,
    '>=4', 5.1
  ) |>  mutate(
    Variable = 'No. of prior births',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    '0',   55.1,
    '1',   24.7,
    '2',   11.6,
    '>=3', 8.6
  ) |>  mutate(
    Variable = 'No. of previous induced abortions',
    .before = everything()
  )
) |> mutate(
  Level = str_squish(Level),
  PctCDC = PctCDC / 100,
  TtlCDC = as.integer(round(guttmacher_APC_2014_total * PctCDC)),
  .before = 'TtlAPC'
)


CDC_2012 <- bind_rows(
  tribble(
    ~name, ~Level, ~PctCDC,
    'abortionslt15',   'Under 15 years',   0.4,
    'abortions1519',   '15-19 years',   12.2,
    'abortions2024',   '20-24 years',   32.8,
    'abortions2529',   '25-29 years',   25.4,
    'abortions3034',   '30-34 years',   16.4,
    'abortions3539',   '35-39 years',   9.1,
    'abortions40plus', '40 years and over', 3.7
  ) |> mutate(
    Variable = 'Age group',
    .before = everything()
  ) |> left_join(
    pivot_longer(
      guttmacher_APC_2012,
      cols = everything(),
      values_to = 'TtlAPC'
    ),
    by = 'name'
  ) |> select(-name),
  tribble(
    ~Level, ~PctCDC,
    '<=8',   65.8,
    '9-13',  25.6,
    '14-15', 3.5,
    '16-17', 1.8,
    '18-20', 1.9,
    '>=21',  1.3
  ) |>  mutate(
    Variable = 'Gestational age',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    'Non-Hispanic White',    37.6,
    'Non-Hispanic Black',    36.7,
    'Non-Hispanic Other',    7.0,
    'Hispanic', 18.7,
  ) |>  mutate(
    Variable = 'Race/ethnicity',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    'Married',    14.7,
    'Unmarried',  85.3
  ) |>  mutate(
    Variable = 'Relationship status',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    '0',   40.3,
    '1',   26.2,
    '2',   19.6,
    '3',   8.7,
    '>=4', 5.3
  ) |>  mutate(
    Variable = 'No. of prior births',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    '0',   55.7,
    '1',   24.6,
    '2',   11.0,
    '>=3', 8.6
  ) |>  mutate(
    Variable = 'No. of previous induced abortions',
    .before = everything()
  )
) |> mutate(
  Level = str_squish(Level),
  PctCDC = PctCDC / 100,
  TtlCDC = as.integer(round(guttmacher_APC_2012_total * PctCDC)),
  .before = 'TtlAPC'
)


CDC_2011 <- bind_rows(
  tribble(
    ~name, ~Level, ~PctCDC,
    'abortionslt15',   'Under 15 years',   0.4,
    'abortions1519',   '15-19 years',   13.5,
    'abortions2024',   '20-24 years',   32.9,
    'abortions2529',   '25-29 years',   24.9,
    'abortions3034',   '30-34 years',   15.8,
    'abortions3539',   '35-39 years',   8.9,
    'abortions40plus', '40 years and over', 3.6
  ) |> mutate(
    Variable = 'Age group',
    .before = everything()
  ) |> left_join(
    pivot_longer(
      guttmacher_APC_2011,
      cols = everything(),
      values_to = 'TtlAPC'
    ),
    by = 'name'
  ) |> select(-name),
  tribble(
    ~Level, ~PctCDC,
    '<=8',   64.5,
    '9-13',  26.9,
    '14-15', 3.5,
    '16-17', 1.9,
    '18-20', 1.9,
    '>=21',  1.4
  ) |>  mutate(
    Variable = 'Gestational age',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    'Non-Hispanic White',    37.2,
    'Non-Hispanic Black',    36.2,
    'Non-Hispanic Other',    7.0,
    'Hispanic', 19.7,
  ) |>  mutate(
    Variable = 'Race/ethnicity',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    'Married',    14.5,
    'Unmarried',  85.5
  ) |>  mutate(
    Variable = 'Relationship status',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    '0',   39.9,
    '1',   26.5,
    '2',   19.6,
    '3',   8.7,
    '>=4', 5.2
  ) |>  mutate(
    Variable = 'No. of prior births',
    .before = everything()
  ),
  tribble(
    ~Level, ~PctCDC,
    '0',   53.7,
    '1',   25.5,
    '2',   11.6,
    '>=3', 9.3
  ) |>  mutate(
    Variable = 'No. of previous induced abortions',
    .before = everything()
  )
) |> mutate(
  Level = str_squish(Level),
  PctCDC = PctCDC / 100,
  TtlCDC = as.integer(round(guttmacher_APC_2011_total * PctCDC)),
  .before = 'TtlAPC'
)


test_harness_abortions <- bind_rows(
  mutate(CDC_2014, Year = 2014, Source = 'CDC', .before = everything()),
  mutate(CDC_2012, Year = 2012, Source = 'CDC', .before = everything()),
  mutate(CDC_2011, Year = 2011, Source = 'CDC', .before = everything())
) |>
  select(-TtlAPC, -PctCDC) |>
  rename(Abortions = TtlCDC) |>
  bind_rows(
    guttmacher_APS_2014 |>
      select(-TtlAPC2014, -PctAPS2014) |>
      mutate(Year = 2014, Source = 'APS', .before = everything()) |>
      rename(Abortions = TtlAPS2014)
  ) |>
  arrange(-Year, Variable, Source)
```


### Compare Births Estimates with Test Harness

```{r compare-births-estimates-with-test-harness, dependson=c("create-births-test-harness", "estimate-births-uncalibrated")}
full_join(
  test_harness_births,
  NSFG_2015_2019_unc_est_births,
  by = c('Year', 'Variable', 'Level')
) |>
  mutate(
    Bias = Est_Births - Births,
    Relative_Bias = Bias / Births,
    `Covers?` = Births <= Est_Births_upp & Births >= Est_Births_low
  ) |>
  gt() |>
  fmt_number(-Year, decimal = 0) |>
  fmt_percent(Relative_Bias, decimal = 1)
```


### Compare Abortions Estimates with Test Harness

```{r compare-abortions-estimates-with-test-harness, dependson=c("create-abortions-test-harness", "estimate-births-uncalibrated")}
test_harness_abortions |>
  mutate(
    Level = case_match(
      Level,
      '<=8' ~ '1st trimester (13 or less weeks)',
      '9-13' ~ '1st trimester (13 or less weeks)',
      '14-15' ~ '2nd or 3rd trimester (14 or more weeks)',
      '16-17' ~ '2nd or 3rd trimester (14 or more weeks)',
      '18-20' ~ '2nd or 3rd trimester (14 or more weeks)',
      '>=21' ~ '2nd or 3rd trimester (14 or more weeks)',
      'Mainline Protestant' ~ 'Protestant',
      'Evangelical Protestant' ~ 'Protestant',
      .default = Level
    )
  ) |> 
  group_by(Year, Source, Variable, Level) |>
  summarize(across(everything(), sum), .groups = 'drop') |>
  arrange(desc(Year)) |>
  left_join(
    NSFG_2015_2019_unc_est_abortions,
    by = c('Year', 'Variable', 'Level')
  ) |>
  mutate(
    Bias = Est_Abortions - Abortions,
    Relative_Bias = Bias / Abortions,
    `Covers?` = Abortions <= Est_Abortions_upp & Abortions >= Est_Abortions_low
  ) |>
  gt() |>
  fmt_number(-Year, decimal = 0) |>
  fmt_percent(Relative_Bias, decimal = 1)
```
