---
title: "NSFG Estimates Compared with Census Counts"
author: "Joshua Born"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(gt)
library(gtExtras)
library(haven)
library(here)
library(lubridate)
library(readr)
library(srvyr)
library(tidyr)

knitr::opts_chunk$set(
  cache = TRUE,
  dpi=200,
  echo = FALSE,
  fig.align = 'center',
  fig.height=9.6,
  fig.retina=1.5,
  fig.width=9.6
)
```

```{r, load-data, message=FALSE}
guttmacher_raw <- read_csv(
  here('data/guttmacher/NationalAndStatePregnancy_PublicUse.csv')
)
vital_births <- read_csv(
  here('data/NCHS_-_Births_and_General_Fertility_Rates__United_States.csv'),
  name_repair = 'universal'
)
preg2017_2019 <- read_sas(
  here('data/d2017_2019fempreg.sas7bdat'),
  catalog_file = here('data/d2017_2019fempreg.sas7bcat')
)
preg2015_2017 <- read_sas(
  here('data/d2015_2017fempreg.sas7bdat'),
  catalog_file = here('data/d2015_2017fempreg.sas7bcat')
)
preg2013_2015 <- read_sas(
  here('data/d2013_2015fempreg.sas7bdat'),
  catalog_file = here('data/d2013_2015fempreg.sas7bcat')
)
preg2011_2013 <- read_sas(
  here('data/d2011_2013fempreg.sas7bdat'),
  catalog_file = here('data/d2011_2013fempreg.sas7bcat')
)
wgt2011_2019 <- read_sas(
  here('data/d2011_2019femwgt.sas7bdat')
)
preg2006_2010 <- read_sas(
  here('data/d2006_2010fempreg.sas7bdat'),
  catalog_file = here('data/d2006_2010fempreg.sas7bcat')
)
preg2002_2003 <- read_sas(
  here('data/d2002_2003fempreg.sas7bdat'),
  catalog_file = here('data/d2002_2003fempreg.sas7bcat')
)
preg1995_1995 <- read_sas(
  here('data/d1995_1995fempreg.sas7bdat'),
  catalog_file = here('data/d1995_1995fempreg.sas7bcat')
)
preg1985_1987 <- read_sas(
  here('data/d1985_1987fempreg.sas7bdat'),
  catalog_file = here('data/d1985_1987fempreg.sas7bcat')
)
preg1982_1983 <- read_sas(
  here('data/d1982_1983fempreg.sas7bdat'),
  catalog_file = here('data/d1982_1983fempreg.sas7bcat')
)
fem1982_1983 <- read_sas(
  here('data/d1982_1983femresp.sas7bdat'),
  catalog_file = here('data/d1982_1983femresp.sas7bcat')
)
```


```{r cycle-3-interview-dates, dependson="load-data"}
cm_to_date <- function(x) {
  x_int <- as.integer(x)
  if_else(
    x_int > 9000 | x_int == 0,
    NA,
    ym(paste0(
      floor((x_int - 1) / 12) + 1900,
      factor(
        x_int %% 12,
        levels = 0:11,
        labels = c('Dec', 'Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov')
      )
    ))
  )
}

fem1982_1983 |>
  mutate(interview_date = cm_to_date(fem1982_1983$R7_CM)) |>
  select(interview_date) |>
  group_by(interview_date) |>
  count()
```


```{r, calculate-census-counts-by-year, dependson="load-data"}
guttmacher_national <- guttmacher_raw |> 
  filter(state == 'US') |>
  select(-state, -notes) |>
  mutate(abortions_interpolated = year %in%
      c(1983, 1986, 1989, 1990, 1993, 1994, 1997, 1998, 2001, 2002, 2003, 2006,
        2009, 2012, 2015)
  )

census_counts <- bind_rows(
  guttmacher_national |>
    filter(!abortions_interpolated) |> 
    rename(Year = year, Total = abortionstotal) |>
    select(Year, Total) |>
    mutate(Outcome = 'Abortions'),
  vital_births |> 
    filter(Year >= 1973) |>
    rename(Total = Birth.Number) |>
    select(Year, Total) |>
    mutate(Outcome = 'Births')
)
```

```{r plot-census-counts-by-year, dependson="calculate-census-counts-by-year", include=FALSE}
ggplot(
  data = census_counts,
  mapping = aes(
    y = Total,
    x = Year,
    color = Outcome,
    shape = Outcome
  )
) +
  geom_point() +
  scale_y_continuous(
    breaks = \(x) seq(x[1], x[2], 5e5),
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA),
    minor_breaks = \(x) seq(x[1], x[2], 25e4)
  ) +
  scale_x_continuous(
    breaks = \(x) seq(x[1]+1, x[2], 4),
    expand = expansion(add = 1),
    minor_breaks = \(x) seq(x[1]+1, x[2], 2)
  ) + 
  theme_bw() +
  theme(
    legend.position = 'bottom',
    legend.title = element_blank(),
    axis.ticks.y = element_blank()
  )
```

```{r print-table-of-census-counts-by-year, dependson="calculate-census-counts-by-year", include=FALSE}
gt(census_counts) |>
  sub_missing(
    missing_text = ''
  ) |> 
  fmt_number(
    starts_with('Total'),
    drop_trailing_zeros = TRUE
  )
```

```{r define-plot-function}
plot_CIs_against_census_counts <- function(these_data, begin, end, outcome) {
  if (outcome == 'Births') {
    y_axis_label <- 'Live Births'
    title_string <- paste('Number of live births in the United States per year')
    subtitle_string <- paste('CIs estimated by the', paste(begin, end, sep = '-'), "NSFG (brackets) compared with the NVSS counts (X's)")
  }
  if (outcome == 'Abortions') {
    y_axis_label <- 'Induced Abortions'
    title_string <- paste('Number of induced abortions in the United States per year')
    subtitle_string <- paste('CIs estimated by the', paste(begin, end, sep = '-'), "NSFG (brackets) compared with the Guttmacher APC counts (X's)")
  }
  ggplot(
    data = filter(
      these_data,
      Begin == begin,
      End == end,
      Outcome == outcome
    )
  ) +
  geom_vline(
    aes(
      xintercept = Begin
    ),
    color = 'red',
    linetype = 2
  ) +
  geom_vline(
    aes(
      xintercept = End
    ),
    color = 'red',
    linetype = 2
  ) +
  geom_segment(
    arrow = arrow(angle = 90, ends = 'both', length = unit(0.075, 'inches')),
    aes(
      x = Year,
      xend = Year,
      y = Estimate_low,
      yend = Estimate_upp
    ),
    color = 'darkblue',
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Estimate
    ),
    color = 'darkblue',
    fill = 'white',
    shape = 21,
    size = 3,
    stroke = 0.75,
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = Year,
      y = Total
    ),
    color = 'darkorange',
    shape = 4,
    size = 4,
    stroke = 1,
    na.rm = TRUE
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  scale_x_continuous(
    breaks = \(x) seq(x[1], x[2], 2),
    expand = expansion(add = 1)
  ) +
  labs(
    title = title_string,
    subtitle = subtitle_string,
    y = y_axis_label
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      color = 'black',
      hjust = 0.5,
      vjust = 0.65
    ),
    axis.ticks.y = element_blank(),
    legend.position = 'bottom',
    legend.title = element_blank()
  )
}
```

```{r create-survey-objects, dependson="load-data"}
cm_to_year_cycle5 <- function(x) {
  x_int <- as.integer(x)
  ifelse(
    x_int %in% c(NA, 0, 9997, 9998, 9999),
    NA,
    floor((x_int - 1) / 12) + 1900
  )
}

count_abortions_and_births <- function(tbl) {
  mutate(
    tbl,
    Abortions = if_else(OUTCOME == 2, 1, 0),
    Births = if_else(OUTCOME == 1, 1, 0),
  )
}


preg_outcomes2015_2019 <- bind_rows(
  select(preg2015_2017, SECU, SEST, CASEID, DATEND, OUTCOME),
  select(preg2017_2019, SECU, SEST, CASEID, DATEND, OUTCOME)
) |>
left_join(
  wgt2011_2019,
  by = join_by(CASEID)
) |>
mutate(
  Year = DATEND
) |> 
count_abortions_and_births() |> 
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2015_2019,
  nest = TRUE
)


preg_outcomes2011_2015 <- bind_rows(
  select(preg2011_2013, SECU, SEST, CASEID, DATEND, OUTCOME) |>
    mutate(
      # Not sure what happened here when the data was exported from SAS, but this adjustment seems to fix it.
      DATEND = as.numeric(DATEND) + 3653
    ), 
  select(preg2013_2015, SECU, SEST, CASEID, DATEND, OUTCOME),
) |>
left_join(
  wgt2011_2019,
  by = join_by(CASEID)
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2011_2015,
  nest = TRUE
)


preg_outcomes2011_2019 <- bind_rows(
  bind_rows(
    select(preg2011_2013, SECU, SEST, CASEID, DATEND, OUTCOME) |>
      mutate(
        # Not sure what happened here when the data was exported from SAS, but this adjustment seems to fix it.
        DATEND = as.numeric(DATEND) + 3653
      ), 
    select(preg2013_2015, SECU, SEST, CASEID, DATEND, OUTCOME),
  ) |> mutate(
    Year = cm_to_year_cycle5(DATEND)
  ),
  bind_rows(
    select(preg2015_2017, SECU, SEST, CASEID, DATEND, OUTCOME),
    select(preg2017_2019, SECU, SEST, CASEID, DATEND, OUTCOME)
  ) |>
    mutate(
      Year = DATEND
    )
) |>
mutate(
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |>
left_join(
  wgt2011_2019,
  by = join_by(CASEID)
) |>
count_abortions_and_births() |>
filter(!is.na(WGT2011_2019)) |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2011_2019,
  nest = TRUE
)


preg_outcomes2011_2013 <- select(
  preg2011_2013, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2011_2013
) |>
mutate(
  DATEND = as.numeric(DATEND) + 3653,
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2011_2013,
  nest = TRUE
)


preg_outcomes2013_2015 <- select(
  preg2013_2015, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2013_2015
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2013_2015,
  nest = TRUE
)


preg_outcomes2015_2017 <- select(
  preg2015_2017, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2015_2017
) |>
mutate(
  Year = DATEND,
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2015_2017,
  nest = TRUE
)


preg_outcomes2017_2019 <- select(
  preg2017_2019, SECU, SEST, CASEID, DATEND, OUTCOME, WGT2017_2019
) |>
mutate(
  Year = DATEND,
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGT2017_2019,
  nest = TRUE
)


preg_outcomes2006_2010 <- select(
  preg2006_2010, SECU, SEST, CASEID, DATEND, OUTCOME, WGTQ1Q16
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |> 
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU,
  weights = WGTQ1Q16,
  nest = TRUE
)


preg_outcomes2002_2003 <- select(
  preg2002_2003, SECU_P, SEST, CASEID, DATEND, OUTCOME, FINALWGT
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  ))
) |>
count_abortions_and_births() |>
count_abortions_and_births() |>
as_survey_design(
  strata = SEST,
  ids = SECU_P,
  weights = FINALWGT,
  nest = TRUE
)


preg_outcomes1995_1995 <- select(
  preg1995_1995, PANEL, COL_STR, CASEID, DATEND, OUTCOME, POST_WT
) |>
mutate(
  Year = cm_to_year_cycle5(DATEND),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "STILLBIRTH" = 3,
    "MISCARRIAGE" = 4,
    "ECTOPIC PREGNANCY" = 5,
    "CURRENT PREGNANCY" = 6
  )),
) |>
count_abortions_and_births() |>
as_survey_design(
  strata = COL_STR,
  ids = PANEL,
  weights = POST_WT,
  nest = TRUE
)


cm_to_year_cycle4 <- function(x) {
  x_int <- as.integer(x)
  specials <- c(NA, 0, 99585, 99696, 99999)
  x_modified <- if_else(
    !is.element(x_int, specials) & x_int > 9000,
    x_int - 90000,
    x_int
  )
  ifelse(
    x_modified %in% specials,
    NA,
    floor((x_modified - 1) / 12) + 1900
  )
}

preg_outcomes1985_1987 <- select(
  preg1985_1987, SEQNUM, OUTCOME, B14MO, W_5, starts_with('REP')
) |>
mutate(
  Year = cm_to_year_cycle4(B14MO),
  W_5 = W_5 / 100, # From Codebook: IMPLIED DECIMAL BETWEEN POSTIONS 370 AND 371
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "MISCARRIAGE" = 3,
    "STILLBIRTH" = 4,
    "CURRENT PREGNANCY" = 5
  )),
  across(
    starts_with('REP'),
    ~ .x / 100 # From Codebook: THERE IS AN IMPLIED DECIMAL BEFORE THE LAST TWO DIGITS OF EACH WEIGHT.
  )
) |>
count_abortions_and_births() |>
as_survey_rep(
  variables = c(OUTCOME, Year, Births, Abortions),
  weights = W_5,
  repweights = starts_with('REP'),
  type = 'BRR'
)


cm_to_year_cycle3 <- function(x) {
  x_int <- as.integer(x)
  specials <- c(NA, 0, 9696, 9797, 9898, 9999)
  x_modified <- if_else(
    !is.element(x_int, specials) & x_int > 9000,
    x_int - 9000,
    x_int
  )
  ifelse(
    x_modified %in% specials,
    NA,
    floor((x_modified - 1) / 12) + 1900
  )
}

preg_outcomes1982_1983 <- left_join(
  select(
    preg1982_1983, CASEID, B22A, OUTCOME
  ),
  select(
    fem1982_1983, CASEID, PSUCODE, PSEUDOST, PAIRNUM, FINALWT
  ),
  by = 'CASEID'
) |>
mutate(
  PAIRNUM = if_else(
    PSEUDOST == 21 & PSUCODE == 62,
    2,
    PAIRNUM
  ),
  FINALWT = FINALWT / 10,
  Year = cm_to_year_cycle3(B22A),
  OUTCOME = labelled(OUTCOME, c(
    "LIVE BIRTH" = 1,
    "INDUCED ABORTION" = 2,
    "MISCARRIAGE" = 3,
    "STILLBIRTH" = 4,
    "CURRENT PREGNANCY" = 5
  ))
) |>
count_abortions_and_births() |>
as_survey_design(
  strata = PSEUDOST,
  ids = PAIRNUM,
  weights = FINALWT,
  nest = TRUE
)
```

```{r generate-estimates, dependson=c("create-survey-objects", "calculate-census-counts-by-year")}
estimate_births_and_abortions <- function(survey, begin, end, census_counts) {
  bind_rows(
    survey |>
      group_by(Year) |>
      summarize(
        Estimate = survey_total(Abortions, vartype = 'ci')
      ) |> 
      filter(!is.na(Year)) |>
      mutate(
        Begin = begin ,
        End = end ,
        Outcome = 'Abortions'
      ),
    survey |>
      group_by(Year) |> 
      summarize(
        Estimate = survey_total(Births, vartype = 'ci')
      ) |> 
      filter(!is.na(Year)) |>
      mutate(
        Begin = begin,
        End = end,
        Outcome = 'Births'
      )
  ) |>
    left_join(
      census_counts,
      by = c('Year', 'Outcome')
    )
}

estimates <- bind_rows(
  estimate_births_and_abortions(preg_outcomes2015_2019, 2015, 2019, census_counts),
  estimate_births_and_abortions(preg_outcomes2011_2015, 2011, 2015, census_counts),
  estimate_births_and_abortions(preg_outcomes2011_2019, 2011, 2019, census_counts),
  estimate_births_and_abortions(preg_outcomes2011_2013, 2011, 2013, census_counts),
  estimate_births_and_abortions(preg_outcomes2013_2015, 2013, 2015, census_counts),
  estimate_births_and_abortions(preg_outcomes2015_2017, 2015, 2017, census_counts),
  estimate_births_and_abortions(preg_outcomes2017_2019, 2017, 2019, census_counts),
  estimate_births_and_abortions(preg_outcomes2006_2010, 2006, 2010, census_counts),
  estimate_births_and_abortions(preg_outcomes2002_2003, 2002, 2003, census_counts),
  estimate_births_and_abortions(preg_outcomes1995_1995, 1995, 1995, census_counts),
  estimate_births_and_abortions(preg_outcomes1985_1987, 1985, 1987, census_counts),
  estimate_births_and_abortions(preg_outcomes1982_1983, 1982, 1983, census_counts)
)
```


## Years with Coverage

These are years for which the census count of live births is covered by at least one 95% confidence interval for live births based on the NSFG.

```{r quantity-coverage, depends="generate-estimates"}
estimates |>
  filter(
    !(Begin == 2011 & End == 2019) &
      !(Begin == 2011 & End == 2013) &
      !(Begin == 2013 & End == 2015) &
      !(Begin == 2015 & End == 2017) &
      !(Begin == 2017 & End == 2019)
  ) |>
  filter(
    Outcome == 'Births' &
    Estimate_low <= Total & Total <= Estimate_upp
  ) |>
  mutate(
    Bias = Total - Estimate,
    CI_Width = Estimate_upp - Estimate_low
  ) |>
  select(Year, Bias, CI_Width, Total, Estimate, Estimate_low, Estimate_upp, Begin, End) |>
  arrange(Year) |>
  gt() |>
  sub_missing(
    missing_text = ''
  ) |>
  fmt_number(
    starts_with(c('Estimate', 'Total', 'Bias', 'CI_Width')),
    drop_trailing_zeros = TRUE,
    decimals = 0
  )
```



## Live Births


### 2015-2019

```{r births-2015-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2015,
  end = 2019,
  outcome = 'Births'
)
```


### 2011-2015

```{r births-2011-2015, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2011,
  end = 2015,
  outcome = 'Births'
)
```


### 2011-2019

```{r births-2011-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2011,
  end = 2019,
  outcome = 'Births'
)
```

While using the combined 2011-2019 data set does result in a slight decrease in estimator variance compared with the 2011-2015 stacked data as reflected in narrower confidence intervals for the number of live births per year, it comes as the expense of a much more bias.


### 2011-2013

```{r births-2011-2013, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2011,
  end = 2013,
  outcome = 'Births'
)
```


### 2013-2015

```{r births-2013-2015, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2013,
  end = 2015,
  outcome = 'Births'
)
```


### 2015-2017

```{r births-2015-2017, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2015,
  end = 2017,
  outcome = 'Births'
)
```


### 2017-2019

```{r births-2017-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2017,
  end = 2019,
  outcome = 'Births'
)
```

Using the 2-year data sets of the NSFG data appears to result in more variance of the estimators of live births, and without any decrease in bias of the estimators. Sometimes the 4-year estimates appear to be less biased, and sometimes the 2-year estimates to appear to be less biased.

The 4-year data sets of the 2011-2019 NSFG appear to produce the best overall estimators of number of live births per year.


### 2006-2010

```{r births-2006-2010, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2006,
  end = 2010,
  outcome = 'Births'
)
```


### 2002-2003 (Cycle 6)

```{r births-2002-2003, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2002,
  end = 2003,
  outcome = 'Births'
)
```


### 1995 (Cycle 5)

```{r births-1995, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 1995,
  end = 1995,
  outcome = 'Births'
)
```


### 1985-1987 (Cycle 4)

```{r births-1985-1987, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 1985,
  end = 1987,
  outcome = 'Births'
)
```


### 1982-1983 (Cycle 3)

```{r births-1982-1983, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 1982,
  end = 1983,
  outcome = 'Births'
)
```



## Induced Abortions


### 2015-2019

```{r abortions-2015-2019, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2015,
  end = 2019,
  outcome = 'Abortions'
)
```


### 2011-2015

```{r abortions-2011-2015, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2011,
  end = 2015,
  outcome = 'Abortions'
)
```


### 2011-2019

```{r abortions-2011-2019, dependson=c("define-plot-function", "generate-estimates"), include=FALSE}
plot_CIs_against_census_counts(
  estimates,
  begin = 2011,
  end = 2019,
  outcome = 'Abortions'
)
```


### 2006-2010

```{r abortions-2006-2010, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2006,
  end = 2010,
  outcome = 'Abortions'
)
```


### 2002-2003 (Cycle 6)

```{r abortions-2002-2003, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 2002,
  end = 2003,
  outcome = 'Abortions'
)
```


### 1995 (Cycle 5)

```{r abortions-1995, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 1995,
  end = 1995,
  outcome = 'Abortions'
)
```


### 1985-1987 (Cycle 4)

```{r abortions-1985-1987, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 1985,
  end = 1987,
  outcome = 'Abortions'
)
```


### 1982-1983 (Cycle 3)

```{r abortions-1982-1983, dependson=c("define-plot-function", "generate-estimates")}
plot_CIs_against_census_counts(
  estimates,
  begin = 1982,
  end = 1983,
  outcome = 'Abortions'
)
```
