---
title: "Data Set and Universe of Census Counts for Post-Stratification"
author: "Joshua Born"
date: "2023-05-14"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(censusapi)
library(dplyr)
library(gt)
library(knitr)
library(tidyr)

opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  tidy = TRUE
)
```


## API Key

```{r check-api-key}
Sys.setenv(CENSUS_KEY='7c06499a2432f6947e76ee1cf0876fc4ee43ce98')
Sys.getenv("CENSUS_KEY")
```


## Demographic Characteristics Estimates by Age Groups (pep/charagegroups)

The pep/charagegroups data set from Census appears to have population totals by needed dimensions for the post-stratification targets of the NSFG. However, it only has population totals by the "Resident population" Census universe, which is larger than the target population for the NSFG.

```{r get-charagegroups-data-2019, dependson="check-api-key"}
results2019 <- getCensus(
  name = 'pep/charagegroups',
  vars = c(
    'DATE_CODE',
    'DATE_DESC',
    'AGEGROUP',
    'HISP',
    'RACE',
    'SEX',
    'POP',
    'UNIVERSE'
  ),
  region = 'us:*',
  vintage = '2019'
)
```

```{r process-charagegroups-data-2019, dependson="get-charagegroups-data-2019"}
tidy_results2019 <- tibble(results2019) |>
  select(-us) |>
  mutate(across(!ends_with(c('_DESC', 'UNIVERSE')), ~ as.integer(.x))) |>
  mutate(
    AGEGROUP_f = factor(
      AGEGROUP,
      levels = 0:31,
      labels = c(
        'All ages',
        'Age 0 to 4 years',
        'Age 5 to 9 years',
        'Age 10 to 14 years',
        'Age 15 to 19 years',
        'Age 20 to 24 years',
        'Age 25 to 29 years',
        'Age 30 to 34 years',
        'Age 35 to 39 years',
        'Age 40 to 44 years',
        'Age 45 to 49 years',
        'Age 50 to 54 years',
        'Age 55 to 59 years',
        'Age 60 to 64 years',
        'Age 65 to 69 years',
        'Age 70 to 74 years',
        'Age 75 to 79 years',
        'Age 80 to 84 years',
        'Age 85 years and older',
        'Under 18 years',
        '5 to 13 years',
        '14 to 17 years',
        '18 to 64 years',
        '18 to 24 years',
        '25 to 44 years',
        '45 to 64 years',
        '65 years and over',
        '85 years and over',
        '16 years and over',
        '18 years and over',
        '15 to 44 years',
        'Median age'
      )
    ),
    HISP_f = factor(
      HISP,
      levels = 0:2,
      labels = c(
        'Both Hispanic Origins',
        'Non-Hispanic',
        'Hispanic'
      )
    ),
    RACE_f = factor(
      RACE,
      levels = 0:11,
      labels =  c(
        'All races',
        'White alone',
        'Black alone',
        'American Indian and Alaska Native alone',
        'Asian alone',
        'Native Hawaiian and Other Pacific Islander alone',
        'Two or more races',
        'White alone or in combination',
        'Black alone or in combination',
        'American Indian and Alaska Native alone or in combination',
        'Asian alone or in combination',
        'Native Hawaiian and Other Pacific Islander alone or in combination'
      )
    ),
    SEX_f = factor(
      SEX,
      levels = 0:2,
      labels = c(
        'Both Sexes',
        'Male',
        'Female'
      )
    ),
    UNIVERSE_f = factor(
      UNIVERSE,
      levels = c('R', 'P', 'H', 'C', 'N'),
      labels = c(
        "Resident population",
        "Resident population plus Armed Forces overseas",
        "Household population",
        "Civilian population",
        "Civilian noninstitutionalized population"
      )
    )
  )
```

```{r compare-with-2017-2019, dependson="process-charagegroups-data-2019"}
nonhispanic_allrace2019 <- tidy_results2019 |>
  filter(
    HISP == 1 & RACE == 0
  ) |>
  select(-starts_with('HISP'), -starts_with('RACE'))

nonhispanic_black2019 <- tidy_results2019 |>
  filter(
    HISP == 1 & RACE == 8
  ) |>
  select(-starts_with('HISP'), -starts_with('RACE'))

nonhispanic_nonblack2019 <- inner_join(
  nonhispanic_allrace2019,
  nonhispanic_black2019 |>
    rename('Black_POP' = 'POP') |>
    select(-AGEGROUP_f, -SEX_f, -DATE_DESC, -UNIVERSE_f),
  by = join_by(DATE_CODE, AGEGROUP, SEX),
  relationship = 'one-to-one'
) |>
  mutate(
    POP = POP - Black_POP
  ) |> 
  select(
    -Black_POP
  )

hispanic2019 <- tidy_results2019 |>
  filter(
    HISP == 2 & RACE == 0
  ) |>
  select(-starts_with('HISP'), -starts_with('RACE'))


race_ethn2019 <- bind_rows(
  mutate(hispanic2019, race_ethn = 2),
  mutate(nonhispanic_black2019, race_ethn = 1),
  mutate(nonhispanic_nonblack2019, race_ethn = 3)
) |>
  mutate(
    race_ethn = factor(
      race_ethn,
      labels = c(
        'Non-Hispanic Black',
        'Hispanic',
        'Non-Hispanic Non-Black'
      )
    )
  )


race_ethn2019 |>
  filter(DATE_CODE == 11 & AGEGROUP %in% 4:10 & SEX > 0) |>
  arrange(SEX_f, race_ethn, AGEGROUP_f) |>
  select(UNIVERSE_f, SEX_f, race_ethn, AGEGROUP_f, POP) |>
  gt() |>
  fmt_number(
    POP,
    decimals = 0
  )
```



## National Monthly Population Estimates (pep/natmonthly)

The pep/natmonthly data set also appears to have population totals for all the necessary dimensions for NSFG post-stratification, and it has population totals for all 5 Census universes.

```{r get-07-2018-monthly-data, dependson="check-api-key"}
raw_monthly_07_18 <- getCensus(
  name = 'pep/natmonthly',
  vars = c(
    'MONTHLY',
    'MONTHLY_DESC',
    'AGE',
    'HISP',
    'RACE',
    'SEX',
    'POP',
    'UNIVERSE'
  ),
  monthly = 101,
  region = 'us:*',
  vintage = '2019'
)
```

```{r process-07-2018-monthly-data, dependson="get-07-2018-monthly-data"}
monthly_07_18 <- tibble(raw_monthly_07_18) |>
  select(-us, -MONTHLY, -MONTHLY_DESC, -MONTHLY_1) |>
  mutate(across(!ends_with(c('_DESC', 'UNIVERSE')), ~ as.integer(.x))) |>
  mutate(
    HISP_f = factor(
      HISP,
      levels = 0:2,
      labels = c(
        'Both Hispanic Origins',
        'Non-Hispanic',
        'Hispanic'
      )
    ),
    RACE_f = factor(
      RACE,
      levels = 0:11,
      labels =  c(
        'All races',
        'White alone',
        'Black alone',
        'American Indian and Alaska Native alone',
        'Asian alone',
        'Native Hawaiian and Other Pacific Islander alone',
        'Two or more races',
        'White alone or in combination',
        'Black alone or in combination',
        'American Indian and Alaska Native alone or in combination',
        'Asian alone or in combination',
        'Native Hawaiian and Other Pacific Islander alone or in combination'
      )
    ),
    SEX_f = factor(
      SEX,
      levels = 0:2,
      labels = c(
        'Both Sexes',
        'Male',
        'Female'
      )
    ),
    UNIVERSE_f = factor(
      UNIVERSE,
      levels = c('H', 'N', 'C', 'R', 'P'),
      labels = c(
        "Household population",
        "Civilian noninstitutionalized population",
        "Civilian population",
        "Resident population",
        "Resident population plus Armed Forces overseas"
      )
    )
  )

decompose_race_ethn <- function(these_data) {
  nonhispanic_allrace <- these_data |>
    filter(
      HISP == 1 & RACE == 0
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  nonhispanic_black <- these_data |>
    filter(
      HISP == 1 & RACE == 8
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  nonhispanic_nonblack <- inner_join(
    nonhispanic_allrace,
    nonhispanic_black |>
      rename('Black_POP' = 'POP') |>
      select(-SEX_f),
    by = join_by(AGE, SEX, UNIVERSE_f),
    relationship = 'one-to-one'
  ) |>
    mutate(POP = POP - Black_POP) |> 
    select(-Black_POP)
  
  hispanic <- these_data |>
    filter(
      HISP == 2 & RACE == 0
    ) |>
    select(-starts_with('HISP'), -starts_with('RACE'))
  
  
  bind_rows(
    mutate(nonhispanic_black, race_ethn = 1),
    mutate(hispanic, race_ethn = 2),
    mutate(nonhispanic_nonblack, race_ethn = 3)
  ) |>
    mutate(
      race_ethn = factor(
        race_ethn,
        labels = c(
          'Non-Hispanic Black',
          'Hispanic',
          'Non-Hispanic Non-Black'
        )
      )
    )
}

processed_07_18 <- decompose_race_ethn(monthly_07_18) |>
  filter(AGE >= 15 & AGE <= 49) |>
  mutate(
    age_group = cut(
      AGE,
      breaks = c(14, 19, 24, 29, 34, 39, 44, 49),
      labels = c('15-19', '20-24', '25-29', '30-34', '35-39', '40-44', '45-49')
    )
  ) |>
  select(-AGE, -SEX) |>
  group_by(UNIVERSE_f, SEX_f, race_ethn, age_group) |>
  summarize(pop = sum(POP), .groups = 'drop')
```


### Total U.S. Population in July 2018 by Universe

```{r verify-populations, dependson="process-07-2018-monthly-data"}
monthly_07_18 |>
  filter(SEX == 0 & AGE == 999 & HISP == 0 & RACE == 0) |>
  select(UNIVERSE_f, POP, SEX_f, HISP_f, RACE_f) |>
  arrange(POP) |>
  gt() |>
  fmt_number(
    POP,
    decimals = 0
  )
```


### Comparisons with Table 7 from NSFG 2017-2019 Weighting Design Documetation by Universe

```{r compare-universes-07-2018, dependson="process-07-2018-monthly-data"}
NSFG_control_totals <- c(
  1517604,
  1503424,
  1699019,
  1350050,
  1259059,
  1101783,
  1138804,
  2447590,
  2355767,
  2452515,
  2316665,
  2259957,
  2042130,
  1854169,
  5925638,
  6223463,
  7393103,
  7137295,
  6943620,
  6339338,
  6960892,
  1501808,
  1588120,
  1832548,
  1561910,
  1502718,
  1354692,
  1402470,
  2364573,
  2292692,
  2319186,
  2145325,
  2147094,
  2006586,
  1857318,
  5572796,
  6070213,
  7228890,
  7164179,
  7059902,
  6488082,
  7150824
)

wide_07_18 <- processed_07_18 |>
  filter(
    SEX_f != 'Both Sexes'
  ) |>
  pivot_wider(
    names_from = UNIVERSE_f,
    values_from = pop
  ) |>
  mutate(
    control_totals = NSFG_control_totals,
    .after = age_group
  )

wide_07_18 |>
  gt() |>
  fmt_number(
    decimals = 0
  ) |>
  cols_width(
    race_ethn ~ px(250)
  )
```


### Differences with Table 7 by Universe

```{r compare-differences-07-2018, dependson="compare-universes-07-2018"}
diffs_07_18 <- wide_07_18 |>
  mutate(
     `Household population` = abs(control_totals - `Household population`),
     `Civilian noninstitutionalized population` = abs(control_totals - `Civilian noninstitutionalized population`),
     `Civilian population` = abs(control_totals - `Civilian population`),
     `Resident population` = abs(control_totals - `Resident population`),
     `Resident population plus Armed Forces overseas` = abs(control_totals - `Resident population plus Armed Forces overseas`)
  )

diffs_07_18 |>
  gt() |>
  fmt_number(
    decimals = 0
  ) |>
  cols_width(
    race_ethn ~ px(250)
  )

diffs_07_18 |>
  summarize(across(ends_with(c('population', 'overseas')), sum)) |>
  pivot_longer(
    everything(),
    values_to = 'Difference',
    names_to = 'Universe'
  ) |>
  gt() |>
  fmt_number(
    decimals = 0
  )
```

The totals used by the 2017-2019 NSFG for post-stratification are closest to the "Household population" universe, which is consistent with the Weighting Design Documentation. However, the actual target population of the NSFG sounds closer to the "Civilian noninstitutionalized population" universe, which is what was used for the 2006-2010 NSFG and earlier.

According to the 2017-2019 NSFG Weighting Design Documentation. The "Household population" universe includes military personnel living in households, but does not include individuals living in noninstitutional group quarters, such as college dormitories, while the "Civilian noninstitutionalized population" universe does include individuals living in noninstitutional group quarters, such as college dormitories.

Based on U.S. Census [definitions](https://www.census.gov/programs-surveys/popest/about/glossary/national.html), it does seem like the "Civilian noninstitutionalized population" does not include military personnel.

Because the "Civilian noninstitutionalized population" of the United States appears to be larger than the "Household population" of the United States, the number of civilians living in group quarters must be greater than the number of military personnel living in households.

It is a judgment call whether to use "Household population" like the NSFG from 2011 to 2019 or the "Civilian noninstitutionalized population" like the NSFG from 2006-2010 and earlier.

