---
title: "Estimator Variance Due to Post-Stratification"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
    css: wide.css
---

```{r setup, include=FALSE}
library(dplyr)
library(haven)
library(here)
library(readr)
library(srvyr)
library(survey)
library(svrep)
library(tidyr)

knitr::opts_chunk$set(
  cache = TRUE,
  dpi=200,
  echo = TRUE,
  fig.align = 'center',
  fig.height=9.6,
  fig.retina=1.5,
  fig.width=9.6
)
```


## Data Preparation

### Load NSFG Data

```{r load-NSFG-data}
NSFG_2017_2019_preg_raw <- read_sas(
  data_file = here('data/NSFG/d2017_2019fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2017_2019fempreg.sas7bcat')
)
NSFG_2015_2017_preg_raw <- read_sas(
  data_file = here('data/NSFG/d2015_2017fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d2015_2017fempreg.sas7bcat')
)
NSFG_2011_2019_wgts <- read_sas(
  data_file = here('data/NSFG/d2011_2019femwgt.sas7bdat')
)
NSFG_2017_2019_fem_raw <- read_sas(
  data_file = here('data/NSFG/d2017_2019femresp.sas7bdat'),
  catalog_file = here('data/NSFG/d2017_2019femresp.sas7bcat')
)
NSFG_2015_2017_fem_raw <- read_sas(
  data_file = here('data/NSFG/d2015_2017femresp.sas7bdat'),
  catalog_file = here('data/NSFG/d2015_2017femresp.sas7bcat')
)
```


### Select NSFG Variables

```{r select-NSFG-variables, dependson="load-NSFG-data"}
NSFG_2015_2019_preg_data <- bind_rows(
  select(
    NSFG_2015_2017_preg_raw,
    SECU, SEST, CASEID,
    DATEND,
    OUTCOME,
    PREGORDR,
    FMAROUT5,
    AGEPREG
  ),
  select(
    NSFG_2017_2019_preg_raw,
    SECU, SEST, CASEID,
    DATEND,
    OUTCOME,
    PREGORDR,
    FMAROUT5,
    AGEPREG
  )
) |>
  left_join(
    select(NSFG_2011_2019_wgts, CASEID, WGT2015_2019),
    by = join_by(CASEID)
  ) |>
  mutate(
    one = 1,
    OUTCOME = as_factor(labelled(OUTCOME, c(
      "LIVE BIRTH" = 1,
      "INDUCED ABORTION" = 2,
      "STILLBIRTH" = 3,
      "MISCARRIAGE" = 4,
      "ECTOPIC PREGNANCY" = 5,
      "CURRENT PREGNANCY" = 6
    ))),
    FMAROUT5 = as_factor(labelled(FMAROUT5, c(
      "MARRIED" = 1,
      "DIVORCED" = 2,
      "WINDOWED" = 3,
      "SEPARATED" = 4,
      "NEVER MARRIED" = 5
    ))),
    marital_status = as.factor(if_else(
      FMAROUT5 == 'MARRIED',
      'Married',
      'Unmarried'
    )),
  ) |>
  rename(
    Year = DATEND
  ) |>
  arrange(
    CASEID, PREGORDR
  )

NSFG_2015_2019_fem_data <- bind_rows(
  select(NSFG_2017_2019_fem_raw, CASEID, SEST, SECU, AGER, HISPRACE2, ABORTION),
  select(NSFG_2015_2017_fem_raw, CASEID, SEST, SECU, AGER, HISPRACE2, ABORTION)
) |>
  mutate(
    one = 1,
    age_group = cut(
      AGER,
      breaks = c(14, 19, 24, 29, 34, 39, 44, 50),
      labels = c('15-19', '20-24', '25-29', '30-34', '35-39', '40-44', '45-49')
    ),
    HISPRACE2 = as_factor(labelled(HISPRACE2, c(
      'Hispanic' = 1,
      'Non-Hispanic White' = 2,
      'Non-Hispanic Black' = 3,
      'Non-Hispanic Other' = 4
    ))),
    any_abortion = !is.na(ABORTION) & ABORTION > 0
  ) |>
  left_join(
    select(NSFG_2011_2019_wgts, CASEID, WGT2015_2019),
    by = join_by(CASEID)
  )
```


### Set Calibration Variables

```{r set-calibration-variables, dependson="select-NSFG-variables"}
NSFG_2015_2019_fem_data_for_calibration <- NSFG_2015_2019_fem_data |>
  mutate(
    poststrat_age_race = factor(gsub(
      '[ -]',
      '_',
      paste(
        'Age',
        age_group,
        case_when(
          grepl('^Hispanic', HISPRACE2) ~ 'Hispanic',
          grepl('Black', HISPRACE2) ~ 'Non-Hispanic Black',
          .default = 'Non-Hispanic Non-Black'
        ),
        sep = '_'
      )
    ))
  ) |>
  left_join(
    NSFG_2015_2019_preg_data |>
      group_by(CASEID) |>
      summarize(
        abortion_age =  case_when(
          any(OUTCOME == 'INDUCED ABORTION' & AGEPREG < 20)  ~ 'Under 20 years',
          any(OUTCOME == 'INDUCED ABORTION' & AGEPREG <= 24) ~ '20-24 years',
          any(OUTCOME == 'INDUCED ABORTION' & AGEPREG <= 29) ~ '25-29 years',
          any(OUTCOME == 'INDUCED ABORTION' & AGEPREG <= 34) ~ '30-34 years',
          any(OUTCOME == 'INDUCED ABORTION' & AGEPREG <= 39) ~ '35-39 years',
          any(OUTCOME == 'INDUCED ABORTION' & AGEPREG >= 40) ~ '40 years and over',
          .default = 'Other'
        ),
        abortion_2013_2014_age = case_when(
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 2013:2014 & AGEPREG < 20) ~ 'Under 20 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 2013:2014 & AGEPREG <= 24) ~ '20-24 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 2013:2014 & AGEPREG <= 29) ~ '25-29 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 2013:2014 & AGEPREG <= 34) ~ '30-34 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 2013:2014 & AGEPREG <= 39) ~ '35-39 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 2013:2014 & AGEPREG >= 40) ~ '40 years and over',
          .default = 'Other'
        )
      ),
    by = 'CASEID'
  ) |>
  mutate(
    across(
      ends_with('_age'),
      function(x) {
        factor(
          coalesce(x, 'Other'),
          ordered = TRUE,
          levels = c(
            'Under 20 years',
            '20-24 years',
            '25-29 years',
            '30-34 years',
            '35-39 years',
            '40 years and over',
            'Other'
          )
        )
      }
    ),
    abortion_age = if_else(
      abortion_age != abortion_2013_2014_age & abortion_2013_2014_age != 'Other',
      abortion_2013_2014_age,
      abortion_age
    )
  )
```


## Before Post-Stratification

### Taylor Linearization

```{r create-NSFG-survey, dependson="select-NSFG-variables"}
NSFG_2015_2019_fem_svy <-  as_survey_design(
    .data = NSFG_2015_2019_fem_data_for_calibration,
    strata = SEST,
    ids = SECU,
    weights = WGT2015_2019,
    nest = TRUE
  )

NSFG_2015_2019_fem_svy
svytotal(~abortion_age, NSFG_2015_2019_fem_svy, deff = 'replace')

NSFG_2015_2019_fem_svy |> 
  group_by(abortion_age) |>
  summarize(
    abortion_age = survey_total(one, vartype = 'ci')
  )
```


### Jackknife

```{r create-jackknife-replicate-weights, dependson="create-NSFG-survey"}
NSFG_2015_2019_fem_svy_jackknife <- as.svrepdesign(
  design = NSFG_2015_2019_fem_svy,
  type = 'JKn'
)

paste(
  ncol(weights(NSFG_2015_2019_fem_svy_jackknife, 'analysis')),
  'replicates'
)

paste(
  'UWE',
  PracTools::deffK(weights(NSFG_2015_2019_fem_svy_jackknife, 'sampling'))
)

svytotal(~abortion_age, NSFG_2015_2019_fem_svy_jackknife, deff = 'replace')

NSFG_2015_2019_fem_svy_jackknife |>
  as_survey_rep() |>
  group_by(abortion_age) |>
  summarize(
    abortion_age = survey_total(one, vartype = 'ci')
  )
```


### Bootstrap

```{r create-bootstrap-replicate-weights, dependson="create-NSFG-survey"}
NSFG_2015_2019_fem_svy_bootstrap <- as.svrepdesign(
  design = NSFG_2015_2019_fem_svy,
  type = 'bootstrap',
  replicates = 100
)

paste(
  ncol(weights(NSFG_2015_2019_fem_svy_bootstrap, 'analysis')),
  'replicates'
)

paste(
  'UWE',
  PracTools::deffK(weights(NSFG_2015_2019_fem_svy_bootstrap, 'sampling'))
)

svytotal(~abortion_age, NSFG_2015_2019_fem_svy_bootstrap, deff = 'replace')

NSFG_2015_2019_fem_svy_bootstrap |>
  as_survey_rep() |>
  group_by(abortion_age) |>
  summarize(
    abortion_age = survey_total(one, vartype = 'ci')
  )
```


## Post-Stratification


### Load Target Totals for Post-Stratification

```{r load-target-totals-for-post-stratification}
census_pop_targets_2015_2019 <- readRDS(
  here('data/Census/prepared/census_for_2015_2019.Rds')
) |>
  filter(SEX_f == 'Female') |>
  mutate(
    poststrat_age_race = gsub(
      '[ -]',
      '_',
      paste0(
        gsub(' to | years', '_', AGEGROUP_f),
        race_ethn_f
      )
    )
  ) |>
  rename(Freq = pop) |>
  select(poststrat_age_race, Freq) |>
  as.data.frame()


census_pop_total_2015_2019 <- sum(census_pop_targets_2015_2019$Freq)


guttmacher_APC_national <- read_csv(
    here('data/Guttmacher/NationalAndStatePregnancy_PublicUse.csv'),
    show_col_types = FALSE
  ) |>
  filter(state == 'US') |>
  mutate(interpolated = year %in%
      c(1983, 1986, 1989, 1990, 1993, 1994, 1997, 1998, 2001, 2002, 2003, 2006,
        2009, 2012, 2015)
  ) |>
  select('year', 'interpolated', 'abortionslt20', 'abortions2024',
    'abortions2529', 'abortions3034', 'abortions3539', 'abortions40plus',
    'abortionstotal')


guttmacher_APC_2013_2014_age <- guttmacher_APC_national |>
  filter(year %in% 2013:2014) |>
  rename(
    `Other` = abortionstotal,
    `Under 20 years` = abortionslt20,
    `20-24 years` = abortions2024,
    `25-29 years` = abortions2529,
    `30-34 years` = abortions3034,
    `35-39 years` = abortions3539,
    `40 years and over` = abortions40plus
  ) |>
  summarize(across(3:9, sum)) |>
  pivot_longer(
    cols = everything(),
    names_to = 'abortion_2013_2014_age',
    values_to = 'Freq'
  ) |>
  mutate(
    abortion_2013_2014_age = factor(
      abortion_2013_2014_age,
      levels = c(
        'Under 20 years',
        '20-24 years',
        '25-29 years',
        '30-34 years',
        '35-39 years',
        '40 years and over',
        'Other'
      ),
      ordered = TRUE
    ),
    Freq = if_else(
      abortion_2013_2014_age == 'Other',
      census_pop_total_2015_2019 - Freq,
      Freq
    )
  ) |>
  arrange(abortion_2013_2014_age)


poststrat_targets_2015_2019 <- c(
  list(census_pop_targets_2015_2019),
  list(guttmacher_APC_2013_2014_age)
)
```


### Run Post-Stratification

```{r post-stratify-jackknife, dependson=c("create-jackknife-replicate-weights", "load-target-totals-for-post-stratification")}
source(here('R/postStratify2.R'))

NSFG_2015_2019_fem_svy_jackknife_post <- postStratify2(
  NSFG_2015_2019_fem_svy_jackknife,
  ~abortion_2013_2014_age,
  poststrat_targets_2015_2019[[2]],
  ~abortion_age,
  verbose=TRUE
)

paste(
  ncol(weights(NSFG_2015_2019_fem_svy_jackknife_post, 'analysis')),
  'replicates'
)

paste(
  'UWE',
  PracTools::deffK(weights(NSFG_2015_2019_fem_svy_jackknife_post, 'sampling'))
)
```

```{r post-stratify-bootstrap, dependson=c("create-bootstrap-replicate-weights", "load-target-totals-for-post-stratification")}
NSFG_2015_2019_fem_svy_bootstrap_post <- postStratify2(
  NSFG_2015_2019_fem_svy_bootstrap,
  ~abortion_2013_2014_age,
  poststrat_targets_2015_2019[[2]],
  ~abortion_age,
  verbose=TRUE
)

paste(
  ncol(weights(NSFG_2015_2019_fem_svy_bootstrap_post, 'analysis')),
  'replicates'
)

paste(
  'UWE',
  PracTools::deffK(weights(NSFG_2015_2019_fem_svy_bootstrap_post, 'sampling'))
)
```


## After Post-Stratification

### Jackknife

```{r, jackknife-after-post-stratification, dependson="post-stratify-jackknife"}
svytotal(~abortion_2013_2014_age, NSFG_2015_2019_fem_svy_jackknife_post, deff = 'replace')

NSFG_2015_2019_fem_svy_jackknife_post |>
  as_survey_rep() |>
  group_by(abortion_2013_2014_age) |>
  summarize(
    Freq = survey_total(one, vartype = 'ci')
  )

svytotal(~abortion_age, NSFG_2015_2019_fem_svy_jackknife_post, deff = 'replace')

NSFG_2015_2019_fem_svy_jackknife_post |>
  as_survey_rep() |>
  group_by(abortion_age) |>
  summarize(
    Freq = survey_total(one, vartype = 'ci')
  )

NSFG_2015_2019_fem_svy_jackknife_post |>
  as_data_frame_with_weights() |>
  group_by(abortion_2013_2014_age) |>
  summarize(across(starts_with('REP_WGT'), sum)) |>
  pivot_longer(-abortion_2013_2014_age) |>
  pivot_wider(
    id_cols = name,
    names_from = abortion_2013_2014_age,
    values_from = value
  ) |>
  summarize(
    mean = across(-name, sd)
  )
```


### Bootstrap

```{r, bootstrap-after-post-stratification, dependson="post-stratify-bootstrap"}
ncol(weights(NSFG_2015_2019_fem_svy_bootstrap_post, 'analysis'))

svytotal(~abortion_age, NSFG_2015_2019_fem_svy_bootstrap_post, deff = 'replace')

NSFG_2015_2019_fem_svy_bootstrap_post |>
  as_survey_rep() |>
  group_by(abortion_age) |>
  summarize(
    abortion_age = survey_total(one, vartype = 'ci')
  )
```


## Ideas

* Reduce variance introduced by post-stratification.
  * Extend the time frame for abortions used to calculate post-stratification adjustment factors to more years before 2013-2014.
  * Reduce number of levels that age.
* Examine standard error and CI of estimate of some quantity I actually care about in my analysis. Perhaps the situation isn't as bad as it seems.
