---
title: "Number of Women of Reproductive Age in the United States"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
    css: wide.css
---


```{r setup, include=FALSE}
library(censusapi)
library(dplyr)
library(ggplot2)
library(gt)
library(here)
library(readr)
library(readxl)

knitr::opts_chunk$set(
  cache = TRUE,
  dpi=200,
  echo = TRUE,
  fig.align = 'center',
  fig.height=6.4,
  fig.retina=1.5,
  fig.width=9.6
)
```


## Loading Data

### For 1940-1979

https://www.census.gov/data/tables/time-series/demo/popest/pre-1980-national.html

Note that this counts the resident population in the United States _plus_ armed forces personnel overseas.

```{r load-data-from-excel-files, message=FALSE, warning=FALSE}
load_Census_year <- function(year) {
  read_xls(
    here(paste0('data/Census/pe-11-', 10 * (year %/% 10), 's.xls')),
    sheet = as.character(year),
    skip = 9,
    col_names = FALSE
  ) |>
  as_tibble() %>%
  mutate(
    Year = year,
    Age = as.integer(.[[1]]),
    Population.Female = as.integer(.[[4]])
  ) |>
  filter(!is.na(Age)) |>
  select(Year, Age, Population.Female)
}

Census_1940_1979 <- Reduce(
  \(x, y) bind_rows(list(x, load_Census_year(y))),
  1941:1979,
  load_Census_year(1940)
)
```


### For 1980-1989

https://www.census.gov/data/datasets/time-series/demo/popest/1980s-national.html

Note that this counts just the resident population in the United States.

```{r load-data-from-fixed-width-file}
load_Census_fixed_width <- function(code) {
  read_fwf(
    here(paste0('data/Census/1980-1989/E', code, 'RQI.TXT')),
    fwf_cols(
      Series = c(1, 2),
      Month = c(3, 4),
      Year = c(5, 6),
      Age = c(7, 10),
      Population = c(12, 21),
      Population.Male = c(22, 31),
      Population.Female = c(32, 41)
    ),
    show_col_types = FALSE
  )
}

Census_1980_1989 <- Reduce(
  \(x, y) bind_rows(list(x, load_Census_fixed_width(y))),
  c(8182, 8283, 8384, 8485, 8586, 8687, 8788, 8889, 8990),
  load_Census_fixed_width(8081)
) |>
 filter(Month == 7) |>
 mutate(Year = Year + 1900)
```


### For 1990-1999

https://www.census.gov/data/developers/data-sets/popest-popproj/popest.1990-2000_Intercensals.html#list-tab-2014455046

```{r load-1990-1999-data-from-API}
Sys.setenv("CENSUS_KEY" = "7c06499a2432f6947e76ee1cf0876fc4ee43ce98")

Census_1990_1999 <- getCensus(
  name = "pep/int_natrespop",
  vintage = 1990,
  vars = c("YEAR", "MONTH", "AGE", "POPULATION_FEMALES")
) |>
  mutate(
    Year = as.numeric(YEAR),
    Age = as.numeric(AGE),
    .keep = "unused"
  ) |>
  rename(
    Month = MONTH,
    Population.Female = POPULATION_FEMALES
  ) |>
  filter(Month == 'July')
```


### For 2000-2009

https://www.census.gov/data/developers/data-sets/popest-popproj/popest.2000-2010_Intercensals.html#list-tab-2014455046

```{r load-2000-2009-data-from-API}
Sys.setenv("CENSUS_KEY" = "7c06499a2432f6947e76ee1cf0876fc4ee43ce98")

Census_2000_2009 <- getCensus(
  name = "pep/int_charage",
  vintage = 2000,
  region = "us:1",
  vars = c("GEONAME", "DATE_", "DATE_DESC", "SEX", "AGE", "AGE_DESC", "RACE", "HISP", "POP")
) |>
  mutate(
    Age = as.numeric(AGE),
    Population.Female = POP,
    Year = case_match(
      DATE_,
      "2" ~ 2000,
      "3" ~ 2001,
      "4" ~ 2002,
      "5" ~ 2003,
      "6" ~ 2004,
      "7" ~ 2005,
      "8" ~ 2006,
      "9" ~ 2007,
      "10" ~ 2008,
      "11" ~ 2009,
      .default = NA
    ),
    .keep = "unused"
  ) |>
  filter(!is.na(Year) & HISP == 0 & RACE == 0 & SEX == 2)
```


### For 2010-2019

The 2010-2020 intercensal estimates are not out yet, so unlike the other data, these are postcensal rather intercensal.

https://www.census.gov/data/developers/data-sets/popest-popproj/popest.Vintage_2019.html#list-tab-2014455046

```{r load-2010-2019-data-from-API}
Sys.setenv("CENSUS_KEY" = "7c06499a2432f6947e76ee1cf0876fc4ee43ce98")

Census_2010_2019 <- getCensus(
  name = "pep/charage",
  vintage = 2019,
  region = "us:1",
  vars = c("DATE_CODE", "DATE_DESC", "SEX", "AGE", "AGE_DESC", "RACE", "HISP", "POP")
) |>
  mutate(
    Age = as.numeric(AGE),
    Population.Female = POP,
    Year = case_match(
      DATE_CODE,
      "3" ~ 2010,
      "4" ~ 2011,
      "5" ~ 2012,
      "6" ~ 2013,
      "7" ~ 2014,
      "8" ~ 2015,
      "9" ~ 2016,
      "10" ~ 2017,
      "11" ~ 2018,
      "12" ~ 2019,
      .default = NA
    ),
    .keep = "unused"
  ) |>
  filter(!is.na(Year) & HISP == 0 & RACE == 0 & SEX == 2)
```


```{r data-processing, dependson=c("load-data-from-excel-files", "load-data-from-fixed-width-file", "load-1990-1999-data-from-API", "load-2000-2009-data-from-API")}
Census_by_age <- bind_rows(list(
  Census_1940_1979,
  select(
    Census_1980_1989,
    Year, Age, Population.Female
  ),
  select(
    Census_1990_1999,
    Year, Age, Population.Female
  ),
  select(
    Census_2000_2009,
    Year, Age, Population.Female
  ),
  select(
    Census_2010_2019,
    Year, Age, Population.Female
  )
)) |>
  mutate(
    across(c(Year, Population.Female), as.integer),
    Age.Group = cut(
      Age,
      c(14, 19, 24, 29, 34, 39, 44),
      labels = c(
        "[15,19]",
        "[20,24]",
        "[25,29]",
        "[30,34]",
        "[35,39]",
        "[40,44]"
      ),
      ordered_result = TRUE
    )
  ) |>
  filter(!is.na(Age.Group)) |>
  group_by(Year, Age.Group) |>
  summarize(Population.Female = sum(Population.Female), .groups = "drop_last")

Census_reproductive_age <- Census_by_age |>
  summarize(Population.Female = sum(Population.Female), .groups = "drop")
```


## Plots

```{r total-women-of-reproductive-age-plot, dependson="data-processing"}
Census_reproductive_age |>
  ggplot(
    mapping = aes(
      x = Year,
      y = Population.Female
    )
  ) +
    geom_vline(
      xintercept = 1980,
      linetype = 2,
      linewidth = 0.25
    ) + 
    geom_line(
      alpha = 0.4,
      color = 'blue',
      linewidth = 1.5
    ) +
    geom_point(
      color = 'blue',
      fill = 'white',
      size = 2,
      shape = 21
    ) +
    scale_shape_manual(
      values = 21
    ) +
    scale_y_continuous(,
      breaks = \(x) seq(0, x[2]+1e6, 2e6),
      labels = scales::label_number(suffix = "M", scale = 1e-6),
    ) +
    scale_x_continuous(
      breaks = \(x) seq(x[1]+1, x[2]-1, 2),
      expand = expansion(add = 1),
      minor_breaks = \(x) seq(x[1]+1, x[2]-1, 1)
    ) +
    labs(
      x = 'Year',
      y = 'Female Population of United States of Reproductive Age (15-44 years)'
    ) +
    theme_classic() +
    theme(
      axis.line.x.bottom = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.line.y.left = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.text.x = element_text(
        angle = 45,
        color = 'black',
        hjust = 0.5,
        vjust = 0.65
      ),
      axis.text.y = element_text(
        color = 'black',
        margin = margin(r = 0)
      ),
      axis.ticks.y = element_blank(),
      axis.ticks.x = element_line(
        color = 'black',
        linewidth = 0.2
      ),
      axis.title.x = element_text(
        margin = margin(t = 2)
      ),
      axis.title.y = element_text(
        margin = margin(r = 10)
      ),
      panel.grid.major = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor.x = element_blank(),
      plot.margin = margin(t = 0)
  )
```

The dashed line denotes where Census estimates represent just the resident population of the United States, not including overseas armed forces population.

```{r women-of-reproductive-age-by-age-group-plot, dependson="data-processing", warning=FALSE}
Census_by_age |>
  ggplot(
    mapping = aes(
      x = Year,
      y = Population.Female,
      color = Age.Group,
      shape = Age.Group
    )
  ) +
    geom_vline(
      xintercept = 1980,
      linetype = 2,
      linewidth = 0.25
    ) + 
    geom_line(
      alpha = 0.4,
      linewidth = 0.75
    ) +
    geom_point(
      fill = 'white',
      size = 1.5
    ) +
    scale_y_continuous(,
      breaks = \(x) seq(0, x[2]+1e6, 5e5),
      expand = expansion(mult = c(0, 0.05)),
      labels = scales::label_number(suffix = "M", scale = 1e-6),
      limits = c(4e6, NA),
      minor_breaks = NULL
    ) +
    scale_x_continuous(
      breaks = \(x) seq(x[1]+1, x[2]-1, 2),
      expand = expansion(add = 1),
      minor_breaks = \(x) seq(x[1]+1, x[2]-1, 1)
    ) +
    scale_color_brewer(
      palette = 2,
      type = "qual"
    ) + 
    labs(
      x = 'Year',
      y = 'Female Population of United States by Age Group'
    ) +
    theme_classic() +
    theme(
      axis.line.x.bottom = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.line.y.left = element_line(
        color = 'black',
        linewidth = 0.3
      ),
      axis.text.x = element_text(
        angle = 45,
        color = 'black',
        hjust = 0.5,
        vjust = 0.65
      ),
      axis.text.y = element_text(
        color = 'black',
        margin = margin(r = 0)
      ),
      axis.ticks.y = element_blank(),
      axis.ticks.x = element_line(
        color = 'black',
        linewidth = 0.2
      ),
      axis.title.x = element_text(
        margin = margin(t = 2)
      ),
      axis.title.y = element_text(
        margin = margin(r = 10)
      ),
      panel.grid.major = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor = element_line(
        color = 'gray',
        linetype = 3,
        linewidth = 0.25
      ),
      panel.grid.minor.x = element_blank(),
      plot.margin = margin(t = 0)
  )
```

The dashed line denotes where Census estimates represent just the resident population of the United States, not including overseas armed forces population.
