---
title: "Use and Nonuse of Contraception in the United States Reported in the 2017-2019 National Survey of Family Growth"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(knitr)
library(survey)

source(here('R/import.R'))

i_am('reports/contraceptive_use_and_nonuse_in_the_united_states_2017-2019.Rmd')

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r helpers}
has_level <- function(x, ids) {
  !is.na(x) & is.element(as.integer(x), ids)
}

get_raw_import_for_sex <- function(sex) {
  if (startsWith(sex, 'f')) {
    fem1719
  } else {
    male1719
  }
}

get_dt_for_sex <- function(sex) {
  if (startsWith(sex, 'f')) {
    setDT(fem1719dt)
  } else {
    setDT(male1719dt)
  }
}

get_idxs <- function(format, skip = NULL, sex = 'fem') {
  raw_import <- get_raw_import_for_sex(sex)
  idxs <- raw_import$Formats[format_name == format, factor_value]
  if (!is.null(skip)) {
    idxs[!is.element(idxs, skip)]
  } else {
    idxs
  }
}

set_mention_count_variable <- function(dt, prefix) {
  if (paste(prefix, 'count', sep = '_') %in% colnames(dt)) {
    dt[
      ,
      (paste(prefix, 'count', sep = '_')) := NULL
    ]
  }
  dt[
    ,
    (paste(prefix, 'count', sep = '_')) := rowSums(.SD),
    .SDcols = patterns(paste('^', prefix, sep = ''))
  ]
}

set_mentions_to_booleans <- function(
  out_prefix,
  in_prefix,
  in_series,
  format,
  skip = NULL,
  sex = 'fem'
) {
  dt <- get_dt_for_sex(sex)
  in_vars <- paste(in_prefix, in_series, sep = '')

  for (x in get_idxs(format, skip, sex)) {
    set(
      dt,
      j = paste(out_prefix, x, sep = ''),
      value = get_raw_import_for_sex(sex)$Data[
        ,
        lapply(.SD, \(col) has_level(col, x)),
        .SDcols = in_vars
      ] |> apply(1, any)
    )
  }

  set_mention_count_variable(dt, out_prefix)
}

set_variables_as_factors <- function(variables, sex = 'fem') {
  raw_import <- get_raw_import_for_sex(sex)
  dt <- get_dt_for_sex(sex)

  for (x in seq(length(variables))) {
    variable_name <- names(variables)[x]
    format_name <- variables[x]
    values <- factorize(
      raw_import$Data[, get(variable_name)],
      format_name,
      raw_import$Formats
    )

    dt[, (variable_name) := ..values]
  }
}

estimate_totals_and_percentages <- function(f, svy) {
  totals <- svytotal(f, svy)
  totals_ci <- as.data.frame(confint(totals, df = degf(svy)))
  names(totals_ci) <- c('total_CI_low', 'total_CI_high')
  percentages <- svymean(f, svy)
  percentages_ci <- as.data.frame(confint(percentages, df = degf(svy)))
  names(percentages_ci) <- c('percentage_CI_low', 'percentage_CI_high')
  level_prefix <- as.character(f)
  level_prefix <- level_prefix[seq(2, length(level_prefix))]

  cbind(
    data.table(
      description = sub(level_prefix, '', names(totals), fixed = TRUE),
      total = as.vector(totals)
    ),
    totals_ci,
    data.table(
      percentage = as.vector(percentages)
    ),
    percentages_ci
  )
}

rounded_estimates <- function(dt) {
  cbind(
    dt[
      ,
      .SD,
      .SDcols = patterns('^[^(total)|(percentage)]')
    ],
    dt[
      ,
      lapply(.SD, \(col) round(col / 10^6, 3)),
      .SDcols = patterns('^total')
    ],
    dt[
      ,
      lapply(.SD, \(col) round(col, 3) * 100),
      .SDcols = patterns('^percentage')
    ]
  )
}

string_estimates <- function(dt) {
  dt[,
    .(
      description,
      total = sprintf("%.3fM", total),
      total_CI = sprintf(
        "(%.3fM, %.3fM)",
        total_CI_low,
        total_CI_high
      ),
      percentage = sprintf("%0.1f%%", percentage),
      percentage_CI = sprintf(
        "(%0.1f%%, %0.1f%%)",
        percentage_CI_low,
        percentage_CI_high
      )
    )
  ]
}

pretty_estimates <- function(dt) {
  dt |>
    rounded_estimates() |>
    string_estimates()
}

svy <- function(dt) {
  svydesign(
    ids = ~SECU,
    strata = ~SEST,
    data = dt,
    nest = TRUE,
    weights = ~WGT2017_2019
  )
}


estimate_mentions <- function(
  svy,
  prefix,
  series,
  format,
  skip = NULL,
  sex = 'fem'
) {
  raw_import <- get_raw_import_for_sex(sex)
  dt <- get_dt_for_sex(sex)

  estimates <- rbindlist(lapply(
    get_idxs(format, skip, sex),
    function(x) {
      row <- estimate_totals_and_percentages(
        as.formula(paste('~', prefix, x, sep = '')),
        svy
      )[
        description == TRUE
      ]
      row[, description := x]
      row
    }
  ))

  estimates[
    , description := factorize(description, format, raw_import$Formats)
  ]

  estimates
}
```

```{r load-data}
fem1719 <- load_NSFG_data('2017_2019', 'FemResp')
male1719 <- load_NSFG_data('2017_2019', 'Male')
```

```{r process-fem-data, dependson=c('helpers', 'load-data')}
fem1719dt <- copy(fem1719$Data[, .(
  CASEID,
  SECU,
  SEST,
  WGT2017_2019
)])

set_variables_as_factors(
  variables = c(
    ATTRACT = 'ATTRACT',
    CONSTAT1 = 'CONSTATF',
    HADSEX = 'HADSEX',
    LSEXPREG = 'Y1N2RECF',
    P1YRELP = 'P1YRELP',
    POSIBLMN = 'Y1N5RDF',
    POSIBLPG = 'Y1N5RDF',
    PSURGSTR = 'Y1N5C',
    REASIMPP = 'REASP',
    REASIMPR = 'REASR',
    SEX3MO = 'Y1N2RECF',
    TUBS = 'Y1N5C',
    WYNOLSTP = 'Y1N5RDF',
    WYNOTUSE = 'Y1N5RDF',
    YNOSEX = 'YNOSEX'
  )
)

set_mentions_to_booleans(
  out_prefix = 'usedlast3m_meth',
  in_prefix = 'METH3M',
  in_series = 1:3,
  format = 'METH3MF',
  skip = 20:21
)

# set any respondents who are sexually active in past 3 months and whose last intercourse was with husband or cohabitating partner who has an active vasectomy as using vasectomy as a contraceptive method
fem1719dt[,
  usedlast3m_meth3 := fifelse(
    SEX3MO == 'YES' & PSURGSTR == 'YES' & (
      P1YRELP == 'Current husband or (current husband from whom she is separated)' |
      P1YRELP == 'Current cohabiting partner'
    ),
    TRUE,
    usedlast3m_meth3
  )
]

# set any respondents with current tubal ligations who are sexually active as currently using tubal ligation as a contraceptive method
fem1719dt[,
  usedlast3m_meth4 := fifelse(
    SEX3MO == 'YES' & TUBS == 'YES',
    TRUE,
    usedlast3m_meth4
  )
]

# reset count of methods used at last intercourse to include overrides of tubal ligations and vasectomies
set_mention_count_variable(fem1719dt, 'usedlast3m_meth')


fem1719dt[,
  usedlast3m := factor(fifelse(
    usedlast3m_meth_count > 0,
    'Currently using contraception',
    'Not currently using contraception'
  ))
]


fem1719dt[,
  usedlast3m_status := factor(
    fcase(
      usedlast3m_meth_count > 0,
        'Used at least one contraceptive method at last intercourse',
      SEX3MO != 'YES',
        'Did not have sexual intercourse with male in past 3 months',
      LSEXPREG == 'YES',
        'Pregnant in month of last intercourse',
      CONSTAT1 == 'Postpartum',
        'Postpartum',
      (WYNOTUSE == 'Yes' | WYNOLSTP == 'Yes'),
        'Currently trying to conceive',
      POSIBLPG == 'No',
        'Physically unable to conceive',
      POSIBLMN == 'No' & (
        P1YRELP == 'Current cohabiting partner' |
        P1YRELP == 'Current husband or (current husband from whom she is separated)'
      ),
        'Last intercourse was with partner who is physically unable to conceive',
      default = 'Otherwise did not use contraception at last intercourse'
    ),
    ordered = TRUE,
    levels = c(
      'Used at least one contraceptive method at last intercourse',
      'Did not have sexual intercourse with male in past 3 months',
      'Pregnant in month of last intercourse',
      'Postpartum',
      'Currently trying to conceive',
      'Physically unable to conceive',
      'Last intercourse was with partner who is physically unable to conceive',
      'Otherwise did not use contraception at last intercourse'
    )
  )
]
```

```{r process-male-data, dependson=c('helpers', 'load-data')}
male1719dt <- copy(male1719$Data[, .(
  CASEID,
  SECU,
  SEST,
  WGT2017_2019
)])

set_variables_as_factors(
  sex = 'male',
  variables = c(
    ATTRACT = 'ATTRACT',
    CWPLMET201 = 'MTHDSV3F',
    CWPLUSE2 = 'Y1N5RDF',
    CWPPOSS = 'Y1N5RDF',
    CWPPRGNW ='Y1N5RDF',
    CWPTRYPG = 'Y1N5RDF',
    FATHPOSS = 'Y1N5RDF',
    HADSEX = 'HADSEX',
    P1COHABIT = 'Y1N5RDF',
    P1CURRWIFE = 'Y1N5RDF',
    PSURGSTR = 'N0Y1CF',
    PXCPREG = 'Y1N5RDF',
    PXLPMETH01 = 'MTHDSV3F',
    PXLPUSE = 'Y1N5RDF',
    PXLSXPRB = 'Y1N5RDF',
    PXTRYING = 'Y1N5RDF',
    RSURGSTR = 'N0Y1RDF',
    SEX3MO = 'SEX3MO',
    YNOSEX = 'YNOSEX'
  )
)

set_mentions_to_booleans(
  sex = 'male',
  out_prefix = 'usedlast3m_meth',
  in_prefix = 'METH3M',
  in_series = 1:4,
  format = 'METH3MF',
  skip = 95:96
)

# count any respondents who are sexually active and who have an active vasectomy as using vasectomy as a contraceptive method
male1719dt[,
  usedlast3m_meth3 := fifelse(
    SEX3MO == 'YES, HAD INTERCOURSE' & RSURGSTR == 'YES',
    TRUE,
    usedlast3m_meth3
  )
]

# count any respondents who are sexually active and who are married or cohabitating and whose last intercourse was with wife or cohabitating partner who has active tubal ligation
male1719dt[,
  usedlast3m_meth5 := fifelse(
    SEX3MO == 'YES, HAD INTERCOURSE' &
      (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') &
      PSURGSTR == 'YES',
    TRUE,
    usedlast3m_meth5
  )
]

# reset count of methods used at last intercourse to include overrides of tubal ligations and vasectomies
set_mention_count_variable(male1719dt, 'usedlast3m_meth')


male1719dt[,
  usedlast3m := factor(fifelse(
    usedlast3m_meth_count > 0,
    'Currently using contraception',
    'Not currently using contraception'
  ))
]


male1719dt[,
  usedlast3m_status := factor(
    fcase(
      usedlast3m_meth_count > 0,
        'Used at least one contraceptive method at last intercourse',
      SEX3MO != 'YES, HAD INTERCOURSE',
        'Did not have sexual intercourse with female in past 3 months',
      PXCPREG == 'Yes' | ((P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPPRGNW == 'Yes'),
        'Last intercourse was with partner who is pregnant',
      PXTRYING == 'Yes' | ((P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPTRYPG == 'Yes'),
        'Last intercourse was with partner who is trying to conceive',
      FATHPOSS == 'No',
        'Physically unable to conceive',
      (P1CURRWIFE == 'Yes' | P1COHABIT == 'Yes') & CWPPOSS == 'No',
        'Last intercourse was with partner who is physically unable to concieve',
      default = 'Otherwise did not use contraception at last intercourse'
    ),
    ordered = TRUE,
    levels = c(
      'Used at least one contraceptive method at last intercourse',
      'Did not have sexual intercourse with female in past 3 months',
      'Last intercourse was with partner who is pregnant',
      'Last intercourse was with partner who is trying to conceive',
      'Physically unable to conceive',
      'Last intercourse was with partner who is physically unable to concieve',
      'Otherwise did not use contraception at last intercourse'
    )
  )
]

male1719dt[,
  not_aware := factor(fifelse(
    CWPLUSE2 == "Don't know" |
      CWPLMET201 == "Don't know" |
      PXLPUSE == "Don't know" |
      PXLPMETH01 == "Don't know" |
      PXLSXPRB == 'Yes' |
      PXLSXPRB == "Don't know",
    'Reports not knowing if partner used contraceptive method at last intercourse',
    'Does not report not knowing if partner used contraceptive method at last intercourse'
  ))
]

```


# Persons in Populations

Description: Persons aged 15-49 living in households in the United States as estimated by the NSFG

**Females**

```{r, dependson='process-fem-data'}
fem1719dt[, person := 1]

(estimate_totals_and_percentages(
  ~person,
  svy(fem1719dt)
) |> pretty_estimates())[
  , .(total, total_CI)
] |> kable()
```

**Males**

```{r, dependson='process-male-data'}
male1719dt[, person := 1]

(estimate_totals_and_percentages(
  ~person,
  svy(male1719dt)
) |> pretty_estimates())[
  , .(total, total_CI)
] |> kable()
```


# Currently Using Contraception

Domain: Persons aged 15-49 living in households in the United States

Description: By whether had intercourse with opposite sex during the 3 months prior to interview and used a method of contraception during last intercourse

**Females**

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~usedlast3m,
  svy(fem1719dt)
) |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~usedlast3m,
  svy(male1719dt)
) |> pretty_estimates() |> kable()
```


# Reason for Nonuse

Domain: Persons not currently using contraception, as defined above

Description: Categorized into exclusive hierarchy, such that individuals categorized earlier in the hierarchy are not counted in later categories

**Females**

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(usedlast3m_status),
  subset(
    svy(fem1719dt),
    usedlast3m_meth_count == 0
  )
) |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~factor(usedlast3m_status),
  subset(
    svy(male1719dt),
    usedlast3m_meth_count == 0
  )
) |> pretty_estimates() |> kable()
```


# Contraception Use Rate

Domain: Persons who had sexual intercourse with someone of the opposite sex in the 3 months preceding the interview, are not nor had last intercourse with someone who is pregnant, postpartum, trying to conceive, or physically infertile

**Females**

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(usedlast3m_status),
  subset(
    svy(fem1719dt),
    usedlast3m_status %in% c(
      'Used at least one contraceptive method at last intercourse',
      'Otherwise did not use contraception at last intercourse'
    )
  )
) |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~factor(usedlast3m_status),
  subset(
    svy(male1719dt),
    usedlast3m_status %in% c(
      'Used at least one contraceptive method at last intercourse',
      'Otherwise did not use contraception at last intercourse'
    )
  )
) |> pretty_estimates() |> kable()
```


# Reported Ignorance of Contraceptive Use

Domain: Males who had intercourse with a female in the past 3 months

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~not_aware,
  subset(
    svy(male1719dt),
    SEX3MO == 'YES, HAD INTERCOURSE'
  )
) |> pretty_estimates() |> kable()
```

Domain: Males who had intercourse with a female in the past 3 months

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~not_aware,
  subset(
    svy(male1719dt),
    usedlast3m_status %in% c(
      'Otherwise did not use contraception at last intercourse'
    )
  )
) |> pretty_estimates() |> kable()
```


# Not Sexually Active

## Never Had Intercourse with Opposite Sex

Domain: Persons who have not had sexual intercourse with someone of the opposite sex in the 3 months preceding the interview

**Females**

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(HADSEX),
  subset(
    svy(fem1719dt),
    SEX3MO != 'YES'
  )
) |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~factor(HADSEX),
  subset(
    svy(male1719dt),
    SEX3MO != 'YES, HAD INTERCOURSE'
  )
) |> pretty_estimates() |> kable()
```

### Reason for Never Having Intercourse

Domain: Persons who have never had sexual intercourse with someone of the opposite sex

**Females**

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(YNOSEX),
  subset(
    svy(fem1719dt),
    HADSEX == 'NO, R NEVER HAD INTERCOURSE'
  )
)[order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~factor(YNOSEX),
  subset(
    svy(male1719dt),
    HADSEX == 'NO, R NEVER HAD INTERCOURSE'
  )
)[order(-total)] |> pretty_estimates() |> kable()
```


## Attraction by Sex

Domain: Females who have not had sexual intercourse with a male in the past 3 months

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(ATTRACT),
  subset(
    svy(fem1719dt),
    SEX3MO != 'YES'
  )
) |> pretty_estimates() |> kable()
```

Domain: Females who had sexual intercourse with a male in the past 3 months

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(ATTRACT),
  subset(
    svy(fem1719dt),
    SEX3MO == 'YES'
  )
) |> pretty_estimates() |> kable()
```

Domain: Males who have not had sexual intercourse with a female in the past 3 months

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~factor(ATTRACT),
  subset(
    svy(male1719dt),
    SEX3MO != 'YES, HAD INTERCOURSE'
  )
) |> pretty_estimates() |> kable()
```

Domain: Males who had sexual intercourse with a female in the past 3 months

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~factor(ATTRACT),
  subset(
    svy(male1719dt),
    SEX3MO == 'YES, HAD INTERCOURSE'
  )
) |> pretty_estimates() |> kable()
```


# Reasons for Physical Inability to Conceive

Domain: Females who report being physically unable to get pregnant

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(REASIMPR),
  subset(
    svy(fem1719dt),
    POSIBLPG == 'No'
  )
)[order(-total)] |> pretty_estimates() |> kable()
```

Domain: Females who report their partner is unable to father a child

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(REASIMPP),
  subset(
    svy(fem1719dt),
    POSIBLMN == 'No'
  )
)[order(-total)] |> pretty_estimates() |> kable()
```


# Method(s) at Last Intercourse

## Number of Methods

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used a method of contraception during last intercourse

Description: By number of methods used at last intercourse

**Females**

```{r, dependson='process-fem-data'}
estimate_totals_and_percentages(
  ~factor(usedlast3m_meth_count),
  subset(
    svy(fem1719dt),
    usedlast3m_meth_count > 0
  )
) |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_totals_and_percentages(
  ~factor(usedlast3m_meth_count),
  subset(
    svy(male1719dt),
    usedlast3m_meth_count > 0
  )
) |> pretty_estimates() |> kable()
```


## Method(s) Used at Last Intercourse

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used a method of contraception during last intercourse

Description: By method(s), with those using multiple methods counted multiple times

**Females**

```{r, dependson='process-fem-data'}
estimate_mentions(
  svy = subset(
    svy(fem1719dt),
    usedlast3m_meth_count > 0
  ),
  prefix = 'usedlast3m_meth',
  series = 1:3,
  skip = 20:21,
  format = 'METH3MF'
)[order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_mentions(
  sex = 'male',
  svy = subset(
    svy(male1719dt),
    usedlast3m_meth_count > 0
  ),
  prefix = 'usedlast3m_meth',
  series = 1:4,
  skip = 95:96,
  format = 'METH3MF'
)[order(-total)] |> pretty_estimates() |> kable()
```


## Method Used Exclusively at Last Intercourse

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used **one and only one** method of contraception during last intercourse

**Females**

```{r, dependson='process-fem-data'}
estimate_mentions(
  svy = subset(
    svy(fem1719dt),
    usedlast3m_meth_count == 1
  ),
  prefix = 'usedlast3m_meth',
  series = 1:3,
  skip = 20:21,
  format = 'METH3MF'
)[order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_mentions(
  sex = 'male',
  svy = subset(
    svy(male1719dt),
    usedlast3m_meth_count == 1
  ),
  prefix = 'usedlast3m_meth',
  series = 1:4,
  skip = 95:96,
  format = 'METH3MF'
)[order(-total)] |> pretty_estimates() |> kable()
```


## Proportion of Method Users Who Also Used Other Method(s)

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used each specific method of contraception during last intercourse

Description: Proportion who also used other method(s)

**Females**

```{r, dependson='process-fem-data'}
(function() {
  estimates <- rbindlist(lapply(
    get_idxs('METH3MF', 20:21, 'fem'),
    function(x) {
      row <- estimate_totals_and_percentages(
        ~I(usedlast3m_meth_count > 1),
        subset(
          svy(fem1719dt),
          fem1719dt[, get(sprintf('usedlast3m_meth%d', x))]
        )
      )[
        description == TRUE
      ]
      row[, description := x]
      row
    }
  ))

  estimates[
    , description := factorize(description, 'METH3MF', fem1719$Formats)
  ]

  estimates
})()[!is.nan(total_CI_low)][order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
(function() {
  estimates <- rbindlist(lapply(
    get_idxs('METH3MF', 95:96, 'male'),
    function(x) {
      row <- estimate_totals_and_percentages(
        ~I(usedlast3m_meth_count > 1),
        subset(
          svy(male1719dt),
          male1719dt[, get(sprintf('usedlast3m_meth%d', x))]
        )
      )[
        description == TRUE
      ]
      row[, description := x]
      row
    }
  ))

  estimates[
    , description := factorize(description, 'METH3MF', male1719$Formats)
  ]

  estimates
})()[order(-total)] |> pretty_estimates() |> kable()
```


## Other Method(s) Used at Last Intercourse with Calendar Method

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used a calendar method during last intercourse

Description: By method(s), with those using multiple methods counted multiple times

**Females**

```{r, dependson='process-fem-data'}
estimate_mentions(
  svy = subset(
    svy(fem1719dt),
    usedlast3m_meth8
  ),
  prefix = 'usedlast3m_meth',
  series = 1:3,
  skip = c(8, 20:21),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_mentions(
  sex = 'male',
  svy = subset(
    svy(male1719dt),
    usedlast3m_meth9
  ),
  prefix = 'usedlast3m_meth',
  series = 1:4,
  skip = c(9, 95:96),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```


## Other Method(s) Used at Last Intercourse with Withdrawal

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used withdrawal during last intercourse

Description: By method(s), with those using multiple methods counted multiple times

**Females**

```{r, dependson='process-fem-data'}
estimate_mentions(
  svy = subset(
    svy(fem1719dt),
    usedlast3m_meth5
  ),
  prefix = 'usedlast3m_meth',
  series = 1:3,
  skip = c(5, 20:21),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_mentions(
  sex = 'male',
  svy = subset(
    svy(male1719dt),
    usedlast3m_meth2
  ),
  prefix = 'usedlast3m_meth',
  series = 1:4,
  skip = c(2, 95:96),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```


## Other Method(s) Used at Last Intercourse with Condoms

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used condoms during last intercourse

Description: By method(s), with those using multiple methods counted multiple times

**Females**

```{r, dependson='process-fem-data'}
estimate_mentions(
  svy = subset(
    svy(fem1719dt),
    usedlast3m_meth2
  ),
  prefix = 'usedlast3m_meth',
  series = 1:3,
  skip = c(2, 20:21),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_mentions(
  sex = 'male',
  svy = subset(
    svy(male1719dt),
    usedlast3m_meth1
  ),
  prefix = 'usedlast3m_meth',
  series = 1:4,
  skip = c(1, 95:96),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```


## Other Method(s) Used at Last Intercourse with Pill

Domain: Persons who had intercourse with opposite sex during the 3 months prior to interview and used the pill during last intercourse

Description: By method(s), with those using multiple methods counted multiple times

**Females**

```{r, dependson='process-fem-data'}
estimate_mentions(
  svy = subset(
    svy(fem1719dt),
    usedlast3m_meth1
  ),
  prefix = 'usedlast3m_meth',
  series = 1:3,
  skip = c(1, 20:21),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```

**Males**

```{r, dependson='process-male-data'}
estimate_mentions(
  sex = 'male',
  svy = subset(
    svy(male1719dt),
    usedlast3m_meth4
  ),
  prefix = 'usedlast3m_meth',
  series = 1:4,
  skip = c(4, 95:96),
  format = 'METH3MF'
)[total > 0][order(-total)] |> pretty_estimates() |> kable()
```



* Physically infertile category (females)
    * By reason in domain
    * By reason generally
    * Partner infertile generally
* Why not using contraception (females)
    * By reason(s)
    * By main or exclusive reason
    * Reasons(s) for those who believe themselves infertile to think themselves infertile
* Method(s) used at last intercourse
* Past 12 months
    * Number of methods
    * By method(s) used
    * By method, among those using only one method
* Consistency
    * Did not use at last intercourse, but used at least one method in past 12 months
    * Did not use at last intercourse, but ever used contraception (females)
    * Frequency of any method use in past 12 months
    * Frequency of condom use in past 12 months
    * Frequency of condom use in past 12 months, among exclusive condom users
    * Frequency of condom use in past 4 weeks (females)
    * Number of missed pills in past 4 weeks (females)
* Ever (females)
    * By method(s)
    * By method(s) among those who have had intercourse with opposite sex
    * Ever discontinued due to dissatisfaction
    * Reason for dissatisfaction with pill
    * Reason for dissatisfaction with condoms
    * Reason for dissatisfaction with IUDs
* Use for reasons other than contraception
    * Condom, pill, IUD, and tubal ligation among females
    * Condom among males



# Things to Do

* Collect estimates into this report
* Style plots and tables for output
* Write text of report
* Review situations in which universes do not match specifications in codebooks and follow up with NSFG administrators
* Compare own report with published reports
* Look into what it takes to publish in an academic journal
* Replicate analysis in SAS
    * Attempt single-factor categorical data analysis, e.g., whether totals of otherwise not using contraception and currently trying to conceive are different
