---
title: "Contraceptive Use and Nonuse in the United States: Based on the 2017-2019 National Survey of Family Growth"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(knitr)

source(here('R/import.R'))

i_am('reports/contraceptive_use_and_nonuse_in_the_united_states_2017-2019.Rmd')

opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  fig.height = 8,
  fig.width = 12,
  warning = FALSE
)
```

```{r load-data}
fem1719 <- load_NSFG_data('2017_2019', 'FemResp')
male1719 <- load_NSFG_data('2017_2019', 'Male')
```

```{r start-data-tables, dependson='load-data'}
fem1719dt <- copy(fem1719$Data[, .(
  CASEID,
  SECU,
  SEST,
  WGT2017_2019
)])

male1719dt <- copy(male1719$Data[, .(
  CASEID,
  SECU,
  SEST,
  WGT2017_2019
)])
```

```{r helpers, dependson='start-data-tables'}
has_level <- function(x, ids) {
  !is.na(x) & is.element(as.integer(x), ids)
}

get_raw_import_for_sex <- function(sex) {
  if (startsWith(sex, 'f')) {
    fem1719
  } else {
    male1719
  }
}

get_dt_for_sex <- function(sex) {
  if (startsWith(sex, 'f')) {
    setDT(fem1719dt)
  } else {
    setDT(male1719dt)
  }
}

get_idxs <- function(format, skip = NULL, sex = 'fem') {
  idxs <- fem1719$Formats[format_name == format, factor_value]
  if (!is.null(skip)) {
    idxs[!is.element(idxs, skip)]
  } else {
    idxs
  }
}

set_count_mention_variables <- function(dt, prefix) {
  dt[
    ,

        (paste(prefix, 'count', sep = '_')) := rowSums(.SD),
    .SDcols = patterns(paste('^', prefix, sep = ''))
  ]
}

set_mentions_to_booleans <- function(
  out_prefix,
  in_prefix,
  in_series,
  format,
  skip = NULL,
  sex = 'fem'
) {
  dt <- get_dt_for_sex(sex)
  in_vars <- paste(in_prefix, in_series, sep = '')

  for (x in get_idxs(format, skip, sex)) {
    set(
      dt,
      j = paste(out_prefix, x, sep = ''),
      value = get_raw_import_for_sex(sex)$Data[
        ,
        lapply(.SD, \(col) has_level(col, x)),
        .SDcols = in_vars
      ] |> apply(1, any)
    )
  }

  set_count_mention_variables(dt, out_prefix)
}

import_variables <- function(variables, formats = c(), sex = 'fem') {
  raw_import <- get_raw_import_for_sex(sex)
  dt <- get_dt_for_sex(sex)

  for (x in seq(length(variables))) {
    variable_name <- variables[x]

    set(
      dt,
      j = variable_name,
      value = raw_import$Data[, ..variable_name]
    )

    if (length(formats) > 0 & !is.na(formats[x])) {
      dt[
        ,
        (variable_name) := factorize(
          get(variable_name),
          formats[x],
          raw_import$Formats
        )
      ]
    }
  }
}
```


# Persons in Populations



# Currently Using Contraception

Domain: Persons aged 15-49 living in households in the United States

Description: By whether had intercourse with opposite sex during the 3 months prior to interview and used a method of contraception during last intercourse

```{r, dependson='helpers'}
set_mentions_to_booleans(
  out_prefix = 'usedlast3m_meth',
  in_prefix = 'METH3M',
  in_series = 1:3,
  format = 'METH3MF',
  skip = 20:21
)

import_variables(
  variables = c(
    'SEX3MO',
    'TUBS'
  ),
  formats = c(
    'Y1N2RECF',
    'Y1N5C'
  )
)
```



* During last intercourse
    * By number of methods used
    * By method(s) used
    * By method, among those using only one method
    * By method, proportion also using another method
    * By other method(s) used, among those using methods commonly used with other methods
* Past 12 months
    * Number of methods
    * By method(s) used
    * By method, among those using only one method
* Consistency
    * Did not use at last intercourse, but used at least one method in past 12 months
    * Did not use at last intercourse, but ever used contraception (females)
    * Frequency of any method use in past 12 months
    * Frequency of condom use in past 12 months
    * Frequency of condom use in past 12 months, among exclusive condom users
    * Frequency of condom use in past 4 weeks (females)
    * Number of missed pills in past 4 weeks (females)
* Ever (females)
    * By method(s)
    * By method(s) among those who have had intercourse with opposite sex
    * Ever discontinued due to dissatisfaction
    * Reason for dissatisfaction with pill
    * Reason for dissatisfaction with condoms
    * Reason for dissatisfaction with IUDs
* Not currently using contraception, overview by category
* Not sexually active currently
    * Never
        * Reason for never
    * Same-sex attractedness
* Physically infertile category (females)
    * By reason in domain
    * By reason generally
    * Partner infertile generally
* Why not using contraception (females)
    * By reason(s)
    * By main or exclusive reason
    * Reasons(s) for those who believe themselves infertile to think themselves infertile
* Overall contraception rate
* Use for reasons other than contraception
    * Condom, pill, IUD, and tubal ligation among females
    * Condom among males



# Things to Do

* Collect estimates into coherent report
* Review situations in which universes do not match specifications in codebooks and follow up with NSFG administrators
* Compare own report with published reports
* Look into what it takes to publish in an academic journal
* Replicate analysis in SAS
    * Attempt single-factor categorical data analysis, e.g., whether totals of otherwise not using contraception and currently trying to conceive are different
