---
title: "NSFG Cycle 3 (1982) Weight Adjustment"
author: "Joshua Born"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
    css: wide.css
---

```{r setup, include=FALSE}
library(PracTools)
library(dplyr)
library(gt)
library(haven)
library(here)
library(lubridate)
library(readr)
library(srvyr)
library(stringr)
library(survey)
library(svrep)
library(tidyr)

knitr::opts_chunk$set(
  cache = TRUE,
  dpi=200,
  echo = TRUE,
  fig.align = 'center',
  fig.height=9.6,
  fig.retina=1.5,
  fig.width=9.6
)
```


## Weight Preparation

### Loading of Raw Data

```{r load-raw-data}
raw_preg_1982_1983 <- read_sas(
  data_file = here('data/NSFG/d1982_1983fempreg.sas7bdat'),
  catalog_file = here('data/NSFG/d1982_1983fempreg.sas7bcat')
)
raw_fem_1982_1983 <- read_sas(
  data_file = here('data/NSFG/d1982_1983femresp.sas7bdat'),
  catalog_file = here('data/NSFG/d1982_1983femresp.sas7bcat')
)
```


### Data Preparation

```{r data-preparation, dependson="load-raw-data"}
data_preg <- select(
    raw_preg_1982_1983,
    
    CASEID,
    
    # Variables used in post-stratification
    AGEPREG,
    YRPREG,
    OUTCOME,
    
    # Variable used in pregnancy counts
    WANTWIFE,
    
    # Variable used in births test harness
    BP19_20B
  ) |>
  mutate(

    # Variables used in post-stratification
    AgeAtEndOfPregnancy = as.integer(AGEPREG / 100),
    Year = if_else(
      !is.na(YRPREG) & YRPREG != 99,
      1900 + YRPREG,
      NA
    ),
    YRPREG = as_factor(labelled(YRPREG, c(
      "CURRENT PREGNANCY" = NA,
      "BEFORE 1970" = 60,
      "YEAR NOT ASCERTAINED" = 99
    ))),
    OUTCOME = as_factor(labelled(OUTCOME, c(
      "LIVE BIRTH" = 1,
      "INDUCED ABORTION" = 2,
      "MISCARRIAGE" = 3,
      "STILLBIRTH" = 4,
      "CURRENTLY PREGNANCY" = 5
    ))),
    
    # Variable used for pregnancy counts
    one = 1,
    intent = as.factor(
      if_else(
        WANTWIFE %in% 1:2,
        'Intended',
        'Unintended'
      )
    ),
    WANTWIFE = as_factor(labelled(WANTWIFE, c(
      "OVERDUE" = 1,
      "TIMED RIGHT" = 2,
      "MISTIMED" = 3,
      "INDIFFERENT" = 4,
      "UNWANTED" = 5,
      "UNDETERMINED" = 6
    ))),
    
    # Variables used for births test harness
    age_group6 = cut(
      as.integer(AGEPREG / 100),
      breaks = c(0, seq(20, 40, 5), Inf),
      right = FALSE,
      labels = c(
        "Under 20 years",
        "20-24 years",
        "25-29 years",
        "30-34 years",
        "35-39 years",
        "40 years and over"
      )
    ),
    
    # Variables used in abortions test harness
    births = if_else(
      OUTCOME == 'LIVE BIRTH',
      BP19_20B,
      0
    ),
    abortion = if_else(
      OUTCOME == 'INDUCED ABORTION',
      1,
      0
    )
  )


data_fem <- select(
    raw_fem_1982_1983,
  
    CASEID,
    FINALWT,
    PSEUDOST,
    PSUCODE,
    PAIRNUM,
  
    AGE5,
    RACEOBS,
    SRRACE,
    FMARITAL
  ) |>
  mutate(
    one = 1,
    
    FINALWT = FINALWT / 10,
    
    PAIRNUM = if_else(
      PSEUDOST == 21 & PSUCODE == 62,
      2,
      PAIRNUM,
    ),
    
    age_coarse = as_factor(if_else(
      AGE5 == 1,
      '15-19 years of age',
      '20-44 years of age'
    )),
    
    # Define the source variables to be used for the post-stratification variables
    AGE5 = as_factor(labelled(AGE5, c(
      "15-19 years" = 1,
      "20-24 years" = 2,
      "25-29 years" = 3,
      "30-34 years" = 4,
      "35-39 years" = 5,
      "40-44 years" = 6
    ))),
    race = if_else(
      SRRACE == 1,
      'Black',
      'White or other'
    ),
    marital_history = if_else(
      FMARITAL == 6,
      'Never married',
      'Ever married'
    ),
    
    # Define the post-stratification variables
    race_age_marital_history = factor(paste(
      race,
      age_coarse,
      marital_history,
      sep = ', '
    ))
  ) |>
  left_join(
    data_preg |>
      group_by(CASEID) |>
      summarize(
        abortion_1974_1979_age = case_when(
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 1974:1979 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 1974:1979 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 1974:1979 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          #any(OUTCOME == 'INDUCED ABORTION' & Year %in% 1974:1979 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 39) ~ '30-39 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 1974:1979 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year %in% 1974:1979 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        ),
        abortion_1974plus_age = case_when(
          any(OUTCOME == 'INDUCED ABORTION' & Year >= 1980 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year >= 1980 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year >= 1980 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          #any(OUTCOME == 'INDUCED ABORTION' & Year >= 1980 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 39) ~ '30-39 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year >= 1980 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'INDUCED ABORTION' & Year >= 1980 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        ),
        birth_1979_age = case_when(
          any(OUTCOME == 'LIVE BIRTH' & Year == 1979 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1979 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1979 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1979 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1979 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        ),
        birth_1978_age = case_when(
          any(OUTCOME == 'LIVE BIRTH' & Year == 1978 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1978 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1978 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1978 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1978 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        ),
        birth_1977_age = case_when(
          any(OUTCOME == 'LIVE BIRTH' & Year == 1977 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1977 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1977 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1977 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1977 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        ),
        birth_1976_age = case_when(
          any(OUTCOME == 'LIVE BIRTH' & Year == 1976 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1976 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1976 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1976 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1976 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        ),
        birth_1975_age = case_when(
          any(OUTCOME == 'LIVE BIRTH' & Year == 1975 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1975 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1975 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1975 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1975 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        ),
        birth_1974_age = case_when(
          any(OUTCOME == 'LIVE BIRTH' & Year == 1974 & AgeAtEndOfPregnancy <= 19) ~ 'Under 20 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1974 & AgeAtEndOfPregnancy >= 20 & AgeAtEndOfPregnancy <= 24) ~ '20-24 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1974 & AgeAtEndOfPregnancy >= 25 & AgeAtEndOfPregnancy <= 29) ~ '25-29 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1974 & AgeAtEndOfPregnancy >= 30 & AgeAtEndOfPregnancy <= 34) ~ '30-34 years',
          any(OUTCOME == 'LIVE BIRTH' & Year == 1974 & AgeAtEndOfPregnancy >= 35 & AgeAtEndOfPregnancy <= 39) ~ '35-39 years',
          .default = 'Other'
        )
      ),
    by = 'CASEID'
  ) |>
  mutate(
    across(
      starts_with('birth_'),
      function(x) {
        factor(
          coalesce(x, 'Other'),
          levels = c(
            'Under 20 years',
            '20-24 years',
            '25-29 years',
            '30-34 years',
            '35-39 years',
            'Other'
          ),
          order = TRUE
        )
      }
    ),
    across(
      starts_with('abortion_'),
      function(x) {
        factor(
          coalesce(x, 'Other'),
          levels = c(
            'Under 20 years',
            '20-24 years',
            '25-29 years',
            '30-34 years',
            '35-39 years',
            #'30-39 years',
            'Other'
          ),
          order = TRUE
        )
      }
    ),
    abortion_1974plus_age = if_else(
      abortion_1974plus_age != abortion_1974_1979_age & abortion_1974_1979_age != 'Other',
      abortion_1974_1979_age,
      abortion_1974plus_age
    )
  )

  
before_fem <- data_fem |>
  as_survey_design(
    strata = PSEUDOST,
    ids = PAIRNUM,
    weights = FINALWT,
    nest = TRUE
  ) |>
  as_survey_rep(
    type = 'BRR'
  )
```


### Recreation of Demographic Post-Stratification Targets

```{r create-demographic-post-stratification-targets, dependson="data-preparation"}
post_stratification_estimates <- before_fem |>
  group_by(race_age_marital_history) |>
  summarize(Freq = survey_total(one)) |>
  mutate(
    Variable = 'race_age_marital_history',
    Level = race_age_marital_history,
    .keep = 'unused',
    .before = everything()
  )


post_stratification_estimates |>
  gt() |>
  fmt_number(decimal = 0)


demographic_targets <- post_stratification_estimates |>
  select(-Freq_se) |>
  group_by(Variable) |>
  group_split() |>
  lapply(function(tbl) {
    variable_name <- first(pull(tbl, Variable))
    tbl <- select(tbl, -Variable)
    names(tbl) <- c(variable_name, 'Freq')
    tbl
  })
```


### Creation of External Post-Stratification Targets

```{r create-post-stratification-targets}
population_total <- before_fem |>
  summarize(Freq = survey_total(one)) |>
  pull(Freq) |>
  sum()


guttmacher_APC_national <- read_csv(
    here('data/Guttmacher/NationalAndStatePregnancy_PublicUse.csv'),
    show_col_types = FALSE
  ) |>
  filter(state == 'US') |>
  mutate(interpolated = year %in%
      c(1983, 1986, 1989, 1990, 1993, 1994, 1997, 1998, 2001, 2002, 2003, 2006,
        2009, 2012, 2015)
  ) |>
  select('year', 'interpolated', 'abortionslt20', 'abortions2024',
    'abortions2529', 'abortions3034', 'abortions3539', 'abortions40plus',
    'abortionstotal')


guttmacher_APC_age <- guttmacher_APC_national |>
  select(-abortions40plus, -abortionstotal) |>
  filter(year %in% 1974:1979) |>
  #mutate(
  #  abortions3039 = abortions3034 + abortions3539,
  #  .keep = 'unused'
  #) |>
  rename(
    `Under 20 years` = abortionslt20,
    `20-24 years` = abortions2024,
    `25-29 years` = abortions2529,
    `30-34 years` = abortions3034,
    `35-39 years` = abortions3539
    #`30-39 years` = abortions3039
  ) |>
  summarize(across(3:7, sum)) |>
  pivot_longer(
    cols = everything(),
    names_to = 'abortion_1974_1979_age',
    values_to = 'Freq'
  ) |>
  select(abortion_1974_1979_age, Freq) %>%
  bind_rows(summarize(
    .,
    abortion_1974_1979_age = 'Other',
    Freq = population_total - sum(Freq)
  )) |>
  mutate(
    abortion_1974_1979_age = factor(
      abortion_1974_1979_age,
      levels = c(
        'Under 20 years',
        '20-24 years',
        '25-29 years',
        #'30-39 years',
        '30-34 years',
        '35-39 years',
        'Other'
      ),
      ordered = TRUE
    )
  )


NVSS_age <- readRDS(here('data/NVSS/natality_by_age_1974-1994us.Rds')) |>
  filter(
    Year %in% 1974:1979 &
      mage8 %in% c(
        'Under 15 years',
        '15-19 years',
        '20-24 years',
        '25-29 years',
        '30-34 years',
        '35-39 years'
      )
  ) |>
  mutate(
    Variable = paste0('birth_', Year, '_age'),
    Level = forcats::fct_recode(
      mage8,
      `Under 20 years` = 'Under 15 years',
      `Under 20 years` = '15-19 years'
    ),
    .keep = 'unused',
    .before = 'n'
  ) |>
  group_by(Variable, Level) |>
  summarize(n = sum(n), .groups = 'drop') |>
  arrange(Variable, Level) |>
  group_by(Variable) |>
  group_split() |>
  lapply(function(tbl) {
    variable_name <- first(pull(tbl, Variable))
    tbl <- select(tbl, -Variable)
    names(tbl) <- c(variable_name, 'Freq')
    tbl %>%
      add_row(
        summarize(
          .,
          across(where(is.factor), ~'Other'),
          Freq = population_total - sum(Freq)
        )
      ) |>
      mutate(
        across(where(is.character), function(x) {
          factor(
            x,
            levels = c(
              'Under 20 years',
              '20-24 years',
              '25-29 years',
              '30-34 years',
              '35-39 years',
              'Other'
            ),
            ordered = TRUE
          )
        })
      )
  })
```


### Post-Stratifications

```{r rake-post-stratifications, dependson=c("data-preparation", "create-post-stratification-targets")}
source(here('R/rake2.R'))
  
post_stratification_targets <- c(
  demographic_targets,
  NVSS_age,
  list(guttmacher_APC_age)
)

weight_adj_strata <- list()
weight_adj_strata[[length(post_stratification_targets)]] <- ~abortion_1974plus_age

post_fem <- rake2(
  before_fem,
  sample.margins = lapply(
    post_stratification_targets,
    \(x) as.formula(paste0('~', colnames(x)[1]))
  ),
  population.margins = post_stratification_targets,
  weight_adj_strata = weight_adj_strata,
  control=list(maxit=50, epsilon=1, verbose=FALSE)
)
```


### Saving of Adjusted Weights

```{r save-adjusted-weights, dependson="rake-post-stratifications"}
post_fem |>
  as_data_frame_with_weights(
    full_wgt_name = "FULL_SAMPLE_WGT",
    rep_wgt_prefix = "REP_WGT_"
  ) |>
  select(CASEID, contains('_WGT')) |>
  arrange(CASEID) |>
  saveRDS(here('data/NSFG/adjusted_cycle_3_weights.Rds'))
```


## Examination of Results of Weight Adjustment

### Creation of Survey Objects

```{r create-survey-objects, dependson=c("save-adjusted-weights")}
after_fem <- data_fem |>
  select(-FINALWT) |>
  left_join(
    readRDS(here('data/NSFG/adjusted_cycle_3_weights.Rds')),
    by = join_by(CASEID)
  ) |>
  as_survey_rep(
    type = 'BRR',
    weights = FULL_SAMPLE_WGT,
    repweights = starts_with('REP_WGT_')
  )

before_preg <- data_preg |>
  left_join(
    before_fem |>
      as_data_frame_with_weights(
        full_wgt_name = "FULL_SAMPLE_WGT",
        rep_wgt_prefix = "REP_WGT_"
      ) |>
      select(CASEID, contains('_WGT')),
    by = join_by(CASEID)
  ) |>
  as_survey_rep(
    type = 'BRR',
    weights = FULL_SAMPLE_WGT,
    repweights = starts_with('REP_WGT_')
  )

after_preg <- data_preg |>
  left_join(
    readRDS(here('data/NSFG/adjusted_cycle_3_weights.Rds')),
    by = join_by(CASEID)
  ) |>
  as_survey_rep(
    type = 'BRR',
    weights = FULL_SAMPLE_WGT,
    repweights = starts_with('REP_WGT_')
  )
```


### UWE

```{r uwe, dependson="create-survey-objects"}
data.frame(
  when = rep(c('Before Raking', 'After Raking'), 2),
  table = rep(c('Female Respondents', 'Pregnancies'), each = 2),
  UWE = c(
    deffK(weights(before_fem, 'sampling')),
    deffK(weights(after_fem, 'sampling')),
    deffK(weights(before_preg, 'sampling')),
    deffK(weights(after_preg, 'sampling'))
  )
) |>
  gt() |>
  fmt_number()
```


### Observations with Extreme Weights

#### By Entropy

```{r extreme-entropy, dependson="create-survey-objects"}
entropy_extreme_before <- before_fem$variables |>
  select(CASEID, FINALWT) |>
  filter(
    FINALWT^2 > 10 * mean(FINALWT^2)
  )

entropy_extreme_after <-  after_fem$variables |>
  select(CASEID, FULL_SAMPLE_WGT) |>
  filter(
    FULL_SAMPLE_WGT^2 > 10 * mean(FULL_SAMPLE_WGT^2)
  )

data.frame(
  Label = c(
    'Total Observations',
    'Extreme Weights Before',
    'Extreme Weights After',
    'Difference'
  ),
  Count = c(
    nrow(after_fem),
    nrow(entropy_extreme_before),
    nrow(entropy_extreme_after),
    nrow(entropy_extreme_after) - nrow(entropy_extreme_before)
  )
) |>
  gt() |>
  fmt_number(decimal = 0)
```


#### By Inter-Quartile Range (IQR)

```{r extreme-iqr, dependson="create-survey-objects"}
iqr_extreme_after <- after_fem$variables |>
  select(CASEID, FULL_SAMPLE_WGT) |>
  filter(
    FULL_SAMPLE_WGT > 5 * IQR(FULL_SAMPLE_WGT) + median(FULL_SAMPLE_WGT)
  )

iqr_extreme_before <- before_fem$variables |>
  select(CASEID, FINALWT) |>
  filter(
    FINALWT > 5 * IQR(FINALWT) + median(FINALWT)
  )

data.frame(
  Label = c(
    'Total Observations',
    'Extreme Weights Before',
    'Extreme Weights After',
    'Difference'
  ),
  Count = c(
    nrow(after_fem),
    nrow(iqr_extreme_before),
    nrow(iqr_extreme_after),
    nrow(iqr_extreme_after) - nrow(iqr_extreme_before)
  )
) |>
  gt() |>
  fmt_number(decimal = 0)
```


### Post-Stratification Targets

```{r verify-post-stratification, dependson="create-survey-objects"}
after_fem |>
  group_by(race_age_marital_history) |>
  summarize(
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[1]],
    by = join_by(race_age_marital_history)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)

after_fem |>
  group_by(birth_1974_age) |>
  summarize(
    Count = n(),
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[2]],
    by = join_by(birth_1974_age)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)

after_fem |>
  group_by(birth_1975_age) |>
  summarize(
    Count = n(),
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[3]],
    by = join_by(birth_1975_age)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)

after_fem |>
  group_by(birth_1976_age) |>
  summarize(
    Count = n(),
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[4]],
    by = join_by(birth_1976_age)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)

after_fem |>
  group_by(birth_1977_age) |>
  summarize(
    Count = n(),
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[5]],
    by = join_by(birth_1977_age)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)

after_fem |>
  group_by(birth_1978_age) |>
  summarize(
    Count = n(),
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[6]],
    by = join_by(birth_1978_age)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)

after_fem |>
  group_by(birth_1979_age) |>
  summarize(
    Count = n(),
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[7]],
    by = join_by(birth_1979_age)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)

after_fem |>
  group_by(abortion_1974_1979_age) |>
  summarize(
    Count = n(),
    Est = survey_total(one)
  ) |>
  full_join(
    post_stratification_targets[[8]],
    by = join_by(abortion_1974_1979_age)
  ) |>
  mutate(
    Diff = Est - Freq
  ) |>
  gt() |>
  fmt_number(c(Est, Freq), decimal = 0) |>
  fmt_number(c(Est_se, Diff), decimal = 2)
```


### Abortion Adjustments

```{r find-average-abortion-adjustment, dependson="create-survey-objects"}
before_fem |>
  group_by(abortion_1974_1979_age) |>
  summarize(
    Count = n(),
    Before = survey_total(one)
  ) |>
  full_join(
    after_fem |>
      group_by(abortion_1974_1979_age) |>
      summarize(
        After = survey_total(one)
      ),
    by = join_by(abortion_1974_1979_age)
  ) |>
  mutate(
    Factor = After / Before
  ) |>
  gt() |>
  fmt_number(decimal = 0) |>
  fmt_number(Factor, decimal = 2)

before_fem |>
  group_by(abortion_1974plus_age) |>
  summarize(
    Count = n(),
    Before = survey_total(one)
  ) |>
  full_join(
    after_fem |>
      group_by(abortion_1974plus_age) |>
      summarize(
        After = survey_total(one)
      ),
    by = join_by(abortion_1974plus_age)
  ) |>
  mutate(
    Factor = After / Before
  ) |>
  gt() |>
  fmt_number(decimal = 0) |>
  fmt_number(Factor, decimal = 2)
```


### Pregnancies by Year

This survey will cover years 1974 to 1979.

```{r pregnancies-per-year, dependson="create-survey-objects"}
bind_rows(list(
  before_preg |>
    group_by(Year) |>
    summarize(
      n = n(),
      Freq = survey_total(one, vartype = 'se')
    ) |>
    mutate(
      RSE = Freq_se / Freq,
      post_strat = 'before'
    ) |>
    filter(Year %in% 1974:1979),
  after_preg |>
    group_by(Year) |>
    summarize(
      n = n(),
      Freq = survey_total(one, vartype = 'se')
    ) |>
    mutate(
      RSE = Freq_se / Freq,
      post_strat = 'after'
    ) |>
    filter(Year %in% 1974:1979)
)) |> inner_join(
  bind_rows(list(
    before_preg |>
      group_by(Year) |>
      summarize(
        n = n(),
        Freq = survey_total(one, vartype = 'ci')
      ) |>
      mutate(
        Freq_CI_range = Freq_upp - Freq_low,
        Rel_CI = Freq_CI_range / Freq,
        post_strat = 'before'
      ) |>
      filter(Year %in% 1974:1979),
    after_preg |>
      group_by(Year) |>
      summarize(
        n = n(),
        Freq = survey_total(one, vartype = 'ci')
      ) |>
      mutate(
        Freq_CI_range = Freq_upp - Freq_low,
        Rel_CI = Freq_CI_range / Freq,
        post_strat = 'after'
      ) |>
      filter(Year %in% 1974:1979)
  )),
  by = join_by(Year, post_strat, Freq, n)
) |>
  select(Year, post_strat, everything()) |>
  mutate(
    post_strat = factor(
      post_strat,
      levels = c('before', 'after'),
      ordered = TRUE
    )
  ) |>
  arrange(Year, post_strat) |>
  group_by(Year) |>
  gt() |>
  fmt_number(
    starts_with('Freq'),
    decimals = 0
  ) |>
  fmt_percent(
    c(RSE, Rel_CI)
  )
```


### Pregnancies by Year and Intent

This survey will cover years 1974 to 1979.

```{r pregnancies-per-year-by-intent, dependson="create-survey-objects"}
bind_rows(list(
  before_preg |>
    group_by(Year, intent) |>
    summarize(
      n = n(),
      Freq = survey_total(one, vartype = 'se')
    ) |>
    mutate(
      RSE = Freq_se / Freq,
      post_strat = 'before'
    ) |>
    filter(Year %in% 1974:1979),
  after_preg |>
    group_by(Year, intent) |>
    summarize(
      n = n(),
      Freq = survey_total(one, vartype = 'se')
    ) |>
    mutate(
      RSE = Freq_se / Freq,
      post_strat = 'after'
    ) |>
    filter(Year %in% 1974:1979)
)) |> inner_join(
  bind_rows(list(
    before_preg |>
      group_by(Year, intent) |>
      summarize(
        n = n(),
        Freq = survey_total(one, vartype = 'ci')
      ) |>
      mutate(
        Freq_CI_range = Freq_upp - Freq_low,
        Rel_CI = Freq_CI_range / Freq,
        post_strat = 'before'
      ) |>
      filter(Year %in% 1974:1979),
    after_preg |>
      group_by(Year, intent) |>
      summarize(
        n = n(),
        Freq = survey_total(one, vartype = 'ci')
      ) |>
      mutate(
        Freq_CI_range = Freq_upp - Freq_low,
        Rel_CI = Freq_CI_range / Freq,
        post_strat = 'after'
      ) |>
      filter(Year %in% 1974:1979)
  )),
  by = join_by(Year, intent, post_strat, Freq, n)
) |>
  select(Year, intent, post_strat, everything()) |>
  mutate(
    post_strat = factor(
      post_strat,
      levels = c('before', 'after'),
      ordered = TRUE
    )
  ) |>
  arrange(Year, intent, post_strat) |>
  gt() |>
  fmt_number(
    starts_with('Freq'),
    decimals = 0
  ) |>
  fmt_percent(
    c(RSE, Rel_CI)
  )
```


### Pregnancies by Year and Outcome

This survey will cover years 1974 to 1979.

```{r pregnancies-per-year-by-outcome, dependson="create-survey-objects"}
bind_rows(list(
  before_preg |>
    group_by(Year, OUTCOME) |>
    summarize(
      n = n(),
      Freq = survey_total(one, vartype = 'se')
    ) |>
    mutate(
      RSE = Freq_se / Freq,
      post_strat = 'before'
    ) |>
    filter(Year %in% 1974:1979),
  after_preg |>
    group_by(Year, OUTCOME) |>
    summarize(
      n = n(),
      Freq = survey_total(one, vartype = 'se')
    ) |>
    mutate(
      RSE = Freq_se / Freq,
      post_strat = 'after'
    ) |>
    filter(Year %in% 1974:1979)
)) |> inner_join(
  bind_rows(list(
    before_preg |>
      group_by(Year, OUTCOME) |>
      summarize(
        n = n(),
        Freq = survey_total(one, vartype = 'ci')
      ) |>
      mutate(
        Freq_CI_range = Freq_upp - Freq_low,
        Rel_CI = Freq_CI_range / Freq,
        post_strat = 'before'
      ) |>
      filter(Year %in% 1974:1979),
    after_preg |>
      group_by(Year, OUTCOME) |>
      summarize(
        n = n(),
        Freq = survey_total(one, vartype = 'ci')
      ) |>
      mutate(
        Freq_CI_range = Freq_upp - Freq_low,
        Rel_CI = Freq_CI_range / Freq,
        post_strat = 'after'
      ) |>
      filter(Year %in% 1974:1979)
  )),
  by = join_by(Year, OUTCOME, post_strat, Freq, n)
) |>
  select(Year, OUTCOME, post_strat, everything()) |>
  mutate(
    post_strat = factor(
      post_strat,
      levels = c('before', 'after'),
      ordered = TRUE
    )
  ) |>
  arrange(Year, OUTCOME, post_strat) |>
  gt() |>
  fmt_number(
    starts_with('Freq'),
    decimals = 0
  ) |>
  fmt_percent(
    c(RSE, Rel_CI)
  )
```


### Births by Year 

This survey will cover years 1974 to 1979.

```{r births-by-year, dependson=c("create-post-stratification-targets", "create-survey-objects")}
readRDS(here('data/NVSS/natality_by_age_1974-1994us.Rds')) |>
  filter(Year %in% 1974:1980) |>
  group_by(Year) |>
  summarize(Births = sum(n)) |>
  left_join(
    before_preg |>
      filter(Year %in% 1974:1980) |>
      group_by(Year) |>
      summarize(
        n = unweighted(sum(births)),
        Est = survey_total(births, vartype = 'ci')
      ) |>
      rename(
        Est_Before = Est,
        Est_Before_low = Est_low,
        Est_Before_upp = Est_upp
      ),
    by = 'Year'
  ) |>
  left_join(
    after_preg |>
      group_by(Year) |>
      summarize(
        n = unweighted(sum(births)),
        Est = survey_total(births, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1980) |>
      rename(
        Est_After = Est,
        Est_After_low = Est_low,
        Est_After_upp = Est_upp
      ),
    by = 'Year',
    suffix = c('.before', '.after')
  ) |>
  mutate(
    unequal_n = n.before != n.after,
    Bias_Before = Est_Before - Births,
    Rel_Bias_Before = Bias_Before / Births,
    `Covers_Before?` = Births <= Est_Before_upp & Births >= Est_Before_low,
    Bias_After = Est_After - Births,
    Rel_Bias_After = Bias_After / Births,
    `Covers_After?` = Births <= Est_After_upp & Births >= Est_After_low,
    Bias_Diff = abs(Bias_After) - abs(Bias_Before),
    Est_Diff = Est_After - Est_Before,
    Rel_Bias_Diff = abs(Rel_Bias_After) - abs(Rel_Bias_Before),
    `Fixed?` = if_else(
        !`Covers_Before?`,
        `Covers_After?`,
        NA
    ),
    `Broke?` = `Covers_Before?` & !`Covers_After?`
  ) |>
  rename(
    n = n.after
  ) |>
  select(
    Year,
    n,
    Births,
    Est_Before,
    #Est_Before_low,
    #Est_Before_upp,
    Est_After,
    Est_After_low,
    Est_After_upp,
    #Est_Diff,
    #Bias_Before,
    #Bias_After,
    #Bias_Diff,
    Rel_Bias_Before,
    Rel_Bias_After,
    #Rel_Bias_Diff,
    #`Unc_Covers?`,
    #`Cal_Covers?`,
    `Fixed?`,
    `Broke?`
  ) |>
  gt() |>
  fmt_number(-Year, decimal = 0) |>
  fmt_percent(contains('Rel_Bias'), decimal = 1)
```


### Births by Year by Age

This survey will cover years 1974 to 1979.

```{r births-by-year-by-age, dependson=c("create-post-stratification-targets", "create-survey-objects")}
readRDS(here('data/NVSS/natality_by_age_1974-1994us.Rds')) |>
  filter(Year %in% 1974:1980) |>
  mutate(
    age_group6 = forcats::fct_recode(
      mage8,
      `Under 20 years` = 'Under 15 years',
      `Under 20 years` = '15-19 years',
      `40 years and over` = '40-44 years',
      `40 years and over` = '45-49 years'
    ),
    .keep = 'unused',
    .before = 'n'
  ) |>
  group_by(Year, age_group6) |>
  summarize(Births = sum(n), .groups = 'drop') |>
  left_join(
    before_preg |>
      group_by(Year, age_group6) |>
      summarize(
        n = unweighted(sum(births)),
        Est = survey_total(births, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1980) |>
      rename(
        Est_Before = Est,
        Est_Before_low = Est_low,
        Est_Before_upp = Est_upp
      ),
    by = c('Year', 'age_group6')
  ) |>
  left_join(
    after_preg |>
      group_by(Year, age_group6) |>
      summarize(
        n = unweighted(sum(births)),
        Est = survey_total(births, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1980) |>
      rename(
        Est_After = Est,
        Est_After_low = Est_low,
        Est_After_upp = Est_upp
      ),
    by = c('Year', 'age_group6'),
    suffix = c('.before', '.after')
  ) |>
  mutate(
    unequal_n = n.before != n.after,
    Bias_Before = Est_Before - Births,
    Rel_Bias_Before = Bias_Before / Births,
    `Covers_Before?` = Births <= Est_Before_upp & Births >= Est_Before_low,
    Bias_After = Est_After - Births,
    Rel_Bias_After = Bias_After / Births,
    `Covers_After?` = Births <= Est_After_upp & Births >= Est_After_low,
    Bias_Diff = abs(Bias_After) - abs(Bias_Before),
    Est_Diff = Est_After - Est_Before,
    Rel_Bias_Diff = abs(Rel_Bias_After) - abs(Rel_Bias_Before),
    `Fixed?` = if_else(
        !`Covers_Before?`,
        `Covers_After?`,
        NA
    ),
    `Broke?` = `Covers_Before?` & !`Covers_After?`
  ) |>
  rename(
    n = n.after
  ) |>
  select(
    Year,
    age_group6,
    n,
    Births,
    Est_Before,
    #Est_Before_low,
    #Est_Before_upp,
    Est_After,
    Est_After_low,
    Est_After_upp,
    #Est_Diff,
    #Bias_Before,
    #Bias_After,
    #Bias_Diff,
    Rel_Bias_Before,
    Rel_Bias_After,
    #Rel_Bias_Diff,
    #`Unc_Covers?`,
    #`Cal_Covers?`,
    `Fixed?`,
    `Broke?`
  ) |>
  group_by(Year) |>
  gt() |>
  fmt_number(-Year, decimal = 0) |>
  fmt_percent(contains('Rel_Bias'), decimal = 1)
```


### Abortions

This survey will cover years 1974 to 1979.

```{r abortions, dependson=c("create-post-stratification-targets", "create-survey-objects")}
guttmacher_APC_national |>
  filter(year %in% 1974:1979) |>
  summarize(Abortions = sum(abortionstotal)) |>
  cbind(
    before_preg |>
      group_by(Year) |>
      summarize(
        n = unweighted(sum(abortion)),
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1979) |>
      select(-Year) |>
      summarize(across(everything(), sum)) |>
      rename(
        Est_Before = Est,
        Est_Before_low = Est_low,
        Est_Before_upp = Est_upp
      )
  ) |>
  cbind(
    after_preg |>
      group_by(Year) |>
      summarize(
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1979) |>
      select(-Year) |>
      summarize(across(everything(), sum)) |>
      rename(
        Est_After = Est,
        Est_After_low = Est_low,
        Est_After_upp = Est_upp
      )
  ) |>
  mutate(
    Bias_Before = Est_Before - Abortions,
    Rel_Bias_Before = Bias_Before / Abortions,
    `Covers_Before?` = Abortions <= Est_Before_upp & Abortions >= Est_Before_low,
    Bias_After = Est_After - Abortions,
    Rel_Bias_After = Bias_After / Abortions,
    `Covers_After?` = Abortions <= Est_After_upp & Abortions >= Est_After_low,
    Bias_Diff = abs(Bias_After) - abs(Bias_Before),
    Est_Diff = Est_After - Est_Before,
    Rel_Bias_Diff = abs(Rel_Bias_After) - abs(Rel_Bias_Before),
    `Fixed?` = if_else(
        !`Covers_Before?`,
        `Covers_After?`,
        NA
    ),
    `Broke?` = `Covers_Before?` & !`Covers_After?`
  ) |>
  select(
    n,
    Abortions,
    Est_Before,
    #Est_Before_low,
    #Est_Before_upp,
    Est_After,
    Est_After_low,
    Est_After_upp,
    #Est_Diff,
    #Bias_Before,
    #Bias_After,
    #Bias_Diff,
    Rel_Bias_Before,
    Rel_Bias_After,
    #Rel_Bias_Diff,
    #`Unc_Covers?`,
    #`Cal_Covers?`,
    `Fixed?`,
    `Broke?`
  ) |>
  gt() |>
  fmt_percent(contains('Rel_Bias'), decimal = 1)
```


### Abortions by Age

This survey will cover years 1974 to 1979.

```{r abortions-by-age, dependson=c("create-post-stratification-targets", "create-survey-objects")}
guttmacher_APC_national |>
  filter(year %in% 1974:1979) |>
  select(-abortionstotal) |>
  summarize(across(starts_with('abortions'), sum)) |>
  rename(
    `Under 20 years` = abortionslt20,
    `20-24 years` = abortions2024,
    `25-29 years` = abortions2529,
    `30-34 years` = abortions3034,
    `35-39 years` = abortions3539,
    `40 years and over` = abortions40plus
  ) |>
  pivot_longer(
    cols = everything(),
    names_to = 'age_group6',
    values_to = 'Abortions'
  ) |>
  inner_join(
    before_preg |>
      group_by(age_group6, Year) |>
      summarize(
        n = unweighted(sum(abortion)),
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1979) |>
      select(-Year) |>
      summarize(across(c('n', starts_with('Est')), sum)) |>
      rename(
        Est_Before = Est,
        Est_Before_low = Est_low,
        Est_Before_upp = Est_upp
      ),
    by = join_by(age_group6)
  ) |>
  inner_join(
    after_preg |>
      group_by(age_group6, Year) |>
      summarize(
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1979) |>
      select(-Year) |>
      summarize(across(starts_with('Est'), sum)) |>
      rename(
        Est_After = Est,
        Est_After_low = Est_low,
        Est_After_upp = Est_upp
      ),
    by = join_by(age_group6)
  ) |>
  mutate(
    Bias_Before = Est_Before - Abortions,
    Rel_Bias_Before = Bias_Before / Abortions,
    `Covers_Before?` = Abortions <= Est_Before_upp & Abortions >= Est_Before_low,
    Bias_After = Est_After - Abortions,
    Rel_Bias_After = Bias_After / Abortions,
    `Covers_After?` = Abortions <= Est_After_upp & Abortions >= Est_After_low,
    Bias_Diff = abs(Bias_After) - abs(Bias_Before),
    Est_Diff = Est_After - Est_Before,
    Rel_Bias_Diff = abs(Rel_Bias_After) - abs(Rel_Bias_Before),
    `Fixed?` = if_else(
        !`Covers_Before?`,
        `Covers_After?`,
        NA
    ),
    `Broke?` = `Covers_Before?` & !`Covers_After?`
  ) |>
  select(
    age_group6,
    n,
    Abortions,
    Est_Before,
    #Est_Before_low,
    #Est_Before_upp,
    Est_After,
    Est_After_low,
    Est_After_upp,
    #Est_Diff,
    #Bias_Before,
    #Bias_After,
    #Bias_Diff,
    Rel_Bias_Before,
    Rel_Bias_After,
    #Rel_Bias_Diff,
    #`Unc_Covers?`,
    #`Cal_Covers?`,
    `Fixed?`,
    `Broke?`
  ) |>
  gt() |>
  fmt_percent(contains('Rel_Bias'), decimal = 1)
```


### Abortions by Year

This survey will cover years 1974 to 1979.

```{r abortions-by-year, dependson=c("create-post-stratification-targets", "create-survey-objects")}
guttmacher_APC_national |>
  filter(year %in% 1974:1980) |>
  rename(Year = year, Abortions = abortionstotal) |>
  select(Year, Abortions) |>
  left_join(
    before_preg |>
      group_by(Year) |>
      summarize(
        n = unweighted(sum(abortion)),
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1980) |>
      rename(
        Est_Before = Est,
        Est_Before_low = Est_low,
        Est_Before_upp = Est_upp
      ),
    by = 'Year'
  ) |>
  left_join(
    after_preg |>
      group_by(Year) |>
      summarize(
        n = unweighted(sum(abortion)),
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1980) |>
      rename(
        Est_After = Est,
        Est_After_low = Est_low,
        Est_After_upp = Est_upp
      ),
    by = 'Year',
    suffix = c('.before', '.after')
  ) |>
  mutate(
    unequal_n = n.before != n.after,
    Bias_Before = Est_Before - Abortions,
    Rel_Bias_Before = Bias_Before / Abortions,
    `Covers_Before?` = Abortions <= Est_Before_upp & Abortions >= Est_Before_low,
    Bias_After = Est_After - Abortions,
    Rel_Bias_After = Bias_After / Abortions,
    `Covers_After?` = Abortions <= Est_After_upp & Abortions >= Est_After_low,
    Bias_Diff = abs(Bias_After) - abs(Bias_Before),
    Est_Diff = Est_After - Est_Before,
    Rel_Bias_Diff = abs(Rel_Bias_After) - abs(Rel_Bias_Before),
    `Fixed?` = if_else(
        !`Covers_Before?`,
        `Covers_After?`,
        NA
    ),
    `Broke?` = `Covers_Before?` & !`Covers_After?`
  ) |>
  rename(
    n = n.after
  ) |>
  select(
    Year,
    n,
    Abortions,
    Est_Before,
    #Est_Before_low,
    #Est_Before_upp,
    Est_After,
    Est_After_low,
    Est_After_upp,
    #Est_Diff,
    #Bias_Before,
    #Bias_After,
    #Bias_Diff,
    Rel_Bias_Before,
    Rel_Bias_After,
    #Rel_Bias_Diff,
    #`Unc_Covers?`,
    #`Cal_Covers?`,
    `Fixed?`,
    `Broke?`
  ) |>
  gt() |>
  fmt_number(-Year, decimal = 0) |>
  fmt_percent(contains('Rel_Bias'), decimal = 1)
```


### Abortions by Year by Age

This survey will cover years 1974 to 1979.

```{r abortions-by-year-by-age, dependson=c("create-post-stratification-targets", "create-survey-objects")}
guttmacher_APC_national |>
  filter(year %in% 1974:1980) |>
  rename(
    Year = year,
    `Under 20 years` = abortionslt20,
    `20-24 years` = abortions2024,
    `25-29 years` = abortions2529,
    `30-34 years` = abortions3034,
    `35-39 years` = abortions3539,
    `40 years and over` = abortions40plus
  ) |>
  pivot_longer(
    cols = contains('years'),
    names_to = 'Level',
    values_to = 'Abortions'
  ) |>
  select(Year, Level, Abortions) |>
  left_join(
    before_preg |>
      group_by(Year, age_group6) |>
      summarize(
        n = unweighted(sum(abortion)),
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1980) |>
      rename(
        Level = age_group6,
        Est_Before = Est,
        Est_Before_low = Est_low,
        Est_Before_upp = Est_upp
      ),
    by = c('Year', 'Level')
  ) |>
  left_join(
    after_preg |>
      group_by(Year, age_group6) |>
      summarize(
        n = unweighted(sum(abortion)),
        Est = survey_total(abortion, vartype = 'ci')
      ) |>
      filter(Year %in% 1974:1980) |>
      rename(
        Level = age_group6,
        Est_After = Est,
        Est_After_low = Est_low,
        Est_After_upp = Est_upp
      ),
    by = c('Year', 'Level'),
    suffix = c('.before', '.after')
  ) |>
  mutate(
    unequal_n = n.before != n.after,
    Bias_Before = Est_Before - Abortions,
    Rel_Bias_Before = Bias_Before / Abortions,
    `Covers_Before?` = Abortions <= Est_Before_upp & Abortions >= Est_Before_low,
    Bias_After = Est_After - Abortions,
    Rel_Bias_After = Bias_After / Abortions,
    `Covers_After?` = Abortions <= Est_After_upp & Abortions >= Est_After_low,
    Bias_Diff = abs(Bias_After) - abs(Bias_Before),
    Est_Diff = Est_After - Est_Before,
    Rel_Bias_Diff = abs(Rel_Bias_After) - abs(Rel_Bias_Before),
    `Fixed?` = if_else(
        !`Covers_Before?`,
        `Covers_After?`,
        NA
    ),
    `Broke?` = `Covers_Before?` & !`Covers_After?`
  ) |>
  rename(
    n = n.after
  ) |>
  select(
    Year,
    Level,
    n,
    Abortions,
    Est_Before,
    #Est_Before_low,
    #Est_Before_upp,
    Est_After,
    Est_After_low,
    Est_After_upp,
    #Est_Diff,
    #Bias_Before,
    #Bias_After,
    #Bias_Diff,
    Rel_Bias_Before,
    Rel_Bias_After,
    #Rel_Bias_Diff,
    #`Unc_Covers?`,
    #`Cal_Covers?`,
    `Fixed?`,
    `Broke?`
  ) |>
  group_by(Year) |>
  gt() |>
  fmt_number(-Year, decimal = 0) |>
  fmt_percent(contains('Rel_Bias'), decimal = 1)
```
